---
title: "8 Conditional Manatees"
format: html
---

```{r}
rm(list = ls())
library(tidyverse)
library(brms)
library(tidybayes)
library(ggthemes)
library(ggdag)
library(ggrepel)
library(patchwork)
```

## 8.1 Building an interaction
```{r}
data(rugged, package = "rethinking")
d <- rugged
rm(rugged)
```

```{r}
theme_set(
  theme_pander() +
    theme(text = element_text(family = "Times"),
          panel.background = element_rect(color = "black"))
)
```

```{r}
dag_coords <-
  tibble(name = c("R", "G", "C", "U"),
         x = c(1, 2, 3, 2),
         y = c(2, 2, 2, 1))

dagify(R ~ U,
       G ~ R + U + C,
       coords = dag_coords) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 alpha = 1/2, size = 6, show.legend = F) +
  geom_point(x = 2, y = 1, size = 6, shape = 1, 
             stroke = 3/4, color = palette_pander(n = 2)[2]) +
  geom_dag_text(color = "black", family = "Times") +
  geom_dag_edges() +
  scale_color_pander() +
  theme_dag()
```

#### 8.1.0.1 Overthinking: Not so simple causation
```{r}
dag_coords <-
  tibble(name = c("G", "R", "H", "C", "U"),
         x    = c(1, 1.5, 2.5, 3.5, 1),
         y    = c(3, 2, 2, 2, 1))

dagify(G ~ R + U + H,
       R ~ U,
       H ~ R + U + C,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 alpha = 1/2, size = 6, show.legend = F) +
  geom_point(x = 1, y = 1, 
             size = 6, shape = 1, stroke = 3/4, color = palette_pander(n = 2)[2]) +
  geom_dag_text(color = "black", family = "Times") +
  geom_dag_edges() +
  scale_colour_pander() +
  theme_dag()
```

### 8.1.1 Making a rugged model
```{r}
# make the log version of criterion
d <-
  d %>% 
  mutate(log_gdp = log(rgdppc_2000))

# extract countries with GDP data
dd <-
  d %>% 
  filter(complete.cases(rgdppc_2000)) %>% 
  # re-scale variables
  mutate(log_gdp_std = log_gdp / mean(log_gdp),
         rugged_std = rugged / max(rugged))
```

```{r}
# African nations
p1 <-
  dd %>% 
  filter(cont_africa == 1) %>% 
  ggplot(aes(x = rugged_std, y = log_gdp_std)) +
  geom_smooth(method = "lm", formula = y ~ x,
              fill = palette_pander(n = 2)[1],
              color = palette_pander(n = 2)[1]) +
  geom_point(color = palette_pander(n = 2)[1]) +
  geom_text_repel(data = . %>% 
                    filter(country %in% c("Lesotho", "Seychelles")),
                  aes(label = country),
                  size = 3, family = "Times", seed = 8) +
  labs(subtitle = "African nations",
       x = "ruggedness (standardized)",
       y = "log GDP (as proportion of the mean)")

# Non_African nations
p2 <-
  dd %>% 
  filter(cont_africa == 0) %>% 
  ggplot(aes(x = rugged_std, y = log_gdp_std)) +
  geom_smooth(method = "lm", formula = y ~ x,
              fill = palette_pander(n = 2)[2],
              color = palette_pander(n = 2)[2]) +
  geom_point(color = palette_pander(n = 2)[2]) +
  geom_text_repel(data = . %>% 
                    filter(country %in% c("Switzerland", "Tajikistan")),  
                  aes(label = country), 
                  size = 3, family = "Times", seed = 8) +
  xlim(0, 1) +
  labs(subtitle = "Non-African nations",
       x = "ruggedness (standardized)",
       y = "log GDP (as proportion of mean)")

# combine
p1 + p2 + plot_annotation(title = "Figure 8.2. Separate linear regressions inside and outside of Africa")
```

Our first Bayesian model will follow the form and we compute rugged_std
```{r}
mean(dd$rugged_std)
```

```{r}
dd <-
  dd %>% 
  mutate(rugged_std_c = rugged_std - mean(rugged_std))
```

```{r}
b8.1 <-
  brm(data = dd,
      family = gaussian,
      log_gdp_std ~ 1 + rugged_std_c,
      prior = c(prior(normal(1, 1), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      sample_prior = T,
      file = "fits/b08.01")
```

```{r}
prior <- prior_draws(b8.1)

set.seed(8)

p1 <-
  prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column() %>% 
  expand_grid(rugged_std_c = c(-2, 2)) %>% 
  mutate(log_gdp_std = Intercept + b * rugged_std_c,
         rugged_std = rugged_std_c + mean(dd$rugged_std)) %>% 
  
  ggplot(aes(x = rugged_std, y = log_gdp_std, group = rowname)) +
  geom_hline(yintercept = range(dd$log_gdp_std), linetype = 2) +
  geom_line(color = palette_pander(n = 2)[2], alpha = .4) +
  geom_abline(intercept = 1.3, slope = -0.6,
              color = palette_pander(n = 2)[1], linewidth = 2) +
  labs(subtitle = "Intercept ~ dnorm(1, 1)\nb ~ dnorm(0, 1)",
       x = "ruggedness",
       y = "log GDP (prop of mean)") +
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(0.5, 1.5))

p1
```

```{r}
prior %>% 
  summarise(a = sum(abs(b) > abs(-0.6)) / nrow(prior))
```

```{r}
b8.1b <-
  brm(data = dd,
      family = gaussian,
      log_gdp_std ~ 1 + rugged_std_c,
      prior = c(prior(normal(1, .1), class = Intercept),
                prior(normal(0, .3), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      sample_prior = T,
      file = "fits/b08.01b")
```

```{r}
set.seed(8)

p2 <-
  prior_draws(b8.1b) %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column() %>% 
  expand_grid(rugged_std_c = c(-2, 2)) %>% 
  mutate(log_gdp_std = Intercept + b * rugged_std_c,
         rugged_std = rugged_std_c + mean(dd$rugged_std)) %>% 
  
  ggplot(aes(x = rugged_std, y = log_gdp_std, group = rowname)) +
  geom_hline(yintercept = range(dd$log_gdp_std), linetype = 2) +
  geom_line(color = palette_pander(n = 2)[2], alpha = .4) +
  scale_y_continuous("", breaks = NULL) +
  labs(subtitle = "Intercept ~ dnorm(1, 0.1)\nb~ dnorm(0, 0.3)",
       x = "ruggedness") +
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(.5, 1.5))

p1 + p2 +
  plot_annotation(title = "Simulating in search of reasonable priors for the terrain ruggedness example.",
                  theme = theme(plot.title = element_text(size = 12)))
```

```{r}
print(b8.1b)
```

### 8.1.2 Adding an indicator variable isn't enough
```{r}
dd <-
  dd %>% 
  mutate(cid = if_else(cont_africa == 1, "1", "2"))
```

```{r}
dd %>% 
  mutate(cid = str_c("cid: ", cid)) %>% 
  arrange(cid, country) %>% 
  group_by(cid) %>% 
  mutate(rank = 1:n()) %>% 
  
  ggplot(aes(x = cid, y = rank, label = country)) +
  geom_text(size = 2, hjust = 0, family = "Times") +
  scale_y_reverse() +
  theme_void() +
  facet_wrap(~ cid, scales = "free_x")
```

```{r}
b8.2 <-
  brm(data = dd,
      family = gaussian,
      log_gdp_std ~ 0 + cid + rugged_std_c,
      prior = c(prior(normal(1, .1), class = b, coef = cid1),
                prior(normal(1, .1), class = b, coef = cid2),
                prior(normal(0, .3), class = b, coef = rugged_std_c),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      file = "fits/b08.02")
```

```{r}
b8.1b <- add_criterion(b8.1b, "waic")
b8.2 <- add_criterion(b8.2, "waic")

loo_compare(b8.1b, b8.2, criterion = "waic") %>% print(simplify = F)
```

```{r}
model_weights(b8.1b, b8.2, weights = "waic") %>% round(digits = 3)
```

```{r}
print(b8.2)
```
Now extract the posterior draws, make a difference score for the two intercepts, and use tidybayes::qi() to compute the percentile-based 89% intervals for the difference.
```{r}
post <-
  as_draws_df(b8.2) %>% 
  mutate(diff = b_cid1 - b_cid2)

qi(post$diff, .width = .89)
```

```{r}
nd <-
  crossing(cid = 1:2,
           rugged_std = seq(from = -.2, to = 1.2, length.out = 30)) %>% 
  mutate(rugged_std_c = rugged_std - mean(dd$rugged_std))

f <-
  fitted(b8.2,
         newdata = nd,
         probs = c(.015, .985)) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  mutate(cont_africa = ifelse(cid == 1, "Africa", "not Africa"))

# what did we do?
head(f)
```

```{r}
dd %>%
  mutate(cont_africa = ifelse(cont_africa == 1, "Africa", "not Africa")) %>%
  
  ggplot(aes(x = rugged_std, fill = cont_africa, color = cont_africa)) +
  geom_smooth(data = f,
              aes(y = Estimate, ymin = Q1.5, ymax = Q98.5),
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(aes(y = log_gdp_std),
             size = 2/3) +
  scale_fill_pander() +
  scale_colour_pander() +
  labs(subtitle = "b8.2",
       x = "ruggedness (standardized)",
       y = "log GDP (as proportion of mean)") +
  coord_cartesian(xlim = c(0, 1)) +
  theme(legend.background = element_blank(),
        legend.direction = "horizontal",
        legend.position = c(.67, .93),
        legend.title = element_blank())
```

#### 8.1.2.1 Rethinking: Why 97%?
```{r}
fitted(b8.2, 
       newdata = nd,
       summary = F) %>%
  data.frame() %>%
  pivot_longer(everything()) %>% 
  bind_cols(expand_grid(draws = 1:4000, nd)) %>%
  mutate(cont_africa = ifelse(cid == 1, "Africa", "not Africa")) %>% 
  
  ggplot(aes(x = rugged_std, y = value, fill = cont_africa, color = cont_africa)) +
  stat_lineribbon(.width = seq(from = .03, to = .99, by = .03),
                  alpha = .1, size = 0) +
  geom_point(data = dd %>% 
               mutate(cont_africa = ifelse(cont_africa == 1, "Africa", "not Africa")),
             aes(y = log_gdp_std),
             size = 2/3) +
  scale_fill_pander() +
  scale_colour_pander() +
  labs(subtitle = "b8.2",
       x = "ruggedness (standardized)",
       y = "log GDP (as proportion of mean)") +
  coord_cartesian(xlim = c(0, 1)) +
  theme(legend.background = element_blank(),
        legend.direction = "horizontal",
        legend.position = c(.67, .93),
        legend.title = element_blank())
```

### 8.1.3 Adding an interaction does work.
```{r}
b8.3 <- 
  brm(data = dd, 
      family = gaussian,
      bf(log_gdp_std ~ 0 + a + b * rugged_std_c, 
         a ~ 0 + cid, 
         b ~ 0 + cid,
         nl = TRUE),
      prior = c(prior(normal(1, 0.1), class = b, coef = cid1, nlpar = a),
                prior(normal(1, 0.1), class = b, coef = cid2, nlpar = a),
                prior(normal(0, 0.3), class = b, coef = cid1, nlpar = b),
                prior(normal(0, 0.3), class = b, coef = cid2, nlpar = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      file = "fits/b08.03")
```

```{r}
print(b8.3)
```

```{r}
b8.1b <- add_criterion(b8.1b, "loo")
b8.2 <- add_criterion(b8.2, "loo")
b8.3 <- add_criterion(b8.3, c("loo", "waic"))

loo_compare(b8.1b, b8.2, b8.3, criterion = "loo") %>% print(simplify = F)
```

```{r}
model_weights(b8.1b, b8.2, b8.3, weights = "loo") %>% round(digits = 2)
```

Pareto k diagnostic plot by feeding the results of the loo() function into plot()
```{r}
loo(b8.3) %>% plot()
```

Rank order the Pareto k value and taking a look at the top values
```{r}
tibble(k = b8.3$criteria$loo$diagnostics$pareto_k,
       row = 1:170) %>% 
  arrange(desc(k))
```

#### 8.1.3.1 Bonus: Give me the Student-t.
```{r}
b8.3t <-
  brm(data = dd,
      family = student,
      bf(log_gdp_std ~ 0 + a + b * rugged_std_c,
         a ~ 0 + cid,
         b ~ 0 + cid,
         nu = 2,
         nl = TRUE),
      prior = c(prior(normal(1, 0.1), class = b, coef = cid1, nlpar = a),
                prior(normal(1, 0.1), class = b, coef = cid2, nlpar = a),
                prior(normal(0, 0.3), class = b, coef = cid1, nlpar = b),
                prior(normal(0, 0.3), class = b, coef = cid2, nlpar = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      file = "fits/b08.03t")
```

```{r}
b8.3t <- add_criterion(b8.3t, c("loo", "waic"))

loo_compare(b8.3, b8.3t, criterion = "loo") %>% print(simplify = F)
```
The PSIS-LOO-CV comparison suggests the robust Student- 
t
  model might be overfit. Just for kicks, we might make our own diagnostic plot to compare the two likelihoods by the Pareto  
k
 values. To get a nice fine-grain sense of the distributions, we’ll employ the handy tidybayes::stat_dots() function which will display each value as an individual dot.
```{r}
tibble(Normal = b8.3$criteria$loo$diagnostics$pareto_k,
       "Student-t" = b8.3t$criteria$loo$diagnostics$pareto_k) %>% 
  pivot_longer(everything(),
               values_to = "pareto_k") %>% 
  
  ggplot(aes(x = pareto_k, y = name)) +
  geom_vline(xintercept = .5, linetype = 2, color = palette_pander(n = 5)[5]) +
  stat_dots(slab_fill = palette_pander(n = 4)[4],
            slab_color = palette_pander(n = 4)[4]) +
  annotate(geom = "text",
           x = .485, y = 1.5, label = "threshold", angle = 90,
           family = "Times", color = palette_pander(n = 5)[5]) +
  ylab(NULL) +
  coord_cartesian(ylim = c(1.5, 2.4))
```

```{r}
fixef(b8.3) %>% round(digits = 2)
```

```{r}
fixef(b8.3t) %>% round(digits = 2)
```

### 8.1.4 Plotting the interaction
```{r}
countries <- c("Equatorial Guinea", "South Africa", "Seychelles", "Swaziland", "Lesotho", "Rwanda", "Burundi", "Luxembourg", "Greece", "Switzerland", "Lebanon", "Yemen", "Tajikistan", "Nepal")

f <-
  fitted(b8.3,
         # we already defined 'nd', above
         newdata = nd,
         probs = c(.015, .985)) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  mutate(cont_africa = ifelse(cid == 1, "African nations", "Non-African nations"))

dd %>% 
  mutate(cont_africa = ifelse(cont_africa == 1, "African nations", "Non-African nations")) %>% 
  
  ggplot(aes(x = rugged_std, y = log_gdp_std, fill = cont_africa, color = cont_africa)) +
  geom_smooth(data = f,
              aes(y = Estimate, ymin = Q1.5, ymax = Q98.5),
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_text_repel(data = . %>% filter(country %in% countries),
                  aes(label = country),
                  size = 3, seed = 8,
                  segment.color = "grey25", min.segment.length = 0) +
  geom_point(aes(y = log_gdp_std),
             size = 2/3) +
  scale_fill_pander() +
  scale_color_pander() +
  labs(x = "ruggedness (standardized)",
       y = "log GDP (as proportion of mean)") +
  coord_cartesian(xlim = c(0, 1)) +
  theme(legend.position = "none") +
  facet_wrap(~ cont_africa)
```

## 8.2 Symmetry of interactions
```{r}
fitted(b8.3, 
       newdata = nd,
       summary = F) %>%
  data.frame() %>%
  pivot_longer(everything()) %>% 
  bind_cols(expand_grid(draws = 1:4000, nd)) %>% 
  select(-name) %>% 
  pivot_wider(names_from = cid, values_from = value) %>% 
  mutate(delta = `1` - `2`) %>% 
  
  ggplot(aes(x = rugged_std, y = delta)) +
  stat_lineribbon(.width = .95, fill = palette_pander(n = 8)[8], alpha = 3/4) +
  geom_hline(yintercept = 0, linetype = 2) +
  annotate(geom = "text",
           x = .2, y = 0,
           label = "Africa higher GDP\nAfrica lower GDP",
           family = "Times") +
  labs(x = "ruggedness (standardized)",
       y = "expected difference log GDP") +
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(-.3, .2))
```

## 8.3 Continuous interactions
### 8.3.1 A winter flower
```{r}
data(tulips, package = "rethinking")
d <- tulips
rm(tulips)

glimpse(d)
```

### 8.3.2 The models
```{r}
d <-
  d %>% 
  mutate(blooms_std = blooms / max(blooms),
         water_cent = water - mean(water),
         shade_cent = shade - mean(shade))
```

```{r}
set.seed(8)

tibble(a = rnorm(1e4, mean = .5, sd = 1)) %>% 
  summarize(proportion_outside_of_the_range = sum(a < 0 | a > 1) / n())
```

```{r}
set.seed(8)

tibble(a = rnorm(1e4, mean = 0.5, sd = 0.25)) %>% 
  summarise(proportion_outside_of_the_range = sum(a < 0 | a > 1) / n())
```

```{r}
range(d$water_cent)
```

```{r}
range(d$shade_cent)
```

```{r}
b8.4 <-
  brm(data = d,
      family = gaussian,
      blooms_std ~ 1 + water_cent + shade_cent,
      prior = c(prior(normal(0.5, 0.25), class = Intercept),
                prior(normal(0, 0.25), class = b, coef = water_cent),
                prior(normal(0, 0.25), class = b, coef = shade_cent),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      file = "fits/b08.04")
```

```{r}
print(b8.4)
```

```{r}
b8.5 <-
  brm(data = d, 
      family = gaussian,
      blooms_std ~ 1 + water_cent + shade_cent + water_cent:shade_cent,
      prior = c(prior(normal(0.5, 0.25), class = Intercept),
                prior(normal(0, 0.25), class = b, coef = water_cent),
                prior(normal(0, 0.25), class = b, coef = shade_cent),
                prior(normal(0, 0.25), class = b, coef = "water_cent:shade_cent"),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8,
      file = "fits/b08.05")
```

```{r}
print(b8.5)
```

```{r}
# loop over values of `water_c` and plot predictions
for(s in -1:1) {
  
  # define the subset of the original data
  dt <- d[d$shade_cent == s, ]
  # defining our new data
  nd <- tibble(shade_cent = s, water_cent = c(-1, 1))
  # use our sampling skills, like before
  f <- 
    fitted(b8.4, 
           newdata = nd,
           summary = F) %>%
    data.frame() %>%
    set_names("-1", "1") %>% 
    slice_sample(n = 20) %>% 
    mutate(row = 1:n()) %>% 
    pivot_longer(-row,
                 names_to = "water_cent",
                 values_to = "blooms_std") %>% 
    mutate(water_cent = as.double(water_cent))
  
  # specify our custom plot
  fig <- 
    ggplot(data = dt,
           aes(x = water_cent, y = blooms_std)) +
    geom_line(data = f,
              aes(group = row),
              color = palette_pander(n = 6)[6], alpha = 1/5, linewidth = 1/2) +
    geom_point(color = palette_pander(n = 6)[6]) +
    scale_x_continuous("Water (centered)", breaks = c(-1, 0, 1)) +
    labs(title = paste("Shade (centered) =", s),
         y = "Blooms (standardized)") +
    coord_cartesian(xlim = c(-1, 1), 
                    ylim = c(0, 1))
  
  # plot that joint
  plot(fig)
  
}
```

```{r}
# augment the data
points <-
  d %>%
  expand_grid(fit = c("b8.4", "b8.5")) %>%
  mutate(x_grid = str_c("shade_cent = ", shade_cent),
         y_grid = fit)

# redefine `nd`
nd <- crossing(shade_cent = -1:1, 
               water_cent = c(-1, 1))

# use `fitted()`
set.seed(8)

rbind(fitted(b8.4, newdata = nd, summary = F, ndraws = 20),
      fitted(b8.5, newdata = nd, summary = F, ndraws = 20)) %>%
  # wrangle
  data.frame() %>%
  set_names(mutate(nd, name = str_c(shade_cent, water_cent, sep = "_")) %>% pull()) %>%
  mutate(row = 1:n(),
         fit = rep(c("b8.4", "b8.5"), each = n() / 2)) %>%
  pivot_longer(-c(row:fit), values_to = "blooms_std") %>%
  separate(name, into = c("shade_cent", "water_cent"), sep = "_") %>%
  mutate(shade_cent = shade_cent %>% as.double(),
         water_cent = water_cent %>% as.double()) %>%
  # these will come in handy for `ggplot2::facet_grid()`
  mutate(x_grid = str_c("shade_cent = ", shade_cent),
         y_grid = fit) %>%
  
  # plot!
  ggplot(aes(x = water_cent, y = blooms_std)) +
  geom_line(aes(group = row),
            color = palette_pander(n = 6)[6], alpha = 1/5, linewidth = 1/2) +
  geom_point(data = points,
             color = palette_pander(n = 6)[6]) +
  scale_x_continuous("Water (centered)", breaks = c(-1, 0, 1)) +
  scale_y_continuous("Blooms (standardized)", breaks = c(0, .5, 1)) +
  ggtitle("Posterior predicted blooms") +
  coord_cartesian(xlim = c(-1, 1),
                  ylim = c(0, 1)) +
  theme(strip.background = element_rect(fill = alpha(palette_pander(n = 2)[2], 1/3))) +
  facet_grid(y_grid ~ x_grid)
```

