---
title: "4 Geocentric Models"
format: html
---

## 4.1. Why normal distributions are normal
### 4.1.1 Normal by addition
```{r}
library(tidyverse)

set.seed(4)

pos <-
  # make data with 100 people, 16 steps each with a starting point of step == 0
  crossing(person = 1:100,
           step = 0:16) %>% 
  # for all steps above step == 0 simulate a deviation
  mutate(deviation = map_dbl(step, 
                             ~if_else(. == 0, 0, 
                                      runif(1, -1, 1)))) %>% 
  # after grouping by person, compute the cumulative sum of the deviations, then ungroup()
  group_by(person) %>% 
  mutate(position = cumsum(deviation)) %>% 
  ungroup()
```

```{r}
glimpse(pos)
```

```{r}
ggplot(data = pos,
       aes(x = step, y = position, group = person)) +
  geom_vline(xintercept = c(4, 8, 16), linetype = 2) +
  geom_line(aes(color = person < 2, alpha = person < 2)) +
  scale_color_manual(values = c("skyblue4", "black")) +
  scale_alpha_manual(values = c(1/5, 1)) +
  scale_x_continuous("step number", breaks = 0:4 * 4) +
  theme(legend.position = "none")
```

```{r}
p1 <-
  pos %>% 
  filter(step == 4) %>% 
  ggplot(aes(x = position)) +
  geom_line(stat = "density", color = "dodgerblue1") +
  labs(title = "4 steps")

p2 <-
  pos %>% 
  filter(step == 8) %>% 
  ggplot(aes(x = position)) +
  geom_line(stat = "density", color = "dodgerblue1") +
  labs(title = "8 steps")

# this is an intermediary step to get an SD value
sd <-
  pos %>% 
  filter(step == 16) %>% 
  summarise(sd = sd(position)) %>% 
  pull(sd)

p3 <-
  pos %>% 
  filter(step == 16) %>% 
  ggplot(aes(x = position)) +
  stat_function(fun = dnorm,
                args = list(mean = 0, sd = sd),
                linetype = 2) +
  geom_density(color = "transparent", fill = "dodgerblue3",
               alpha = 1/2) +
  labs(title = "16 steps", y = "density")

library(patchwork)

( p1 | p2 | p3) & coord_cartesian(xlim = c(-6, 6))
```

### 4.1.2 Normal by multiplication
```{r}
set.seed(4)

prod(1 + runif(12, 0, 0.1))
```

A more explicit variant of the above code
```{r}
set.seed(4)

tibble(a = 1,
       b = runif(12, 0, .1)) %>% 
  mutate(c = a + b) %>% 
  summarise(p = prod(c))
```

Do this many times using purrr:map_dbl()
```{r}
set.seed(4)

growth <-
  tibble(growth = map_dbl(1:10000, 
                          ~prod(1 + runif(12, 0, .1))))

ggplot(data = growth, aes(x = growth)) + geom_density()
```
```{r}
# simulate
set.seed(4)

samples <-
  tibble(big = map_dbl(1:1e4, ~prod(1 + runif(12, 0, .5))),
         small = map_dbl(1:1e4, ~prod(1 + runif(12, 0, .01))))

# wrangle
samples %>% 
  pivot_longer(everything(), values_to = "samples") %>% 
  
  # plot
  ggplot(aes(x = samples)) +
  geom_density(fill = "transparent") +
  facet_wrap(~name, scales = "free")
  
```

### 4.1.3 Normal by log-multiplication
```{r}
samples %>% 
  mutate(log_big = log(big)) %>% 
  
  ggplot(aes(x = log_big)) +
  geom_density(fill = "gray33") +
  xlab("the log of the big")
```

### 4.1.4 Using Gaussian distributions
#### 4.1.4.1 Ontological justification
#### 4.1.4.2 Epistemological justification
#### 4.1.4.3 Rethinking: Heavy tails
#### 4.1.4.4 Overthinking: Gaussian distribution
```{r}
tibble(y = seq(from = -1, to = 1, by = .01),
       mu = 0,
       sigma = .1) %>% 
  # compute p(y) with a hand-made Gaussian probability density function
  mutate(p = (1 + sqrt(2 * pi * sigma^2)) * exp(-((y - mu)^2 / (2 * sigma^2)))) %>% 
  
  ggplot(aes(x = y, y = p)) +
  geom_line() +
  ylab("p(y | mu = 0, sigma = 0.1)")
```

## 4.2 A language for describing models
### 4.2.1 Re-describing the globe tossing model
#### 4.2.1.1 Overthinking: From model definition to Bayes' theorem
```{r}
# how many p_grid points would you like?
n_points <- 100

d <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = n_points),
         w = 6,
         n = 9) %>% 
  mutate(prior = dunif(p_grid, min = 0, max = 1),
         likelihood = dbinom(w, size = n, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))

head(d)
```

```{r}
d %>% 
  pivot_longer(prior:posterior) %>% 
  # this line allows us to dictate the order in which the panels will appear
  mutate(name = factor(name,
                       levels = c("prior", "likelihood",
                                  "posterior"))) %>% 
  
  ggplot(aes(x = p_grid, y = value, fill = name)) +
  geom_area() +
  scale_fill_manual(values = c("blue", "red", "purple")) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(legend.position = "none") +
  facet_wrap(~ name, scales = "free")
```


## 4.3 A Gaussian model of height
### 4.3.1 The data
```{r}
library(rethinking)
data(Howell1)
d <- Howell1
```

```{r}
rm(Howell1)
detach(package:rethinking, unload = T)
library(brms)
```

```{r}
d %>% str()
```

```{r}
d %>% summary()
```
Make histograms
```{r}
d %>% 
  pivot_longer(everything()) %>% 
  mutate(name = factor(name, levels = c("height", "weight",
                                        "age", "male"))) %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ name, scales = "free", ncol = 1)
```

McElreath made those tiny histograms with help from Wichkahm's histospark() function.
```{r}
sparks <- c("\u2581", "\u2582", "\u2583", "\u2585", "\u2587")

histospark <- function(x, width = 10){
  bins <- graphics::hist(x, breaks = width, plot = FALSE)
  
  factor <- cut(
    bins$counts / max(bins$counts),
    breaks = seq(0, 1, length = length(sparks) + 1),
    labels = sparks,
    include.lowest = TRUE
  )
  
  paste0(factor, collapes = "")
}
```

```{r}
histospark(d$weight)
```

```{r}
d %>% 
  select(height) %>% 
  glimpse()
```

If you want the values in a numeric vector rather than in a data fame, try pull(d, height).
```{r}
d2 <-
  d %>% 
  filter(age >= 18)
```

```{r}
d2 %>% count()
```

### 4.3.2 The model

$$
\begin{align}
\text{heights}_{i} & \sim \text{Normal}(\mu, \sigma)\\
\mu & \sim \text{Normal}(178, 20)\\
\sigma & \sim \text{Uniform}(0, 50)
\end{align}
$$


```{r}
p1 <-
  tibble(x = seq(from = 100, to = 250, by = .1)) %>% 
  ggplot(aes(x = x, y = dnorm(x, mean = 178, sd = 20))) +
  geom_line() + 
  scale_x_continuous(breaks = seq(from = 100, to = 250, by = 75)) +
  labs(title = "mu ~ dnorm(178, 20)", y = "density")

p1
```

```{r}
p2 <-
  tibble(x = seq(from = -10, to = 60, by = .1)) %>% 
  ggplot(aes(x = x, y = dunif(x, 0, 50))) +
  geom_line() +
  scale_x_continuous(breaks = c(0, 50)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("sigma ~ dunif(0, 50)")

p2
```

We can simulate from both priors at once to get a prior probability distribution of heights
```{r}
n <- 1e4

set.seed(4)

sim <-
  tibble(sample_mu = rnorm(n, mean = 178, sd = 20),
         sample_sigma = runif(n, min = 0, max = 50)) %>% 
  mutate(height = rnorm(n, mean = sample_mu, sd = sample_sigma))

p3 <-
  sim %>% 
  ggplot(aes(x = height)) +
  geom_density(fill = "grey33") +
  scale_x_continuous(breaks = c(0, 73, 178, 283)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("height ~ dnorm(mu, sigma)") +
  theme(panel.grid = element_blank())

p3
  
```

Compute the mean and 3 standard deviations above and below
```{r}
sim %>% 
  summarise(ll = mean(height) - sd(height) * 3,
            mean = mean(height),
            ul = mean(height) + sd(height) * 3) %>% 
  mutate_all(round, digits = 1)
```

Make the lower right panel of Figure 4.3
```{r}
# simulate
set.seed(4)

sim <-
  tibble(sample_mu = rnorm(n, mean = 178, sd = 100),
         sample_sigma = runif(n, min = 0, max = 50)) %>% 
  mutate(height = rnorm(n, mean = sample_mu, sd = sample_sigma))
  
# compute the values we'll use to break on our x axis
breaks <-
  c(mean(sim$height) - 3 * sd(sim$height), 
    0, 
    mean(sim$height), 
    mean(sim$height) + 3 * sd(sim$height)) %>% 
  round(digits = 0)

# this is just for aesthetics
text <-
  tibble(height = 272 - 25,
         y = .0013,
         label = "tallest man",
         angle = 90)

# plot
p4 <-
  sim %>% 
  ggplot(aes(x = height)) +
  geom_density(fill = "black", linewidth = 0) +
  geom_vline(xintercept = 0, color = "grey92") +
  geom_vline(xintercept = 272, color = "grey92", linetype = 3) +
  geom_text(data = text,
            aes(y = y, label = label, angle = angle),
            color = "grey92") +
  scale_x_continuous(breaks = breaks) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("height ~ dnorm(mu, sigma)\nmu ~ dnorm(178, 100)") +
  theme(panel.grid = element_blank())

p4
```

Combine all plots
```{r}
(p1 + xlab("mu") | p2 + xlab("sigma")) / (p3 | p4)
```
On page 84, McElreath said his prior simulation indicated 4% of the heights would be below zero. Here’s how we might determe that percentage for our simulation.
```{r}
sim %>% 
  count(height < 0) %>% 
  mutate(percent = 100 * n / sum(n))
```

Here's the breakdown compared to the tallest man on record
```{r}
sim %>%
  count(height < 272) %>% 
  mutate(percent = 100 * n / sum(n))
```

### 4.3.3 Grid approximation of the posterior distribution
```{r}
n <- 200

d_grid <-
  # we'll accomplish with tidyr::crossing() what McElreath did with base R expand.grid()
  crossing(mu = seq(from = 140, to = 160, length.out = n),
           sigma = seq(from = 4, to = 9, length.out = n))

glimpse(d_grid)
```

```{r}
grid_function <- function(mu, sigma){
  dnorm(d2$height, mean = mu, sd = sigma, log = T) %>% 
    sum()
}
```

```{r}
d_grid <-
  d_grid %>% 
  mutate(log_likelihood = map2(mu, sigma, grid_function)) %>% 
  unnest(log_likelihood) %>% 
  mutate(prior_mu = dnorm(mu, mean = 178, sd = 20, log = T),
         prior_sigma = dunif(sigma, min = 0, max = 50, log = T)) %>% 
  mutate(product = log_likelihood + prior_mu + prior_sigma) %>% 
  mutate(probability = exp(product - max(product)))

head(d_grid)
```
```{r}
d_grid %>% 
  ggplot(aes(x = mu, y = sigma, z = probability)) +
  geom_contour() +
  labs(x = expression(mu),
       y = expression(sigma)) +
  coord_cartesian(xlim = range(d_grid$mu),
                  ylim = range(d_grid$sigma)) +
  theme(panel.grid = element_blank())
```

```{r}
d_grid %>% 
  ggplot(aes(x = mu, y = sigma, fill = probability)) + 
  geom_raster(interpolate = T) +
  scale_fill_viridis_c(option = "B") +
  labs(x = expression(mu),
       y = expression(sigma)) +
  theme(panel.grid = element_blank())
```

### 4.3.4 Sampling from the posterior
```{r}
set.seed(4)

d_grid_samples <-
  d_grid %>% 
  sample_n(size = 1e4, replace = T, weight = probability)

d_grid_samples %>% 
  ggplot(aes(x = mu, y = sigma)) +
  geom_point(size = .9, alpha = 1/15) +
  scale_fill_viridis_c() +
  labs(x = expression(mu),
       y = expression(sigma)) +
  theme(panel.grid = element_blank())
```

```{r}
d_grid_samples %>% 
  pivot_longer(mu:sigma) %>% 
  
  ggplot(aes(x = value)) +
  geom_density(fill = "grey33") +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free", labeller = label_parsed)
```

```{r}
library(tidybayes)

d_grid_samples %>% 
  pivot_longer(mu:sigma) %>% 
  group_by(name) %>% 
  mode_hdi(value)
```

#### 4.3.4.1 Overthinking: Sample size and the normality of sigma's posterior.
```{r}
set.seed(4)
(d3 <- sample(d2$height, size = 20))
```

```{r}
n <- 200

# note we've redefined the ranges of mu and sigma
d_grid <-
  crossing(mu = seq(from = 150, to = 170, length.out = n),
           sigma = seq(from = 4, to = 20, length.out = n))
```

```{r}
grid_function <- function(mu, sigma){
  dnorm(d3, mean = mu, sd = sigma, log = T) %>% 
    sum()
}
```

```{r}
d_grid <-
  d_grid %>% 
  mutate(log_likelihood = map2_dbl(mu, sigma, grid_function)) %>% 
  mutate(prior_mu = dnorm(mu, mean = 178, sd = 20, log = T),
         prior_sigma = dunif(sigma, min = 0, max = 50, log = T)) %>% 
  mutate(product = log_likelihood + prior_mu + prior_sigma) %>% 
  mutate(probability = exp(product - max(product)))
```

```{r}
set.seed(4)

d_grid_samples <-
  d_grid %>% 
  sample_n(size = 1e4, replace = T, weight = probability)

d_grid_samples %>% 
  ggplot(aes(x = mu, y = sigma)) +
  geom_point(size = .9, alpha = 1/15) +
  labs(x = expression(mu[samples]),
       y = expression(sigma[samples])) +
  theme(panel.grid = element_blank())
```

```{r}
d_grid_samples %>% 
  pivot_longer(mu:sigma) %>% 
  
  ggplot(aes(x = value)) +
  geom_density(fill = "grey33", linewidth = 0) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free", labeller = label_parsed)
```

### 4.3.5 Finding the posterior distribution with brm()
```{r}
b4.1 <-
  brm(data = d2,
      family = gaussian,
      height ~ 1,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(uniform(0, 50), class = sigma,
                      ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.01")
```

```{r}
plot(b4.1)
```

```{r}
print(b4.1)
```
Stan-like summary with a little indexing
```{r}
b4.1$fit
```

If you want to use 89% intervals using brms models
```{r}
summary(b4.1, prob = .89)
```

```{r}
b4.2 <-
  brm(data = d2,
      family = gaussian,
      height ~ 1,
      prior = c(prior(normal(178, .1), class = Intercept),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.02")

plot(b4.2, widths = c(1, 2))
```

```{r}
summary(b4.2)
```

Subsetting the summary() output with $fixed provides a convenient way to compare the Intercept summaries between b4.1_hc and b4.2
```{r}
rbind(summary(b4.1)$fixed,
      summary(b4.2)$fixed)
```

### 4.3.6 Sampling from a brm() fit
```{r}
vcov(b4.1)
```

```{r}
post <- as_draws_df(b4.1)

head(post)
```

Now select() the columns containing the draws from the desired parameters and feed them into cov()
```{r}
select(post, b_Intercept:sigma) %>% 
  cov()
```

```{r}
# variances
select(post, b_Intercept:sigma) %>% 
  cov() %>% 
  diag()
```

```{r}
# correlation
post %>% 
  select(b_Intercept, sigma) %>% 
  cor()
```

```{r}
str(post)
```

```{r}
summary(post[, 1:2])
```

```{r}
t(apply(post[, 1:2], 2, quantile, prob = c(.5, .025, .75)))
```

```{r}
post %>% 
  pivot_longer(b_Intercept:sigma) %>% 
  group_by(name) %>% 
  summarise(mean = mean(value),
            sd = sd(value),
            `2.5%` = quantile(value, prob = .025),
            `97.5%` = quantile(value, probs = .975)) %>% 
  mutate_if(is.numeric, round, digits = 2)
```

You can always get pretty similar information by just putting the brm() fit object into posterior summary
```{r}
posterior_summary(b4.1)
```

And if you're willing to drop the posterior sds, you can use tidybayes::mean_hdi()
```{r}
post %>% 
  pivot_longer(b_Intercept:sigma) %>% 
  group_by(name) %>% 
  mean_qi(value)
```

```{r}
rbind(histospark(post$b_Intercept),
      histospark(post$sigma))
```
## 4.4 Linear prediction
Scatter plot of weight and height
```{r}
ggplot(data = d2,
       aes(x = weight, y = height)) +
  geom_point(shape = 1, size = 2) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

### 4.4.1 The linear model strategy
#### 4.4.1.1 Probability of the data
#### 4.4.1.2 Linear model
#### 4.4.1.3 Priors
```{r}
set.seed(2971)
# how many lines would you like?
n_lines <- 100

lines <-
  tibble(n = 1:n_lines,
         a = rnorm(n_lines, mean = 170, sd = 20),
         b = rnorm(n_lines, mean = 0, sd = 10)) %>% 
  expand_grid(weight = range(d2$weight)) %>% 
  mutate(height = a + b * (weight - mean(d2$weight)))

head(lines)
```

```{r}
lines %>% 
  ggplot(aes(x = weight, y = height, group = n)) +
  geom_hline(yintercept = c(0, 272), linetype = 2:1, linewidth = 1/3) +
  geom_line(alpha = 1/10) +
  coord_cartesian(ylim = c(-100, 400)) +
  ggtitle("b ~ dnorm(0, 10)") +
  theme_classic()
```

To remove the absurd negative correlation between weight and height we createan always positive slope using Log-Normal distribution. Here's what Log-Normal(0, 1) looks like
```{r}
set.seed(4)

tibble(b = rlnorm(1e4, mean = 0, sd = 1)) %>% 
  ggplot(aes(x = b)) +
  geom_density(fill = "grey92") +
  coord_cartesian(xlim = c(0, 5)) +
  theme_classic()
```

Log-Normal is the distribution whose log is normally distributed. Comparison of Normal(0, 1) and Log(Log-Normal(0, 1))
```{r}
set.seed(4)

tibble(rnorm = rnorm(1e5, mean = 0, sd = 1),
       `log(rlognorm)` = log(rlnorm(1e5, mean = 0, sd = 1))) %>% 
  pivot_longer(everything()) %>% 
  
  ggplot(aes(x = value)) +
  geom_density(fill = "grey92") +
  coord_cartesian(xlim = c(-3, 3)) +
  theme_classic() +
  facet_wrap(~ name, nrow = 2)
```

Prior predictive simulation using Log-Normal prior
```{r}
# make a tibble to annotate the plot
text <-
  tibble(weight = c(34, 43),
       height = c(0 - 25, 272 + 25),
       label = c("Embryo", "World's tallest persion (272 cm)"))

# simulate
set.seed(2971)

tibble(n = 1:n_lines,
       a = rnorm(n_lines, mean = 178, sd = 20),
       b = rlnorm(n_lines, mean = 0, sd = 1)) %>% 
  expand_grid(weight = range(d2$weight)) %>% 
  mutate(height = a + b * (weight - mean(d2$weight))) %>% 
  
  # plot
  ggplot(aes(x = weight, y = height, group = n)) +
  geom_hline(yintercept = c(0, 272), linetype = 2:1,
             linewidth = 1/3) +
  geom_line(alpha = 1/10) +
  geom_text(data = text,
            aes(label = label),
            size = 3) +
  coord_cartesian(ylim = c(-100, 400)) +
  ggtitle("log(b) ~ dnorm(0, 1)") +
  theme_classic()
```

### 4.4.2 Finding the posterior distribution
```{r}
d2 <-
  d2 %>% 
  mutate(weight_c = weight - mean(weight))
```

```{r}
b4.3 <-
  brm(data = d2,
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03")
```

```{r}
plot(b4.3, widths = c(1, 2))
```

#### 4.4.2.1 Overthinking: Logs and exps, oh my

Non-linear syntax of brms
```{r}
b4.3b <-
  brm(data = d2,
      family = gaussian,
      bf(height ~ a + exp(lb) * weight_c,
         a ~ 1,
         lb ~ 1,
         nl = TRUE),
      prior = c(prior(normal(178, 20), class = b, nlpar = a),
                prior(normal(0, 1), class = b, nlpar = lb),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03b")
```

```{r}
fixef(b4.3)["weight_c", "Estimate"]
```

```{r}
fixef(b4.3b)["lb_Intercept", "Estimate"] %>% exp()
```

### 4.4.3 Interpreting the posterior distribution
#### 4.4.3.1 Tables of marginal distributions
```{r}
posterior_summary(b4.3)[1:3,] %>% 
  round(digits = 2)
```

If we put our brms into vcov() function, we'll get the variance/covariance matrix of the intercept and weight_c coefficient
```{r}
vcov(b4.3) %>% round(4)
```

To get sigma
```{r}
as_draws_df(b4.3) %>% 
  select(b_Intercept:sigma) %>% 
  cov() %>% 
  round(digits = 3)
```

The pairs function will work for brms like it would one from rethinking. It will show both the marginal posterios and the covariance
```{r}
pairs(b4.3)
```

#### 4.4.3.2 Plotting posterior inference against the data
```{r}
d2 %>% 
  ggplot(aes(x = weight_c, y = height)) +
  geom_abline(intercept = fixef(b4.3)[1],
              slope = fixef(b4.3)[2]) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  theme_classic()
```

Note how the breaks on x-axis looks off. That's because we fit the model with weight_c and weight_c is subtracted by the mean of weight. We can adjust back the x-axis break point labels by adding that value back
```{r}
labels <-
  c(-10, 0, 10) + mean(d2$weight) %>% 
  round(digits = 0)

d2 %>% 
  ggplot(aes(x = weight_c, y = height)) +
  geom_abline(intercept = fixef(b4.3)[1],
              slope = fixef(b4.3)[2]) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  scale_x_continuous("weight",
                     breaks = c(-10, 0, 10),
                     labels = labels) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

### 4.4.3.3 Adding uncertainty around the mean
```{r}
post <- as_draws_df(b4.3)

post %>% 
  slice(1:5) # this serves a similar function as head()
```

```{r}
N <- 10

b4.3_010 <-
  brm(data = d2 %>% slice(1:N),
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03_010")

N <- 50

b4.3_050 <-
  brm(data = d2 %>% slice(1:N),
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03_050")

N <- 150

b4.3_150 <-
  brm(data = d2 %>% slice(1:N),
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03_150")

N <- 352

b4.3_352 <-
  brm(data = d2 %>% slice(1:N),
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.03_352")
```

```{r}
plot(b4.3_010)
print(b4.3_010)

plot(b4.3_050)
print(b4.3_050)

plot(b4.3_150)
print(b4.3_150)

plot(b4.3_352)
print(b4.3_352)
```
Put the chains of each model into data frames
```{r}
post010 <- as_draws_df(b4.3_010)
post050 <- as_draws_df(b4.3_050)
post150 <- as_draws_df(b4.3_150)
post352 <- as_draws_df(b4.3_352)
```

```{r}
p1 <-
  ggplot(data = d2[1:10,],
         aes(x = weight_c, y = height)) +
  geom_abline(data = post010 %>% slice(1:20),
              aes(intercept = b_Intercept, slope = b_weight_c),
              linewidth = 1/3, alpha = .3) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  coord_cartesian(xlim = range(d2$weight_c),
                  ylim = range(d2$height)) +
  labs(subtitle = "N = 10")

p2 <-
  ggplot(data = d2[1:20,],
         aes(x = weight_c, y = height)) +
  geom_abline(data = post050 %>% slice(1:20),
              aes(intercept = b_Intercept,
                  slope = b_weight_c),
              linewidth = 1/3, alpha = .3) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  coord_cartesian(xlim = range(d2$weight_c),
                  ylim = range(d2$height)) +
  labs(subtitle = "N = 50")

p3 <-
  ggplot(data = d2[1:150,],
         aes(x = weight_c, y = height)) +
  geom_abline(data = post150 %>% slice(1:150),
              aes(intercept = b_Intercept,
                  slope = b_weight_c),
              linewidth = 1/3, alpha = .3) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  coord_cartesian(xlim = range(d2$weight_c),
                  ylim = range(d2$height)) +
  labs(subtitle = "N = 150")

p4 <-
  ggplot(data = d2,
         aes(x = weight_c, y = height)) +
  geom_abline(data = post352 %>% slice(1:20),
              aes(intercept = b_Intercept,
                  slope = b_weight_c),
              linewidth = 1/3, alpha = .3) +
  geom_point(shape = 1, size = 2, color = "royalblue") +
  coord_cartesian(xlim = range(d2$weight_c),
                  ylim = range(d2$height)) +
  labs(subtitle = "N = 352")
```

Combine the ggplot with patchwork syntax to make the full version
```{r}
(p1 + p2 + p3 + p4) &
  scale_x_continuous("weight",
                     breaks = c(-10, 0, 10),
                     labels = labels) &
  theme_classic()
```

#### 4.4.3.4 Plotting regression intervals and contours
```{r}
mean(d2$weight)
```

Just a hair under 45. If we're interested in $\mu$ at ```weight``` = 50, that implies we're also interested in $\mu$ at ```weight_c``` + 5.01. Within the context of our model, we compute this with $\alpha + \beta \dot 5.01.$ Here's what that looks like with post
```{r}
mu_at_50 <-
  post %>% 
  transmute(mu_at_50 = b_Intercept + b_weight_c * 5.01)

head(mu_at_50)
```

Density plot
```{r}
mu_at_50 %>% 
  ggplot(aes(x = mu_at_50)) +
  geom_density(linewidth = 0, fill = "royalblue") +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(mu["height | weight = 50"])) +
  theme_classic()
```

We'll use mean_hdi() to get both 89% and 95% HPDIs along with the mean
```{r}
mean_hdi(mu_at_50[, 1], .width = c(.89, .95))
```

If you wanted to express those sweet 95% HPDIs on your density plot, you might use tidybayes::stat_halfeye(). Since stat_halfeye() also returns a point estimate, we'll throw in the mode
```{r}
mu_at_50 %>% 
  ggplot(aes(x = mu_at_50, y = 0)) +
  stat_halfeye(point_interval = mode_hdi, .width = .95,
               fill = "royalblue") +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(mu["height | weight - 50"])) +
  theme_classic()
```

With brms, you would use fitted() to do what McElreath accomplished with link().
```{r}
mu <- fitted(b4.3, summary = F)

str(mu)
```

Much like rethinking's link(), brms::fitted() can accommodate custom predictor values with its newdata argument
```{r}
weight_seq <-
  tibble(weight = 25:70) %>% 
  mutate(weight_c = weight - mean(d2$weight))

mu <-
  fitted(b4.3,
         summary = F,
         newdata = weight_seq) %>% 
  data.frame() %>% 
  # here we name the columns after the weight values from which they were computed
  set_names(25:70) %>% 
  mutate(iter = 1:n())
```

```{r}
mu <-
  mu %>% 
  pivot_longer(-iter,
               names_to = "weight",
               values_to = "height") %>% 
  # we might reformat weight to numberals
  mutate(weight = as.numeric(weight))

head(mu)
```

Reproduce McElreath's Figure 4.9.a.
```{r}
d2 %>% 
  ggplot(aes(x = weight, y = height)) +
  geom_point(data = mu %>% filter(iter < 101),
             color = "navyblue", alpha = .05) +
  coord_cartesian(xlim = c(30, 65)) +
  theme(panel.grid = element_blank())
```

With fitted(), it's quite easy to plot a regression line and its intervals. Just omit the summary = T argument
```{r}
mu_summary <-
  fitted(b4.3,
         newdata = weight_seq) %>% 
  data.frame() %>% 
  bind_cols(weight_seq)

head(mu_summary)
```

Here's our analogue to Figure 4.9.b.
```{r}
d2 %>% 
  ggplot(aes(x = weight, y = height)) +
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity", fill = "grey70",
              color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(color = "navyblue", shape = 1, linewidth = 1.5,
             alpha = 2/3) +
  coord_cartesian(xlim = range(d2$weight)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())
```

#### 4.4.3.5 Prediction intervals
Even though our statistical model (omitting priors for the sake of simplicity) is
$$
\text{height}_{i} \sim \text{Normal}(\mu_{i} = \alpha + \beta x, \sigma),
$$
we've only been plotting the $\mu$ part. In order to bring the variability expressed by $\sigma$, we'll have to switch to the predict() function. Much as brms::fitted() was our analogue to rethinking::link(), brms::predict() is our analogue to rethinking::sim()

```{r}
weight_seq <-
  tibble(weight = 25:70) %>% 
  mutate(weight_c = weight - mean(d2$weight))
```

```{r}
pred_height <-
  predict(b4.3,
          newdata = weight_seq) %>% 
  data.frame() %>% 
  bind_cols(weight_seq)

pred_height %>% 
  slice(1:6)
```

This time the summary information in our data frame is for, as McElreath put it, "simulated heights, not distributions of plausible average hgieht, $\mu$". Another way of saying that is that these simulations are the joint consequence of buth $\mu$ and $\sigma$, unlike the results of fitted(), which only reflect $\mu$. Figure 4.10 shows how you might visualize them.
```{r}
d2 %>% 
  ggplot(aes(x = weight)) +
  geom_ribbon(data = pred_height,
              aes(ymin = Q2.5, ymax = Q97.5),
              fill = "grey83") +
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity", fill = "grey70",
              color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(aes(y = height),
             shape = 1, size = 2, color = "navyblue") +
  coord_cartesian(xlim = range(d2$weight),
                  ylim = range(d2$height)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())
```
##### 4.4.3.5.1 Overthinking: Rolling your own sim predict()
```{r}
# predict() by hand
set.seed(4)

post %>% 
  expand_grid(weight = 25:70) %>% 
  mutate(weight_c = weight - mean(d2$weight)) %>% 
  mutate(sim_height = rnorm(n(),
                            mean = b_Intercept + b_weight_c * weight_c,
                            sd = sigma)) %>% 
  group_by(weight) %>% 
  summarise(mean = mean(sim_height),
            ll = quantile(sim_height, prob = .025),
            ul = quantile(sim_height, prob = .975)) %>%
  
  # plot
  ggplot(aes(x = weight)) +
  geom_smooth(aes(y = mean, ymin = ll, ymax = ul),
              stat = "identity", fill = "grey83",
              color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(data = d2,
             aes(y = height),
             color = "navyblue", shape = 1, size = 1.5,
             alpha = 2/3) +
  coord_cartesian(xlim = range(d2$weight),
                  ylim = range(d2$height)) +
  theme(panel.grid = element_blank())
```

## 4.5 Curves from lines
### 4.5.1 Polynomial regression
```{r}
d %>% glimpse()
```

Plot height and weight using the full sample
```{r}
d %>% 
  ggplot(aes(x = weight, y = height)) +
  geom_point(shape = 1, size = 2, color = "navyblue",
             alpha = 2/3) +
  annotate(geom = "text",
           x = 42, y = 115,
           label = "This relation is \nvisibly curved.",
           family = "Times") +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())
```

Quadratic equation
```{r}
d <-
  d %>% 
  mutate(weight_s = (weight - mean(weight)) / sd(weight)) %>% 
  mutate(weight_s2 = weight_s^2)
```

```{r}
b4.5 <-
  brm(data = d,
      family = gaussian,
      height ~ 1 + weight_s + weight_s2,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b, coef = "weight_s"),
                prior(normal(0, 1), class = b, coef = "weight_s2"),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.05")
```

```{r}
plot(b4.5, widths = c(1, 2))
```

```{r}
print(b4.5)
```

Our quadratic plot requires new fitted() - and predict() -oriented wrangling
```{r}
weight_seq <-
  tibble(weight_s = seq(from = -2.5, to = 2.5, length.out = 30)) %>% 
  mutate(weight_s2 = weight_s^2)

fitd_quad <-
  fitted(b4.5,
         newdata = weight_seq) %>% 
  data.frame() %>% 
  bind_cols(weight_seq)

pred_quad <-
  predict(b4.5,
          newdata = weight_seq) %>% 
  data.frame() %>% 
  bind_cols(weight_seq)
```

```{r}
p2 <-
  ggplot(data = d,
         aes(x = weight_s)) +
  geom_ribbon(data = pred_quad,
              aes(ymin = Q2.5, ymax = Q97.5),
              fill = "grey83") +
  geom_smooth(data = fitd_quad,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black",
              alpha = 1, linewidth = 1/2) +
  geom_point(aes(y = height),
             color = "navyblue", alpha = .5, linewidth = 1/2,
             shape = 1) +
  labs(subtitle = "quadratic", y = "height") +
  coord_cartesian(xlim = range(d$weight_s),
                  ylim = range(d$height)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())

p2
```

```{r}
d <-
  d %>% 
  mutate(weight_s3 = weight_s^3)
```

```{r}
b4.6 <-
  brm(data = d,
      family = gaussian,
      height ~ 1 + weight_s + weight_s2 + weight_s3,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b, coef = "weight_s"),
                prior(normal(0, 1), class = b, coef = "weight_s2"),
                prior(normal(0, 1), class = b, coef = "weight_s3"),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.06")
```

Good old linear model
```{r}
b4.7 <- 
  brm(data = d, 
      family = gaussian,
      height ~ 1 + weight_s,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b, coef = "weight_s"),
                prior(uniform(0, 50), class = sigma, ub = 50)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.07")
```

```{r}
weight_seq <-
  weight_seq %>% 
  mutate(weight_s3 = weight_s^3)

fitd_cub <-
  fitted(b4.6,
         newdata = weight_seq) %>% 
  as_tibble() %>% 
  bind_cols(weight_seq)

pred_cub <-
  predict(b4.6,
          newdata = weight_seq) %>% 
  as_tibble() %>% 
bind_cols(weight_seq)

p3 <-
  ggplot(data = d,
         aes(x = weight_s)) +
  geom_ribbon(data = pred_cub,
              aes(ymin = Q2.5, ymax = Q97.5),
              fill = "grey83") +
  geom_smooth(data = fitd_cub,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity", fill = "grey70", 
              color = "black", alpha = 1, linewidth = 1/4) +
  geom_point(aes(y = height),
             color = "navyblue", shape = 1, size = 1.5,
             alpha = 1/3) +
  labs(subtitle = "cubic",
       y = "height") +
  coord_cartesian(xlim = range(d$weight_s),
                  ylim = range(d$height)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())

p3
```

Fitted, predict, and ggplot 2 for Figure 4.11.a
```{r}
fitd_line <-
  fitted(b4.7, 
         newdata = weight_seq) %>%
  as_tibble() %>%
  bind_cols(weight_seq)

pred_line <-
  predict(b4.7, 
          newdata = weight_seq) %>%
  as_tibble() %>%
  bind_cols(weight_seq) 

p1 <-
  ggplot(data = d, 
       aes(x = weight_s)) +
  geom_ribbon(data = pred_line, 
              aes(ymin = Q2.5, ymax = Q97.5),
              fill = "grey83") +
  geom_smooth(data = fitd_line,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/4) +
  geom_point(aes(y = height),
             color = "navyblue", shape = 1, size = 1.5, alpha = 1/3) +
  labs(subtitle = "linear",
       y = "height") +
  coord_cartesian(xlim = range(d$weight_s),
                  ylim = range(d$height)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank())

p1
```

```{r}
p1 | p2 | p3
```

##### 4.5.1.0.1 Overthinking: Converting back to natural scale.
```{r}
at <- c(-2, -1, 0, 1, 2)

ggplot(data = d,
       aes(x = weight_s)) +
  geom_ribbon(data = pred_cub,
              aes(ymin = Q2.5, ymax = Q97.5),
              fill = "grey83") +
  geom_point(aes(y = height),
             color = "navyblue", shape = 1, size = 1.5,
             alpha = 1/3) +
  coord_cartesian(xlim = range(d$weight_s)) +
  theme(text = element_text(family = "Times"),
        panel.grid = element_blank()) +
  
  # here it is !
  scale_x_continuous("standardized weight converted back",
                     breaks = at,
                     labels = round(at * sd(d$weight) +
                                      mean(d$weight), 1))
```

### 4.5.2 Splines
Load the cherry_blossoms data
```{r}
library(rethinking)

data(cherry_blossoms)
d <- cherry_blossoms
rm(cherry_blossoms)
detach(package:rethinking, unload = T)
```

```{r}
d %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(mean = mean(value, na.rm = T),
            sd = sd(value, na.rm = T),
            ll = quantile(value, prob = .055, na.rm = T),
            ul = quantile(value, prob = .945, na.rm = T)) %>% 
  mutate_if(is.double, round, digits = 2)
```

Plot doy against year
```{r}
d %>% 
  ggplot(aes(x = year, y = doy)) +
  # color from here: https://www.colorhexa.com/ffb7c5
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        # color from here: https://www.colordic.org/w/, inspired by https://chichacha.netlify.com/2018/11/29/plotting-traditional-colours-of-japan/
        panel.background = element_rect(fill = "#4f455c"))
```

There are missing data. To check execute
```{r}
d %>% 
  count(is.na(doy)) %>% 
  mutate(percent = 100 * n / sum(n))
```

```{r}
d2 <-
  d %>% 
  drop_na(doy)
```

```{r}
num_knots <- 15
knot_list <- quantile(d2$year,
                      prob = seq(from = 0, to = 1, 
                                 length.out = num_knots))
```

```{r}
knot_list
```

Here what it looks like if we use knot_list values to chop up our year / doy scatter plot
```{r}
d %>% 
  ggplot(aes(x = year, y = doy)) +
  geom_vline(xintercept = knot_list, color = "white",
             alpha = 1/2) +
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

Next choice is polynomial degree
```{r}
library(splines)

B <- bs(d2$year,
        knots = knot_list[-c(1, num_knots)],
        degree = 3,
        intercept = TRUE)
```

```{r}
B %>% str()
```

```{r}
knot_list[c(1, num_knots)]
```

By the degree = 3 argument, we indicated we wanted a cubic splite. McElreath used degree = 1 for Figure 4.12.
```{r}
# wrangle a bit
b <-
  B %>% 
  data.frame() %>% 
  set_names(str_c(0, 1:9), 10:17) %>% 
  bind_cols(select(d2, year)) %>% 
  pivot_longer(-year,
               names_to = "bias_function",
               values_to = "bias")

# plot
b %>% 
  ggplot(aes(x = year, y = bias, group = bias_function)) +
  geom_vline(xintercept = knot_list, color = "white",
             alpha = 1/2) +
  geom_line(color = "#ffb7c5", alpha = 1/2, linewidth = 1.5) +
  ylab("bias value") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

To elucidate what's goin on, we might break it up with facet_wrap()
```{r}
b %>% 
  mutate(bias_function = str_c("bias function ", bias_function)) %>% 
  
  ggplot(aes(x = year, y = bias)) +
  geom_vline(xintercept = knot_list, color = "white", alpha = 1/2) +
  geom_line(color = "#ffb7c5", linewidth = 1.5) +
  ylab("bias value") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = scales::alpha("#ffb7c5", .25), color = "transparent"),
        strip.text = element_text(size = 8, margin = margin(0.1, 0, 0.1, 0, "cm"))) +
  facet_wrap(~ bias_function, ncol = 1)
```
Our model will follow the form
$$
\begin{align}
\text{day_in_year}_{i} & \sim \text{Normal}(\mu_{i}, \sigma)\\
\mu_{i} & = \alpha + \Sigma_{k = 1}^{K}w_{k}B_{k, i}\\
\alpha & \sim \text{Normal}(100, 10) \\
w_{j} & \sim \text{Normal}(0, 10) \\
\sigma & \sim \text{Exponential}(1)
\end{align}
$$
```{r}
tibble(x = seq(from = 0, to = 10, by = .1)) %>% 
  mutate(d = dexp(x, rate = 1)) %>% 
  
  ggplot(aes(x = x, y = d)) +
  geom_area(fill = "#ffb7c5") +
  scale_y_continuous(NULL, breaks = NULL) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

We'll use exponential priors for the rest of the book, in place of uniform priors. it is much more common to have a sense of the average deviation than of the maximum.

Before fitting this model in brms, we'll take a minor detour on the data structure. In his R code 4.76, McElreath defined his data in a list, list(D = d2$doy, B = B). Our approach will be a little different. Here, we'll add the B matrix to our d2 data frame and name the results as d3
```{r}
d3 <-
  d2 %>% 
  mutate(B = B)

# take a look at the structure of d3
d3 %>% glimpse()
```
In our d3 data, columns year through temp_lower are all standard data columns. The B column is a matrix column, which contains the same number of rows as the others, but also smuggled in 17 columns within that column. Each of those 17 columns corresponds to one of our synthetic $B_{k}$ variables. The advantage 

```{r}
b4.8 <-
  brm(data = d3,
      family = gaussian,
      doy ~ 1 + B,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.08")
```

```{r}
print(b4.8)
```

```{r}
post <- as_draws_df(b4.8)

glimpse(post)
```

```{r}
post %>% 
  select(b_B1:b_B17) %>% 
  set_names(c(str_c(0, 1:9), 10:17)) %>% 
  pivot_longer(everything(), names_to = "bias_function") %>% 
  group_by(bias_function) %>% 
  summarise(weight = mean(value)) %>% 
  full_join(b, by = "bias_function") %>% 
  
  # plot
  ggplot(aes(x = year, y = bias * weight, group = bias_function)) +
  geom_vline(xintercept = knot_list, color = "white", alpha = 1/2) +
  geom_line(color = "#ffb7c5", alpha = 1/2, linewidth = 1.5) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

```{r}
f <- fitted(b4.8)

f %>% 
  data.frame() %>% 
  bind_cols(d2) %>% 
  
  ggplot(aes(x = year, y = doy, ymin = Q2.5, ymax = Q97.5)) +
  geom_vline(xintercept = knot_list, color = "white",
             alpha = 1/2) +
  geom_hline(yintercept = fixef(b4.8)[1, 1], color = "white",
             linetype = 2) +
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  geom_ribbon(fill = "white", alpha = 2/3) +
  labs(x = "year", y = "day in year") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

Let's create a simpler model using 5 knots
```{r}
# redo the B splines
num_knots <- 5
knot_list <- quantile(d2$year, 
                      probs = seq(from = 0, to = 1, length.out = num_knots))

B <- bs(d2$year,
        knots = knot_list[-c(1, num_knots)],
        # this makes the splines liner rater than cubic
        degree = 1,
        intercept = TRUE)

# define a new d4 data
d4 <-
  d2 %>% 
  mutate(B = B)

b4.9 <-
  brm(data = d4,
      family = gaussian,
      formula = doy ~ 1 + B,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "fits/b04.09")
```

```{r}
print(b4.9)
```

```{r}
## top
# wrangle a bit
b <-
  invoke(data.frame, d4) %>% 
  pivot_longer(starts_with("B"),
               names_to = "bias_function",
               values_to = "bias")

# plot
p1 <-
  b %>% 
  ggplot(aes(x = year, y = bias, group = bias_function)) +
  geom_vline(xintercept = knot_list, color = "white",
             alpha = 1/2) +
  geom_line(color = "#ffb7c5", alpha = 1/2, linewidth = 1.5) +
  scale_x_continuous(NULL, breaks = NULL) +
  ylab("bias value")

## middle
# wrangle
p2 <-
  as_draws_df(b4.9) %>% 
  select(b_B1:b_B5) %>% 
  set_names(str_c("B.", 1:5)) %>% 
  pivot_longer(everything(), names_to = "bias_function") %>% 
  group_by(bias_function) %>% 
  summarise(weight = mean(value)) %>% 
  full_join(b, by = "bias_function") %>% 
  
  # plot
  ggplot(aes(x = year, y = bias * weight, group = bias_function)) +
  geom_vline(xintercept = knot_list, color = "white", alpha = 1/2) +
  geom_line(color = "#ffb7c5", alpha = 1/2, linewidth = 1.5) +
  scale_x_continuous(NULL, breaks = NULL)

## bottom
# wrangle
f <- fitted(b4.9)

p3 <-
  f %>% 
  data.frame() %>% 
  bind_cols(d2) %>% 
  
  # plot
  ggplot(aes(x = year, y = doy, ymin = Q2.5, ymax = Q97.5)) +
  geom_vline(xintercept = knot_list, color = "white",
             alpha = 1/2) +
  geom_hline(yintercept = fixef(b4.9)[1, 1], color = "white",
             linetype = 2) +
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  geom_ribbon(fill = "white", alpha = 2/3) +
  labs(x = "year", y = "day in year")
```

```{r}
(p1 / p2 / p3) &
  theme_bw() &
  theme(panel.background = element_rect(fill = "#4f455c"))
```

## 4.6 First bonus: Smooth functions with brms::s()


```{r}
get_prior(data = d2,
          family = gaussian,
          doy ~ 1 + s(year))
```

```{r}
b4.10 <-
  brm(data = d2,
      family = gaussian,
      doy ~ 1 + s(year),
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(student_t(3, 0, 5.9), class = sds),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      control = list(adapt_delta = .99),
      file = "fits/b04.10")
```

```{r}
print(b4.10)
```

Our intercept and $\sigma$ summaries are similar to those we got from b4.8. The rest looks different and maybe a little disorienting. Here's what happens when we use brms::fitted to plot the implications of the model
```{r}
fitted(b4.10) %>% 
  data.frame() %>% 
  bind_cols(select(d2, year, doy)) %>% 
  
  ggplot(aes(x = year, y = doy, ymin = Q2.5, ymax = Q97.5)) +
  geom_hline(yintercept = fixef(b4.10)[1, 1], 
             color = "white", linetype = 2) +
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  geom_ribbon(fill = "white", alpha = 2/3) +
  labs(subtitle = "b4.7 using s(year)",
       y = "day in year") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

The smooth looks different because by default s() use thin plate regression spline and we didn't use a B-spline. To fit a B-spline we have to set bs = "bs"
```{r}
b4.11 <-
  brm(data = d2,
      family = gaussian,
      doy ~ 1 + s(year, bs = "bs", k = 19),
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(student_t(3, 0, 5.9), class = sds),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      control = list(adapt_delta = .99),
      file = "fits/b04.11")
```

```{r}
print(b4.11)
```

Now here's the depiction of our s() -based B-spline model
```{r}
fitted(b4.11) %>% 
  data.frame() %>% 
  bind_cols(select(d2, year, doy)) %>% 
  
  ggplot(aes(x = year, y = doy, ymin = Q2.5, ymax = Q97.5)) +
  geom_hline(yintercept = fixef(b4.11)[1, 1], color = "white", linetype = 2) +
  geom_point(color = "#ffb7c5", alpha = 1/2) +
  geom_ribbon(fill = "white", alpha = 2/3) +
  labs(subtitle = 'b4.7_bs using s(year, bs = "bs")',
       y = "day in year") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#4f455c"),
        panel.grid = element_blank())
```

## 4.7 Second bonus: Group predictors with matrix columns