---
title: "Chapter 4 - Exercise"
format: html
---

### Medium problems

:::question
> **4M1.**  For the model definition below, simulate observed y values from the prior (not the posterior).

$$
\begin{align}
y_{i} & \sim \text{Normal}(\mu, \sigma) \\
\mu & \sim \text{Normal}(0, 10) \\
\sigma & \sim \text{Exponential}(1)
\end{align}
$$

```{r}
n <- 1e4

sim <-
  tibble(mu = rnorm(n, 0, 10),
         sigma = rexp(n, 1)) %>%  
  mutate(y = rnorm(n, mean = mu, sd = sigma))

ggplot(sim, aes(x = y)) +
  geom_density() +
  labs(x = "y", y = "Density") +
  theme_bw()
```

:::question
> 4M4. A sample of students is measured for height each year for 3 years. After the third year, you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.

```{r}
n <- 50

sim <-
  tibble(group = seq_len(n),
         alpha = rnorm(n, 100, 10),
         beta = rnorm(n, 0,  10),
         sigma = rexp(n, 1)) %>% 
  expand(nesting(group, alpha, beta, sigma),
         year = c(1, 2, 3)) %>% 
  mutate(height = rnorm(n(), 
                        alpha + beta * (year - mean(year)), 
                        sigma)) 

sim %>% 
  ggplot(aes(x = year, y = height, group = group)) +
  geom_line() +
  labs(x = "Year", y = "Height") +
  theme_minimal()
```

:::question
> 4M5. Now suppose I remind you that every student got taller each year. Does this information lead you to change your choice of priors? How?

```{r}
library(tidybayes)

set.seed(123)
samples <- rlnorm(1e4, 1, .5)
bounds <- mean_hdi(samples, .width = .89)

ggplot() +
  stat_function(data = tibble(x = c(0, 10)), mapping = aes(x = x),
                geom = "line", fun = dlnorm,
                args = list(meanlog = 1, sdlog = .5)) +
  geom_ribbon(data = tibble(x = seq(bounds$ymin, bounds$ymax,
                                    0.01)),
              aes(x = x, ymin = 0, ymax = dlnorm(x, 1, .5)),
              alpha = .8) +
  scale_x_continuous(breaks = seq(0, 10, 2)) +
  labs(x = expression(beta), y = "Density") +
  theme_minimal()
```

```{r}
n <- 50

sim <-
  tibble(group = seq_len(n),
         alpha = rnorm(n, 100, 10),
         beta = rlnorm(n, 1, 0.5),
         sigma = rexp(n, 1)) %>% 
  expand(nesting(group, alpha, beta, sigma),
         year = c(1, 2, 3)) %>% 
  mutate(height = rnorm(n(),
                        alpha + beta * (year - mean(year)),
                        sigma))

sim %>% 
  ggplot(aes(x = year, y = height, group = group)) +
  geom_line() +
  labs(x = "Year", y = "Height") +
  theme_minimal()
```

:::question
> 4M7. Refit model m4.3 from the chapter, but omit the mean weight xbar this time. Compare the new model’s posterior to that of the original model. In particular, look at the covariance among the parameters. What is different? Then compare the posterior predictions of both models.

```{r}
library(rethinking)
library(brms)

data(Howell1)
how_dat <- Howell1 %>% 
  filter(age >= 18) %>% 
  mutate(weight_c = weight - mean(weight))

# first, duplicate model with quap
m4.3 <-
  quap(alist(height ~ dnorm(mu, sigma),
             mu <- a + b * (weight_c),
             a ~ dnorm(178, 20),
             b ~ dlnorm(0, 1),
             sigma ~ dunif(0, 50)),
       data = how_dat)

round(vcov(m4.3), 3)

# brms
b4.3 <-
  brm(height ~ 1 + weight_c,
      data = how_dat,
      family = gaussian,
      prior = c(prior(normal(178, 20), class = Intercept),
                prior(lognormal(0, 1), class = b, lb = 0),
                prior(uniform(0, 50), class = sigma)),
      iter = 28000, warmup = 27000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4.3-0")

as_draws_df(b4.3) %>% 
  as_tibble() %>% 
  select(b_Intercept, b_weight_c, sigma) %>% 
  cov() %>% 
  round(digits = 3)
```

Let's use the non-centered parametrization. This time, we'll only use brm()
```{r}
b4.3_nc <-
  brm(height ~ 1 + weight_c,
      data = how_dat,
      family = gaussian,
      prior = c(prior(normal(178, 20), class = Intercept),
                         prior(lognormal(0, 1), class = b, lb = 0),
                         prior(uniform(0, 50), class = sigma)),
               iter = 28000, warmup = 27000, chains = 4, cores = 4, seed = 1234,
               file = "fits/b4.3_nc")

as_draws_df(b4.3_nc) %>% 
  as_tibble() %>% 
  select(b_Intercept, b_weight, sigma) %>% 
  cov() %>% 
  round(digits = 3)
```

```{r}
library(gganimate)

weight_seq <-
  tibble(weight = seq(25, 70, length.out = 100)) %>% 
  mutate(weight_c = weight - mean(how_dat$weight))

predictions <- bind_rows(
  predict(b4.3, newdata = weight_seq) %>% 
    as_tibble() %>% 
    bind_cols(weight_seq) %>% 
    mutate(type = "Centered"),
  predict(b4.3_nc, newdata = weight_seq) %>% 
    as_tibble() %>% 
    bind_cols(weight_seq) %>% 
    mutate(type = "Non-centered")
)

fits <- bind_rows(
  weight_seq %>% 
    add_epred_draws(b4.3) %>% 
    mutate(type = "Centered"),
  weight_seq %>% 
    add_epred_draws(b4.3_nc) %>% 
    mutate(type = "Non-centered")
) %>% 
  ungroup()

bands <-
  fits %>% 
  group_by(type, weight) %>% 
  median_qi(.epred, .width = c(.67, .89, .97))

lines <-
  fits %>% 
  filter(.draw %in% sample(unique(.data$.draw), size = 50))

ggplot(lines, aes(x = weight)) +
  facet_wrap(~type, nrow = 1) +
  geom_ribbon(data = predictions, aes(ymin = Q2.5, ymax = Q97.5), alpha = .3) +
  geom_lineribbon(data = bands, aes(y = .epred, ymin = .lower, ymax = .upper), color = NA) +
  scale_fill_brewer(palette = "Blues",
                    breaks = c(.67, .89, .97)) +
  geom_line(aes(y = .epred, group = .draw)) +
  geom_point(data = how_dat, aes(y = height), shape = 1,
             alpha = .7) +
  labs(x = "Weight", y = "Height", fill = "Interval") +
  transition_states(.draw, 0, 1)
```

:::question
> 4M8. In the chapter, we used 15 knots with the cherry blossom spline. Increase the number of knots and observe what happens to the resulting spline. Then adjust also the width of the prior on the weights—change the standard deviation of the prior and watch what happens. What do you think the combination of know number and the prior on the weights controls?

Duplicate the 15-knot spline model from the chapter
```{r}
library(splines)

data(cherry_blossoms, package = "rethinking")
cb_dat <- cherry_blossoms %>% 
  drop_na(doy)

# original m4.7 model
knots_15 <- quantile(cb_dat$year, prob = seq(0, 1, length.out = 15))
B_15 <- bs(cb_dat$year, knots = knots_15[-c(1, 15)],
           degree = 3, intercept = TRUE)

cb_dat_15 <- 
  cb_dat %>% 
  mutate(B = B_15)

b4.7 <-
  brm(doy ~ 1 + B,
      data = cb_dat_15,
      family = gaussian,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4.7")
```

```{r}
original_draws <-
  cb_dat_15 %>% 
  add_epred_draws(b4.7) %>% 
  summarize(mean_hdi(.epred, .width = .89), .groups = "drop")
```


```{r}
ggplot(original_draws, aes(x = year, y = doy)) +
  geom_point(alpha = .5) +
  geom_vline(xintercept = knots_15, alpha = .5) +
  geom_hline(yintercept = fixef(b4.7)[1, 1], 
             linetype = "dashed") +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), 
              fill = "#009FB7", alpha = .8) +
  labs(x = "Year", y = "Day in Year") +
  theme_minimal()
```

Fit two additional models. The first use 30 knots, the second uses a tighter prior
```{r}
# double the number of the knots
knots_30 <- quantile(cb_dat$year, probs = seq(0, 1, length.out = 30))
B_30 <- bs(cb_dat$year, knots = knots_30[-c(1, 30)],
           degree = 3, intercept = TRUE)

cb_dat_30 <- 
  cb_dat %>% 
  mutate(B = B_30)

b4.7_30 <-
  brm(doy ~ 1 + B,
      family = gaussian,
      data = cb_dat_30,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4.7_30")

# and modify the prior
b4.7_30p <-
  brm(doy ~ 1 + B,
      family = gaussian,
      data = cb_dat_30,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4.7_30p")
```

```{r}
# create plot data
spline_15 <-
  original_draws %>% 
  select(-B) %>% 
  mutate(knots = "15 knots (original model)")

spline_30 <-
  cb_dat_30 %>% 
  add_epred_draws(b4.7_30) %>% 
  summarize(mean_hdi(.epred, .width = .89),
            .groups = "drop") %>% 
  select(-B) %>% 
  mutate(knots = "30 knots")
  
spline_30p <-
  cb_dat_30 %>% 
  add_epred_draws(b4.7_30p) %>% 
  summarize(mean_hdi(.epred, .width = .89),
            .groups = "drop") %>% 
  select(-B) %>% 
  mutate(knots = "30 knots; Tight prior")

all_splines <-
  bind_rows(spline_15, spline_30, spline_30p)
```


```{r}
# make plot
ggplot(all_splines, aes(x = year, y = doy)) +
  geom_point(alpha = .5) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), fill = "#009FB7") +
  theme_minimal() +
  facet_wrap(~ knots, ncol = 1) +
  labs(x = "Year", y = "Day in Year")

```

### Hard problems

:::question
> 4H1. The weights listed below were recored in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals for each of these individuals. That is, fill in the table, below, using model-based predictions.

```{r}
tibble(individual = 1:5,
       weight = c(47, 43.7, 64.8, 32.6, 54.6)) %>%
  mutate(weight_c = weight - mean(how_dat$weight)) %>% 
  add_predicted_draws(b4.3) %>% 
  group_by(individual, weight) %>% 
  mean_qi(.prediction, .width = .89) %>% 
  mutate(range = glue("[{sprintf('%0.1f', .lower)}--",
                      "{sprintf('%0.1f', .upper)}]"),
         .prediction = sprintf("%0.1f", .prediction)) %>% 
  select(individual, weight, exp = .prediction, range) %>% 
  kbl(align = "c", booktabs = TRUE,
      col.names = c("Individual", "weight", 
                    "expected height", "89% interval"))
```
:::question
> 4H2. Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it.

```{r}
young_how <-
  Howell1 %>% 
  dplyr::filter(age < 18) %>% 
  mutate(weight_c = weight - mean(weight))

nrow(young_how)
```

> a. Fit a linear regression to these data, using quap. Present and interpret the estimates. For every 10 units of increase in weight, how much taller does the model predict a child gets?

```{r}
b4h2 <-
  brm(data = young_how,
      family = gaussian,
      height ~ 1 + weight_c,
      prior = c(prior(normal(138, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4h2")

summary(b4h2)
```

> b. Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Superimpose the MAP regression line and 89% interval for the mean. Also superimpose the 89% interval for predicted heights.

```{r}
mod_fits <-
  tibble(weight = seq(from = range(young_how$weight)[1], 
                      to = range(young_how$weight)[2],
                      length.out = 100)) %>% 
  mutate(weight_c = weight - mean(young_how$weight)) %>% 
  add_epred_draws(b4h2) %>% 
  group_by(weight) %>% 
  mean_qi(.epred, .width = .89)

mod_preds <-
  tibble(weight = seq(from = range(young_how$weight)[1], 
                      to = range(young_how$weight)[2],
                      length.out = 100)) %>% 
  mutate(weight_c = weight - mean(young_how$weight)) %>% 
  add_predicted_draws(b4h2) %>% 
  group_by(weight) %>% 
  mean_qi(.prediction, .width = .89)
```


```{r}
ggplot(young_how, aes(x = weight)) +
  geom_point(aes(y = height), alpha = .5) +
  geom_ribbon(data = mod_preds, 
                aes(ymin = .lower, ymax = .upper),
                alpha = .2) +
  geom_lineribbon(data = mod_fits,
                  aes(y = .epred, ymin = .lower, 
                      ymax = .upper), fill = "grey60",
                  size = 1) +
  labs(x = "Weight", y = "Height") +
  theme_minimal()
```

:::question
> 4H3.Suppose a colleauge of yours, who works on allometry, glances at the practice problems just above. Your colleague exclaims, “That’s silly. Everyone knows that it’s only the logarithm of body weight that scales with height!” Let’s take your colleague’s advice and see what happens.
a. Model the relationship between height (cm) and the natural 
logarithm of weight (log-kg). Use the entire Howell1 data frame, all 544 rows, adults and non-adults. Can you interpret the resulting estimates?

```{r}
full_how <-
  Howell1 %>% 
  mutate(log_weight = log(weight),
         log_weight_c = log_weight - mean(log_weight))

b4h3 <-
  brm(data = full_how,
      family = gaussian,
      height ~ 1 + log_weight_c,
      prior = c(prior(normal(150, 20), class = Intercept),
                prior(lognormal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4, 
      seed = 1234,
      file = "fits/b4h3")

summary(b4h3)
```
> b. Begin with this plot: plot( height ~ weight , data = Howell1 ). Then use samples from the quadratic approximate posterior of the model in (a) to superimpose on the plot: (1) the predicted mean height as a function of weight, (2) the 97% interval for the mean, and (3) the 97% interval for predicted heights.

```{r}
how_fits <-
  tibble(weight = seq(from = range(full_how$weight)[1],
                      to = range(full_how$weight)[2],
                      length.out = 100)) %>%
  mutate(log_weight = log(weight),
         log_weight_c = log_weight - mean(full_how$log_weight)) %>% 
  add_epred_draws(b4h3) %>% 
  group_by(weight) %>% 
  mean_qi(.epred, .width = .97)

how_preds <-
  tibble(weight = seq(from = range(full_how$weight)[1],
                      to = range(full_how$weight)[2],
                      length.out = 100)) %>%
  mutate(log_weight = log(weight),
         log_weight_c = log_weight - mean(full_how$log_weight)) %>% 
  add_predicted_draws(b4h3) %>% 
  group_by(weight) %>% 
  mean_qi(.prediction, .width = .97)
```

```{r}
ggplot(full_how, aes(x = weight)) +
  geom_point(aes(y = height), alpha = 0.4) +
  geom_ribbon(data = how_preds, aes(ymin = .lower, ymax = .upper),
              alpha = 0.2) +
  geom_lineribbon(data = how_fits,
                  aes(y = .epred, ymin = .lower, ymax = .upper),
                  fill = "grey60", size = 1) +
  labs(x = "Weight", y = "Height") +
  theme_minimal()
```

4H4. Plot the prior predictive distribution for the parabolic polynomial regression model in the chapter. You can modify the code that plots the linear regression prior predictive distribution. Can you modify the prior distributions of $\alpha$, $\beta_{1}$, and $\beta_{2}$ so that the prior predictions stay within the biologically reasonable outcome space? That is to say: Do not try to fit the data by hand. But do try to keep the curves consistent with what you know about height and weight, before seeing these exact data.

```{r}
n <- 1000

tibble(group = seq_len(n),
       alpha = rnorm(n, 178, 20),
       beta1 = rlnorm(n, 0, 1),
       beta2 = rnorm(n, 0, 1)) %>% 
  expand(nesting(group, alpha, beta1, beta2),
         weight = seq(25, 70, length.out = 100)) %>% 
  mutate(height = alpha + (beta1 * weight) + (beta2 * (weight ^ 2))) %>% 
  
  ggplot(aes(x = weight, y = height, group = group)) +
  geom_line(alpha = 1/20) +
  geom_hline(yintercept = c(0, 272), linetype = 2:1,
             color = "red") +
  annotate(geom = "text", x = 25, y = 0, hjust = 0, vjust = 1,
           label = "Embryo") +
  annotate(geom = "text", x = 25, y = 272, hjust = 0, 
           vjust = 0, 
           label = "World's tallest person (272cm)") +
  coord_cartesian(ylim = c(-25, 300)) +
  labs(x = "Weight", y = "Height")
```

```{r}
n <- 1000

tibble(group = seq_len(n),
       alpha = rnorm(n, -190, 5),
       beta1 = rnorm(n, 13, 0.2),
       beta2 = runif(n, -0.13, -0.1)) %>% 
  expand(nesting(group, alpha, beta1, beta2),
         weight = seq(25, 70, length.out = 100)) %>% 
  mutate(height = alpha + (beta1 * weight) + (beta2 * (weight ^ 2))) %>% 
  
  ggplot(aes(x = weight, y = height, group = group)) +
  geom_line(alpha = 1/20) +
  geom_hline(yintercept = c(0, 272), linetype = 2:1,
             color = "black") +
  annotate(geom = "text", x = 25, y = -3, hjust = 0, vjust = 1,
           label = "Embryo") +
  annotate(geom = "text", x = 25, y = 275, hjust = 0, 
           vjust = 0, 
           label = "World's tallest person (272cm)") +
  coord_cartesian(ylim = c(-25, 300)) +
  labs(x = "Weight", y = "Height") +
  theme_minimal()
```

:::question
> 4H5. Return to data(cherry_blossoms) and model the association between blossom date (doy) and March temperature (temp). Note that there are many missing values in both variables. You may consider a linear model, a polynomial, or a spline on temperature. How well does temperature trend predict the blossom trend?

```{r}
data(cherry_blossoms, package = "rethinking")

cb_temp <- cherry_blossoms %>% 
  drop_na(doy, temp) %>% 
  mutate(temp_c = temp - mean(temp),
         temp_s = temp_c / sd(temp),
         temp_s2 = temp_s ^ 2,
         temp_s3 = temp_s ^ 3)

# linear model
lin_mod <-
  brm(data = cb_temp,
      family = gaussian,
      doy ~ 1 + temp_c,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      file = "fits/b4h5-linear")

# quadratic model
qad_mod <-
  brm(data = cb_temp,
      family = gaussian,
      doy ~ 1 + temp_s + temp_s2,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b, coef = "temp_s"),
                prior(normal(0, 1), class = b, coef = "temp_s2"),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      file = "fits/b4h5-quadratic")

# cubic model
cub_mod <-
  brm(data = cb_temp,
      family = gaussian,
      doy ~ 1 + temp_s + temp_s2 + temp_s3,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b, coef = "temp_s"),
                prior(normal(0, 1), class = b, coef = "temp_s2"),
                prior(normal(0, 1), class = b, coef = "temp_s3"),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      file = "fits/b4h5-cubic")

# spline model
knots_30 <- quantile(cb_temp$temp, 
                     seq(from = 0, to = 1, length.out = 30))
B_30 <- bs(cb_temp$temp, knots = knots_30[-c(1, 30)],
           degree = 3, intercept = TRUE)

cb_temp_30 <-
  cb_temp %>% 
  mutate(B = B_30)

spl_mod <-
  brm(data = cb_temp_30,
      family = gaussian,
      doy ~ 1 + B,
      prior = c(prior(normal(100, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4h5-spline")

```


```{r}
grid <-
  tibble(temp = seq(from = range(cb_temp_30$temp)[1],
                    to = range(cb_temp_30$temp)[2],
                    length.out = 100)) %>% 
  mutate(temp_c = temp - mean(cb_temp$temp),
         temp_s = temp_c / sd(cb_temp$temp),
         temp_s2 = temp_s ^ 2,
         temp_s3 = temp_s ^ 3)

knots_30 = quantile(grid$temp, 
                    probs = seq(0,1,length.out = 30))
B_30 <-
  bs(grid$temp, knots = knots_30[-c(1, 30)],
     degree = 3, intercept = TRUE)

grid <-
  grid %>% 
  mutate(B = B_30)

fits <- bind_rows(
  add_epred_draws(grid, lin_mod) %>%
    summarize(mean_qi(.epred, .width = c(0.67, 0.89, 0.97)),
              .groups = "drop") %>% 
    select(-B) %>% 
    rename(.epred = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Linear"),
  add_epred_draws(grid, qad_mod) %>%
    summarize(mean_qi(.epred, .width = c(0.67, 0.89, 0.97)),
              .groups = "drop") %>% 
    select(-B) %>% 
    rename(.epred = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Quadratic"),
  add_epred_draws(grid, cub_mod) %>%
    summarize(mean_qi(.epred, .width = c(0.67, 0.89, 0.97)),
              .groups = "drop") %>% 
    select(-B) %>% 
    rename(.epred = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Cubic"),
  add_epred_draws(grid, spl_mod) %>%
    summarize(mean_qi(.epred, .width = c(0.67, 0.89, 0.97)),
              .groups = "drop") %>% 
    select(-B) %>% 
    rename(.epred = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Spline")
) %>%
  ungroup() %>%
  mutate(model = factor(model, levels = c("Linear", "Quadratic", "Cubic",
                                          "Spline")))

preds <- bind_rows(
  add_predicted_draws(grid, lin_mod) %>%
    summarize(mean_qi(.prediction, .width = 0.89), .groups = "drop") %>% 
    select(-B) %>% 
    rename(.prediction = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Linear"),
  add_predicted_draws(grid, qad_mod) %>%
    summarize(mean_qi(.prediction, .width = 0.89), .groups = "drop") %>% 
    select(-B) %>% 
    rename(.prediction = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Quadratic"),
  add_predicted_draws(grid, cub_mod) %>%
    summarize(mean_qi(.prediction, .width = 0.89), .groups = "drop") %>% 
    select(-B) %>% 
    rename(.prediction = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Cubic"),
  add_predicted_draws(grid, spl_mod) %>%
    summarize(mean_qi(.prediction, .width = 0.89), .groups = "drop") %>% 
    select(-B) %>% 
    rename(.prediction = y, .lower = ymin, .upper = ymax) %>% 
    mutate(model = "Spline")
) %>%
  ungroup() %>%
  mutate(model = factor(model, levels = c("Linear", "Quadratic", "Cubic",
                                          "Spline")))
```

```{r}
ggplot(cb_temp, aes(x = temp)) +
  facet_wrap(~model, nrow = 2) +
  geom_point(aes(y = doy), alpha = 0.2) +
  geom_ribbon(data = preds, aes(ymin = .lower, ymax = .upper),
              alpha = 0.2) +
  geom_lineribbon(data = fits, aes(y = .epred, ymin = .lower, ymax = .upper),
                  size = .6) +
  scale_fill_brewer(palette = "Blues", breaks = c(0.67, 0.89, 0.97)) +
  labs(x = "March Temperature", y = "Day in Year")
```

:::question
> 4H6. Simulate the prior predictive distribution for the cherry blossom spline in the chapter. Adjust the prior on the weights and observe what happens. What do you think the prior on the weights is doing?

Recreate the basis functions
```{r}
cb_dat <-
  cherry_blossoms %>% 
  drop_na(doy)

knot_list <-
  quantile(cb_dat$year, 
           probs = seq(0, 1, length.out = 15))

B <- bs(cb_dat$year,
        knots = knot_list[-c(1, 15)],
        degree = 3, intercept = TRUE)
```


```{r}
n <- 50

tibble(
  .draw = seq_len(n),
  alpha = rnorm(n, 100, 10),
  w = purrr::map(seq_len(n),
                 function(x, knots){
                   w <- rnorm(n = knots + 2, 0, 10)
                   return(w)
                 },
                 knots = 15)) %>% 
  mutate(
    mu = map2(alpha, w,
              function(alpha, w, b){
                res <- b %*% w
                res <- res + alpha
                res <- res %>% 
                  as_tibble(.name_repair = ~".value") %>% 
                  mutate(year = cb_dat$year, .before = 1)
                return(res)
              },
              b = B)) %>% 
  unnest(cols = mu) %>% 
  
  ggplot(aes(x = year, y = .value)) +
  geom_vline(xintercept = knot_list, alpha = .5) +
  geom_line(aes(group = .draw)) +
  expand_limits(y = c(60, 140)) +
  labs(x = "Year", y = "Day in Year") +
  theme_minimal()
```

Tighter prior
```{r}
n <- 50
tibble(.draw = seq_len(n),
       alpha = rnorm(n, 100, 10),
       w = purrr::map(seq_len(n),
                      function(x, knots){
                        w <- rnorm(n = knots + 2, 0, 1)
                        return(w)
                      },
                      knots = 15)) %>% 
  mutate(mu = map2(alpha, w,
                   function(alpha, w, b){
                     res <- b %*% w
                     res <- res + alpha
                     res <- res %>% 
                       as_tibble(.name_repair = ~".value") %>% 
                       mutate(year = cb_dat$year, .before = 1)
                     return(res)
                   },
                   b = B)) %>% 
  unnest(cols = mu) %>% 
  ggplot(aes(x = year, y = .value)) +
  geom_vline(xintercept = knot_list, alpha = .5) +
  geom_line(aes(group = .draw)) +
  expand_limits(y = c(60, 140)) +
  labs(x = "Year", y = "Day in Year") +
  theme_minimal()
```
:::question
> 4H8. (sic; there is no 4H7 in the text, so I’ve kept this labeled 4H8 to be consistent with the book) The cherry blossom spline in the chapter used an intercept $\alpha$
, but technically it doesn’t require one. The first basis functions could substitute for the intercept. Try refitting the cherry blossom spline without the intercept. What else about the model do you need to change to make this work?

```{r}
cb_dat <-
  cherry_blossoms %>% 
  drop_na(doy)

knots_15 <-
  quantile(cb_dat$year, probs = seq(0, 1, length.out = 15))

B_15 <-
  bs(cb_dat$year, knots = knots_15[-c(1, 15)],
     degree = 3, intercept = TRUE)

cb_dat_15 <-
  cb_dat %>% 
  mutate(B = B_15)
```

```{r}
b4h8 <-
  brm(data = cb_dat_15,
      family = gaussian,
      doy ~ 0 + B,
      prior = c(prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4h8")
```

```{r}
epred_draws(b4h8, cb_dat_15) %>% 
  summarise(mean_qi(.epred, .width = .89),
            .groups = "drop") %>% 
  ggplot(aes(x = year, y = doy)) +
  geom_point(alpha = .2) +
  geom_hline(yintercept = mean(cb_dat$doy), 
             linetype = "dashed") +
  geom_lineribbon(aes(y = y, ymin = ymin, ymax = ymax),
                  alpha = .8, fill = "#009FB7") +
  labs(x = "Year", y = "Day in Year") +
  theme_minimal()
```

```{r}
b4h8_2 <-
  brm(data = cb_dat_15,
      family = gaussian,
      doy ~ 0 + B,
      prior = c(prior(normal(100, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/b4h8_2")
```

```{r}
epred_draws(b4h8_2, cb_dat_15) %>% 
  summarise(mean_qi(.epred, .width = .89),
            .groups = "drop") %>% 
  ggplot(aes(x = year, y = doy)) +
  geom_point(alpha = 2) +
  geom_hline(yintercept = mean(cb_dat$doy), linetype = "dashed") +
  geom_lineribbon(aes(y = y, ymin = ymin, ymax = ymax),
                  alpha = .8, fill = "#009FB7") +
  labs(x = "Year", y = "Day in Year") +
  theme_minimal()
```

## 2.3 Homework
:::question
> 1. Construct a linear regression of weight as predicted by height, using the adults (age 18 or greater) from the Howell1 dataset. The heights listed below were recorded in the !Kung census, but weights were not recorded for these individuals. Provide predicted weights and 89% compatibility intervals for each of these individuals. That is, fill in the table below, using model-based predictions.

```{r}
data(Howell1, package = "rethinking")

how_dat <- 
  Howell1 %>% 
  dplyr::filter(age >= 18)

rm(Howell1)

how_dat <-
  how_dat %>% 
  mutate(height_c = height - mean(height),
         height_s = height_c / sd(height),
         weight_c = weight - mean(weight),
         weight_s = weight_c / sd(weight))
```


```{r}
w2h1 <-
  brm(data = how_dat,
      family = gaussian,
      weight_s ~ 1 + height_s,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w2h1")
```

```{r}
plot(w2h1)
```

```{r}
summary(w2h1)
```

```{r}
tibble(individual = 1:3,
         height = c(140, 160, 175)) %>% 
  mutate(height_c = height - mean(how_dat$height),
         height_s = height_c / sd(how_dat$height)) %>% 
  add_predicted_draws(w2h1) %>% 
  summarise(mean_qi(.prediction, .width = .89)) %>% 
  mutate(exp_weight = 
           (y * sd(how_dat$weight)) + mean(how_dat$weight),
         exp_weight_min = 
           (ymin * sd(how_dat$weight)) + mean(how_dat$weight),
         exp_weight_max = 
           (ymax * sd(how_dat$weight)) + mean(how_dat$weight)) %>% 
  select(individual, height, exp_weight, exp_weight_min,
         exp_weight_max)
```
:::question
> 2. From the Howell1 dataset, consider only the people younger than 13 years old. Estimate the causal association between age and weight. Assume that age influences weight through two paths. First, age influences height, and height influences weight. Second, age directly influences weight through age-related changes in muscle growth and body proportions. All of this implies this causal model (DAG):

> Use a linear regression to estimate the total (not just direct) causal effect of each year of growth on weight. Be sure to carefully consider the priors. Try using prior predictive simulation to assess what they imply.

```{r}
data(Howell1, package = "rethinking")

how_u13 <-
  Howell1 %>% 
  dplyr::filter(age < 13)

rm(Howell1)
```

Prior predictive simulation
```{r}
set.seed(123)
n <- 50

tibble(
  group = seq_len(n),
  alpha = rnorm(n, 4, 1),
  beta = rlnorm(n, 0, 1)) %>% 
  expand(nesting(group, alpha, beta), age = 0:12) %>% 
  mutate(weight = alpha + beta * age) %>% 
  
  ggplot(aes(x = age, y = weight, group = group)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  coord_cartesian(xlim = c(0, 12), ylim = c(0, 30)) +
  labs(x = "Age", y = "Weight") +
  theme_minimal()
```


```{r}
w2h2 <-
  brm(data = how_u13,
      family = gaussian,
      weight ~ 1 + age,
      prior = c(prior(normal(4, 1), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w2h2")
  
```

```{r}
plot(w2h2)
```
```{r}
summary(w2h2)
```


```{r}
draws <- gather_draws(w2h2, b_Intercept, b_age, sigma)

mean_qi(draws, .width = .89)

ggplot(draws, aes(x = .value, y = .variable)) +
  stat_halfeye(.width = .89) +
  labs(x = "Parameter value", y = "Parameter") +
  theme_minimal()
```

:::question
> 3. Now suppose the causal association between age and weight might be different for boys and girls. Use a single linear regression, with a categorical variable for sex, to estimate the total causal effect of age on weight separately for boys and girls. How do girls and boys differ? Provide one or more posterior contrasts as a summary.

```{r}
how_u13 <-
  how_u13 %>% 
  mutate(sex = male + 1,
         sex = factor(sex))

w2h3 <-
  brm(data = how_u13,
      family = gaussian,
      bf(weight ~ 0 + a + b * age,
         a ~ 0 + sex,
         b ~ 0 + sex,
         nl = TRUE),
      prior = c(prior(normal(4, 1), class = b, coef = sex1, nlpar = a),
                prior(normal(4, 1), class = b, coef = sex2, nlpar = a),
                prior(normal(0, 1), class = b, coef = sex1, nlpar = b),
                prior(normal(0, 1), class = b, coef = sex2, nlpar = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w2h3")
```

```{r}
plot(w2h3)
```

```{r}
summary(w2h3)
```
```{r}
draws <- gather_draws(
  w2h3, b_a_sex1, b_a_sex2, b_b_sex1, b_b_sex2, sigma
)

mean_qi(draws, .width = .89)

ggplot(draws, aes(x = .value, y = .variable)) +
  stat_halfeye(.width = .89) +
  labs(x = "Parameter value", y = "Parameter") +
  theme_minimal()
```


```{r}
new_age <-
  expand_grid(age = 0:12, sex = factor(c(1, 2)))

all_lines <-
  new_age %>% 
  add_epred_draws(w2h3) %>% 
  ungroup() %>% 
  mutate(group = paste0(sex, "_", .draw))

plot_lines <-
  all_lines %>% 
  dplyr::filter(.draw %in% sample(unique(.data$.draw), size = 1000)) %>% 
  select(-.draw)

animate_lines <-
  all_lines %>% 
  dplyr::filter(.draw %in% sample(unique(.data$.draw), 
                                  size = 50))

okabe_ito_palette <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999"
)

ggplot(animate_lines, 
       aes(x = age, y = .epred, color = sex, group = group)) +
  geom_line(data = plot_lines, alpha = .01, 
            show.legend = FALSE) +
  geom_point(data = how_u13, aes(x = age, y = weight,
                                  color = sex),
             inherit.aes = FALSE) +
  geom_line(alpha = 1, show.legend = FALSE, color = "black") +
  scale_color_manual(values = okabe_ito_palette, labels = c("Girls", "Boys")) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  labs(x = "Age", y = "Weight (kg)", color = NULL) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  transition_states(.draw, 0, 1) +
  theme_minimal()
```

```{r}
contrast <- new_age %>% 
  add_predicted_draws(w2h3) %>% 
  ungroup() %>% 
  select(age, sex, .draw, .prediction) %>% 
  mutate(sex = fct_recode(sex,
                          "Girls" = "1",
                          "Boys" = "2")) %>% 
  pivot_wider(names_from = sex, values_from = .prediction) %>% 
  mutate(diff = Boys - Girls)

ggplot(contrast, aes(x = age, y = diff)) +
  stat_lineribbon(aes(fill_remp = stat(.width)),
                  .width = ppoints(50),
                  fill = "#009FB7", show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  #scale_fill_gradientn(colors = c("blue", "green", "yellow", "red")) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  labs(x = "Age", y = "Weight difference (kg; Boys-Girls)") +
  theme_minimal()
```

:::question
> 4 - OPTIONAL CHALLENGE. The data in data(Oxboys) (rethinking package) are growth records for 26 boys measured over 9 periods. I want you to model their growth. Specifically, model the increments in growth from one period (Occasion in the data table) to the next. Each increment is simply the difference between height in one occasion and height in the previous occasion. Since none of these boys shrunk during the study, all of the growth increments are greater than zero. Estimate the posterior distribution of these increments. Constrain the distribution so it is always positive—it should not be possible for the model to think that boys can shrink from year to year. Finally compute the posterior distribution of the total growth over all 9 occasions.

```{r}
data(Oxboys, package = "rethinking")

increments <- Oxboys %>% 
  group_by(Subject) %>% 
  mutate(previous_height = dplyr::lag(height),
         change = height - previous_height) %>% 
  dplyr::filter(!is.na(change)) %>% 
  ungroup()
```

```{r}
set.seed(123)

n <- 1000
tibble(alpha = rnorm(n, mean = 0, sd = .3),
       sigma = rexp(n, rate = 4)) %>% 
  mutate(sim_change = rlnorm(n, meanlog = alpha, sdlog = sigma)) %>% 
  ggplot(aes(x = sim_change)) +
  geom_density() +
  labs(x = "Prior Expectation for Incremental Growth (cm)",
       y = "Density") +
  theme_minimal()
```

```{r}
w2h4 <-
  brm(data = increments,
      family = lognormal,
      change ~ 1,
      prior = c(prior(normal(0, .3), class = Intercept),
                prior(exponential(4), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w2h4")
```

Posterior distribution
```{r}
set.seed(234)

as_draws_df(w2h4) %>% 
  mutate(post_change = rlnorm(n(), meanlog = b_Intercept,
                              sdlog = sigma)) %>% 
  ggplot(aes(x = post_change)) +
  geom_density() +
  labs(x = "Posterior Distribution for Incremental Growth (cm)",
       y = "Density") +
  theme_minimal()
```

```{r}
set.seed(345)

as_draws_df(w2h4) %>% 
  mutate(all_change = map2_dbl(b_Intercept, sigma, ~sum(rlnorm(8, .x, .y)))) %>% 
  ggplot(aes(x = all_change)) +
  stat_halfeye(aes(fill = stat(between(x, 8.6, 18.5))), color = NA) +
  scale_fill_manual(values = c("#009FB7", NA),
                    labels = "89% Compatibility Interval",
                    breaks = "TRUE", na.value = "#F0F0F0") +
  labs(x = "Posterior Distribution for Cumulative Growth (cm)",
       y = "Density", fill = NULL)
```

