---
title: "13 Models With Memory"
format: html
---
Benefits of multilevel model:

- better estimates for repeated sampling (9.e, in longitudinal data),
- better estimates when there are imbalances among subsamples,
- estimates of the variation across subsamples, and
- avoiding simplistic averaging by retaining variation across subsamples.

## 13.1 Example: Multilevel tadpoles
```{r}
library(brms)
data(reedfrogs, package = "rethinking")
d <- reedfrogs
rm(reedfrogs)
```

```{r}
library(tidyverse)

d %>% glimpse()
```

```{r}
d <-
  d %>% mutate(tank = 1:nrow(d))
```

```{r}
d %>% count(density)
```

```{r}
b13.1 <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 0 + factor(tank),
      prior(normal(0, 1.5), class = b),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13.01")
```

```{r}
print(b13.1)
```

```{r}
library(ggthemes)
library(tidybayes)

# change the default
theme_set(theme_gray() + theme_fivethirtyeight())

tibble(estimate = fixef(b13.1)[, 1]) %>% 
  mutate(p = inv_logit_scaled(estimate)) %>% 
  pivot_longer(estimate:p) %>% 
  mutate(name = if_else(name == "p",
                        "expected survival probability",
                        "expected survival log-odds")) %>% 
  
  ggplot(aes(x = value, fill = name)) +
  stat_dots(size = 0) +
  scale_fill_manual(values = c("orange1", "orange4")) +
  scale_y_continuous(breaks = NULL) +
  labs(title = "Tank-level intercepts from the no-pooling model",
       subtitle = "Notice now inspecting the distributions of the posterior means can offer insights you\nmight not get if you looked at them one at a time") +
  theme(legend.position = "none",
        panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free_x")
```

Even though it seems like we can derive important insights from how the tank-level intercepts are distributed, that information is not explicitly encoded in the statistical model. Keep that in mind as we now consider the multilevel alternative. Its formula is
$$
\begin{align}
\text{surv}_i & \sim \text{Binomial}(n_i, p_i) \\
\text{logit}(p_i) & = \alpha_{\text{tank}[i]} \\
\alpha_j & \sim \text{Normal}(\bar{\alpha}, \sigma) \\
\bar{\alpha} & \sim \text{Normal}(0, 1.5) \\
\sigma & \sim \text{Exponential}(1),
\end{align}
$$

```{r}
b13.2 <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank),
      prior = c(prior(normal(0, 1.5), class = Intercept), # alpha bar
                prior(exponential(1), class = sd)), # sigma
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      sample_prior = "yes",
      seed = 13,
      file = "fits/b13.02")
```

```{r}
b13.1 <- add_criterion(b13.1, "waic")
b13.2 <- add_criterion(b13.2, "waic")

w <- loo_compare(b13.1, b13.2, criterion = "waic")

print(w, simplify = F)
```

```{r}
cbind(waic_diff = w[, 1] * -2,
      se = w[, 2] * 2)
```

```{r}
model_weights(b13.1, b13.2, weights = "waic") %>% 
  round(digits = 2)
```

```{r}
w[, "p_waic"]
```

```{r}
print(b13.2)
```

```{r}
post <- as_draws_df(b13.2)

post_mdn <-
  coef(b13.2, robust = T)$tank[, , ] %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  mutate(post_mdn = inv_logit_scaled(Estimate))

head(post_mdn)
```

```{r}
post_mdn %>% 
  ggplot(aes(x = tank)) +
  geom_hline(yintercept =
               inv_logit_scaled(median(post$b_Intercept)),
             linetype = 2, linewidth = 1/4) +
  geom_vline(xintercept = c(16.5, 32.5), linewidth = 1/4,
             color = "grey25") +
  geom_point(aes(y = propsurv), color = "orange2") +
  geom_point(aes(y = post_mdn), shape = 1) +
  annotate(geom = "text",
           x = c(8, 16 + 8, 32 + 8), y = 0,
           label = c("small tanks", "medium tanks", "large tanks")) +
  scale_x_continuous(breaks = c(1, 16, 32, 48)) +
  scale_y_continuous(breaks = 0:5 / 5, limits = c(0, 1)) +
  labs(title = "Multilevel shrinkage!",
        subtitle = "The empirical proportions are in orange while the model\n-implied proportions are the black circles. The dashed line is\nthe model-implied average survival proportion.") +
  theme(panel.grid.major = element_blank())
```

```{r}
set.seed(13)

p1 <-
  post %>% 
  slice_sample(n = 100) %>% 
  expand_grid(x = seq(from = -4, to = 5, length.out = 100)) %>% 
  mutate(density = dnorm(x, mean = b_Intercept, sd = sd_tank__Intercept)) %>% 
  
  ggplot(aes(x = x, y = density, group = .draw)) +
  geom_line(alpha = .2, color = "orange2") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Population survival distribution",
       subtitle = "log-odds scale") +
  coord_cartesian(xlim = c(-3, 4))
```

```{r}
set.seed(13)

p2 <-
  post %>% 
  slice_sample(n = 8000, replace = T) %>% 
  mutate(sim_tanks = rnorm(n(),
                           mean = b_Intercept,
                           sd = sd_tank__Intercept)) %>% 
  
  ggplot(aes(x = inv_logit_scaled(sim_tanks))) +
  geom_density(linewidth = 0, fill = "orange2", adjust = 0.1) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Probability of survival",
       subtitle = "transformed by the inverse-logit function")
```

```{r}
library(patchwork)

(p1 + p2) & 
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 10))
```

#### 13.1.0.2 Overthinking: Prior for variance components.
```{r}
b13.2b <-
  update(b13.2,
         prior = c(prior(normal(0, 1.5), class = Intercept),
                   prior(normal(0, 1), class = sd)),
         iter = 5000, warmup = 1000, chains = 4, cores = 4,
         sample_prior = "yes",
         seed = 13,
         file = "fits/b13.02b")
```

```{r}
print(b13.2b)
```

If you're curious how the exponential and half-Normal priors compare to one another and to their posterios, you might just plot
```{r}
# for annotation
text <-
  tibble(value = c(0.5, 2.4),
         density = c(1, 1.85),
         distribution = factor(c("prior", "posterior"),
                               levels = c("prior", "posterior")),
         prior = "Exponential(1)")

# gather and wrangle the prior and posterior draws
tibble(`prior_Exponential(1)` = prior_draws(b13.2) %>% pull(sd_tank),
       `posterior_Exponential(1)` = as_draws_df(b13.2) %>% pull(sd_tank__Intercept),
       `prior_Half-Normal(0, 1)` = prior_draws(b13.2b) %>% pull(sd_tank),
       `posterior_Half-Normal(0, 1)` = as_draws_df(b13.2b) %>% pull(sd_tank__Intercept)) %>% 
  pivot_longer(everything(),
               names_sep = "_",
               names_to = c("distribution", "prior")) %>% 
  mutate(distribution = factor(distribution,
                               levels = c("prior", "posterior"))) %>% 
  
  # plot!
  ggplot(aes(x = value, fill = distribution)) +
  geom_density(linewidth = 0, alpha = 2/3, adjust = 0.25) +
  geom_text(data = text,
            aes(y = density, label = distribution, color = distribution)) +
  scale_fill_manual(NULL, values = c("orange4", "orange2")) +
  scale_color_manual(NULL, values = c("orange4", "orange2")) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = expression(Hierarchical~sigma~parameter)) +
  coord_cartesian(xlim = c(0, 4)) +
  theme(legend.position = "none") +
  facet_wrap(~ prior)
```

## 13.2 Varying effects and the underfitting/overfitting trade-off

### 13.2.1 The model.
The simulation formula should look familiar
$$
\begin{align}
\text{surv}_{i} & \sim \text{Binomial}(n_i, p_i) \\
\text{logit}(p_i) & = \alpha_{\text{pond}[i]} \\
\alpha_j & \sim \text{Normal}(\bar{\alpha}, \sigma) \\
\bar{\alpha} & \sim \text{Normal}(0, 1.5) \\
\sigma & \sim \text{Exponential}(1)
\end{align}
$$

### 13.2.2 Assign values to the parameters.

```{r}
a_bar <- 1.5
sigma <- 1.5
n_ponds <- 60

set.seed(5005)

dsim <-
  tibble(pond = 1:n_ponds,
         ni = rep(c(5, 10, 25, 35), each = n_ponds / 4) %>%
           as.integer(),
         true_a = rnorm(n = n_ponds, mean = a_bar, sd = sigma))

head(dsim)
```
McElreath twice urged us to inspect the contents of this simulation. In addition to looking at the data with head(), we might well plot
```{r}
dsim %>% 
  mutate(ni = factor(ni)) %>% 
  
  ggplot(aes(x = true_a, y = ni)) +
  stat_dotsinterval(fill = "orange2", slab_size = 0, .width = .5) +
  ggtitle("Log-odds varying by # tadpoles per pond") +
  theme(plot.title = element_text(size = 14))
```

### 13.2.3 Simulate survivors
> Each pond $i$ has $n_i$ potential survivors, and nature flips each tadpole's coin, so to speak, with probability of survival $p_i$. This probability $p_i$ is implied by the model definition, and is equal to:
$$
p_i = \frac{\text{exp}(\alpha_i)}{1 + \text{exp}(\alpha_i)}
$$
The model uses a logit link, and so the probability is defined by the ```[inv_logit_scaled()]``` function.

```{r}
set.seed(5005)

(
  dsim <-
    dsim %>% 
    mutate(si = rbinom(n = n(), 
                       prob = inv_logit_scaled(true_a),
                       size = ni))
)
```

### 13.2.4 Compute the no-pooling estimates.
```{r}
(
  dsim <-
    dsim %>% 
    mutate(p_nopool = si / ni)
)
```

### 13.2.5 Compute the partial-pooling estimates
Fit the multilevel partial-pooling model.
```{r}
b13.3 <-
  brm(data = dsim,
      family = binomial,
      si | trials(ni) ~ 1 + (1 | pond),
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(exponential(1), class = sd)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13.03")
```

```{r}
print(b13.3)
```

```{r}
b13.3$fit
```

```{r}
p_partpool <-
  coef(b13.3)$pond[, , ] %>% 
  data.frame() %>% 
  transmute(p_partpool = inv_logit_scaled(Estimate))

dsim <-
  dsim %>% 
  bind_cols(p_partpool) %>% 
  mutate(p_true = inv_logit_scaled(true_a)) %>% 
  mutate(nopool_error = abs(p_nopool - p_true),
         partpool_error = abs(p_partpool - p_true))

dsim %>% glimpse()
```

```{r}
dfline <-
  dsim %>% 
  select(ni,nopool_error:partpool_error) %>% 
  pivot_longer(-ni) %>% 
  group_by(name, ni) %>% 
  summarise(mean_error = mean(value)) %>% 
  mutate(x = c(1, 16, 31, 46),
         xend = c(15, 30, 45, 60))

dsim %>% 
  ggplot(aes(x = pond)) +
  geom_vline(xintercept = c(15.5, 30.5, 45.4), 
             color = "white", linewidth = 2/3) +
  geom_point(aes(y = nopool_error), color = "orange2") +
  geom_point(aes(y = partpool_error), shape = 1) +
  geom_segment(data = dfline,
               aes(x = x, xend = xend,
                   y = mean_error, yend = mean_error),
               color = rep(c("orange2", "black"), each = 4),
               linetype = rep(1:2, each = 4)) +
  annotate(geom = "text",
           x = c(15 - 7.5, 30 - 7.5, 45 - 7.5, 60 - 7.5),
           y = .45,
           label = c("tiny (5)", "small (10)", "medium (25)",
                     "large (35)")) +
  scale_x_continuous(breaks = c(1, 10, 20, 30, 40, 50, 60)) +
  labs(title = "Estimate error by model type",
       subtitle = "The horizontal axis displays pond number. The vertical axis measures\nthe absolute error in the predicted proportion of survivors, compared to\nthe true value used in the simulation. The higher the point, the worse\nthe estimate. No-pooling shown in orange. Partial pooling shown in black.\nThe orange and dashed black lines show the average error for each kind\nof estimate, across each initial density of tadpoles (pond size).") +
  theme(panel.grid.major = element_blank(),
        plot.subtitle = element_text(size = 10))
```

If you wanted to quantify the difference in simple summaries, you might execute something like this.
```{r}
dsim %>% 
  select(ni, nopool_error:partpool_error) %>% 
  pivot_longer(-ni) %>% 
  group_by(name) %>% 
  summarise(mean_error = mean(value) %>% round(digits = 3),
            median_error = median(value) %>% round(digits = 3))
```

#### 13.2.5.1 Overthinking: Repeating the pond simulation.
```{r}
a_bar <- 1.5
sigma <- 1.5
n_ponds <- 60

set.seed(1999)

new_dsim <-
  tibble(pond = 1:n_ponds,
         ni = rep(c(5, 10, 25, 35), each = n_ponds / 4) %>% 
           as.integer(),
         true_a = rnorm(n = n_ponds, mean = a_bar, sd = sigma)) %>% 
  mutate(si = rbinom(n = n(), prob = inv_logit_scaled(true_a), size = ni)) %>% 
  mutate(p_nopool = si / ni)

glimpse(new_dsim)
```
Fit the new model
```{r}
b13.3_new <-
  update(b13.3,
         newdata = new_dsim,
         chains = 4, cores = 4,
         seed = 13,
         file = "fits/b13.03_new")
```

```{r}
print(b13.3_new)
```

Plot the first simulation versus the second one
```{r}
bind_rows(as_draws_df(b13.3),
          as_draws_df(b13.3_new)) %>% 
  mutate(model = rep(c("b13.3", "b13.3_new"),
                     each = n() / 2)) %>% 
  ggplot(aes(x = b_Intercept, y = sd_pond__Intercept)) +
  stat_density_2d(geom = "raster",
                  aes(fill = after_stat(density)),
                  contour = F, n = 200) +
  geom_vline(xintercept = a_bar, color = "orange3", linetype = 3) +
  geom_hline(yintercept = sigma, color = "orange3", linetype = 3) +
  scale_fill_gradient(low = "grey25", high = "orange3") +
  ggtitle("Our simulation posterior contrast a bit",
          subtitle = expression(alpha*" is on the x and "*sigma*" is on the y, both in log-odds. The dotted lines intersect at the true values.")) +
  coord_cartesian(xlim = c(.7, 2),
                  ylim = c(.8, 1.9)) +
  theme(legend.position = "none",
        panel.grid.major = element_blank()) +
  facet_wrap(~ model, ncol = 2)
```

```{r}
b13.3$fit@stanmodel
```
## 13.3 More than one type of cluster
#### 13.3.0.1 Rethinking: Cross-classification and hierarchy.
## 13.3.1 Multilevel chimpanzees.

The initial multilevel update from model b11.4 follows the statistical formula
$$
\begin{align}
\text{left_pull}_i & \sim \text{Binomial}(n_i = 1, p_i) \\
\text{logit}(p_i) & = \alpha_{\text{actor}[i]} + \gamma_{\text{block}[i]} + \beta_{\text{treatment}[i]} \\
\beta_j & \sim \text{Normal}(0, 0.5) \hspace{1em} ,\text{for} j = 1, ..., 4 \\
\alpha_j & \sim \text{Normal}(\bar{\alpha}, \sigma_\alpha) \hspace{1em} ,\text{for} j = 1, ..., 7 \\
\gamma_j & \sim \text{Normal}(0, \sigma_\gamma) \hspace{1em} ,\text{for} j = 1, ..., 6 \\
\bar{\alpha} & \sim \text{Normal}(0, 1.5) \\
\sigma_\alpha & \sim \text{Exponential}(1) \\
\sigma_\gamma & \sim \text{Exponential}(1).
\end{align}
$$

The statistical model using non-centered parameterization more faithful to the way it will be expressed with brms::brm():

$$
\begin{align}
\text{left_pull} & \sim \text{Binomial}(n_i = 1, p_i) \\
\text{logit}(p_i) & = \bar{\alpha} + \beta_{\text{treatment}[i]} + z_{\text{actor}[i]}\sigma_\alpha + x_{\text{block}[i]}\sigma_\gamma \\
\bar{\alpha} & \sim \text{Normal}(0, 1.5) \\
\beta_j & \sim \text{Normal}(0, 0.5) \hspace{1em} ,\text{for} j = 1,...,4 \\
z_j & \sim \text{Normal}(0, 1) \\
x_j & \sim \text{Normal}(0, 1) \\
\sigma_\alpha & \sim \text{Exponential}(1) \\
\sigma_\gamma & \sim \text{Exponential}(1).
\end{align}
$$

```{r}
rm(list = ls())

data(chimpanzees, package = "rethinking")
d <- chimpanzees
rm(chimpanzees)
```

```{r}
d <-
  d %>% 
  mutate(actor = factor(actor),
         block = factor(block),
         treatment = factor(1 + prosoc_left + 2 * condition))

glimpse(d)
```

```{r}
b13.4 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + b,
         a ~ 1 + (1 | actor) + (1 | block),
         b ~ 0 + treatment,
         nl = TRUE),
      prior = c(prior(normal(0, .5), nlpar = b),
                prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
                prior(exponential(1), class = sd, group = actor, nlpar = a),
                prior(exponential(1), class = sd, group = block, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13.04")
```

```{r}
library(bayesplot)

color_scheme_set("orange")

as_draws_df(b13.4) %>% 
  mcmc_trace(pars = vars(b_a_Intercept:`r_block__a[6,Intercept]`),
             facet_args = list(ncol = 4),
             linewidth = 0.15) +
  theme(legend.position = "none")
```

```{r}
print(b13.4)
```

```{r}
posterior_summary(b13.4) %>% round(digits = 2)
```

Make a coefficient plot using ```bayesplot::mcmc_plot()```
```{r}
b13.4 %>% 
  mcmc_plot(variable = c("^r_", "^b_", "^sd_"), regex = T) +
  theme(axis.text.y = element_text(hjust = 0))
```

For a little more control, we might switch to a tidybayes-oriented approach.
```{r}
# extract the posterior draws
post <- as_draws_df(b13.4)

# this is all stylistic fluff
levels <-
  c("sd_block__a_Intercept", "sd_actor__a_Intercept",
    "b_a_Intercept",
    str_c("r_block__a[", 6:1, ",Intercept]"),
    str_c("r_actor__a[", 7:1, ",Intercept]"),
    str_c("b_b_treatment", 4:1))

text <-
  tibble(x = posterior_summary(b13.4, probs = c(0.055, 0.955),)["r_actor__a[2,Intercept]", c(3,1)],
         y = c(13.5, 16.5),
         label = c("89% CI", "mean"),
         hjust = c(.5, 0))

arrow <-
  tibble(x = posterior_summary(b13.4, probs = c(.055, .955),)["r_actor__a[2,Intercept]", c(3,1)] + c(-0.3, 0.2),
         xend = posterior_summary(b13.4, probs = c(.055, .955),)["r_actor__a[2,Intercept]", c(3,1)],
         y = c(14, 16),
         yend = c(14.8, 15.35))

# here's the main event
post %>% 
  pivot_longer(b_a_Intercept:`r_block__a[6,Intercept]`) %>% 
  mutate(name = factor(name, levels = levels)) %>% 
  
  ggplot(aes(x = value, y = name)) +
  stat_pointinterval(point_interval = mean_qi,
                     .width = .89, shape = 21, size = 1,
                     point_size = 2, point_fill = "blue") +
  geom_text(data = text,
            aes(x = x, y = y, label = label, hjust = hjust)) +
  geom_segment(data = arrow,
               aes(x = x, xend = xend, y = y, yend = yend),
               arrow = arrow(length = unit(0.15, "cm"))) +
  theme(axis.text.y = element_text(hjust = 0),
        panel.grid.major.y = element_line(linetype = 3))
```

```{r}
post %>% 
  pivot_longer(starts_with("sd")) %>% 
  
  ggplot(aes(x = value, fill = name)) +
  geom_density(linewidth = 0, alpha = 3/4, adjust = 2/3,
               show.legend = F) +
  annotate(geom = "text", x = .67, y = 2, label = "block",
           color = "orange4") +
  annotate(geom = "text", x = 2.725, y = .5, label = "actor",
           color = "orange1") +
  scale_fill_manual(values = str_c("orange", c(1, 4))) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle(expression(sigma["<group>"])) +
  coord_cartesian(xlim = c(0, 4))
```

Since both the coefficient plots and the density plots indicate there is much more variability among the actor parameters than in the block parameters, we might fit a model that ignores the variation among the levels of block.
```{r}
b13.5 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + b,
         a ~ 1 + (1 | actor),
         b ~ 0 + treatment,
         nl = TRUE),
      prior = c(prior(normal(0, .5), nlpar = b),
                prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
                prior(exponential(1), class = sd, group = actor, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13.05")
```

```{r}
b13.4 <- add_criterion(b13.4, "waic")
b13.5 <- add_criterion(b13.5, "waic")

loo_compare(b13.4, b13.5, criterion = "waic") %>% print(simplify = F)
```

```{r}
model_weights(b13.4, b13.5, weights = "waic") %>% 
  round(digits = 2)
```

### 13.3.2 Even more clusters.
We can extend partial pooling to the treatment conditions, too. With brms, it will be more natural to revert to the conventional formula syntax.

```{r}
b13.6 <-
  brm(data = d,
      family = binomial,
      pulled_left | trials(1) ~ 1 + (1 | actor) + (1 | block) + (1 | treatment),
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(exponential(1), class = sd)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13.06")
```

```{r}
tibble(parameter = str_c("b[", 1:4, "]"),
       `b13.4` = fixef(b13.4)[2:5, 1],
       `b13.6` = ranef(b13.6)$treatment[, 1, "Intercept"]) %>% 
  mutate_if(is.double, round, digits = 2)
```

Compare group-level $\sigma$ parameters with a plot
```{r}
as_draws_df(b13.6) %>% 
  pivot_longer(starts_with("sd")) %>% 
  mutate(group = str_remove(name, "sd_") %>% str_remove(., "__Intercept")) %>% 
  mutate(parameter = str_c("sigma[", group, "]")) %>% 
  
  ggplot(aes(x = value, y = parameter)) +
  stat_halfeye(.width = .95, size = 1, fill = "orange", adjust = .1) +
  labs(subtitle = "The variation among treatment levels is small, but the\nvariation among the levels of block is still the smallest.") +
  theme(axis.text.y = element_text(hjust = 0))
```

Among the three $\sigma_{<group>}$ parameters, $\sigma_{block}$ is the smallest. Now we'll compare b13.6 to the last two models with the WAIC.
```{r}
b13.6 <- add_criterion(b13.6, "waic")

loo_compare(b13.4, b13.5, b13.6, criterion = "waic") %>% 
  print(simplify = F)
```
```{r}
model_weights(b13.4, b13.5, b13.6, weights = "loo") %>% 
  round(digits = 2)
```

## 13.4 Divergent transitions and non-centered priors
### 13.4.1 The Devil's Funnel
McElreath posed a joint distribution
$$
\begin{align}
v & \sim \text{Normal}(0, 3) \\
x & \sim \text{Normal}(0, \text{exp}(v)),
\end{align}
$$
where the scale of $x$ depends on another variable, $v$. In R code 13.26, McElreath then proposed fifting the following model with ```rethinking::ulam()````
```{r}
m13.7 <-
  rethinking::ulam(
    data = list(N = 1),
    alist(
      v ~ normal(0, 3),
      x ~ normal(0, exp(v))
    ),
    chains = 4)
```

```{r}
set.seed(13)

tibble(v = rnorm(1e3, mean = 0, sd = 3)) %>% 
  mutate(x = rnorm(1e3, mean = 0, sd = exp(v))) %>% 
  
  ggplot(aes(x = x)) +
  geom_histogram(binwidth = 1, fill = "orange2") +
  annotate(geom = "text",
           x = -100, y = 490, hjust = 0,
           label = expression(italic(v)%~%Normal(0, 3))) +
  annotate(geom = "text",
           x = -100, y = 440, hjust = 0,
           label = expression(italic(x)%~%Normal(0, exp(italic(v))))) +
  coord_cartesian(xlim = c(-100, 100)) +
  scale_y_continuous(breaks = NULL)
```

The distribution looks something like a Student-t with a very low v parameter. We can express the joint likelihood of $p(v, x)$ as
$$
p(v, x) = p(x|v)p(v).
$$

Here that is in a plot
```{r}
# define the parameter space
parameter_space <- seq(from = -4, to = 4, length.out = 200)

# simulate
crossing(v = parameter_space,
         x = parameter_space) %>% 
  mutate(likelihood_v = dnorm(v, mean = 0, sd = 3),
         likelihood_x = dnorm(x, mean = 0, sd = exp(v))) %>% 
  mutate(joint_likelihood = likelihood_v * likelihood_x) %>% 
  
  # plot!
  ggplot(aes(x = x, y = v, fill = joint_likelihood)) +
  geom_raster(interpolate = T) +
  scale_fill_viridis_c(option = "B") +
  labs(subtitle = "Centered parameterization") +
  theme(legend.position = "none")
```

To avoid the divergent transitions that can arise from steep valleys like this, we can switch from our original formula to a non-centered parametrization, such as:
$$
\begin{align}
v & \sim \text{Normal}(0, 3) \\
z & \sim \text{Normal}(0, 1) \\
x & = z\text{exp}(v),
\end{align} 
$$

where $x$ is now the product of two independent distributions, $v$ and $z$. With this parameterization, we can express the joint likelihood $p(v, z)$ as
$$
p(v, z) = p(z)p(v),
$$

where $p(z)$ is not conditional on $v$ and $p(v)$ is not conditional on $z$. Here's what that looks like on a plot.
```{r}
# simulate
crossing(v = parameter_space,
         z = parameter_space / 2) %>% 
  mutate(likelihood_v = dnorm(v, mean = 0, sd = 3),
         likelihood_z = dnorm(z, mean = 0, sd = 1)) %>% 
  mutate(joint_likelihood = likelihood_v * likelihood_z) %>% 
  
  # plot!
  ggplot(aes(x = z, y = v, fill = joint_likelihood)) +
  geom_raster(interpolate = T) +
  scale_fill_viridis_c(option = "B") +
  labs(subtitle = "Non-centered parameterization") +
  theme(legend.position = "none")
```

### 13.4.2 Non-centered chimpanzees
```{r}
print(b13.4)
```

```{r}
library(posterior)

as_draws_df(b13.4) %>% 
  summarise_draws()
```

```{r}
as_draws_df(b13.4) %>% 
  summarise_draws() %>% 
  
  ggplot(aes(x = ess_bulk, y = ess_tail)) +
  geom_abline(linetype = 2) +
  geom_point(color = "blue") +
  xlim(0, 4700) +
  ylim(0, 4700) +
  ggtitle("Effective sample size summaries for b13.4",
          subtitle = "ess_bulk is on the x and ess_tail is on the y") +
  theme(plot.subtitle = element_text(size = 10),
        plot.title = element_text(size = 11.5),
        plot.title.position = "plot")
```

## 13.5 Multilevel posterior predictions
### 13.5.1 Posterior prediction for same clusters.
```{r}
chimp <- 2

nd <-
  d %>% 
  distinct(treatment) %>% 
  mutate(actor = chimp,
         block = 1)

labels <- c("R/N", "L/N", "R/P", "L/P")

f <-
  nd %>% 
  add_epred_draws(b13.4) %>% 
  mean_qi() %>% 
  mutate(treatment = factor(treatment, labels = labels))
```

Here are the empirical probabilities computed directly from the data (i.e., the no-pooling model).
```{r}
(
  chimp_2_d <-
    d %>% 
    filter(actor == chimp) %>% 
    group_by(treatment) %>% 
    summarise(prob = mean(pulled_left)) %>% 
    ungroup() %>% 
    mutate(treatment = factor(treatment, labels = labels))
)
```

```{r}
f %>% 
  # if you want to use `geom_line()` or `geom_ribbon()` with a factor on the x-axis,
  # you need to code something like `group = 1` in `aes()`
  ggplot(aes(x = treatment, y = .epred, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "orange1") +
  geom_line(color = "blue") +
  geom_point(data = chimp_2_d,
             aes(y = prob),
             color = "grey25") +
  ggtitle("Chimp #2",
          subtitle = "The posterior mean and 95%\nintervals are the blue line\nand orange band, respectively.\nThe empirical means are the charcoal dots.") +
  coord_cartesian(ylim = c(.75, 1)) +
  theme(plot.subtitle = element_text(size = 10))
```

```{r}
post <- as_draws_df(b13.4)

glimpse(post)
```

```{r}
post %>% 
  transmute(actor_5 = `r_actor__a[5,Intercept]`) %>% 
  
  ggplot(aes(x = actor_5)) +
  geom_density(linewidth = 0, fill = "blue") +
  scale_y_continuous(breaks = NULL) +
  ggtitle("Chimp $5's density")
```

Because we made the density only using the ```r_actor__a[5,Intercept]```` values (i.e., we didn't add ```b_Intercept``` to them), the density is in a deviance-score metric.

McElreath built his own link function. With this particular model, it will be easiest for us to just work directly with post.
```{r}
f <-
  post %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(fitted = inv_logit_scaled(b_a_Intercept + value + `r_actor__a[1,Intercept]` + `r_block__a[1,Intercept]`)) %>% 
  mutate(treatment = factor(str_remove(name, "b_b_treatment"),
                            labels = labels)) %>% 
  select(name:treatment)

f
```

Now we'll summarize those values and compute their empirical analogues directly from the data.
```{r}
# the posterior summaries
(
  f <-
    f %>% 
    group_by(treatment) %>% 
    mean_qi(fitted)
)
```

```{r}
# the empirical summaries
chimp <- 5

(
  chimp_5_d <-
    d %>% 
    filter(actor == chimp) %>% 
    group_by(treatment) %>% 
    summarise(prob = mean(pulled_left)) %>% 
    ungroup() %>% 
    mutate(treatment = factor(treatment, labels = labels))
)
```

Let's see how good we are retrodicting the pulled_left probabilities for actor == 5.
```{r}
f %>% 
  ggplot(aes(x = treatment, y = fitted, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "orange1") +
  geom_line(color = "blue") +
  geom_point(data = chimp_5_d,
             aes(y = prob),
             color = "grey25") +
  ggtitle("Chimp #5",
          subtitle = "This plot is like the last except\nwe did more by hand.") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.subtitle = element_text(size = 10))
```

### 13.5.2 Posterior prediction for new clusters
By average actor, McElreath referred to a chimp with an intercept exactly at the population mean $\bar{\alpha}$. Given our non-centered parameterization for b13.4, this means we’ll leave out the random effects for actor. Since we’re predicting what might happen in new experimental blocks, we’ll leave out the random effects for block, too. When doing this by hand, the workflow is much like is was before, just with fewer columns added together within the first mutate() line.
```{r}
f <-
  post %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(fitted = inv_logit_scaled(b_a_Intercept + value)) %>% 
  mutate(treatment = factor(str_remove(name, "b_b_treatment"),
                            labels = labels)) %>% 
  select(name:treatment) %>% 
  group_by(treatment) %>% 
  # note we're using 80% intervals
  mean_qi(fitted, .width = .8)

f
```

```{r}
p1 <-
  f %>% 
  ggplot(aes(x = treatment, y = fitted, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.title = element_text(size = 14, hjust = .5))

p1
```

If we want to depict the variability across the chimps, we need to include ```sd_actor__a_Intercept``` into the calculations. In the first block of code, below, we simulate a bundle of new intercepts defined by
$$
\text{simulated chimpanzees} \sim \text{Normal}(\bar{\alpha}, \sigma_\alpha).
$$

As before, we are also averaging over block
```{r}
set.seed(13)

f <-
  post %>% 
  # simulated chimpanzees
  mutate(a_sim = rnorm(n(), mean = b_a_Intercept,
                       sd = sd_actor__a_Intercept)) %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(fitted = inv_logit_scaled(a_sim + value)) %>% 
  mutate(treatment = factor(str_remove(name, "b_b_treatment"),
                            labels = labels)) %>% 
  group_by(treatment) %>% 
  # note we're using 80% intervals
  mean_qi(fitted, .width = .8)

f
```

```{r}
p2 <-
  f %>% 
  ggplot(aes(x = treatment, y = fitted, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "orange1") +
  geom_line(color = "blue") +
  ggtitle("Marginal of actor") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.title = element_text(size = 14, hjust = .5))

p2
```

```{r}
# how many simulated chimps would you like?
n_chimps <- 100

set.seed(13)

f <-
  post %>% 
  slice_sample(n = n_chimps) %>% 
  # simulated chimpanzees
  mutate(a_sim = rnorm(n(), mean = b_a_Intercept,
                       sd = sd_actor__a_Intercept)) %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(fitted = inv_logit_scaled(a_sim + value)) %>% 
  mutate(treatment = factor(str_remove(name, "b_b_treatment"),
                            labels = labels)) %>% 
  select(.draw:treatment)

f
```

```{r}
p3 <-
  f %>% 
  ggplot(aes(x = treatment, y = fitted, group = .draw)) +
  geom_line(alpha = 1/2, color = "orange3") +
  ggtitle("100 simulated actors") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.title = element_text(size = 14, hjust = .5))

p3
```

```{r}
library(patchwork)

p1 | p2 | p3
```

#### 13.5.2.1 Bonus: Let's use fitted this time.
```{r}
nd <- distinct(d, treatment)

(
  f <-
    nd %>% 
    add_epred_draws(b13.4, re_formula = NA) %>% 
    mean_qi(.width = .9) %>% 
    mutate(treatment = factor(treatment, labels = labels))
)
```

You should notice a few things. Since b13.4 is a cross-classified multilevel model, it had three predictors: treatment, block, and actor. However, our nd data only included the first of those three. The reason fitted() permitted that was because we set re_formula = NA. When you do that, you tell fitted() to ignore group-level effects (i.e., focus only on the fixed effects). This was our fitted() version of ignoring the r_ vectors returned by as_draws_df(). Here's the plot. 
```{r}
p4 <-
  f %>% 
  ggplot(aes(x = treatment, y = .epred, group = 1)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), fill = "blue") +
  geom_line(color = "orange1") +
  ggtitle("Average actor") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.title = element_text(size = 14, hjust = .5))

p4
```

```{r}
(
  f <-
    fitted(b13.4,
           newdata = nd,
           probs = c(.1, .9),
           allow_new_levels = T,
           sample_new_levels = "gaussian") %>% 
    data.frame() %>% 
    bind_cols(nd) %>% 
    mutate(treatment = factor(treatment, labels = labels))
)
```

```{r}
p5 <-
  f %>% 
  ggplot(aes(x = treatment, y = Estimate, group = 1)) +
  geom_ribbon(aes(ymin = Q10, ymax = Q90), fill = "blue") +
  geom_line(color = "orange1") +
  labs(title = "Marginal of actor") +
  coord_cartesian(ylim = 0:1) +
  theme(plot.title = element_text(size = 14, hjust = .5))

p5
```

```{r}
# how many simulated chimps would you like?
n_chimps <- 100

nd <-
  distinct(d, treatment) %>% 
  # define 100 new actors
  expand_grid(actor = str_c("new", 1:n_chimps)) %>% 
  arrange(actor, treatment) %>% 
  # this adds a row number, which will come in handy, later
  mutate(row = 1:n())

# fitted
set.seed(13)

f <-
  fitted(b13.4,
         newdata = nd,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         summary = F,
         ndraws = n_chimps)
```

```{r}
p6 <-
  f %>% 
  data.frame() %>% 
  # name the columns by the row values in nd
  set_names(pull(nd, row)) %>% 
  # add a draw index
  mutate(draw = 1:n()) %>% 
  # make it long
  pivot_longer(-draw, names_to = "row") %>% 
  mutate(row = as.double(row)) %>% 
  # add the new data
  left_join(nd, by = "row") %>% 
  # extract the numbers from the names of the new actors
  mutate(actor_number = str_extract(actor, "\\d+") %>% 
           as.double()) %>% 
  # only keep the posterior iterations that match the `actor_number` values
  filter(actor_number == draw) %>% 
  # add the `treatment` labels
  mutate(treatment = factor(treatment, labels = labels)) %>% 
  
  # plot!
  ggplot(aes(x = treatment, y = value, group = actor)) +
  geom_line(alpha = 1/2, color = "blue") +
  ggtitle("100 simulated actors") +
  theme(plot.title = element_text(size = 14, hjust = .5))

p6
```

```{r}
p4 | p5 | p6
```

### 13.5.3 Post-stratification.
If you have estimates $p_i$ for each relevant demographic category $i$, the post-stratified prediction for the whole population just re-weights these estimates using the number of individuals $N_i$ in each category with the formula
$$
\frac{\Sigma_iN_ip_i}{\Sigma_iN_i}
$$

You can find a more comprehensive introduction to post-stratification in Chapter 17 of Gelman et al. Within the multilevel context, this approach is called multilevel regression an post-stratification. Gelman is a long-time advocate for MRP.

## 13.6 Bonus: Post-stratification in an example
### 13.6.1 Meet the data.
```{r}
rm(list = ls())

load("data/mrp_data_ch13.rds")

glimpse(d)
```

```{r}
glimpse(cell_counts)
```

```{r}
cell_counts %>% count()
```

```{r}
cell_counts %>% 
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 2000, fill = "blue") +
  scale_x_continuous(breaks = 0:3 * 100000,
                     labels = c(0, "100K", "200K", "300K"))
```

### 13.6.2 Settle the MR part of MRP.

Look at the scheme of prior of $\bar{\alpha} = \text{Normal}(-1,1)$

```{r}
set.seed(13)

tibble(n = rnorm(1e6, mean = -1, sd = 1)) %>% 
  mutate(p = inv_logit_scaled(n)) %>% 
  
  ggplot(aes(x = p)) +
  geom_histogram(fill = "blue", binwidth = .02, boundary = 0) +
  scale_y_continuous(breaks = NULL)
```

```{r}
b13.7 <-
  brm(data = d,
      family = binomial,
      kept_name | trials(1) ~ 1 + (1 | age_group) + (1 | decade_married) + (1 | educ_group) + (1 | state_name),
      prior = c(prior(normal(-1, 1), class = Intercept),
                prior(exponential(1), class = sd)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      control = list(adapt_delta = .95),
      file = "fits/b13.07")
```

```{r}
print(b13.7)
```
Even with 4373 cases in the data, the uncertainty around $\bar{\alpha}$ is massive, -0.72[-2, 0.46], suggesting a lot of the action is lurking in the $\sigma_{<group>}$ parameters. It might be easier to compare the $\sigma_{<group>}$ parameters with an interval plot.
```{r}
as_draws_df(b13.7) %>% 
  select(starts_with("sd_")) %>% 
  set_names(str_c("sigma[", c("age", "decade~married",
                              "education", "state"),
                  "]")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(.width = seq(from = .5, to = .9, by = .1)) %>% 
  
  ggplot(aes(x = value, xmin = .lower, xmax = .upper,
             y = reorder(name, value))) +
  geom_interval(aes(alpha = .width), color = "orange3") +
  scale_alpha_continuous("CI width", range = c(.7, .15)) +
  scale_y_discrete(labels = ggplot2:::parse_safe) +
  xlim(0, NA) +
  theme(axis.text.y = element_text(hjust = 0),
        panel.grid.major.y = element_blank())
```

### 13.6.3 Post-stratify to put the P in MRP.
#### 13.6.3.1 Estimates by age group
```{r}
levels <- c("raw data", "multilevel", "MRP")

p1 <-
  # compute the proportions from the data
  d %>% 
  group_by(age_group, kept_name) %>% 
  summarise(n = n()) %>% 
  group_by(age_group) %>% 
  mutate(prop = n / sum(n),
         type = factor("raw data", levels = levels)) %>% 
  filter(kept_name == 1, age_group < 80, age_group > 20) %>% 
  
  # plot!
  ggplot(aes(x = prop, y = age_group)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, .5, 1), limits = 0:1) +
  facet_wrap(~ type)

p1
```

We'll combine the plot with the next two, in a bit. I just wanted to give a preview of what we're doing. The second plot will showcase the typical multilevel estimates for the same. The most straightforward way to do this with brms is with the fitted() function. We'll use the re_formula argument to average over the levels of all grouping variables other than age_group. Relatedly, we'll feed in the unique levels of age_group into the newdata argument. Then we just wrangle and plot.
```{r}
nd <- distinct(d, age_group) %>% arrange(age_group)

p2 <-
  fitted(b13.7,
         re_formula = ~ (1 | age_group),
         newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  mutate(prop = Estimate,
         type = factor("multilevel", levels = levels)) %>% 
  
  ggplot(aes(x = prop, xmin = Q2.5, xmax = Q97.5, y = age_group)) + 
  geom_pointrange(color = "blue2", linewidth = 0.8, fatten = 2) +
  scale_x_continuous(breaks = c(0, .5, 1), limits = c(0, 1)) +
  scale_y_discrete(labels = NULL) +
  facet_wrap(~ type)
```

Compute the MRP estimates
```{r}
age_prop <-
  cell_counts %>% 
  group_by(age_group) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

age_prop
```

```{r}
p <-
  add_predicted_draws(b13.7,
                      newdata = age_prop %>% 
                        filter(age_group > 20,
                               age_group < 80,
                               decade_married > 1969),
                      allow_new_levels = T)

glimpse(p)
```

Next comes the MRP magic. If we group the results by age_group and .draw, we can sum the product of the posterior predictions and the weights, which will leave us with 4000 stratified posterior draws for each of the 11 levels of age_group. This is the essence of the post-stratification equation McElreath presented in Section 13.5.3
$$
\frac{\Sigma_iN_ip_i}{\Sigma_iN_i}
$$

We will follow Alexander and call these summary values kept_name_predict. We then complete the porject by grouping by age_group and summarizing each stratified posterior predictive distribution by its mean and 95% interval.
```{r}
p <-
  p %>% 
  group_by(age_group, .draw) %>% 
  summarise(kept_name_predict = sum(.prediction * prop)) %>% 
  group_by(age_group) %>% 
  mean_qi(kept_name_predict)

p
```

```{r}
# MRP plot
p3 <-
  p %>% 
  mutate(type = factor("MRP", levels = levels)) %>% 
  
  ggplot(aes(x = kept_name_predict, xmin = .lower,
             xmax = .upper, y = age_group)) +
  geom_pointrange(color = "orange2", linewidth = .8,
                  fatten = 2) +
  scale_x_continuous(breaks = c(0, .5, 1), limits = 0:1) +
  scale_y_discrete(labels = NULL) +
  facet_wrap(~ type)

# combine!
(p1 | p2 | p3) +
  plot_annotation(title = "Proportion of women keeping name after marriage, by age",
                  subtitle = "Proportions are on the x-axis and age groups are on the y-axis.")
```

#### 13.6.3.2 Estimates by US state.
```{r}
library(statebins)

p1 <-
  d %>% 
  group_by(state_name, kept_name) %>% 
  summarise(n = n()) %>% 
  group_by(state_name) %>% 
  mutate(prop = n / sum(n)) %>% 
  filter(kept_name == 1,
         state_name != "puerto rico") %>% 
  mutate(type = factor("raw data", levels = levels),
         statename = str_to_title(state_name)) %>% 
  
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins(lbl_size = 2.5, border_size = 1/4, radius = grid::unit(2, "pt")) +
  scale_fill_viridis_c("proportion\nkeeping\nname", option = "B", limits = c(0, 0.8)) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  theme(legend.position = "none") +
  facet_wrap(~ type)

p1
```

```{r}
nd <- distinct(d, state_name)

p2 <-
  fitted(b13.7,
         re_formula = ~ (1 | state_name),
         newdata = nd,
         allow_new_levels = T) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  filter(state_name != "puerto rico") %>% 
  mutate(prop = Estimate,
         type = factor("multilevel", levels = levels),
         statename = str_to_title(state_name)) %>% 
  
  ggplot(aes(fill = prop, state = statename)) + 
  geom_statebins(lbl_size = 2.5, border_size = 1/4, radius = grid::unit(2, "pt")) +
  scale_fill_viridis_c("proportion\nkeeping\nname", option = "B", limits = c(0, 0.8)) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  facet_wrap(~ type)
```

```{r}
state_prop <- 
  cell_counts %>% 
  group_by(state_name) %>% 
  mutate(prop = n/sum(n))  %>% 
  ungroup()

state_prop
```

```{r}
p3 <-
  add_predicted_draws(b13.7,
                      newdata = state_prop %>% 
                        filter(age_group > 20, 
                               age_group < 80, 
                               decade_married > 1969),
                      allow_new_levels = T) %>%
  group_by(state_name, .draw) %>% 
  summarise(kept_name_predict = sum(.prediction * prop)) %>% 
  group_by(state_name) %>% 
  mean_qi(kept_name_predict) %>% 
  mutate(prop      = kept_name_predict,
         type      = factor("MRP", levels = levels),
         statename = str_to_title(state_name)) %>% 
  
  ggplot(aes(fill = kept_name_predict, state = statename)) + 
  geom_statebins(lbl_size = 2.5, border_size = 1/4, radius = grid::unit(2, "pt")) +
  scale_fill_viridis_c("proportion\nkeeping\nname", option = "B", limits = c(0, 0.8)) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  theme(legend.position = "none") +
  facet_wrap(~ type)
```

```{r}
(p1 | p2 | p3) +
  plot_annotation(title = "Proportion off women keeping name after marriage, by state",
                  theme = theme(plot.margin = margin(0.2, 0, 0.01, 0, "cm")))
```

