---
title: "Chapter 13 Practice"
format: html
---

```{r}
library(tidyverse)
library(brms)
library(tidybayes)
```

```{r}
rm(list = ls())
```

:::question
13M1. Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size treatment variables to the varying intercepts model. Consider models with either main effect alone, both main effects, as well as a model including both and their interaction. Instead of focusing on inferences about these two predictor variables, focus on the inferred variation across tanks. Explain why it changes as it does across models.

```{r}
data(reedfrogs, package = "rethinking")
d <- reedfrogs 
d <- d %>%
  mutate(tank = 1:nrow(d))
rm(reedfrogs)
```

```{r}
# model with tank variation alone
b13m1_0 <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank),
      prior = c(prior(normal(0, 10), class = Intercept), # alpha bar
                prior(exponential(1), class = sd)), # sigma
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m1_0")
```

```{r}
# model with tank variation and predation
b13m1_p <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank) + factor(pred),
      prior = c(prior(normal(0, 10), class = Intercept), # alpha bar
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd, group = tank)), # sigma
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m1_p")
```

```{r}
# model with tank variation and size
b13m1_s <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank) + factor(size),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m1_s")
```

```{r}
b13m1_p_s <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank) + factor(pred) + factor(size),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m1_p_s")
```

```{r}
b13m1_p_s_ps <-
  brm(data = d,
      family = binomial,
      surv | trials(density) ~ 1 + (1 | tank) + factor(pred) + factor(size) + factor(pred):factor(size),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m1_p_s_ps")
```

```{r}
post_0 <- as_draws_df(b13m1_0)
post_p <- as_draws_df(b13m1_p)
post_p_s <- as_draws_df(b13m1_p_s)
post_p_s_ps <- as_draws_df(b13m1_p_s_ps)
post_s <- as_draws_df(b13m1_s)
```

```{r}
post_0 <-
  post_0 %>% mutate(sd_prop = inv_logit_scaled(sd_tank__Intercept))

post_p <-
  post_p %>% mutate(sd_prop = inv_logit_scaled(sd_tank__Intercept))

post_p_s <-
  post_p_s %>% mutate(sd_prop = inv_logit_scaled(sd_tank__Intercept))

post_p_s_ps <-
  post_p_s_ps %>% mutate(sd_prop = inv_logit_scaled(sd_tank__Intercept))

post_s <-
  post_s %>% mutate(sd_prop = inv_logit_scaled(sd_tank__Intercept))
```


```{r}
get_mean <- function(post){
  output <- 
    post %>% 
    select(starts_with("r_")) %>% 
    pivot_longer(everything()) %>% 
    group_by(name) %>% 
    mean_hdci(value) %>% 
    mutate(tank = as.numeric(str_extract(name, "\\d+"))) %>% 
    mutate(prop = inv_logit_scaled(value))
  
  return(output)
}


post_0_mean <- get_mean(post_0)
post_p_mean <- get_mean(post_p)
post_p_s_mean <- get_mean(post_p_s)
post_p_s_ps_mean <- get_mean(post_p_s_ps)
post_s_mean <- get_mean(post_s)
```

```{r}
ggplot() +
  geom_density(data = post_0,
               aes(x = sd_prop),
               color = "blue") +
  geom_density(data = post_p,
               aes(x = sd_prop),
               color = "lightblue") +
  geom_density(data = post_p_s,
               aes(x = sd_prop),
               color = "orange1") +
  geom_density(data = post_p_s_ps,
               aes(x = sd_prop),
               color = "lightpink") +
  geom_density(data = post_s,
               aes(x = sd_prop),
               color = "lightgreen") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

:::question
> 13M2. Compare the models you fit just above, using WAIC. Can you reconcile the differences in WAIC with the posterior distributions of the model?

```{r}
b13m1_0 <- add_criterion(b13m1_0, criterion = "waic")
b13m1_p <- add_criterion(b13m1_p, criterion = "waic")
b13m1_p_s <- add_criterion(b13m1_p_s, criterion = "waic")
b13m1_p_s_ps <- add_criterion(b13m1_p_s_ps, criterion = "waic")
b13m1_s <- add_criterion(b13m1_s, criterion = "waic")
```

```{r}
loo_compare(b13m1_0, b13m1_p, b13m1_p_s, b13m1_p_s_ps, b13m1_s, criterion = "waic")
```

Model without size differs only a bit in terms of WAIC comparison suggesting that size doesn't really impact survival. We can check the parameters for size further.

```{r}
b13m1_s$fit
```

We can see that size parameters overlap with zero as it's 95% credibility interval [-0.53, 1.25].

:::question
> 13M3. Re-estimate the basic Reed frog varying intercept model, but now using a Cauchy distribution in place of the Gaussian distribution for the varying intercepts. That is, fit this model:

$$
\begin{align}
s_i & \sim \text{Binomial}(n_i, p_i) \\
\text{logit}(p_i) & = \alpha_{TANK[i]} \\
\alpha_{TANK} & \sim \text{Cauchy}(\alpha, \sigma) \\
\alpha & \sim \text{Normal}(0, 1) \\
\sigma & \sim \text{Exponential}(1)
\end{align}
$$

(You are likely to see many divergent transitions for this model. Can you figure out why? Can you fix them?) Compare the posterior means of the intercepts, $\alpha_{tank}$, to the posterior means produced in the chapter, using the customary Gaussian prior. Can you explain the pattern of differences? Take note of any change in the mean $\alpha$ as well.

```{r}
b13m3 <-
  brm(data = d,
      family = binomial(link = "logit"),
      surv | trials(density) ~ 1 + (1 | tank),
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(exponential(1), class = sd, group = tank)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13m3")
```

:::question
>   13H1. In 1980, a typical Bengali woman could have 5 or more children in her lifetime. By the year 2000, a typical Bengali woman had only 2 or 4. You're going to look at a historical set of data, when contraception was widely available but many families chose not to use it. These data reside in data(bangladesh) and come from the 1988 Bangladesh Fertility Survey. Each row is one of 1934 women. There are six variables, but you can focus on two of them for this practice problem:
(1) district: ID number of administrative district each woman resided in
(2) use.contraception: An indicator (0/1) of whether the woman was using contraception

The first thing to do is ensure that the cluster variable, district, is a contiguous set of integers. Recall that these values will be index values inside the model. If there are gaps, you'll have parameters for which there is no data to inform them. Worse, the model probably won't run. Look at the unique values of the district variable:

```{r}
rm(list = ls())

data(bangladesh, package = "rethinking")
d <- bangladesh
rm(bangladesh)
```

```{r}
sort(unique(d$district))
```
District 54 is absent. So district isn't yet a good index variable, because it's not contiguous. This is easy to fix. Just make a new variable that is contiguous. This is enough to do it:
```{r}
d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))
```
Now there are 60 values, contiguous integers 1 to 60. Now, focus on predicting use.contraception, clustered by district_id. Fit both 
1. A traditional fixed-effects model that uses an index variable for district
2. Multilevel model with varying intercepts for district.
Plot the predicted proportions of women in each district using contraception, for both the fixed-effects model and the varying-effects model. That is, make a plot in which district ID is on the horizontal axis and expected proportion using contraception is on the vertical. Make one plot for each model, or layer them on the same plot, as you prefer. How do the models disagree? Can you explain the pattern of disagreement? In particular, can you explain the most extreme cases of disagreement, both why they happen where they do and why the models reach different inferences?

```{r}
b13h1_fixed <-
  brm(data = d,
      family = binomial,
      use.contraception | trials(1) ~ 0 + factor(district_id),
      prior = c(prior(normal(0, 10), class = b)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13h1_fixed")
```

```{r}
b13h1_varying <-
  brm(data = d,
      family = binomial,
      use.contraception | trials(1) ~ 1 + (1 | district_id),
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(student_t(1, 0, 1), class = sd)),
      iter = 5000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13h1_varying")
```


```{r}
print(b13h1_fixed)
```

```{r}
print(b13h1_varying)
```

```{r}
new_dat <- tibble(district_id = 1:60)
```

```{r}
post_fixed <- 
  new_dat %>% 
  add_epred_draws(b13h1_fixed) %>% 
  mean_hdci()

post_varying <-
  new_dat %>% 
  add_epred_draws(b13h1_varying) %>% 
  mean_hdci()

d_district <-
  d %>% 
  group_by(district_id) %>% 
  summarise(use_contraception = sum(use.contraception),
            woman = n()) %>% 
  mutate(p_contraception = use_contraception / woman)
```

```{r}
ggplot() +
  geom_point(data = d_district,
             aes(x = district_id,
                 y = p_contraception),
             shape = 1) +
  geom_point(data = post_fixed,
             aes(x = district_id,
                 y = .epred),
             color = "lightblue",
             shape = 1) +
  geom_point(data = post_varying,
             aes(x = district_id,
                 y = .epred),
             color = "lightgreen",
             shape = 1) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

:::question
> 13H2. Return to data(Trolley) from Chapter 12. Define and fit a varying intercepts model for these data. Cluster intercepts on individual participants, as indicated by the unique values in the id variable. Include action, intention, and contact as ordinary terms. Compare the varying intercepts model and a model that ignores individuals, using both WAIC and posterior predictions. What is the impact of individual variation in these data?

```{r}
rm(list = ls())
data(Trolley, package = "rethinking")

d <- 
  Trolley %>% 
  select(response, action, intention, contact, id, story) %>% 
  mutate(id = dense_rank(id),
         story = dense_rank(story))
```

```{r}
b13h2_f <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + action + contact + intention,
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b13h2_f")
```

```{r}
b13h2_v <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + (1 | id) + action + contact + intention,
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b13h2_v")
```

```{r}
print(b13h2_f)
```

```{r}
print(b13h2_v)
```

```{r}
b13h2_f <- add_criterion(b13h2_f, criterion = "waic")
b13h2_v <- add_criterion(b13h2_v, criterion = "waic")
```

```{r}
loo_compare(b13h2_f, b13h2_v, criterion = "waic")
```

```{r}
sigma_id <-
  b13h2_v %>% 
  as_draws_df() %>% 
  pull(sd_id__Intercept) %>% 
  as_tibble()
  
sigma_id %>% 
  ggplot() +
  geom_density(aes(x = value)) +
  geom_vline(xintercept = mean(sigma_id$value),
             linetype = "dashed",
             color = "gray") +
  theme_bw() +
  theme(panel.grid = element_blank())
  
```

There is a lot of variation because the sigma value (which is on the log-odd scale) is 1.9. To check the individual variation let's simulate a one scenario but for multiple person
```{r}
d_pred <- tibble(
  response = 7,
  action = 0,
  intention = 1,
  contact = 0,
  id = 1:331)
```

```{r}
pred <-
  d_pred %>% 
  add_epred_draws(b13h2_v, ndraws = 1000)

exp_pred <-
  pred %>% 
  mutate(response = as.integer(.category) * .epred) %>%  # Multiply category by probability %>% 
  group_by(.draw, id) %>% 
  summarize(avg_response = sum(response), .groups = "drop")

final_results <- exp_pred %>% 
  group_by(id) %>% 
  summarise(mean_response = mean(avg_response), .groups = "drop")
```

```{r}
final_results %>% 
  ggplot(aes(x = id, y = mean_response)) +
  geom_point(shape = 1) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(y = "average response")
```

:::question
> 13H3. The Trolley data are also clustered by story, which indicates a unique narrative for each vignette. Define and fit a cross-classified varying intercepts model with both id and story. Use the same ordinary terms as in the previous problem. Compare this model to the previous models. What do you infer about the impact of different stories on response.

```{r}
b13h3 <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + (1 | id) + (1 | story) + action + intention + contact,
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/b13h3")
```
```{r}
print(b13h3)
```

```{r}
b13h2_v <- add_criterion(b13h2_v, criterion = "waic")
b13h3 <- add_criterion(b13h3, criterion = "waic")
```

```{r}
post <- as_draws_df(b13h3)

sd <-
  post$sd_story__Intercept %>% 
  mean_hdi(.width = .95)

as_tibble(post$sd_story__Intercept) %>% 
  ggplot() +
  geom_density(aes(x = value)) +
  geom_vline(xintercept = mean(post$sd_story__Intercept),
             linetype = "dashed",
             color = "gray") +
  theme_bw() +
  theme(panel.grid = element_blank())
```


```{r}
loo_compare(b13h2_v, b13h3, criterion = "waic")
```

```{r}
new_dat <-
  tibble(
    response = 7,
    action = 0,
    intention = 1,
    contact = 0,
    id = 1,
    story = 1:12)
```

```{r}
pred_new_dat <-
  new_dat %>% 
  add_epred_draws(b13h3, ndraws = 1000) %>% 
  mutate(response = as.integer(.category) * .epred) %>%  # Multiply category by probability %>% 
  group_by(.draw, story) %>% 
  summarize(avg_response = sum(response), .groups = "drop") %>%
  group_by(story) %>% 
  summarise(mean_response = mean(avg_response), .groups = "drop")

pred_new_dat %>% 
  ggplot(aes(x = story, y = mean_response)) +
  geom_point(shape = 1) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(y = "average response") +
  scale_x_continuous(breaks = 1:12)
```

:::question
> 13H4. Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size treatment variables to the varying intercepts model. Consider models with either predictor alone, both predictors, as well as a model including their interaction. What do you. infer about the causal influence of these predictor variables? Also focus on the inferred variation across tanks (the $\sigma$ across tanks). Explain why it changes as it does across models with different predictors included.

```{r}
rm(list = ls())
data(reedfrogs, package = "rethinking")
d <- reedfrogs %>% 
  mutate(tank = 1:n())
rm(reedfrogs)
```

```{r}
b13h4_p <-
  brm(
    data = d,
    family = binomial,
    formula = surv | trials(density) ~ 1 + (1 | tank) + pred,
    prior = c(prior(normal(0, 1), class = Intercept),
              prior(normal(0, 1), class = b),
              prior(exponential(1), class = sd)),
    iter = 4000, warmup = 2000, chains = 8, cores = 8,
    seed = 13,
    file = "fits/b13h4_p"
  )
```

```{r}
print(b13h4_p)
```

```{r}
b13h4_s <-
  brm(
    data = d,
    family = binomial,
    formula = surv | trials(density) ~ 1 + (1 | tank) + size,
    prior = c(prior(normal(0, 1), class = Intercept),
              prior(normal(0, 1), class = b),
              prior(exponential(1), class = sd)),
    iter = 4000, warmup = 2000, chains = 8, cores = 8,
    seed = 13,
    file = "fits/b13h4_s"
  )
```
```{r}
print(b13h4_s)
```

```{r}
b13h4_p_s <-
  brm(
    data = d,
    family = binomial,
    formula = surv | trials(density) ~ 1 + (1 | tank) + pred + size,
    prior = c(prior(normal(0, 1), class = Intercept),
              prior(normal(0, 1), class = b),
              prior(exponential(1), class = sd)),
    iter = 4000, warmup = 2000, chains = 8, cores = 8,
    seed = 13,
    file = "fits/b13h4_p_s"
  )
```
```{r}
print(b13h4_p_s)
```

```{r}
b13h4_p_s_ps <-
  brm(
    data = d,
    family = binomial,
    formula = surv | trials(density) ~ 1 + (1 | tank) + pred + size + pred:size,
    prior = c(prior(normal(0, 1), class = Intercept),
              prior(normal(0, 1), class = b),
              prior(exponential(1), class = sd)),
    iter = 4000, warmup = 2000, chains = 8, cores = 8,
    seed = 13,
    file = "fits/b13h4_p_s_ps"
  )
```

```{r}
print(b13h4_p_s_ps)
```

```{r}
b13h4_p <- add_criterion(b13h4_p, criterion = "waic")
b13h4_s <- add_criterion(b13h4_s, criterion = "waic")
b13h4_p_s <- add_criterion(b13h4_p_s, criterion = "waic")
b13h4_p_s_ps <- add_criterion(b13h4_p_s_ps, criterion = "waic")
```

```{r}
loo_compare(b13h4_p, b13h4_s, b13h4_p_s, b13h4_p_s_ps, criterion = "waic")
```

```{r}
sd_p <- as_draws_df(b13h4_p) %>% pull(sd_tank__Intercept) %>% as_tibble()
sd_s <- as_draws_df(b13h4_s) %>% pull(sd_tank__Intercept) %>% as_tibble()
sd_p_s <- as_draws_df(b13h4_p_s) %>% pull(sd_tank__Intercept) %>% as_tibble()
sd_p_s_ps <- as_draws_df(b13h4_p_s_ps) %>% pull(sd_tank__Intercept) %>% as_tibble()
```

```{r}
sd_summary <-
  bind_rows(
  sd_p %>% mean_hdci() %>% mutate(model = "b13h4_p"),
  sd_s %>% mean_hdci() %>% mutate(model = "b13h4_s"),
  sd_p_s %>% mean_hdci() %>% mutate(model = "b13h4_p_s"),
  sd_p_s_ps %>% mean_hdci() %>% mutate(model = "b13h4_p_s_ps"),
)

sd_summary
```


```{r}
ggplot() +
  geom_density(data = sd_p, aes(x = value), color = "blue") +
  geom_density(data = sd_s, aes(x = value), color = "lightblue") +
  geom_density(data = sd_p_s, aes(x = value), color = "red") +
  geom_density(data = sd_p_s_ps, aes(x = value), color = "yellow") +
  theme_bw()
```

