---
title: "7 Ulysses' Compass"
format: html
---

## 7.1 The problem with parameters
### 7.1.1 More parameters (almost) always improve fit
```{r}
library(tidyverse)

(
  d <- 
  tibble(species = c("afarensis", "africanus", "habilis", "boisei", "rudolfensis", "ergaster", "sapiens"), 
         brain   = c(438, 452, 612, 521, 752, 871, 1350), 
         mass    = c(37.0, 35.5, 34.5, 41.5, 55.5, 61.0, 53.5))
)
```

```{r}
library(rcartocolor)
```

```{r}
carto_pal(7, "BurgYl")
```

```{r}
display_carto_pal(7, "BurgYl")
```

```{r}
library(ggrepel)

theme_set(
  theme_classic() +
    theme(text = element_text(family = "Courier"),
          panel.background = element_rect(fill = alpha(carto_pal(7, "BurgYl")[3], 1/4)))
)

d %>% 
  ggplot(aes(x = mass, y = brain, label = species)) +
  geom_point(color = carto_pal(7, "BurgYl")[5]) +
  geom_text_repel(size = 3, color = carto_pal(7, "BurgYl")[7],
                  family = "Courier", seed = 438) +
  labs(subtitle = "Average brain volume by body\nmass for six hominin species",
       x = "body mass (kg)",
       y = "brain volume (cc)") +
  xlim(30, 65)
```

```{r}
d <-
  d %>% 
  mutate(mass_std = (mass - mean(mass)) / sd(mass),
         brain_std = brain / max(brain))
```

```{r}
library(brms)
```

```{r}
b7.1 <-
  brm(data = d,
      family = gaussian,
      brain_std ~ 1 + mass_std,
      prior = c(prior(normal(0.5, 1), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(lognormal(0, 1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 7,
      file = "fits/b07.01")
```

```{r}
print(b7.1)
```

```{r}
R2_is_bad <- function(brm_fit, seed = 7, ...){
  
  set.seed(seed)
  p <- predict(brm_fit, summary = F, ...)
  # in my experience, it's more typical to define residuals as the criterion minus the predictions
  r <- d$brain_std - apply(p, 2, mean)
  1 - rethinking::var2(r) / rethinking::var2(d$brain_std)
}
```

```{r}
R2_is_bad(b7.1)
```

```{r}
# quadratic
b7.2 <-
  update(b7.1,
         newdata = d,
         formula = brain_std ~ 1 + mass_std + I(mass_std^2),
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 7,
         file = "fits/b07.02")

# cubic
b7.3 <-
  update(b7.1,
         newdata = d,
         formula = brain_std ~ 1 + mass_std + I(mass_std^2) +
           I(mass_std^3),
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 7,
         control = list(adapt_delta = .9),
         file = "fits/b07.03")

# fourth-order
b7.4 <-
  update(b7.1,
         newdata = d,
         formula = brain_std ~ 1 + mass_std + I(mass_std^2) +
           I(mass_std^3) + I(mass_std^4),
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 7,
         control = list(adapt_delta = .995),
         file = "fits/b07.04")

# fifth-order
b7.5 <- 
  update(b7.1,
         newdata = d, 
         formula = brain_std ~ 1 + mass_std + I(mass_std^2) + I(mass_std^3) + I(mass_std^4) + I(mass_std^5),
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 7,
         control = list(adapt_delta = .99999),
         file = "fits/b07.05")
```

```{r}
custom_normal <- custom_family(
  "custom_normal", dpars = "mu",
  links = "identity",
  type = "real"
)

stan_funs  <- "real custom_normal_lpdf(real y, real mu) {
  return normal_lpdf(y | mu, 0.001);
}
real custom_normal_rng(real mu) {
  return normal_rng(mu, 0.001);
}
" 

stanvars <- stanvar(scode = stan_funs, block = "functions")
```

```{r}
b7.6 <- 
  brm(data = d, 
      family = custom_normal,
      brain_std ~ 1 + mass_std + I(mass_std^2) + I(mass_std^3) + I(mass_std^4) + I(mass_std^5) + I(mass_std^6),
      prior = c(prior(normal(0.5, 1), class = Intercept),
                prior(normal(0, 10), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 7,
      stanvars = stanvars,
      file = "fits/b07.06")
```

```{r}
expose_functions(b7.6, vectorize = TRUE)

posterior_epred_custom_normal <- function(prep) {
  mu <- prep$dpars$mu
  mu 
}

posterior_predict_custom_normal <- function(i, prep, ...) {
  mu <- prep$dpars$mu
  mu 
  custom_normal_rng(mu)
}

log_lik_custom_normal <- function(i, prep) {
  mu <- prep$dpars$mu
  y <- prep$data$Y[i]
  custom_normal_lpdf(y, mu)
}
```

```{r}
library(tidybayes)

nd <- tibble(mass_std = seq(from = -2, to = 2, length.out = 100))

nd %>% 
  add_epred_draws(b7.1) %>% 
  mean_qi(.epred, .width = .89) %>% 
  
  ggplot(aes(x = mass_std, y = .epred)) +
  geom_lineribbon(aes(ymin = .lower, ymax = .upper),
              color = carto_pal(7, "BurgYl")[7],
              linewidth = 1/2, 
              fill = alpha(carto_pal(7, "BurgYl")[6], 1/3)) +
  geom_point(data = d,
              aes(y = brain_std),
              color = carto_pal(7, "BurgYl")[7]) +
  labs(subtitle = bquote(italic(R)^2 == .(round(R2_is_bad(b7.1), digits = 2))),
       x = "body mass (standardized)",
       y = "brain volume (standardized)") +
  coord_cartesian(xlim = range(d$mass_std))
```

```{r}
make_figure7.3 <- function(brms_fit, ylim = range(d$brain_std)){
  
  # compute the R2
  r2 <- R2_is_bad(brms_fit)

  # define the new data
  nd <- tibble(mass_std = seq(from = -2, to = 2, length.out = 200))
  
  # simulate and wrangle
  add_epred_draws(nd, brms_fit) %>% 
    mean_qi(.epred, .width = .89) %>% 
    
    # plot!
    ggplot(aes(x = mass_std)) +
    geom_lineribbon(aes(y = .epred, ymin = .lower, ymax = .upper),
                    color = carto_pal(7, "BurgYl")[7],
                    linewidth = 1/2, 
                    fill = alpha(carto_pal(7, "BurgYl")[6], 1/3)) +
    geom_point(data = d,
               aes(y = brain_std),
               color = carto_pal(7, "BurgYl")[7]) +
    labs(subtitle = bquote(italic(R)^2==.(round(r2, digits = 2))),
         x = "body mass (std)",
         y = "brain volume (std)") +
    coord_cartesian(xlim = c(-1.2, 1.5),
                    ylim = ylim)
  
}
```

```{r}
p1 <- make_figure7.3(b7.1)
p2 <- make_figure7.3(b7.2)
p3 <- make_figure7.3(b7.3)
p4 <- make_figure7.3(b7.4, ylim = c(.25, 1.1))
p5 <- make_figure7.3(b7.5, ylim = c(.1, 1.4))
p6 <- make_figure7.3(b7.6, ylim = c(-0.25, 1.5)) +
  geom_hline(yintercept = 0, color = carto_pal(7, "BurgYl")[2], linetype = 2) 
```

```{r}
library(patchwork)

((p1 | p2) / (p3 | p4) / (p5 | p6)) +
  plot_annotation(title = "Figure7.3. Polynomial linear models of increasing\ndegree for the hominin data.")
```

### 7.1.2 Too few parameters hurts, too.
```{r}
d %>% 
  mutate(row = 1:n()) %>% 
  filter(row_number() != 2)
```

```{r}
library(rethinking)
brain_loo_plot
```

```{r}
b7.1.1 <-
  update(b7.1,
         newdata = filter(d, row_number() != 1),
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 7,
         file = "fits/b07.01.1")

print(b7.1.1)
```

```{r}
nd %>% 
  add_epred_draws(b7.1.1) %>% 
  mean_qi(.epred, .width = .89) %>% 
  
  ggplot(aes(x = mass_std, y = .epred)) +
  geom_line(color = carto_pal(7, "BurgYl")[7], 
            linewidth = 1/2, alpha = 1/2) +
  geom_point(data = d,
             aes(y = brain_std),
             color = carto_pal(7, "BurgYl")[7]) +
  labs(subtitle = "b7.1.1",
       x = "body mass (std)",
       y = "brain volume (std)") +
  coord_cartesian(xlim = range(d$mass_std),
                  ylim = range(d$brain_std))
```

```{r}
brain_loo_lines <- function(brms_fit, row, ...){
  
  # refit the model
  new_fit <-
    update(brms_fit,
           newdata = filter(d, row_number() != row),
           iter = 2000, warmup = 1000, chains = 4, cores = 4,
           seed = 7,
           refresh = 0,
           ...)
  
  # pull the lines values
  add_epred_draws(nd, new_fit) %>% 
    mean_qi(.epred) %>% 
    select(.epred) %>% 
    bind_cols(nd) %>% 
    rename(Estimate = .epred)
}
```

```{r}
brain_loo_lines(b7.1, row = 1) %>% 
  glimpse()
```

```{r}
b7.1_fits <-
  tibble(row = 1:7) %>% 
  mutate(post = purrr::map(row, ~brain_loo_lines(brms_fit = b7.1, row = .))) %>% 
  unnest(post)

b7.4_fits <-
  tibble(row = 1:7) %>% 
  mutate(post = purrr::map(row, ~brain_loo_lines(brms_fit = b7.4, 
                                                 row = ., 
                                                 control = list(adapt_delta = .999)))) %>% 
  unnest(post)
```

```{r}
# left
p1 <-
  b7.1_fits %>% 
  
  ggplot(aes(x = mass_std)) +
  geom_line(aes(y = Estimate, group = row),
            color = carto_pal(7, "BurgYl")[7], linewidth = 1/2, alpha = 1/2) +
  geom_point(data = d,
             aes(y = brain_std),
             color = carto_pal(7, "BurgYl")[7]) +
  labs(subtitle = "b7.1",
       x = "body mass (std)",
       y = "brain volume (std)") +
  coord_cartesian(xlim = range(d$mass_std),
                  ylim = range(d$brain_std))

# right
p2 <-
  b7.4_fits %>%  
  
  ggplot(aes(x = mass_std, y = Estimate)) +
  geom_line(aes(group = row),
            color = carto_pal(7, "BurgYl")[7], linewidth = 1/2, alpha = 1/2) +
  geom_point(data = d,
             aes(y = brain_std),
             color = carto_pal(7, "BurgYl")[7]) +
  labs(subtitle = "b7.4",
       x = "body mass (std)",
       y = "brain volume (std)") +
  coord_cartesian(xlim = range(d$mass_std),
                  ylim = c(-0.1, 1.4))

# combine
p1 + p2
```

## 7.2 Entropy and accuracy
### 7.2.1 Firing the weatherperson.
```{r}
weatherperson <-
  tibble(day = 1:10,
         prediction = rep(c(1, .6), times = c(3, 7)),
         observed = rep(1:0, times = c(3, 7)))

weatherperson %>% 
  pivot_longer(-day) %>% 
  
  ggplot(aes(x = day, y = name, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = value, color = value == 0)) +
  scale_x_continuous(breaks = 1:10, expand = c(0, 0)) +
  scale_y_discrete(NULL, expand = c(0, 0)) +
  scale_fill_viridis_c(direction = -1) +
  scale_color_manual(values = c("white", "black")) +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
newcomer <-
  tibble(day        = 1:10,
         prediction = 0,
         observed   = rep(1:0, times = c(3, 7)))

newcomer %>% 
  pivot_longer(-day) %>%
  
  ggplot(aes(x = day, y = name, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = value, color = value == 0)) +
  scale_x_continuous(breaks = 1:10, expand = c(0, 0)) +
  scale_y_discrete(NULL, expand = c(0, 0)) +
  scale_fill_viridis_c(direction = -1) +
  scale_color_manual(values = c("white", "black")) +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
weatherperson %>% 
  bind_rows(newcomer) %>% 
  mutate(person = rep(c("weatherperson", "newcomer"),
                      each = n()/2),
         hit = ifelse(prediction == observed, 1, 
                      1 - prediction - observed)) %>% 
  group_by(person) %>% 
  summarise(hit_rate = mean(hit))
```

#### 7.2.1.1 Costs and benefits.
```{r}
bind_rows(weatherperson, newcomer) %>% 
  mutate(person = rep(c("weatherperson", "newcomer"),
                      each = n()/2),
         points = ifelse(observed == 1 & prediction != 1,
                         -5, ifelse(observed == 1 & prediction == 1, -1, -1 * prediction))) %>% 
  group_by(person) %>% 
  summarise(happiness = sum(points))
```

#### 7.2.1.2 Measuring accuracy.
```{r}
bind_rows(weatherperson, newcomer) %>% 
  mutate(person = rep(c("weatherperson", "newcomer"),
                      each = n() / 2),
         hit = ifelse(prediction == observed, 1, 
                      1 - prediction - observed)) %>% 
  count(person, hit) %>% 
  mutate(power = hit ^ n,
         term = rep(letters[1:2], times = 2)) %>% 
  select(person, term, power) %>% 
  pivot_wider(names_from = term,
              values_from = power) %>% 
  mutate(probability_correct_sequence = a * b)
```

### 7.2.2 Information and uncertainty.
```{r}
tibble(place = c("McElreath's house", "Abu Dhabi"),
       p_rain = c(.3, .01)) %>% 
  mutate(p_shine = 1 - p_rain) %>% 
  group_by(place) %>% 
  mutate(h_p = (p_rain * log(p_rain) + p_shine * log(p_shine)) %>% mean() * -1) %>% 
  mutate_if(is.double, round, digits = 3)
```
```{r}
p <- c(.7, .15, .15)
-sum(p * log(p))
```

### 7.2.3 From entropy to accuracy.
```{r}
tibble(p_1 = .3,
       p_2 = .7,
       q_1 = .25,
       q_2 = .75) %>% 
  mutate(d_kl = (p_1 * log(p_1 / q_1)) + (p_2 * log(p_2 / q_2)))
```

```{r}
tibble(p_1 = .3) %>% 
  mutate(p_2 = 1 - p_1,
         q_1 = p_1) %>% 
  mutate(q_2 = 1 - q_1) %>% 
  mutate(d_kl = (p_1 * log(p_1 / q_1) + (p_2 * log(p_2 / q_2))))
```

```{r}
t <-
  tibble(p_1 = .3,
         p_2 = .7,
         q_1 = seq(from = .01, to = .99, by = .01)) %>% 
  mutate(q_2 = 1 - q_1) %>% 
  mutate(d_kl = (p_1 * log(p_1 / q_1)) + (p_2 * log(p_2 / q_2)))

head(t)
```

```{r}
t %>% 
  ggplot(aes(x = q_1, y = d_kl)) +
  geom_vline(xintercept = .3, color = carto_pal(7, "BurgYl")[5], linetype = 2) +
  geom_line(color = carto_pal(7, "BurgYl")[7], linewidth = 1.5) +
  annotate(geom = "text", x = .4, y = 1.5, label = "q = p",
           color = carto_pal(7, "BurgYl")[5], family = "Courier", size = 3.5) +
  labs(x = "q[1]", y = "Divergence of q from p")
```

#### 7.2.3.1 Rethinking: Divergence depends upon direction.
```{r}
tibble(direction = c("Earth to Mars", "Mars to Earth"),
       p_1 = c(.01, .7),
       q_1 = c(.7, .01)) %>% 
  mutate(p_2 = 1 - p_1,
         q_2 = 1 - q_1) %>% 
  mutate(d_kl = (p_1 * log(p_1 / q_1)) +
           (p_2 * log(p_2 / q_2)))
```

### 7.2.4 Estimating divergence.
```{r}
lm(data = d, brain_std ~ mass_std) %>% 
  logLik() * -2
```

```{r}
log_lik(b7.1) %>% 
  data.frame() %>% 
  set_names(pull(d, species)) %>% 
  pivot_longer(everything(),
               names_to = "species",
               values_to = "logprob") %>% 
  mutate(prob = exp(logprob)) %>% 
  group_by(species) %>% 
  summarise(log_probability_score = mean(prob) %>% log())
```

```{r}
log_lik(b7.1) %>% 
  data.frame() %>% 
  set_names(pull(d, species)) %>% 
  pivot_longer(everything(),
               names_to = "species",
               values_to = "logprob") %>% 
  mutate(prob = exp(logprob)) %>% 
  group_by(species) %>% 
  summarise(log_probability_score = mean(prob) %>% log()) %>% 
  summarise(total_log_probability_score = sum(log_probability_score))
```
#### 7.2.4.1 Overthinking: Computing the lppd.
```{r}
log_prob <- log_lik(b7.1)

log_prob %>% glimpse()
```

```{r}
prob <-
  log_prob %>% 
  # make it a data frame
  data.frame() %>% 
  # add case names, for convenience
  set_names(pull(d, species)) %>% 
  # add an s iteration index, for convenience
  mutate(s = 1:n()) %>% 
  # make it long
  pivot_longer(-s,
               names_to = "species",
               values_to = "logprob") %>% 
  # compute the probability scores
  mutate(prob = exp(logprob))

prob
```

```{r}
prob <-
  prob %>% 
  group_by(species) %>% 
  summarise(log_probability_score = mean(prob) %>% log())

prob
```

### 7.2.5 Scoring the right data.
```{r}
my_lppd <- function(brms_fit){
  log_lik(brms_fit) %>% 
    data.frame() %>% 
    pivot_longer(everything(),
                 values_to = "logprob") %>% 
    mutate(prob = exp(logprob)) %>% 
    group_by(name) %>% 
    summarise(log_probability_score = mean(prob) %>% log()) %>% 
    summarise(total_log_probability_score = sum(log_probability_score))
}
```

```{r}
tibble(name = str_c("b7.", 1:6)) %>% 
  mutate(brms_fit = purrr::map(name, get)) %>% 
  mutate(lppd = purrr::map(brms_fit, ~ my_lppd(.))) %>% 
  unnest(lppd)
```
#### 7.2.5.1 Overthinking: Simulated training and testing.
```{r}
sim_train_test
```
```{r}
# I've reduced this number by one order of magnitude to reduce computation time
n_sim <- 10
n_cores <- 8
kseq <- 1:5

# define the simulation function
my_sim <- function(k){
  
  print(k);
  r <- mcreplicate(n_sim, sim_train_test(N = n, k = k),
                   mc.cores = n_cores);
  c(mean(r[1, ]), mean(r[2, ]), sd(r[1, ]), sd(r[2, ]))
}

# here's our dev object based on N <- 20
n <- 20
dev_20 <-
  sapply(kseq, my_sim)

# here's our dev object based on N <- 100
n <- 100
dev_100 <- sapply(kseq, my_sim)
```

```{r}
dev_tibble <-
  rbind(dev_20, dev_100) %>% 
  data.frame() %>% 
  mutate(statistic = rep(c("mean", "sd"), each = 2) %>% rep(., times = 2),
         sample    = rep(c("in", "out"), times = 2) %>% rep(., times = 2),
         n         = rep(c("n = 20", "n = 100"), each = 4)) %>% 
  pivot_longer(-(statistic:n)) %>% 
  pivot_wider(names_from = statistic, values_from = value) %>%
  mutate(n     = factor(n, levels = c("n = 20", "n = 100")),
         n_par = str_extract(name, "\\d+") %>% as.double()) %>% 
  mutate(n_par = ifelse(sample == "in", n_par - .075, n_par + .075))

head(dev_tibble)
```

```{r}
# for the annotation
text <-
  dev_tibble %>% 
  filter(n_par > 1.5, 
         n_par < 2.5) %>% 
  mutate(n_par = ifelse(sample == "in", n_par - 0.2, n_par + 0.29))
  
# plot!
dev_tibble %>% 
  ggplot(aes(x = n_par, y = mean,
             ymin = mean - sd, ymax = mean + sd,
             group = sample, color = sample, fill  = sample)) +
  geom_pointrange(shape = 21) +
  geom_text(data = text,
            aes(label = sample)) +
  scale_fill_manual(values  = carto_pal(7, "BurgYl")[c(5, 7)]) +
  scale_color_manual(values = carto_pal(7, "BurgYl")[c(7, 5)]) +
  labs(title = "Figure 7.6. Deviance in and out of sample.",
       x = "number of parameters",
       y = "deviance") +
  theme(legend.position = "none",
        strip.background = element_rect(fill = alpha(carto_pal(7, "BurgYl")[1], 1/4), color = "transparent")) +
  facet_wrap(~ n, scale = "free_y")
```

## 7.3 Golem taming: regularization
```{r}
tibble(x = seq(from = - 3.5, to = 3.5, by = 0.01)) %>%
  mutate(a = dnorm(x, mean = 0, sd = 0.2),
         b = dnorm(x, mean = 0, sd = 0.5),
         c = dnorm(x, mean = 0, sd = 1.0)) %>% 
  pivot_longer(-x) %>% 
  
  ggplot(aes(x = x, y = value,
             fill = name, color = name, linetype = name)) +
  geom_area(alpha = 1/2, linewidth = 1/2, position = "identity") +
  scale_fill_manual(values = carto_pal(7, "BurgYl")[7:5]) +
  scale_color_manual(values = carto_pal(7, "BurgYl")[7:5]) +
  scale_linetype_manual(values = 1:3) +
  scale_x_continuous("parameter value", breaks = -3:3) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(legend.position = "none")
```

```{r}
library(rethinking)

# I've reduced this number by one order of magnitude to reduce computation time
n_sim   <- 10
n_cores <- 8

make_sim <- function(n, b_sigma) {
  sapply(kseq, function(k) {
    print(k);
    r <- mcreplicate(n_sim, sim_train_test(N = n, k = k, b_sigma = b_sigma),  # this is an augmented line of code
                     mc.cores = n_cores);
    c(mean(r[1, ]), mean(r[2, ]), sd(r[1, ]), sd(r[2, ])) 
    }
    ) %>% 
    
    # this is a new line of code
    data.frame()
}

s <-
  crossing(n       = c(20, 100),
           b_sigma = c(1, 0.5, 0.2)) %>% 
  mutate(sim = map2(n, b_sigma, make_sim)) %>% 
  unnest(sim)
```

```{r}
# wrangle the simulation data
s %>% 
  mutate(statistic = rep(c("mean", "sd"), each = 2) %>% rep(., times = 3 * 2),
         sample    = rep(c("in", "out"), times = 2) %>% rep(., times = 3 * 2)) %>% 
  pivot_longer(-c(n:b_sigma, statistic:sample)) %>% 
  pivot_wider(names_from = statistic, values_from = value) %>%
  mutate(n     = str_c("n = ", n) %>% factor(., levels = c("n = 20", "n = 100")),
         n_par = str_extract(name, "\\d+") %>% as.double()) %>% 
  
  # plot
  ggplot(aes(x = n_par, y = mean,
             group = interaction(sample, b_sigma))) +
  geom_line(aes(color = sample, size = b_sigma %>% as.character())) +
  # this function contains the data from the previous simulation
  geom_point(data = dev_tibble, 
             aes(group = sample, fill = sample),
             color = "black", shape = 21, size = 2.5, stroke = .1) +
  scale_size_manual(values = c(1, 0.5, 0.2)) +
  scale_fill_manual(values = carto_pal(7, "BurgYl")[c(7, 5)]) +
  scale_color_manual(values = carto_pal(7, "BurgYl")[c(7, 5)]) +
  labs(x = "number of parameters",
       y = "deviance") +
  theme(legend.position = "none",
        strip.background = element_rect(fill = alpha(carto_pal(7, "BurgYl")[1], 1/4), 
                                        color = "transparent")) +
  facet_wrap(~ n, scale = "free_y")
```

## 7.4 Predicting predictive accuracy
### 7.4.1 Cross-validation
### 7.4.2 Information criteria
#### 7.4.2.1 Overthinking: WAIC calculations.
```{r}
data(cars)

b7.m <-
  brm(data = cars,
      family = gaussian,
      dist ~ 1 + speed,
      prior = c(prior(normal(0, 100), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 7,
      file = "fits/b07.0m")
```

```{r}
print(b7.m)
```

Now use the brms::log_lik() function to return the log-likelihood for each observation i at each posterior draw s, where S = 4000
```{r}
n_cases <- nrow(cars)

ll <- 
  log_lik(b7.m) %>% 
  data.frame() %>% 
  set_names(c(str_c(0, 1:9), 10:n_cases))

dim(ll)
```

```{r}
log_mu_l <-
  ll %>% 
  pivot_longer(everything(),
               names_to = "i",
               values_to = "loglikelihood") %>% 
  mutate(likelihood = exp(loglikelihood)) %>% 
  group_by(i) %>% 
  summarise(log_mean_likelihood = mean(likelihood %>% log()))

(
  lppd <-
    log_mu_l %>% 
    summarise(lppd = sum(log_mean_likelihood)) %>% 
    pull(lppd)
)
```

```{r}
v_i <-
  ll %>% 
  pivot_longer(everything(),
               names_to = "i",
               values_to = "loglikelihood") %>% 
  group_by(i) %>% 
  summarise(var_loglikelihood = var(loglikelihood))

pwaic <-
  v_i %>% 
  summarise(pwaic = sum(var_loglikelihood)) %>% 
  pull()

pwaic
```

```{r}
-2 * (lppd - pwaic)
```

```{r}
waic(b7.m)
```

```{r}
(lppd - pwaic)
```

```{r}
tibble(lppd   = pull(log_mu_l, log_mean_likelihood),
       p_waic = pull(v_i, var_loglikelihood)) %>% 
  mutate(waic_vec = -2 * (lppd - p_waic)) %>% 
  summarise(waic_se = sqrt(n_cases * var(waic_vec)))
```

```{r}
waic(b7.m)$pointwise %>% 
  head()
```

### 7.4.3 Comparing CV, PSIS, and WAIC
```{r}
make_sim <- function(n, k, b_sigma) {
  
  r <- mcreplicate(n = n_sim, 
                   expr = sim_train_test(
                     N       = n,
                     k       = k,
                     b_sigma = b_sigma,
                     WAIC    = T,
                     LOOCV   = T, 
                     LOOIC   = T),
                   mc.cores = n_cores)
  
  t <- 
    tibble(
      deviance_os = mean(unlist(r[2, ])),
      deviance_w  = mean(unlist(r[3, ])),
      deviance_p  = mean(unlist(r[11, ])),
      deviance_c  = mean(unlist(r[19, ])),
      error_w     = mean(unlist(r[7, ])),
      error_p     = mean(unlist(r[15, ])),
      error_c     = mean(unlist(r[20, ]))
    )
  
  return(t)
  
}
```

```{r}
n_sim <- 10
n_cores <- 8

s <-
  crossing(n       = c(20, 100),
           k       = 1:5,
           b_sigma = c(0.5, 100)) %>%
  mutate(sim = pmap(list(n, k, b_sigma), make_sim)) %>% 
  unnest(sim)
```
## 7.5 Model comparison
### 7.5.1 Model mis-selection.
```{r}
b6.6 <- readRDS("fits/b06.06.rds")
b6.7 <- readRDS("fits/b06.07.rds")
b6.8 <- readRDS("fits/b06.08.rds")
```

```{r}
waic(b6.7)
```

```{r}
b6.7 <- add_criterion(b6.7, criterion = "waic")
```

```{r}
b6.7$criteria$waic
```

```{r}
# compute and save the WAIC information for the next three models
b6.6 <- add_criterion(b6.6, criterion = "waic")
b6.8 <- add_criterion(b6.8, criterion = "waic")

# compare the WAIC estimates
w <- loo_compare(b6.6, b6.7, b6.8, criterion = "waic")

print(w)
```
```{r}
print(w, simplify = F)
```

```{r}
cbind(waic_diff = w[, 1] * -2,
      se = w[, 2] * 2)
```

```{r}
str(w)
```

```{r}
print(w, simplify = F)
```

```{r}
b6.6 <- add_criterion(b6.6, criterion = "loo")
b6.7 <- add_criterion(b6.7, criterion = "loo")
b6.8 <- add_criterion(b6.8, criterion = "loo")

# compare the WAIC estimates
loo_compare(b6.6, b6.7, b6.8, criterion = "loo") %>% 
  print(simplify = F)
```
```{r}
n <- length(b6.7$criteria$waic$pointwise[, "waic"])

tibble(waic_b6.7 = b6.7$criteria$waic$pointwise[, "waic"],
       waic_b6.8 = b6.8$criteria$waic$pointwise[, "waic"]) %>% 
  mutate(diff = waic_b6.7 - waic_b6.8) %>% 
  summarise(diff_se = sqrt(n * var(diff)))
```

```{r}
w[2, 2] * 2
```

Presuming the difference is Gaussian distributed, here’s our 99% interval.
```{r}
(w[2, 1] * -2) + c(-1, 1) * (w[2, 2] * 2) * 2.6
```

```{r}
w[, 7:8] %>% 
  data.frame() %>% 
  rownames_to_column("model_name") %>% 
  mutate(model_name = fct_reorder(model_name, waic, .desc = T)) %>% 
  
  ggplot(aes(x = waic, y = model_name, 
             xmin = waic - se_waic, 
             xmax = waic + se_waic)) +
  geom_pointrange(color = carto_pal(7, "BurgYl")[7], 
                  fill = carto_pal(7, "BurgYl")[5], shape = 21) +
  labs(title = "My custom WAIC plot",
       x = NULL, y = NULL) +
  theme(axis.ticks.y = element_blank())
```

```{r}
tibble(waic_b6.6 = waic(b6.6)$pointwise[, "waic"],
       waic_b6.8 = waic(b6.8)$pointwise[, "waic"]) %>% 
  mutate(diff = waic_b6.6 - waic_b6.8) %>% 
  summarise(diff_se = sqrt(n * var(diff)))
```

```{r}
loo_compare(b6.6, b6.7, b6.8, criterion = "waic") %>% 
  str()
```

```{r}
loo_compare(b6.6, b6.8, criterion = "waic")
```

```{r}
loo_compare(b6.6, b6.8, criterion = "waic")[2, 2] * 2
```

```{r}
model_weights(b6.6, b6.7, b6.8, weights = "waic") %>% 
  round(digits = 2)
```

### 7.5.2 Outliers and other illusions.
```{r}
data(WaffleDivorce, package = "rethinking")

d <-
  WaffleDivorce %>% 
  mutate(d = rethinking::standardize(Divorce),
         m = rethinking::standardize(Marriage),
         a = rethinking::standardize(MedianAgeMarriage))

rm(WaffleDivorce)
```

```{r}
b5.1 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + a,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      sample_prior = T,
      file = "fits/b05.01")

b5.2 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + m,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "fits/b05.02")

b5.3 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + m + a,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "fits/b05.03")
```

```{r}
b5.1 <- add_criterion(b5.1, criterion = "loo")
b5.2 <- add_criterion(b5.2, criterion = "loo")
b5.3 <- add_criterion(b5.3, criterion = "loo")
```

```{r}
loo_compare(b5.1, b5.2, b5.3, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
loo(b5.3)
```

```{r}
library(loo)

loo(b5.3) %>% 
  pareto_k_ids(threshold = 0.4)
```

```{r}
d %>% 
  slice(13) %>% 
  select(Location:Loc)
```

```{r}
pareto_k_values(loo(b5.3))[13]
```

```{r}
b5.3 <- add_criterion(b5.3, "waic", file = "fits/b05.03")
```

```{r}
tibble(pareto_k = b5.3$criteria$loo$diagnostics$pareto_k,
       p_waic = b5.3$criteria$waic$pointwise[, "p_waic"],
       Loc = pull(d, Loc)) %>% 
  
  ggplot(aes(x = pareto_k, y = p_waic, color = Loc == "ID")) +
  geom_point(aes(shape = Loc == "ID")) +
  geom_vline(xintercept = 0.4, linetype = "dashed",
             color = "black", alpha = 1/2) +
  geom_text(data = . %>% dplyr::filter(p_waic > 0.4),
            aes(x = pareto_k - 0.03, label = Loc),
            hjust = 1) +
  scale_color_manual(values = carto_pal(7, "BurgYl")[c(5, 7)]) +
  scale_shape_manual(values = c(1, 19)) +
  labs(subtitle = "Gaussian model (b5.3)") +
  theme(legend.position = "none")
```

```{r}
waic(b5.3)
```

```{r}
tibble(x = seq(from = -6, to = 6, by = .01)) %>% 
  mutate(Gaussian = dnorm(x),
         `Student-t` = dstudent(x)) %>% 
  pivot_longer(-x,
               names_to = "likelihood",
               values_to = "density") %>% 
  mutate(`minus log density` = -log(density)) %>% 
  pivot_longer(contains("density")) %>% 
  
  ggplot(aes(x = x, y = value, group = likelihood,
             color = likelihood)) +
  geom_line() +
  scale_color_manual(values = c(carto_pal(7, "BurgYl")[6],
                                "black")) +
  ylim(0, NA) +
  labs(x = "value", y = NULL) +
  theme(strip.background = element_blank()) +
  facet_wrap(~ name, scales = "free_y")
```

```{r}
b5.3t <-
  brm(data = d,
      family = student,
      bf(d ~ 1 + m + a, nu = 2),
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "fits/b05.03t")
```

```{r}
print(b5.3t)
```

```{r}
b5.3t <- add_criterion(b5.3t, criterion = c("loo", "waic"))
```

```{r}
tibble(pareto_k = b5.3t$criteria$loo$diagnostics$pareto_k,
       p_waic = b5.3t$criteria$waic$pointwise[, "p_waic"],
       Loc = pull(d, Loc)) %>% 
  
  ggplot(aes(x = pareto_k, y = p_waic, color = Loc == "ID")) +
  geom_point(aes(shape = Loc == "ID")) +
  geom_text(data = . %>% dplyr::filter(Loc %in% c("ID", "ME")),
            aes(x = pareto_k - 0.005, label = Loc),
            hjust = 1) +
  scale_color_manual(values = carto_pal(7, "BurgYl")[c(5, 7)]) +
  scale_shape_manual(values = c(1, 19)) +
  labs(subtitle = "Student-t model (b5.3t)") +
  theme(legend.position = "none")
```

```{r}
loo_compare(b5.3, b5.3t, criterion = "waic") %>% print(simplify = F)
```

```{r}
loo_compare(b5.3, b5.3t, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
bind_rows(as_draws_df(b5.3),
          as_draws_df(b5.3t)) %>% 
  mutate(fit = rep(c("Gaussian (b5.3)", "Student-t (b5.3t)"), each = n() / 2)) %>% 
  pivot_longer(b_Intercept:sigma) %>% 
  mutate(name = factor(name,
                       levels = c("b_Intercept", "b_a", "b_m", "sigma"),
                       labels = c("alpha", "beta[a]", "beta[m]", "sigma"))) %>% 
  
  ggplot(aes(x = value, y = fit, color = fit)) +
  stat_pointinterval(.width = .95, size = 1) +
  scale_color_manual(values = c(carto_pal(7, "BurgYl")[6], "black")) +
  labs(x = "posterior", y = NULL) +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = alpha(carto_pal(7, "BurgYl")[1], 1/4), color = "transparent"),
        strip.text = element_text(size = 12)) +
  facet_wrap(~ name, ncol = 1, labeller = label_parsed)
```

```{r}
# for the annotation
text <-
  tibble(nu    = c(1.35, 20),
         sd    = c(sqrt(2.1 / (2.1 - 2)), 0.875),
         angle = c(90, 0),
         hjust = 1,
         label = "asymptote")

# wrangle
tibble(nu = seq(from = 2.1, to = 20, by = 0.01)) %>% 
  mutate(var = nu / (nu - 2)) %>% 
  mutate(sd = sqrt(var)) %>% 
  
  # plot
  ggplot(aes(x = nu, y = sd)) +
  geom_hline(yintercept = 1, color = carto_pal(7, "BurgYl")[2], linetype = 2) +
  geom_vline(xintercept = 2, color = carto_pal(7, "BurgYl")[2], linetype = 2) +
  geom_text(data = text,
            aes(label = label, angle = angle, hjust = hjust),
            color = carto_pal(7, "BurgYl")[3]) +
  geom_line(color = carto_pal(7, "BurgYl")[7]) +
  scale_x_continuous(expression(nu), breaks = c(2, 1:4 * 5)) +
  labs(subtitle = expression(Student-t(nu*', '*0*', '*1)),
       y = expression(standard~deviation*", "*~sqrt(nu/(nu-2))))
```

## 7.6 Bonus: R2 talk
```{r}
bayes_R2(b5.3) %>% round(digits = 3)
```

```{r}
rbind(bayes_R2(b5.1),
      bayes_R2(b5.2),
      bayes_R2(b5.3)) %>% 
  data.frame() %>% 
  mutate(model = c("b5.1", "b5.2", "b5.3"),
         r_square_posterior_mean = round(Estimate, digits = 3)) %>% 
  select(model, r_square_posterior_mean)
```

```{r}
r2_b5.1 <- bayes_R2(b5.1, summary = F)

r2_b5.1 %>% glimpse()
```

```{r}
r2 <-
  cbind(bayes_R2(b5.1, summary = F),
        bayes_R2(b5.2, summary = F)) %>% 
  data.frame() %>% 
  set_names(str_c("b5.", 1:2))

r2 %>% 
  ggplot() +
  geom_density(aes(x = b5.1), alpha = 3/4, linewidth = 0,
               fill = carto_pal(7, "BurgYl")[4]) +
  geom_density(aes(x = b5.2), alpha = 3/4, linewidth = 0,
               fill = carto_pal(7, "BurgYl")[6]) +
  annotate(geom = "text", x = c(.1, .34), y = 3.5,
           label = c("b5.2", "b5.1"),
           color = alpha("white", 3/4), family = "Courier") +
  scale_x_continuous(NULL, limits = c(0, 1)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle(expression(italic(R)^2~distributions))
```

```{r}
r2 %>% 
  mutate(diff = b5.2 - b5.1) %>% 
  
  ggplot(aes(x = diff, y = 0)) +
  stat_halfeye(point_interval = median_qi, .width = .95,
               fill = carto_pal(7, "BurgYl")[5],
               color = carto_pal(7, "BurgYl")[7]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = expression(Turns~out~b5.2~had~a~lower~italic(R)^2~than~b5.1),
       x = expression(Delta*italic(R)^2))
```

