---
title: "15 Missing Data and Other Opportunities"
format: html
---

For opening example, we're playing with the conditional probabilty

$$
\Pr(\text{burnt down} | \text{burnt up}) = \frac{\Pr(\text{burnt up, burnt down})}{\Pr(\text{burnt up})}.
$$
$$
\Pr(\text{burnt down} | \text{burnt up}) = \frac{1/3}{1/2} = \frac{2}{3}.
$$

```{r}
library(tidyverse)

p_pancake <- 1/3
(
  d <-
    tibble(pancake = c("BB", "BU", "UU"),
           p_burnt = c(1, .5, 0)) %>% 
    mutate(p_burnt_up = p_burnt * p_pancake)
)
```

```{r}
d %>% 
  summarise(`p (burnt_down | burnt_up)` = p_pancake / sum(p_burnt_up))
```

```{r}
set.seed(1)
sample(x = 1:3, size = 1)
```

```{r}
take_sample <- function(seed){
  set.seed(seed)
  sample(x = 1:3, size = 1)
}

tibble(seed = 2:11) %>% 
  mutate(value_returned = map_dbl(seed, take_sample))
```

```{r}
matrix(c(1, 1, 1, 0, 0, 0), nrow = 2, ncol = 3)
```

```{r}
matrix(c(1, 1, 1, 0, 0, 0), nrow = 2, ncol = 3)[, 2]
```

```{r}
matrix(c(1, 1, 1, 0, 0, 0), nrow = 2, ncol = 3)[, 2] %>% str()
```

```{r}
sim_pancake <- function(seed){
  set.seed(seed)
  
  pancake <- sample(x = 1:3, size = 1)
  sides <- matrix(c(1, 1, 1, 0, 0, 0), nrow = 2, ncol = 3)[, pancake]
  sample(sides)
}
```

```{r}
# how many simulations would you like?
n_sim <- 1e4

d <-
  tibble(seed = 1:n_sim) %>% 
  mutate(burnt = purrr::map(seed, sim_pancake)) %>% 
  unnest(burnt) %>% 
  mutate(side = rep(c("up", "down"), times = n() / 2))
```

```{r}
head(d, n = 10)
```

```{r}
d %>% 
  pivot_wider(names_from = side, values_from = burnt) %>% 
  summarise(`p (burnt_down | burnt_up)` = sum(up == 1 & down == 1) / (sum(up)))
```

## 15.1 Measurement error
```{r}
data(WaffleDivorce, package = "rethinking")
d <- WaffleDivorce
rm(WaffleDivorce)
```

```{r}
d <-
  d %>% 
  mutate(D_obs = (Divorce - mean(Divorce)) / sd(Divorce),
         D_sd = Divorce.SE / sd(Divorce),
         M = (Marriage - mean(Marriage)) / sd(Marriage),
         A = (MedianAgeMarriage - mean(MedianAgeMarriage)) / sd(MedianAgeMarriage),
         M_obs = M,
         M_sd = Marriage.SE - sd(Marriage))
```

```{r}
library(ggdark)

theme_set(
  dark_theme_bw() +
    theme(legend.position = "none",
          panel.grid = element_blank())
)
```

```{r}
library(viridis)
```

```{r}
viridis_pal(option = "C")(7)
```

```{r}
tibble(factor = "a",
       number = factor(1:7),
       color_number = str_c(1:7, ". ", viridis_pal(option = "C")(7))) %>% 
  
  ggplot(aes(x = factor, y = number)) +
  geom_tile(aes(fill = number)) +
  geom_text(aes(label = color_number, color = number %in% c("5", "6", "7"))) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_viridis(option = "C", discrete = T, direction = -1) +
  scale_x_discrete(NULL, breaks = NULL, expand = c(0, 0)) +
  scale_y_discrete(NULL, breaks = NULL, expand = c(0, 0)) +
  ggtitle("Behold: viridis C!")
```

```{r}
color <- viridis_pal(option = "C")(7)[7]

p1 <- d %>% 
  ggplot(aes(x = MedianAgeMarriage,
             y = Divorce,
             ymin = Divorce - Divorce.SE,
             ymax = Divorce + Divorce.SE)) +
  geom_pointrange(shape = 20, alpha = 2/3, color = color) +
  labs(x = "Median age marriage", y = "Divorce rate")
```

```{r}
color <- viridis_pal(option = "C")(7)[6]

p2 <-
  d %>% 
  ggplot(aes(x = log(Population),
             y = Divorce,
             ymin = Divorce - Divorce.SE,
             ymax = Divorce + Divorce.SE)) +
  geom_pointrange(shape = 20, alpha = 2/3, color = color) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab("log population")

library(patchwork)
p1 | p2
```

### 15.1.1 Error on the outcome
```{r}
library(ggdag)

dag_coords <-
  tibble(name = c("A", "M", "D", "Dobs", "eD"),
         x    = c(1, 2, 2, 3, 4),
         y    = c(2, 3, 1, 1, 1))

dagify(M    ~ A,
       D    ~ A + M,
       Dobs ~ D + eD,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("D", "eD"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("A", "D", expression(D[obs]), "M", expression(italic(e)[D]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
d %>% 
  mutate(Divorce_distribution = str_c("Divorce ~ Normal(", Divorce, ", ", Divorce.SE, ")")) %>% 
  select(Loc, Divorce_distribution) %>% 
  head()
```

Our model will follow the form
$$
\begin{align*}
\color{#5D01A6FF}{\text{Divorce}_{\text{OBS}, i}} & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Normal}(\text{Divorce}_{\text{TRUE}, i}, \text{Divorce}_{\text{SE}, i})} \\
\color{#5D01A6FF}{\text{Divorce}_{\text{TRUE}, i}} & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu     & = \alpha + \beta_1 \text A_i + \beta_2 \text M_i \\
\alpha  & \sim \operatorname{Normal}(0, 0.2) \\
\beta_1 & \sim \operatorname{Normal}(0, 0.5) \\
\beta_2 & \sim \operatorname{Normal}(0, 0.5) \\
\sigma  & \sim \operatorname{Exponential}(1).
\end{align*}
$$

```{r}
library(brms)
```

```{r}
# put the data into a list()
dlist <- list(
  D_obs = d$D_obs,
  D_sd = d$D_sd,
  M = d$M,
  A = d$A
)

b15.1 <-
  brm(data = dlist,
      family = gaussian,
      D_obs | mi(D_sd) ~ 1 + A + M,
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, cores = 8, chains = 4,
      seed = 15,
      # note this line
      save_pars = save_pars(latent = TRUE),
      file = "fits/b15.01")
```

```{r}
print(b15.1)
```

To return the summaries for the `D_true[i]` parameters, you might execute `posterior_summary(b15.1)` or `b15.1$fit`.
```{r}
posterior_summary(b15.1) %>% 
  round(digits = 2) %>% 
  data.frame()
```

```{r}
library(ggrepel)

states <- c("AL", "AR", "ME", "NH", "RI", "DC", "VT", "AK", "SD", "UT", "ID", "ND", "WY")

d_est <-
  posterior_summary(b15.1) %>% 
  data.frame() %>% 
  rownames_to_column("term") %>% 
  mutate(D_est = Estimate) %>% 
  select(term, D_est) %>% 
  filter(str_detect(term, "Yl")) %>% 
  bind_cols(d)

color <- viridis_pal(option = "C")(7)[5]

p1 <-
  d_est %>% 
  ggplot(aes(x = D_sd, y = D_est - D_obs)) +
  geom_hline(yintercept = 0, linetype = 2, color = "white") +
  geom_point(alpha = 2/3, color = color) +
  geom_text_repel(data = . %>% filter(Loc %in% states),
                  aes(label = Loc),
                  size = 3, seed = 15, color = "white")
```

```{r}
library(tidybayes)

states <- c("AR", "ME", "RI", "ID", "WY", "ND", "MN")

color <- viridis_pal(option = "C")(7)[4]

p2 <-
  as_draws_df(b15.1) %>% 
  expand_grid(A = seq(from = -3.5, to = 3.5, length.out = 50)) %>% 
  mutate(fitted = b_Intercept + b_A * A) %>% 
  
  ggplot(aes(x = A)) +
  stat_lineribbon(aes(y = fitted), .width = .95, size = 1/3,
                  color = "grey50", fill = "grey20") +
  geom_segment(data = d_est, aes(y = D_obs, yend = D_est),
               linewidth = 1/5) +
  geom_point(data = d_est, aes(y = D_obs), color = color) +
  geom_point(data = d_est, aes(y = D_est), shape = 1,
            stroke = 1/3) +
  geom_text_repel(data = d %>% filter(Loc %in% states),
                  aes(y = D_obs, label = Loc),
                  size = 3, seed = 15, color = "white") +
  labs(x = "median age marriage (std)", 
       y = "divorce rate (std)") +
  coord_cartesian(xlim = range(d$A),
                  ylim = range(d$D_obs))
```

```{r}
p1 | p2
```

### 15.1.2 Error on both outcome and predictor.
```{r}
dag_coords <-
  tibble(name = c("A", "M", "Mobs", "eM", "D", "Dobs", "eD"),
         x    = c(1, 2, 3, 4, 2, 3, 4),
         y    = c(2, 3, 3, 3, 1, 1, 1))

dagify(M    ~ A,
       D    ~ A + M,
       Mobs ~ M + eM,
       Dobs ~ D + eD,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("A", "Mobs", "Dobs"), "b", "a")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("A", "D",
                                     expression(D[obs]), 
                                     "M",
                                     expression(M[obs]), 
                                     expression(italic(e)[D]),
                                     expression(italic(e)[M]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

$$
\begin{align*}
\text{Divorce}_{\text{OBS}, i}  & \sim \operatorname{Normal}(\text{Divorce}_{\text{TRUE}, i}, \text{Divorce}_{\text{SE}, i}) \\
\text{Divorce}_{\text{TRUE}, i} & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i & = \alpha + \beta_1 \text A_i + \beta_2 \color{#5D01A6FF}{\text{Marriage}_{\text{TRUE}, i}} \\
\color{#5D01A6FF}{\text{Marriage}_{\text{OBS}, i}} & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Normal}(\text{Marriage}_{\text{TRUE}, i}, \text{Marriage}_{\text{SE}, i})} \\
\color{#5D01A6FF}{\text{Marriage}_{\text{TRUE}, i}} & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Normal}(0, 1)} \\
\alpha  & \sim \operatorname{Normal}(0, 0.2) \\
\beta_1 & \sim \operatorname{Normal}(0, 0.5) \\
\beta_2 & \sim \operatorname{Normal}(0, 0.5) \\
\sigma  & \sim \operatorname{Exponential}(1).
\end{align*}
$$

```{r}
# put the data into a `list()`
dlist <- list(
  D_obs = d$D_obs,
  D_sd = d$D_sd,
  M_obs = d$M_obs,
  M_sd = d$M_sd,
  A = d$A)

b15.2 <-
  brm(data = dlist, 
      family = gaussian,
      D_obs | mi(D_sd) ~ 1 + A + me(M_obs, M_sd),
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(normal(0, 1), class = meanme),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      # note this line
      save_pars = save_pars(latent = TRUE),
      file = "fits/b15.02")
```

```{r}
posterior_summary(b15.2) %>% round(digits = 2)
```

```{r}
color_y <- viridis_pal(option = "C")(7)[7]
color_p <- viridis_pal(option = "C")(7)[2]

# wrangle
full_join(
  # D
  tibble(Loc   = d %>% pull(Loc),
         D_obs = d %>% pull(D_obs),
         D_est = posterior_summary(b15.2) %>% 
           data.frame() %>% 
           rownames_to_column("term") %>% 
           filter(str_detect(term, "Yl")) %>% 
           pull(Estimate)) %>% 
    pivot_longer(-Loc, values_to = "d") %>% 
    mutate(name = if_else(name == "D_obs", "observed", "posterior")),
  # M
  tibble(Loc   = d %>% pull(Loc),
         M_obs = d %>% pull(M_obs),
         M_est = posterior_summary(b15.2) %>% 
           data.frame() %>% 
           rownames_to_column("term") %>% 
           filter(str_detect(term, "Xme_")) %>% 
           pull(Estimate)) %>% 
    pivot_longer(-Loc, values_to = "m") %>% 
    mutate(name = if_else(name == "M_obs", "observed", "posterior")),
  by = c("Loc", "name")
)  %>% 
  
  # plot!
  ggplot(aes(x = m, y = d)) +
  geom_line(aes(group = Loc),
            linewidth = 1/4) +
  geom_point(aes(color = name)) +
  scale_color_manual(values = c(color_p, color_y)) +
  labs(subtitle = "Shrinkage of both divorce rate and marriage rate", 
       x = "Marriage rate (std)" , 
       y = "Divorce rate (std)")
```

#### 15.1.3 Measurement terrors.
```{r}
dag_coords <-
  tibble(name = c("A", "M", "Mobs", "eM", "D", "Dobs", "eD", "P"),
         x    = c(1, 2, 3, 4, 2, 3, 4, 5),
         y    = c(2, 3, 3, 3, 1, 1, 1, 2))

dagify(M    ~ A,
       D    ~ A + M,
       Mobs ~ M + eM,
       Dobs ~ D + eD,
       eM   ~ P,
       eD   ~ P,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("A", "Mobs", "Dobs", "P"), "b", "a")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("A", "D",
                                     expression(D[obs]), "M", 
                                     expression(M[obs]),
                                     "P",
                                     expression(italic(e)[D]), 
                                     expression(italic(e)[M]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
dag_coords <-
  tibble(name = c("A", "M", "Mobs", "eM", "D", "Dobs", "eD"),
         x    = c(1, 2, 3, 4, 2, 3, 4),
         y    = c(2, 3, 3, 3, 1, 1, 1))

dagify(M    ~ A,
       D    ~ A + M,
       Mobs ~ M + eM,
       Dobs ~ D + eD,
       eD   ~ M,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("A", "Mobs", "Dobs"), "b", "a")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("A", "D",
                                     expression(D[obs]), "M", 
                                     expression(M[obs]),
                                     expression(italic(e)[D]), 
                                     expression(italic(e)[M]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
dag_coords <-
  tibble(name = c("eA", "Aobs", "A", "M", "D"),
         x    = c(1, 2, 3, 4, 4),
         y    = c(2, 2, 2, 3, 1))

dagify(Aobs ~ A + eA,
       M    ~ A,
       D    ~ A,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("A", "eA"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("A",
                                     expression(A[obs]),
                                     "D", 
                                     "M", 
                                     expression(italic(e)[A]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
n <- 500

set.seed(15)

dat <-
  tibble(A = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(M = rnorm(n, mean = -A, sd = 1),
         D = rnorm(n, mean = A, sd = 1),
         A_obs = rnorm(n, mean = A, sd = 1))
```

```{r}
# the model with A containing no measurement error
b15.2b <-
  brm(data = dat,
      family = gaussian,
      D ~ 1 + A + M,
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, cores = 8, chains = 4,
      seed = 15,
      # note this line
      save_pars = save_pars(latent = TRUE),
      file = "fits/b15.02b")

# The model where A has measurement error, but we ignore it
b15.2c <-
  brm(data = dat,
      family = gaussian,
      D ~ 1 + A_obs + M,
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, cores = 8, chains = 4,
      seed = 15,
      # note this line
      save_pars = save_pars(latent = TRUE),
      file = "fits/b15.02c")
```

```{r}
print(b15.2b)
```

```{r}
print(b15.2c)
```

```{r}
# for annotation
text <-
  tibble(fit      = "b15.2b",
         term     = "beta[0]",
         Estimate = fixef(b15.2b, probs = .99)["Intercept", 3],
         label    = "In this plot, we like the yellow posteriors.")

# wrangle
bind_rows(
  posterior_summary(b15.2b) %>% 
    data.frame() %>% 
    rownames_to_column("term"),
  posterior_summary(b15.2c) %>% 
    data.frame() %>% 
    rownames_to_column("term")) %>% 
  filter(term != "lp__") %>% 
  filter(term != "lprior") %>% 
  filter(term != "Intercept") %>% 
  mutate(term = rep(c(str_c("beta[", 0:2, "]"), "sigma"), times = 2),
         fit = rep(c("b15.2b", "b15.2c"), each = n() / 2)) %>% 
  
  # plot!
  ggplot(aes(x = Estimate, y = fit)) +
  geom_vline(xintercept = 0, linetype = 3, alpha = 1/2) +
  geom_pointrange(aes(xmin = Q2.5, xmax = Q97.5, color = fit)) +
  geom_text(data = text,
            aes(label = label),
            hjust = 0, color = color_y) +
  scale_color_manual(values = c(color_y, "white")) +
  labs(x = "marginal posterior", y = NULL) +
  theme(axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "transparent",
                                        fill = "transparent")) +
  facet_wrap(~ term, labeller = label_parsed, ncol = 1)
```

## 15.2 Missing data
### 15.2.1 DAG ate my homework.
```{r}
# panel a
dag_coords <-
  tibble(name = c("S", "H", "Hs", "D"),
         x    = c(1, 2, 2, 1),
         y    = c(2, 2, 1, 1))

p1 <-
  dagify(H  ~ S,
         Hs ~ H + D,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "H", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(label = c("D", "H", "H*", "S")) +
  geom_dag_edges(edge_colour = "#FCF9F0")

# panel b
p2 <-
  dagify(H  ~ S,
         Hs ~ H + D,
         D  ~ S,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "H", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(label = c("D", "H", "H*", "S")) +
  geom_dag_edges(edge_colour = "#FCF9F0")

# panel d
p4 <-
  dagify(H  ~ S,
         Hs ~ H + D,
         D  ~ H,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "H", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(label = c("D", "H", "H*", "S")) +
  geom_dag_edges(edge_colour = "#FCF9F0")
```

```{r}
dag_coords <-
  tibble(name = c("S", "H", "Hs", "D", "X"),
         x    = c(1, 2, 2, 1, 1.5),
         y    = c(2, 2, 1, 1, 1.5))

p3 <-
  dagify(H  ~ S + X,
         Hs ~ H + D,
         D  ~ X,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("H", "X"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(label = c("D", "H", "H*", "S", "X")) +
  geom_dag_edges(edge_colour = "#FCF9F0")
```

```{r}
(p1 + p2 + p3 + p4) +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) &
  dark_theme_void() +
  theme(panel.background = element_rect(fill = "grey8"),
        plot.margin = margin(0.15, 0.15, 0.15, 0.15, "in"))
```

```{r}
n <- 100

set.seed(15)

d <-
  tibble(s = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(h = rbinom(n, size = 10, inv_logit_scaled(s)),
         d_a = rbinom(n, size = 1, prob = .5),
         d_b = ifelse(s > 0, 1, 0)) %>% 
  mutate(hm_a = ifelse(d_a == 1, NA, h),
         hm_b = ifelse(d_b == 1, NA, h))

d
```

```{r}
p1 <-
  d %>% 
  ggplot(aes(x = s, y = h)) + 
  geom_point(color = viridis_pal(option = "C")(7)[7], alpha = 2/3) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "true distribution")

p2 <-
  d %>% 
  ggplot(aes(x = s, y = hm_a)) + 
  geom_point(color = viridis_pal(option = "C")(7)[6], alpha = 2/3) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "missing completely at random")

p3 <-
  d %>% 
  ggplot(aes(x = s, y = hm_b)) + 
  geom_point(color = viridis_pal(option = "C")(7)[6], alpha = 2/3) +
  scale_y_continuous(breaks = 1:10, limits = c(1, 10)) +
  labs(subtitle = "missing conditional on s")

p1 + p2 + p3
```

```{r}
n <- 1000

set.seed(501)

d <-
  tibble(x = rnorm(n, mean = 0, sd = 1),
         s = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(h = rbinom(n, size = 10, inv_logit_scaled(2 + s - 2 * x)),
         d = ifelse(x > 1, 1, 0)) %>% 
  mutate(hm = ifelse(d == 1, NA, h))

d
```

```{r}
p1 <-
  d %>% 
  ggplot(aes(x = s, y = h)) + 
  geom_point(color = viridis_pal(option = "C")(7)[7], alpha = 1/4) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "true distribution")

p2 <-
  d %>% 
  ggplot(aes(x = s, y = hm)) + 
  geom_point(color = viridis_pal(option = "C")(7)[6], alpha = 1/4) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "missing conditional on x")

p1 + p2
```

```{r}
b15.3 <-
  brm(data = d,
      family = binomial,
      h | trials(10) ~ 1 + s,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, .5), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.03")
```

```{r}
print(b15.3)
```

```{r}
b15.4 <-
  brm(data = d %>% filter(d == 0),
      family = binomial,
      h | trials(10) ~ 1 + s,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, .5), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.04")
```

```{r}
print(b15.4)
```

```{r}
d <-
  d %>% 
  mutate(d = ifelse(abs(x) < 1, 1, 0)) %>% 
  mutate(hm = ifelse(d == 1, NA, h))

d
```

```{r}
p1 <-
  d %>% 
  ggplot(aes(x = s, y = h)) + 
  geom_point(color = viridis_pal(option = "C")(7)[7], alpha = 1/4) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "true distribution")

p2 <-
  d %>% 
  ggplot(aes(x = s, y = hm)) + 
  geom_point(color = viridis_pal(option = "C")(7)[6], alpha = 1/4) +
  scale_y_continuous(breaks = 1:10) +
  labs(subtitle = "missing conditional on x")

p1 + p2
```

```{r}
b15.4b <-
  brm(data = d %>% filter(d == 0),
      family = binomial,
      h | trials(10) ~ 1 + s,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, .5), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.04b")
```

```{r}
print(b15.4b)
```
### 15.2.2 Imputing primates.
```{r}
data(milk, package = "rethinking")
d <- milk
rm(milk)

# transform
d <-
  d %>% 
  mutate(neocortex.prop = neocortex.perc / 100,
         logmass = log(mass)) %>% 
  mutate(k = (kcal.per.g - mean(kcal.per.g)) / sd(kcal.per.g),
         b = (neocortex.prop - mean(neocortex.prop, na.rm = T)) / sd(neocortex.prop, na.rm = T),
         m = (logmass - mean(logmass)) / sd(logmass))
```

```{r}
d %>% 
  count(is.na(neocortex.prop))
```

```{r}
dag_coords <-
  tibble(name = c("M", "U", "K", "B"),
         x    = c(1, 2, 2, 3),
         y    = c(2, 2, 1, 2))

dagify(M ~ U,
       B ~ U,
       K ~ M + B,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "U", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text() +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
dag_coords <-
  tibble(name = c("M", "U", "K", "B", "RB", "Bs"),
         x    = c(1, 2, 2, 3, 2, 3),
         y    = c(2, 2, 1, 2, 3, 3))

# left
p1 <-
  dagify(M  ~ U,
         B  ~ U,
         K  ~ M + B,
         Bs ~ RB + B,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("U", "B"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, 
                label = c("B", 
                          expression(B^'*'),
                          "K", 
                          "M", 
                          expression(R[B]),
                          "U")) +
  geom_dag_edges(edge_colour = "#FCF9F0")

# middle
p2 <-
  dagify(M  ~ U,
         B  ~ U,
         K  ~ M + B,
         Bs ~ RB + B,
         RB ~ M,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("U", "B"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, 
                label = c("B", 
                          expression(B^'*'),
                          "K", 
                          "M", 
                          expression(R[B]),
                          "U")) +
  geom_dag_edges(edge_colour = "#FCF9F0")

# right
p3 <-
  dagify(M  ~ U,
         B  ~ U,
         K  ~ M + B,
         Bs ~ RB + B,
         RB ~ B,
         coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("U", "B"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, 
                label = c("B", 
                          expression(B^'*'),
                          "K", 
                          "M", 
                          expression(R[B]),
                          "U")) +
  geom_dag_edges(edge_colour = "#FCF9F0")

# combine!
(p1 + p2 + p3) &
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) &
  dark_theme_void() &
  theme(panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill = "grey8", color = "grey8"),
        plot.margin = margin(0.1, 0.1, 0.1, 0.1, "in"))
```

```{r}
dag_coords <-
  tibble(name = c("M", "U", "K", "B", "Bs", "RB", "V"),
         x    = c(1, 2, 2, 3, 4, 4, 3.4),
         y    = c(2, 2, 1, 2, 2, 1, 1.45))

dagify(M  ~ U,
       B  ~ U + V,
       K  ~ M + B,
       Bs ~ RB + B,
       RB ~ V,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("U", "B", "V"), "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, 
                label = c("B", 
                          expression(B^'*'),
                          "K", 
                          "M", 
                          expression(R[B]),
                          "U",
                          "V")) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

$$
\begin{align*}
K_i     & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i   & = \alpha + \beta_1 \color{#5D01A6FF}{B_i} + \beta_2 \log M_i \\
\color{#5D01A6FF}{B_i} & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Normal}(\nu, \sigma_B)} \\
\alpha  & \sim \operatorname{Normal}(0, 0.5) \\
\beta_1 & \sim \operatorname{Normal}(0, 0.5) \\
\beta_2 & \sim \operatorname{Normal}(0, 0.5) \\
\sigma  & \sim \operatorname{Exponential}(1) \\
\color{#5D01A6FF}\nu & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Normal}(0, 0.5)} \\
\color{#5D01A6FF}{\sigma_B} & \color{#5D01A6FF}\sim \color{#5D01A6FF}{\operatorname{Exponential}(1)},
\end{align*}
$$

```{r}
b_model <-
  # here's the primary `k` model
  bf(k ~ 1 + mi(b) + m) +
  # here's the model for the missing `b` data
  bf(b | mi() ~ 1) +
  # here we set the residual correlations for the two models to zero
  set_rescor(FALSE)
```

```{r}
get_prior(data = d,
          family = gaussian,
          b_model)
```

```{r}
b15.5 <-
  brm(data = d,
      family = gaussian,
      b_model, # here we insert the model
      prior = 
        c(prior(normal(0, .5), class = Intercept, resp = k),
          prior(normal(0, .5), class = Intercept, resp = b),
          prior(normal(0, .5), class = b, resp = k),
          prior(exponential(1), class = sigma, resp = k),
          prior(exponential(1), class = sigma, resp = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.05")
```

```{r}
print(b15.5)
```

```{r}
posterior_summary(b15.5) %>% 
  round(digits = 2)
```
```{r}
as_draws_df(b15.5) %>% 
  select(starts_with("Ymi_b")) %>% 
  set_names(filter(d, is.na(b)) %>% pull(species)) %>% 
  pivot_longer(everything(),
               names_to = "species") %>% 
  
  ggplot(aes(x = value, y = reorder(species, value))) +
  stat_slab(fill = viridis_pal(option = "C")(7)[4],
            alpha = 3/4, height = 1.5, slab_color = "black",
            slab_size = 1/4) +
  labs(x = "imputed values for b", y = NULL) +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank())
```
Here's the model that drops the cases with NAs on `b`.
```{r}
b15.6 <-
  brm(data = d,
      family = gaussian,
      k ~ 1 + b + m,
      prior = c(prior(normal(0, .5), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.06")
```

```{r}
print(b15.6)
```

```{r}
bind_rows(fixef(b15.5) %>% data.frame() %>% rownames_to_column("term"),
          fixef(b15.6) %>% data.frame() %>% rownames_to_column("term")) %>% 
  slice(c(4:3, 6:7)) %>% 
  mutate(term = str_c("beta[", c(1:2, 1:2), "]"),
         fit = rep(c("b15.5", "b15.6"), each = n() / 2)) %>% 
  
  # plot!
  ggplot(aes(x = Estimate, y = fit)) +
  geom_vline(xintercept = 0, linetype = 3, alpha = 1/2) +
  geom_pointrange(aes(xmin = Q2.5, xmax = Q97.5)) +
  labs(x = "marginal posterior", y = NULL) +
  theme(axis.ticks.y = element_blank(),
        strip.background = 
          element_rect(color = "transparent",
                       fill = "transparent")) +
  facet_wrap(~ term, labeller = label_parsed, ncol = 1)
```

```{r}
d <-
  d %>% 
  mutate(row = 1:n()) %>% 
  left_join(
    posterior_summary(b15.5) %>% 
      data.frame() %>% 
      rownames_to_column("term") %>% 
      filter(str_detect(term, "Ymi")) %>% 
      mutate(row = str_extract(term, "(\\d)+") %>%
               as.integer()),
    by = "row")

d %>% 
  select(species, k:Q97.5) %>% 
  glimpse()
```

```{r}
color <- viridis_pal(option = "D")(7)[4]

p1 <-
  d %>% 
  ggplot() +
  geom_point(aes(x = b, y = k), color = color) +
  geom_pointrange(aes(xmin = Q2.5, xmax = Q97.5,
                      x = Estimate, y = k),
                  shape = 1, linewidth = 1/4, 
                  fatten = 4, stroke = 1/4) +
  labs(x = "neocortex percent (std)", y = "kcal milk (std)") +
  coord_cartesian(xlim = range(d$b, na.rm = T))

p2 <-
  d %>% 
  ggplot() +
  geom_point(aes(x = m, y = b), color = color) +
  geom_pointrange(aes(ymin = Q2.5, ymax = Q97.5,
                      y = Estimate, x = m),
                  shape = 1, linewidth = 1/4,
                  fatten = 4, stroke = 1/4) +
  labs(x = "log body mass (std)", 
       y = "neocortex percent (std)") +
  coord_cartesian(ylim = range(d$b, na.rm = T))

# combine and plot!
p1 + p2
```

$$
\begin{align*}
K_i   & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i & = \alpha + \beta_1 \color{#5D01A6FF}{B_i} + \beta_2 \log M_i \\
\color{#5D01A6FF}{\begin{bmatrix} M_i \\ B_i \end{bmatrix}} & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{MVNormal}
\begin{pmatrix}
\begin{bmatrix} \mu_M \\\mu_B \end{bmatrix},
\mathbf \Sigma \end{pmatrix}} \\
\alpha  & \sim \operatorname{Normal}(0, 0.5) \\
\beta_1 & \sim \operatorname{Normal}(0, 0.5) \\
\beta_2 & \sim \operatorname{Normal}(0, 0.5) \\
\sigma  & \sim \operatorname{Exponential}(1) \\
\color{#5D01A6FF}{\mu_M} & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Normal}(0, 0.5)} \\
\color{#5D01A6FF}{\mu_B} & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Normal}(0, 0.5)} \\
\color{#5D01A6FF}{\mathbf \Sigma} & \color{#5D01A6FF} = \color{#5D01A6FF}{\operatorname{\mathbf S \mathbf R \mathbf S}} \\
\color{#5D01A6FF}{\mathbf S} & \color{#5D01A6FF} = \color{#5D01A6FF}{\begin{bmatrix} \sigma_M & 0 \\ 0 & \sigma_B \end{bmatrix}} \\
\color{#5D01A6FF}{\mathbf R} & \color{#5D01A6FF} = \color{#5D01A6FF}{\begin{bmatrix} 1 & \rho \\ \rho & 1 \end{bmatrix}} \\
\color{#5D01A6FF}{\sigma_M}  & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Exponential}(1)} \\
\color{#5D01A6FF}{\sigma_B}  & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Exponential}(1)} \\
\color{#5D01A6FF} \rho       & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{LKJ}(2)},
\end{align*}
$$

Currently brms does not support this multivariate for only some of the variables. Until the update we can fit the alternative model

$$
\begin{align*}
K_i   & \sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i & = \alpha + \beta_1 \color{#5D01A6FF}{B_i} + \beta_2 \log M_i \\
\color{#5D01A6FF}{B_i}   & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Normal}(\nu_i, \sigma_B)} \\
\color{#5D01A6FF}{\nu_i} & \color{#5D01A6FF} = \color{#5D01A6FF}{\gamma + \delta_1 \log M_i} \\
\alpha  & \sim \operatorname{Normal}(0, 0.5) \\
\beta_1 & \sim \operatorname{Normal}(0, 0.5) \\
\beta_2 & \sim \operatorname{Normal}(0, 0.5) \\
\sigma  & \sim \operatorname{Exponential}(1) \\
\color{#5D01A6FF}\gamma     & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Normal}(0, 0.5)} \\
\color{#5D01A6FF}{\delta_1} & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Normal}(0, 0.5)} \\
\color{#5D01A6FF}{\sigma_B} & \color{#5D01A6FF} \sim \color{#5D01A6FF}{\operatorname{Exponential}(1)},
\end{align*}
$$

```{r}
b_model <-
  mvbf(bf(k ~ 1 + mi(b) + m),
       bf(b | mi() ~ 1 + m),
       rescor = FALSE)

b15.7 <-
  brm(data = d,
      family = gaussian,
      b_model,
      prior = c(prior(normal(0, .5), class = Intercept, resp = k),
                prior(normal(0, .5), class = Intercept, resp = b),
                prior(normal(0, .5), class = b, resp = k),
                prior(normal(0, .5), class = b, resp = b),
                prior(exponential(1), class = sigma, resp = k),
                prior(exponential(1), class = sigma, resp = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.07")
```

```{r}
print(b15.7)
```

```{r}
d %>% 
  select(-(term:Q97.5)) %>% 
  left_join(
    posterior_summary(b15.7) %>% 
      data.frame() %>% 
      rownames_to_column("term") %>% 
      filter(str_detect(term, "Ymi")) %>% 
      mutate(row = str_extract(term, "(\\d)+") %>% as.integer()),
    by = "row"
  )

d %>% 
  select(species, k:Q97.5) %>% 
  glimpse()
```

```{r}
color <- viridis_pal(option = "D")(7)[3]

p1 <-
  d %>% 
  ggplot(aes(y = k)) +
  geom_point(aes(x = b),
             color = color) +
  geom_pointrange(aes(x = Estimate, xmin = Q2.5, xmax = Q97.5),
                  shape = 1, linewidth = 1/4, fatten = 4, stroke = 1/4) +
  labs(x = "neocortex percent (std)",
       y = "kcal milk (std)") +
  coord_cartesian(xlim = range(d$b, na.rm = T))

p2 <-
  d %>% 
  ggplot(aes(x = m)) +
  geom_point(aes(y = b),
             color = color) +
  geom_pointrange(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
                  shape = 1, linewidth = 1/4, fatten = 4, stroke = 1/4) +
  labs(x = "log body mass (std)",
       y = "neocortex percent (std)") +
  coord_cartesian(ylim = range(d$b, na.rm = T))

p1 + p2
```

### 15.2.3 Where is your god now?
```{r}
data(Moralizing_gods, package = "rethinking")
d <- Moralizing_gods
rm(Moralizing_gods)
```

```{r}
glimpse(d)
```

```{r}
d %>% 
  count(moralizing_gods) %>% 
  mutate(`%` = 100 * n / sum(n))
```

```{r}
d %>% 
  mutate(mg = factor(ifelse(is.na(moralizing_gods), 2, 1 - moralizing_gods),levels = 0:2, labels = c("present", "absent", "unknown"))) %>% 
  
  ggplot(aes(x = year, y = log(population), color = mg)) +
  geom_point(alpha = 2/3) +
  scale_color_manual("Moralizing gods",
                     values = viridis_pal(option = "D")(7)[c(7, 4, 1)]) +
  labs(subtitle = '"This is a highly non-random missingness pattern" (p. 514).',
       x = "Time (year)", y = "Population size (log)") +
  theme(legend.position = c(.125, .67))
```

Here are the counts broken down by gods and literacy status
```{r}
d %>% 
  mutate(gods = moralizing_gods,
         literacy = writing) %>% 
  count(gods, literacy) %>% 
  mutate(`%` = 100 * n / sum(n))
```

```{r}
dag_coords <-
  tibble(name = c("P", "W", "G", "RG", "Gs"),
         x    = c(1, 2.33, 4, 5.67, 7),
         y    = c(2, 1, 2, 1, 2))

dagify(P  ~ G,
       W  ~ P,
       RG ~ W,
       Gs ~ G + RG,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "G", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("G", 
                                     expression(G^'*'),
                                     "P", 
                                     expression(R[G]), 
                                     "W")) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

```{r}
d %>% 
  filter(polity == "Big Island Hawaii") %>% 
  select(year, writing, moralizing_gods)
```
## 15.3 Categorical errors and discrete absences
### 15.3.1 Discrete cats.
```{r}
dag_coords <-
  tibble(name = c("Rc", "Cs", "C", "N"),
         x    = 1:4,
         y    = 1)

dagify(Cs ~ Rc + C,
       N  ~ C,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name == "C", "a", "b")) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 size = 7, show.legend = F) +
  geom_dag_text(parse = T, label = c("C", expression(C^"*"), "N",expression(R[C]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c(viridis_pal(option = "C")(7)[2], "black")) +
  dark_theme_void()
```

$$
\begin{align*}
N_i & \sim \operatorname{Poisson}(\lambda_i) \\
\log \lambda_i & = \alpha + \beta C_i \\
C_i & \sim \operatorname{Bernoulli}(k) \\
R_{C, i} & \sim \operatorname{Bernoulli}(r).
\end{align*}
$$

```{r}
set.seed(9)

n_houses <- 1000L
alpha <- 5
beta <- -3
k <- .5
r <- .2

dat <-
  tibble(cat = rbinom(n_houses, size = 1, prob = k)) %>% 
  mutate(notes = rpois(n_houses, lambda = alpha + beta * cat),
         r_c = rbinom(n_houses, size = 1, prob = r)) %>% 
  mutate(cat_obs = if_else(r_c == 1, (-9L), cat))

glimpse(dat)
```

## 15.5 Bonus: Bayesian meta-analysis with odds ratios
### 15.5.2 We need some data
```{r}
d <- 
  readr::read_csv("https://raw.githubusercontent.com/ASKurz/Statistical_Rethinking_with_brms_ggplot2_and_the_tidyverse_2_ed/master/data/Hauser_1_study_by_order_all_CLEAN_CASE.csv")

d <- 
  d %>% 
  mutate(y   = ifelse(variable == "Yes", 1, 0),
         loc = factor(Location,
                      levels = distinct(d, Location) %>% pull(Location),
                      labels = 1:59))

glimpse(d)
```

```{r}
d %>% 
  distinct(uID) %>% 
  count()
```

```{r}
d %>% 
  count(case.include)
```

```{r}
d %>% 
  count(Location) %>% 
  arrange(desc(n))
```

### 15.5.3 Our effect size will be an odds ratio.
```{r}
d %>% 
  count(variable, factor) %>% 
  group_by(factor) %>% 
  mutate(percent = 100 * n / sum(n))
```

```{r}
glm0 <- glm(data = d %>% 
              filter(loc == 1),
            y ~ factor, family = binomial(logit))

summary(glm0)
```

```{r}
coef(glm0)[2] %>% exp()
```

```{r}
d %>% 
  filter(loc == 1) %>% 
  count(variable, factor) %>% 
  group_by(factor) %>% 
  mutate(percent = 100 * n / sum(n)) %>% 
  filter(variable == "Yes")
```

### 15.5.5 Compute the study-specific effect sizes.
```{r}
library(broom)

glms <-
  d %>% 
  select(loc, y, factor) %>% 
  nest(data = c(y, factor)) %>% 
  mutate(glm = map(data, ~update(glm0, data = .))) %>% 
  mutate(coef = map(glm, tidy)) %>% 
  select(-data, -glm) %>% 
  unnest(coef) %>% 
  filter(term == "factorSideEffect")

# what did we do?
glms %>% 
  mutate_if(is.double, round, digits = 3)
```

```{r}
color <- viridis_pal(option = "C")(7)[5]

glms %>% 
  ggplot(aes(x = std.error, y = estimate)) +
  geom_point(color = color) +
  labs(x = expression(sigma[italic(j)]~("log-odds")),
       y = expression(italic(y[j])~("log-odds")))
```

### 15.5.6 Fit the Bayesian meta-analysis
$$
\begin{align*}
\text{estimate}_j & \sim \operatorname{Normal}(\theta_j, \; \text{std.error}_j) \\
\theta_j   & \sim \operatorname{Normal}(\mu, \tau) \\
\mu        & \sim \operatorname{Normal}(0, 1.5) \\
\tau       & \sim \operatorname{Exponential}(1),
\end{align*}
$$

```{r}
b15.10 <-
  brm(data = glms,
      family = gaussian,
      estimate | se(std.error) ~ 1 + (1 | loc),
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(exponential(1), class = sd)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.10")
```

```{r}
print(b15.10)
```

```{r}
fixef(b15.10) %>% exp()
```

### 15.5.7 Fit the Bayesian multilevel alternative
$$
\begin{align*}
\text{y}_{ij} & \sim \operatorname{Binomial}(n = 1, p_{ij}) \\
\operatorname{logit}(p_{ij}) & \sim \alpha + \beta \text{factor}_{ij} + u_{\alpha j} + u_{\beta j} \text{factor}_{ij} \\

\begin{bmatrix} u_{\alpha j} \\ u_{\beta j} \end{bmatrix} & \sim \operatorname{MVNormal} \begin{pmatrix} \begin{bmatrix} 0 \\ 0 \end{bmatrix}, \mathbf{SRS} \end{pmatrix} \\

\mathbf S & = \begin{bmatrix} \sigma_\alpha & 0 \\ 0 & \sigma_\beta \end{bmatrix} \\
\mathbf R & = \begin{bmatrix} 0 & \rho_{\alpha \beta} \\ \rho_{\beta \alpha} & 0 \end{bmatrix} \\

\alpha & \sim \operatorname{Normal}(0, 1.5) \\
\beta  & \sim \operatorname{Normal}(0, 1.5) \\
\sigma_\alpha & \sim \operatorname{Exponential}(1) \\
\sigma_\beta  & \sim \operatorname{Exponential}(1) \\
\mathbf R & \sim \operatorname{LKJ}(2),
\end{align*}
$$

```{r}
b15.11 <-
  brm(data = d,
      family = binomial,
      y | trials(1) ~ 0 + Intercept + factor + (1 + factor | loc),
      prior = c(prior(normal(0, 1.5), class = b),
                prior(exponential(1), class = sd),
                prior(lkj(2), class = cor)),
      iter = 2000, warmup = 1000, chains = 4, cores = 8,
      seed = 15,
      file = "fits/b15.11")
```

```{r}
print(b15.11)
```

```{r}
fixef(b15.11)[2, -2] %>% exp()
```

```{r}
color <- viridis_pal(option = "C")(7)[3]

# how many levels are there?
n_loc <- distinct(d, loc) %>% count() %>% pull(n)

# rank by meta-analysis
ranks <-
  tibble(Estimate = coef(b15.10)$loc[, 1, "Intercept"],
         index = 1:n_loc) %>% 
  arrange(Estimate) %>% 
  mutate(rank = 1:n_loc)

# combine parameter summaries
rbind(coef(b15.10)$loc[, , "Intercept"],
      coef(b15.11)$loc[, , "factorSideEffect"]) %>% 
  data.frame() %>% 
  mutate(index = rep(1:n_loc, times = 2),
         type = rep(c("meta-analysis", "multilevel model"), each = n_loc)) %>% 
  # add the ranks
  left_join(select(ranks, - Estimate),
            by = "index") %>% 
  
  ggplot(aes(x = Estimate, xmin = Q2.5, xmax = Q97.5, y = rank)) +
  geom_pointrange(fatten = 1, color = color) +
  scale_x_continuous(expression(log-odds~effect~size~(theta[italic(j)])), limits = c(0, 4.5)) +
  scale_y_continuous(NULL, breaks = NULL) +
  facet_wrap(~type)
```

```{r}
labs <-
  d %>% 
  mutate(lab = case_when(
    Location == "Social and Work Psychology Department, University of Brasilia, DF, Brazil" ~ "University of Brasilia",
    Location == "Wilfrid Laurier University, Waterloo, Ontario, Canada" ~ "Wilfrid Laurier University",
    Location == "University of British Columbia, Vancouver, Canada" ~ "University of British Columbia",
    Location == "University of Toronto, Scarborough" ~ "University of Toronto",
    Location == "Division of Social Science, The Hong Kong University of Science and Technology, Hong Kong, China" ~ "Hong Kong University of Science and Technology",
    Location == "Chinese Academy of Science, Beijing, China" ~ "Chinese Academy of Science",
    Location == "Shanghai International Studies University, SISU Intercultural Institute, Shanghai, China" ~ "Shanghai International Studies University",
    Location == "Guangdong Literature & Art Vocational College, Guangzhou, China" ~ "Guangdong Literature & Art Vocational College",
    Location == "The University of J. E. Purkyně, Ústí nad Labem, Czech Republic" ~ "The University of J. E. Purkyně",
    Location == "University of Leuven, Belgium" ~ "University of Leuven",
    Location == "Department of Experimental and Applied Psychology, VU Amsterdam, 1081BT, Amsterdam, The Netherlands" ~ "VU Amsterdam",
    Location == "Department of Social Psychology, Tilburg University, P.O. Box 90153, Tilburg, 5000 LE, Netherlands" ~ "Department of Social Psychology, Tilburg University",
    Location == "Eindhoven University of Technology, Eindhoven, Netherlands" ~ "Eindhoven University of Technology",
    Location == "Department of Communication and Information Sciences, P.O. Box 90153, Tilburg, 5000 LE, Netherlands" ~ "Department of Communication and Information Sciences, Tilburg University",
    Location == "University of Navarra, Spain" ~ "University of Navarra",
    Location == "University of Lausanne, Switzerland" ~ "University of Lausanne",
    Location == "Université de Poitiers, France" ~ "Université de Poitiers",
    Location == "Eotvos Lorand University, in Budapest, Hungary" ~ "Eotvos Lorand University",
    Location == "MTurk India Workers" ~ "MTurk India Workers",
    Location == "University of Winchester, Winchester, Hampshire, England" ~ "University of Winchester",
    Location == "Doshisha University, Kyoto, Japan" ~ "Doshisha University",
    Location == "Victoria University of Wellington, New Zealand" ~ "Victoria University of Wellington",
    Location == "University of Social Sciences and Humanities, Wroclaw, Poland" ~ "University of Social Sciences and Humanities",
    Location == "Department of Psychology, SWPS University of Social Sciences and Humanities Campus Sopot, Sopot, Poland" ~ "SWPS University of Social Sciences and Humanities Campus Sopot",
    Location == "badania.net" ~ "badania.net",
    Location == "Universidade do Porto, Portugal" ~ "Universidade do Porto",
    Location == "University of Belgrade, Belgrade, Serbia" ~ "University of Belgrade",
    Location == "University of Johannesburg, Johanneburg, South Africa" ~ "University of Johannesburg",
    Location == "Santiago, Chile" ~ "Santiago, Chile",
    Location == "Universidad de Costa Rica, Costa Rica" ~ "Universidad de Costa Rica",
    Location == "National Autonomous University of Mexico in Mexico City" ~ "National Autonomous University of Mexico",
    Location == "University of the Republic, Montevideo, Uruguay" ~ "University of the Republic",
    Location == "Lund University, Lund, Sweden" ~ "Lund University",
    Location == "Academia Sinica, Taiwan National Taiwan Normal University, Taiwan" ~ "Taiwan National Taiwan Normal University",
    Location == "Bilgi University, Istanbul, Turkey" ~ "Bilgi University",
    Location == "Koç University, Istanbul, Turkey" ~ "Koç University",
    Location == "American University of Sharjah, United Arab Emirates" ~ "American University of Sharjah",
    Location == "University of Hawaii, Honolulu, HI" ~ "University of Hawaii",
    Location == "Social Science and Policy Studies Department, Worcester Polytechnic Institute, Worcester, MA 01609" ~ "Worcester Polytechnic Institute",
    Location == "Department of Psychology, Washington and Lee University, Lexington, VA 24450" ~ "Washington and Lee University",
    Location == "Department of Psychology, San Diego State University, San Diego, CA 92182" ~ "San Diego State University",
    Location == "Tufts" ~ "Tufts",
    Location == "University of Florida, Florida" ~ "University of Florida",
    Location == "University of Illinois at Urbana-Champaign, Champaign, IL" ~ "University of Illinois at Urbana-Champaign",
    Location == "Pacific Lutheran University, Tacoma, WA" ~ "Pacific Lutheran University",
    Location == "University of Virginia, VA" ~ "University of Virginia",
    Location == "Marian University, Indianapolis, IN" ~ "Marian University",
    Location == "Department of Psychology, Ithaca College, Ithaca, NY 14850" ~ "Ithaca College",
    Location == "University of Michigan" ~ "University of Michigan",
    Location == "Department of Psychology, Pennsylvania State University Abington, Abington, PA 19001" ~ "Pennsylvania State University Abington",
    Location == "Department of Psychology, Texas A&M University, College Station, TX 77843" ~ "Texas A&M University",
    Location == "William Paterson University, Wayne, NJ" ~ "William Paterson University",
    Location == "Department of Cognitive Science, Occidental College, Los Angeles, CA" ~ "Occidental College",
    Location == "The Pennsylvania State University" ~ "The Pennsylvania State University",
    Location == "MTurk US Workers" ~ "MTurk US Workers",
    Location == "University of Graz AND the Universty of Vienna" ~ "University of Graz and the Universty of Vienna",
    Location == "University of Potsdam, Germany" ~ "University of Potsdam",
    Location == "Open University of Hong Kong" ~ "Open University of Hong Kong",
    Location == "Concepción, Chile" ~ "Concepción"
  )) %>% 
  distinct(loc, lab)

# what is this?
labs %>% 
  glimpse()
```

```{r}
# this will help us format the labels on the secondary y-axis
my_format <- function(number){
  formatC(number, digits = 2, format = "f")
}

# grab the theta_j summaries
groups <-
  coef(b15.11)$loc[, , "factorSideEffect"] %>% 
  data.frame() %>% 
  mutate(loc = distinct(d, loc) %>% pull()) %>% 
  arrange(Estimate)

# grab the mu summary
average <-
  fixef(b15.11) %>% 
  data.frame() %>% 
  slice(2) %>% 
  mutate(loc = "Average")

# combine and wrangle
post <-
  bind_rows(groups, average) %>% 
  mutate(rank = c(1:59, 0),
         Estimate = exp(Estimate),
         Q2.5 = exp(Q2.5),
         Q97.5 = exp(Q97.5)) %>% 
  left_join(labs, by = "loc") %>% 
  arrange(rank) %>% 
  mutate(label = ifelse(is.na(lab), "POPULATION AVERAGE", lab),
         summary = str_c(my_format(Estimate), " [",
                         my_format(Q2.5), ", ",
                         my_format(Q97.5), "]"))

# what have we dont?
post %>% glimpse()
```
```{r}
post %>% 
  ggplot(aes(x = Estimate, xmin = Q2.5, xmax = Q97.5,
             y = rank)) +
  geom_interval(aes(color = label == "POPULATION AVERAGE"),
                size = 1/2) +
  geom_point(aes(size = 1 - Est.Error, 
                 color = label == "POPULATION AVERAGE"),
             shape = 15) +
  scale_color_viridis_d(option = "C", begin = .33, end = .67) +
  scale_size_continuous(range = c(1, 3.5)) +
  scale_x_continuous("odds ratio", breaks = 1:6 * 10, expand = expansion(mult = c(0.005, 0.005))) +
  scale_y_continuous(NULL, breaks = 0:59, limits = c(-1, 60), expand = c(0, 0),
                     labels = pull(post, label),
                     sec.axis = dup_axis(labels = pull(post, summary))) +
  theme(text = element_text(family = "Times"),
        axis.text.y = element_text(hjust = 0, color = "white", size = 7),
        axis.text.y.right = element_text(hjust = 1, size = 7),
        axis.ticks.y = element_blank(),
        panel.background = element_rect(fill = "grey8"),
        panel.border = element_rect(color = "transparent"))
```

```{r}
post <- as_draws_df(b15.10)

post %>% 
  select(b_Intercept:sd_loc__Intercept) %>% 
  head()
```

```{r}
color <- viridis_pal(option = "C")(7)[6]
set.seed(15)

post %>% 
  transmute(lo = rnorm(n(), mean = b_Intercept, sd = sd_loc__Intercept),
            or = rnorm(n(), mean = b_Intercept, sd = sd_loc__Intercept) %>% exp()) %>% 
  slice(1:1e3) %>% 
  pivot_longer(lo:or, values_to = "effect size") %>% 
  mutate(name = factor(name, labels = c("log-odds", "odds ratio"))) %>% 
  
  ggplot(aes(x = `effect size`, y = 0)) +
  geom_dots(color = color, fill = color) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(Normal(mu*', '*tau))) +
  theme(text = element_text(family = "Times"),
        strip.background = element_rect(color = "transparent")) +
  facet_wrap(~ name, scales = "free")
```

### 15.5.8 Relative to the multilevel model, meta-analysis have limitations.
```{r}
post <- as_draws_df(b15.11)
```

```{r}
color <- viridis_pal(option = "C")(7)[7]

c <-
  full_join(
  # intercepts
  post %>% 
    select(`r_loc[1,Intercept]`:`r_loc[59,Intercept]`) %>% 
    set_names(1:59) %>% 
    mutate(iter = 1:n()) %>% 
    pivot_longer(-iter, values_to = "intercepts"),
  # slopes
  post %>% 
    select(`r_loc[1,factorSideEffect]`:`r_loc[59,factorSideEffect]`) %>% 
    set_names(1:59) %>% 
    mutate(iter = 1:n()) %>% 
    pivot_longer(-iter, values_to = "slopes"),
  by = c("iter", "name")
) 

# summarize by the means
c %>% 
  group_by(name) %>% 
  summarise(intercepts = mean(intercepts),
            slopes = mean(slopes)) %>% 
  
  # plot!
  ggplot(aes(x = intercepts, y = slopes)) +
  geom_point(color = color) +
  labs(x = expression(alpha[italic(j)]),
       y = expression(beta[italic(j)]))
```

```{r}
c %>% 
  ggplot(aes(x = intercepts, y = slopes)) +
  stat_ellipse(aes(group = name),
               geom = "polygon", level = .05, linewidth = 0,
               fill = color, alpha = 1/2) +
  stat_ellipse(level = .95, color = color) +
  labs(x = expression(alpha[italic(j)]),
       y = expression(beta[italic(j)]))
```

```{r}
breaks <- c(0, inv_logit_scaled(c(fixef(b15.11)[1, 1], fixef(b15.11)[1, 1] + fixef(b15.11)[2, 1])), 1)

nd <- d %>% distinct(loc, factor)

fitted(b15.11, newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  mutate(factor = if_else(factor == "SideEffect",
                          "Side effect", "Greater good")) %>% 
  
  ggplot(aes(x = factor, y = Estimate, group = loc)) +
  geom_line(alpha = 1/2, color = color) +
  scale_y_continuous("probability", limits = 0:1,
                     breaks = breaks, 
                     labels = round(breaks, digits = 2)) +
  xlab(NULL) +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.x = element_blank())
```

