---
title: "Chapter 2 BRMS"
format: html
---

# 2.1 The farden of forking data

## 2.1.1 Counting possibilities
```{r}
rm(list = ls())
library(tidyverse)

d <- tibble(p1 = 0,
            p2 = rep(1:0, times = c(1, 3)),
            p3 = rep(1:0, times = c(2, 2)),
            p4 = rep(1:0, times = c(3, 1)),
            p5 = 1)

head(d)
```
```{r}
d %>%
  set_names(1:5) %>%
  mutate(x = 1:4) %>%
  pivot_longer(-x, names_to = "possibility") %>%
  mutate(value = value %>% as.character()) %>%
  
  ggplot(aes(x = x, y = possibility, fill = value)) +
  geom_point(shape = 21, size = 5) +
  scale_fill_manual(values = c("white", "navy")) +
  scale_x_discrete(NULL, breaks = NULL) +
  theme(legend.position = "none")
```

```{r}
library(flextable)

tibble(draw = 1:3,
       marbles = 4) %>%
  mutate(possibilities = marbles ^ draw) %>%
  flextable() 
```

```{r}
(
  d <- 
    tibble(position = c((1:4^1) / 4^0,
                        (1:4^2) / 4^1,
                        (1:4^3) / 4^2),
           draw = rep(1:3, times = c(4^1, 4^2, 4^3)),
           fill = rep(c("b", "w"), times = c(1, 3)) %>%
             rep(., times = c(4^0 + 4^1 + 4^2)))
)

d %>%
  ggplot(aes(x = position, y = draw, fill = fill)) +
  geom_point(shape = 21, size = 3) +
  scale_fill_manual(values = c("navy", "white")) +
  scale_y_continuous(breaks = 1:3) +
  theme(legend.position = "none",
        panel.grid.minor = element_blank())
```
```{r}
(
  lines_1 <-
    tibble(x = rep(1:4, each = 4),
           xend = ((1:4^2) / 4),
           y = 1,
           yend = 2)
)

(
  lines_2 <-
    tibble(x = rep((1:4^2) / 4, each = 4),
           xend = (1:4^3) / (4^2),
           y = 2,
           yend = 3)
)

d %>%
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_segment(data = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_fill_manual(values = c("navy", "white")) +
  scale_y_continuous(breaks = 1:3) +
  theme(legend.position = "none",
        panel.grid.minor = element_blank())
```
Make the dot centered
```{r}
d <- 
  d %>%
  mutate(denominator = ifelse(draw == 1, .5,
                              ifelse(draw == 2, .5 / 4,
                                     .5 / 4^2))) %>%
  mutate(position = position - denominator)

d

(
  lines_1 <-
    lines_1 %>%
    mutate(x = x - 0.5,
           xend = xend - 0.5 / 4^1)
)

(
  lines_2 <-
    lines_2 %>%
    mutate(x = x - 0.5 / 4^1,
           xend = xend - 0.5 / 4^2)
)

d %>%
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_segment(data = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_fill_manual(values = c("navy", "white")) +
  scale_y_continuous(breaks = 1:3) +
  theme(legend.position = "none",
        panel.grid.minor = element_blank())
```
Change to polar coordinate
```{r}
d %>%
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_segment(data = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               linewidth = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_fill_manual(values = c("navy", "white")) +
  scale_y_continuous(breaks = 1:3) +
  coord_polar() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank())
```

Show path that remain logically valid
```{r}
lines_1 <-
  lines_1 %>%
  mutate(remain = c(rep(0:1, times = c(1, 3)),
                    rep(0, times = 4 * 3)))

lines_2 <-
  lines_2 %>%
  mutate(remain = c(rep(0, times = 4),
                    rep(1:0, times = c(1, 3)) %>% rep(., times = 3),
                    rep(0, times = 12 * 4)))

d <-
  d %>%
    mutate(remain = c(rep(1:0, times = c(1, 3)),
                    rep(0:1, times = c(1, 3)),
                    rep(0,   times = 4 * 4),
                    rep(1:0, times = c(1, 3)) %>% rep(., times = 3),
                    rep(0,   times = 12 * 4))) 

d %>%
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               linewidth = 1/3) +
   geom_segment(data = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               linewidth = 1/3) +
  geom_point(aes(fill = fill, alpha = remain %>% as.character()),
             shape = 21, size = 4) +
  # it's the alpha parameter that makes elements semitransparent
  scale_fill_manual(values = c("navy", "white")) +
  scale_alpha_manual(values = c(1/5, 1)) +
  scale_x_continuous(NULL, limits = c(0, 4), breaks = NULL) +
  scale_y_continuous(NULL, limits = c(0.75, 3), breaks = NULL) +
  coord_polar() +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

Create table to show the count of marbles
```{r}
# if we make two custom functions, here,
# it will simplify the 'mutate()' code, below
n_blue <- function(x) rowSums(x == "b")
n_white <- function(x) rowSums(x == "w")

# make the data
t <-
  tibble(d1 = rep(c("w", "b"), times = c(1, 4)),
         d2 = rep(c("w", "b"), times = c(2, 3)),
         d3 = rep(c("w", "b"), times = c(3, 2)),
         d4 = rep(c("w", "b"), times = c(4, 1))) %>%
  mutate(blue1 = n_blue(.),
         white = n_white(.),
         blue2 = n_blue(.)) %>%
  mutate(product = blue1 * white * blue2)

# format the table
t %>%
  transmute(conjecture = str_c("[", d1, " ", d2, " ", d3, " ", d4, "]"),
            'Ways to produce [w b w]' = str_c(blue1, " * ", white, " * ", blue2, " = ", product)) %>%
  flextable() %>%
  width(j = 1:2, width = c(1, 2)) %>%
  align(align = "center", part = "all")
```
```{r}
d <-
  tibble(position = c((1:4^1) / 4^0,
                      (1:4^2) / 4^1,
                      (1:4^3) / 4^2),
         draw = rep(1:3, times = c(4^1, 4^2, 4^3)))

(
  d <-
    d %>%
    bind_rows(
      d, d
    ) %>%
    # here are the fill colors
    mutate(fill = c(rep(c("w", "b"), times = c(1, 3)) %>% rep(., times = c(4^0 + 4^1 + 4^2)),
                    rep(c("w", "b"), each = 2) %>% rep(., times = c(4^0 + 4^1 + 4^2)),
                    rep(c("w", "b"), times = c(3, 1)) %>% rep(., times = c(4^0 + 4^1 + 4^2)))) %>%
    mutate(denominator = ifelse(draw == 1, .5,
                                ifelse(draw == 2, .5 / 4,
                                       .5 / 4^2))) %>%
    mutate(position = position - denominator) %>%
    # here we'll add an index for which pie wedge we're working with
     mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for pie_index == "b" or "c", we'll need to offset
  mutate(position = ifelse(pie_index == "a", position,
                           ifelse(pie_index == "b", position + 4,
                                  position + 4 * 2)))
)
```
```{r}
move_over <- function(position, index){
  ifelse(
    index == "a", position,
    ifelse(
      index == "b", position + 4, position + 4 + 2
    )
  )
}
```

```{r}
(
  lines_1 <-
  tibble(x    = rep(1:4, each = 4) %>% rep(., times = 3),
         xend = ((1:4^2) / 4)      %>% rep(., times = 3),
         y    = 1,
         yend = 2) %>% 
  mutate(x    = x - .5,
         xend = xend - .5 / 4^1) %>% 
  # here we'll add an index for which pie wedge we're working with
  mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for `pie_index == "b"` or `"c"`, we'll need to offset
  mutate(x    = move_over(position = x,    index = pie_index),
         xend = move_over(position = xend, index = pie_index))
)
```

```{r}
(
  lines_2 <-
  tibble(x    = rep((1:4^2) / 4, each = 4) %>% rep(., times = 3),
         xend = (1:4^3 / 4^2)              %>% rep(., times = 3),
         y    = 2,
         yend = 3) %>% 
  mutate(x    = x - .5 / 4^1,
         xend = xend - .5 / 4^2) %>% 
  # here we'll add an index for which pie wedge we're working with
  mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for `pie_index == "b"` or `"c"`, we'll need to offset
  mutate(x    = move_over(position = x,    index = pie_index),
         xend = move_over(position = xend, index = pie_index))
)
```

```{r}
d <- 
  d %>% 
  mutate(remain = c(# pie_index == "a"
                    rep(0:1, times = c(1, 3)),
                    rep(0,   times = 4),
                    rep(1:0, times = c(1, 3)) %>% rep(., times = 3),
                    rep(0,   times = 4 * 4),
                    rep(c(0, 1, 0), times = c(1, 3, 4 * 3)) %>% rep(., times = 3),
                    # pie_index == "b"
                    rep(0:1, each = 2),
                    rep(0,   times = 4 * 2),
                    rep(1:0, each = 2) %>% rep(., times = 2),
                    rep(0,   times = 4 * 4 * 2),
                    rep(c(0, 1, 0, 1, 0), times = c(2, 2, 2, 2, 8)) %>% rep(., times = 2),
                    # pie_index == "c",
                    rep(0:1, times = c(3, 1)),
                    rep(0,   times = 4 * 3),
                    rep(1:0, times = c(3, 1)), 
                    rep(0,   times = 4 * 4 * 3),
                    rep(0:1, times = c(3, 1)) %>% rep(., times = 3),
                    rep(0,   times = 4)
                    )
         )

lines_1 <-
  lines_1 %>% 
  mutate(remain = c(rep(0,   times = 4),
                    rep(1:0, times = c(1, 3)) %>% rep(., times = 3),
                    rep(0,   times = 4 * 2),
                    rep(1:0, each  = 2) %>% rep(., times = 2),
                    rep(0,   times = 4 * 3),
                    rep(1:0, times = c(3, 1))
                    )
         )

lines_2 <-
  lines_2 %>% 
  mutate(remain = c(rep(0,   times = 4 * 4),
                    rep(c(0, 1, 0), times = c(1, 3, 4 * 3)) %>% rep(., times = 3),
                    rep(0,   times = 4 * 8),
                    rep(c(0, 1, 0, 1, 0), times = c(2, 2, 2, 2, 8)) %>% rep(., times = 2),
                    rep(0,   times = 4 * 4 * 3),
                    rep(0:1, times = c(3, 1)) %>% rep(., times = 3),
                    rep(0,   times = 4)
                    )
         )
```

```{r}
d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_vline(xintercept = c(0, 4, 8), color = "white", linewidth = 2/3) +
  geom_segment(data = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               linewidth = 1/3) +
  geom_segment(data = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               linewidth = 1/3) +
  geom_point(aes(fill = fill, size = draw, alpha = remain %>% as.character()),
             shape = 21) +
  scale_fill_manual(values = c("navy", "white")) +
  scale_size_continuous(range = c(3, 1.5)) +
  scale_alpha_manual(values = c(0.2, 1)) +
  scale_x_continuous(NULL, limits = c(0, 12), breaks = NULL) +
  scale_y_continuous(NULL, limits = c(0.75, 3.5), breaks = NULL) +
  coord_polar() +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

## 2.1.2 Combining other information
```{r}
# update t
t <-
  t %>% 
  mutate(nc = blue1 * product)

# format the table
t %>%
  transmute(Conjecture            = str_c("[", d1, " ", d2, " ", d3, " ", d4, "]"),
            `Ways to produce [b]` = blue1,
            `Prior counts`        = product,
            `New count`           = str_c(blue1, " * ", product, " = ", nc)) %>% 
  flextable() %>% 
  width(width = c(1, 1, 0.8, 1)) %>% 
  align(align = "center", part = "all") %>% 
  align(j = 4, align = "left", part = "all") %>% 
  valign(valign = "bottom", part = "header")
```

```{r}
# update t
t <-
  t %>% 
  rename(pc = nc) %>% 
  mutate(fc = c(0, 3:0)) %>% 
  mutate(nc = pc * fc) 

# format the table
t %>% 
  transmute(Conjecture      = str_c("[", d1, " ", d2, " ", d3, " ", d4, "]"),
            `Prior count`   = pc,
            `Factory count` = fc,
            `New count`     = str_c(pc, " * ", fc, " = ", nc)) %>% 
  flextable() %>% 
  width(width = c(1, 1, 0.8, 1)) %>% 
  align(align = "center", part = "all") %>% 
  align(j = 4, align = "left", part = "all") %>% 
  valign(valign = "bottom", part = "header")
```
## 2.1.3 From counts to probability
```{r}
# update t
t %>% 
  rename(ways = product) %>% 
  mutate(p = blue1 / 4) %>% 
  mutate(pl = ways / sum(ways)) %>% 
  transmute(`Possible composition` = str_c("[", d1, " ", d2, " ", d3, " ", d4, "]"),
            p                      = p,
            `Ways to produce data` = ways,
            `Plausibility`         = pl) %>% 
  
  # format for the table
  flextable() %>% 
  width(width = c(1.8, 1, 1.2, 1)) %>% 
  align(align = "center", part = "all") %>% 
  valign(valign = "bottom", part = "header") %>% 
  italic(j = 2, part = "header")
```
```{r}
ways <- c(0, 3, 8, 9, 0)

ways / sum(ways)
```


# 2.2 Building a model
```{r}
(d <- tibble(toss = c("w", "l", "w", "w", "w", "l", "w", "l", "w")))
```
## 2.2.2 Bayesian updating
```{r}
(
  d <- 
    d %>%
    mutate(n_trials = 1:9,
           n_success = cumsum(toss == "w"))
)
```

```{r}
sequence_length <- 50

d %>% 
  expand_grid(p_water = seq(from = 0, to = 1, length.out = sequence_length)) %>% 
  group_by(p_water) %>% 
  # to learn more about lagging, go to:
  # https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lag
  # https://dplyr.tidyverse.org/reference/lead-lag.html
  mutate(lagged_n_trials  = lag(n_trials, n = 1),
         lagged_n_success = lag(n_success, n = 1)) %>% 
  ungroup() %>% 
  mutate(prior      = ifelse(n_trials == 1, .5,
                             dbinom(x    = lagged_n_success, 
                                    size = lagged_n_trials, 
                                    prob = p_water)),
         likelihood = dbinom(x    = n_success, 
                             size = n_trials, 
                             prob = p_water),
         strip      = str_c("n = ", n_trials)) %>% 
  # the next three lines allow us to normalize the prior and the likelihood, 
  # putting them both in a probability metric 
  group_by(n_trials) %>% 
  mutate(prior      = prior / sum(prior),
         likelihood = likelihood / sum(likelihood)) %>% 
  
  # plot!
  ggplot(aes(x = p_water)) +
  geom_line(aes(y = prior), 
            linetype = 2) +
  geom_line(aes(y = likelihood)) +
  scale_x_continuous("proportion water", breaks = c(0, .5, 1)) +
  scale_y_continuous("plausibility", breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ strip, scales = "free_y")
```

# 2.3 Components of the model

### 2.3.2.1 Observed variables
```{r}
tibble(prob = seq(from = 0, to = 1, by= .01)) %>%
  ggplot(aes(x = prob, y = dbinom(x = 6, size = 9, prob = prob))) +
  geom_line() +
  labs(x = "probability",
       y = "binomial likelihood") +
  theme(panel.grid = element_blank())
```

```{r}
tibble(a = 0,
       b = c(1, 1.5, 2, 3, 9)) |>
  expand_grid(parameter_space = seq(from = 0, to = 9, length.out = 500)) |>
  mutate(prob = dunif(parameter_space, a, b),
         b = str_c("b = ", b)) %>%
  
  ggplot(aes(x = parameter_space, y = prob)) +
  geom_area() +
  scale_x_continuous(breaks = c(0, 1:3, 9)) +
  scale_y_continuous(breaks = c(0, 1/9, 1/3, 1/2, 2/3, 1),
                     labels = c("0", "1/9", "1/3", "1/2", "2/3", "1")) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())+
  facet_wrap(~ b, ncol = 5)
```

# 2.4 Making the model go
## 2.4.1 Bayes' theorem
```{r}
sequence_length <- 1e3

d <-
  tibble(probability = seq(from = 0, to = 1, length.out = sequence_length)) %>%
  expand_grid(row = c("flat", "stepped", "Laplace")) %>%
  arrange(row, probability) %>%
  mutate(prior = ifelse(row == "flat", 1,
                        ifelse(row == "stepped", rep(0:1, each = sequence_length / 2),
                               exp(-abs(probability - 0.5) / .25) / (2 * 0.25))),
         likelihood = dbinom(x = 6, size = 9, prob = probability)) %>%
  group_by(row) %>%
  mutate(posterior = prior * likelihood / sum(prior * likelihood)) %>%
  pivot_longer(prior:posterior) %>%
  ungroup() %>%
  mutate(name = factor(name, levels = c("prior", "likelihood", "posterior")),
         row = factor(row, levels = c("flat", "stepped", "Laplace")))
```

```{r}
p1 <-
  d %>%
  filter(row == "flat") %>%
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

p2 <-
  d %>%
  filter(row == "stepped") %>%
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

p3 <-
  d %>%
  filter(row == "Laplace") %>%
  ggplot(aes(x = probability, y = value)) +
  geom_line() +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ name, scales = "free_y")

library(patchwork)

p1 / p2 / p3
```

## 2.4.3 Grid approximation
```{r}
(
  d <- 
    tibble(p_grid = seq(from = 0, to = 1, length.out = 20), # define grid
           prior = 1) %>% # define prior
    mutate(likelihood = dbinom(6, size = 9, prob = p_grid)) %>% # compute likelihood at each value in grid
    mutate(unstd_posterior = likelihood * prior) %>% # compute product of likelihood and prior
    mutate(posterior = unstd_posterior / sum(unstd_posterior)) # standardize the posterior, so it sums to 1
)
```

```{r}
p1 <-
  d %>%
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(subtitle = "20 points",
       x = "probability of water",
       y = "posterior probability") +
  theme(panel.grid = element_blank())

p2 <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = 5),
         prior = 1) %>%
  mutate(likelihood = dbinom(6, size = 9, prob = p_grid)) %>%
  mutate(unstd_posterior = likelihood * prior) %>%
  mutate(posterior = unstd_posterior / sum(unstd_posterior)) %>%
  
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(subtitle = "5 points",
       x = "probability of water",
       y = "posterior probability") +
  theme(panel.grid = element_blank())
  
p2 + p1 + plot_annotation(title = "More grid points make for smoother approximation")
```

```{r}
# make the data
tibble(n_points = c(5, 20)) %>%
  mutate(p_grid = map(n_points, ~seq(from = 0, to = 1, length.out = .))) %>%
  unnest(p_grid) %>%
  expand_grid(priors = c("ifelse(p_grid < 0.5, 0, 1)", "exp(-5 * abs(p_grid = 0.5))")) %>%
  mutate(prior = ifelse(priors == "ifelse(p_grid < 0.5, 0, 1)",
                        ifelse(p_grid < 0.5, 0, 1),
                        exp(-5 * abs(p_grid - 0.5)))) %>%
  mutate(likelihood = dbinom(6, size = 9, prob = p_grid)) %>%
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>%
  mutate(n_points = str_c("# points = ", n_points),
         priors = str_c("prior = ", priors)) %>%
  
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  geom_point() +
  labs(x = "probability of water",
       y = "posterior probability") +
  theme(panel.grid = element_blank()) +
  facet_grid(n_points ~ priors, scales = "free")
```

## 2.4.4 Quadratic approximation
```{r}
library(rethinking)

globe_qa <- quap(
  alist(
    W ~ dbinom(W + L, p), # binomial likelihood
    p ~ dunif(0, 1) # uniform prior
  ),
  data = list(W = 6, L = 3)
)

precis(globe_qa)
```

```{r}
globe_qa_18 <-
  quap(
    alist(
      w ~ dbinom(9 * 2, p),
      p ~ dunif(0, 1)
    ), data = list(w = 6 * 2))

globe_qa_36 <-
  quap(
    alist(
      w ~ dbinom(9 * 4, p),
      p ~ dunif(0, 1)
    ), data = list(w = 6 * 4))

precis(globe_qa_18)
```

```{r}
n_grid <- 100

# wrangle
tibble(w = c(6, 12, 24),
       n = c(9, 18, 36),
       s = c(.16, .11, .08)) %>% 
  expand_grid(p_grid = seq(from = 0, to = 1, length.out = n_grid)) %>% 
  mutate(prior = 1,
         m     = .67)  %>%
  mutate(likelihood = dbinom(w, size = n, prob = p_grid)) %>%
  mutate(unstd_grid_posterior = likelihood * prior,
         unstd_quad_posterior = dnorm(p_grid, m, s)) %>%
  group_by(w) %>% 
  mutate(grid_posterior = unstd_grid_posterior / sum(unstd_grid_posterior),
         quad_posterior = unstd_quad_posterior / sum(unstd_quad_posterior),
         n              = str_c("n = ", n)) %>% 
  mutate(n = factor(n, levels = c("n = 9", "n = 18", "n = 36"))) %>% 
  
  # plot
  ggplot(aes(x = p_grid)) +
  geom_line(aes(y = grid_posterior)) +
  geom_line(aes(y = quad_posterior),
            color = "grey50") +
  labs(x = "proportion water",
       y = "density") +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ n, scales = "free")
```

## 2.4.5 Markov chain Monte Carlo
```{r}
library(brms)

b2.1 <-
  brm(data = list(w = 24),
      family = binomial(link = "identity"),
      w | trials(36) ~ 0 + Intercept,
      prior(beta(1, 1), class = b, lb = 0, ub = 1),
      seed = 2,
      file = "fits/b02.01")

print(b2.1)
```

```{r}
posterior_summary(b2.1) %>%
  round(digits = 2)
```

```{r}
as_draws_df(b2.1) %>% 
  mutate(n = "n = 36") %>%
  
  ggplot(aes(x = b_Intercept)) +
  geom_density(fill = "black") +
  scale_x_continuous("proportion water", limits = c(0, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ n)
```

