---
title: "14 Adventures in Covariance"
format: html
---

## 14.1 Varying slopes by construction
### 14.1.1 Simulate the population.
```{r}
a <- 3.5 # average morning wait time
b <- -1 # average difference afternoon wait time
sigma_a <- 1 # std dev in intercepts
sigma_b <- 0.5 # std dev in slopes
rho <- -.7 # correlation between intercepts and slopes

# the next three lilnes of code simply combine the terms, above
mu <- c(a, b)

cov_ab <- sigma_a * sigma_b * rho
sigma <- matrix(c(sigma_a^2, cov_ab,
                  cov_ab, sigma_b^2), ncol = 2)
```

Covariance matrix is commonly mentioned as $\Sigma$
$$
\begin{align}
\begin{bmatrix}
\sigma_\alpha^2 & \sigma_\alpha\sigma_\beta\rho \\
\sigma_\alpha\sigma_\beta\rho & \sigma_\beta^2 \\
\end{bmatrix}
\end{align}
$$
```{r}
matrix(1:4, nrow = 2, ncol = 2)
```


```{r}
library(tidyverse)

sigmas <- c(sigma_a, sigma_b) # standard deviation
rho <- matrix(c(1, rho,
                rho, 1), nrow = 2)

# now matrix multiply to get covariance matrix
sigma <- diag(sigmas) %*% rho %*% diag(sigmas)

# how many cafes would you like?
n_cafes <- 20

set.seed(5) # used to replicate example

vary_effects <-
  MASS::mvrnorm(n_cafes, mu, sigma) %>% 
  data.frame() %>% 
  set_names("a_cafe", "b_cafe")

head(vary_effects)
```

```{r}
library(dutchmasters)

dutchmasters$pearl_earring
```

```{r}
scales::show_col(dutchmasters$pearl_earring)
```

```{r}
theme_pearl_earring <- function(light_color = "#E8DCCF", 
                                dark_color = "#100F14", 
                                my_family = "Courier",
                                ...) {
  
  theme(line = element_line(color = light_color),
        text = element_text(color = light_color, family = my_family),
        axis.line = element_blank(),
        axis.text = element_text(color = light_color),
        axis.ticks = element_line(color = light_color),
        legend.background = element_rect(fill = dark_color, color = "transparent"),
        legend.key = element_rect(fill = dark_color, color = "transparent"),
        panel.background = element_rect(fill = dark_color, color = light_color),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = dark_color, color = dark_color),
        strip.background = element_rect(fill = dark_color, color = "transparent"),
        strip.text = element_text(color = light_color, family = my_family),
        ...)
  
}

# now set `theme_pearl_earring()` as the default theme
theme_set(theme_pearl_earring())
```

```{r}
vary_effects %>% 
  ggplot(aes(x = a_cafe, y = b_cafe)) +
  geom_point(color = "#80a0c7") +
  geom_rug(color = "#8b9daf", linewidth = 1/7)
```

```{r}
cor(vary_effects$a_cafe, vary_effects$b_cafe)
```

### 14.1.2 Simulate observations
```{r}
n_visits <- 10
sigma <- 0.5 # std dev within cafes

set.seed(22) # used to replicate example

d <-
  vary_effects %>% 
  mutate(cafe = 1:n_cafes) %>% 
  expand_grid(visit = 1:n_visits) %>% 
  mutate(afternoon = rep(0:1, times = n() / 2)) %>% 
  mutate(mu = a_cafe + b_cafe * afternoon) %>% 
  mutate(wait = rnorm(n = n(), mean = mu, sd = sigma)) %>% 
  select(cafe, everything())
```

```{r}
d %>% glimpse()
```
```{r}
d %>% 
  mutate(afternoon = ifelse(afternoon == 0, "M", "A")) %>% 
  mutate(day = rep(rep(1:5, each = 2), times = n_cafes)) %>% 
  filter(cafe %in% c(3, 5)) %>% 
  mutate(cafe = str_c("cafe #", cafe)) %>% 
  
  ggplot(aes(x = visit, y = wait, group = day)) +
  geom_point(aes(color = afternoon), size = 2) +
  geom_line(color = "#8b9daf") +
  scale_color_manual(values = c("#80a0c7", "#eeda9d")) +
  scale_x_continuous(NULL, breaks = 1:10,
                     labels = rep(c("M", "A"), times = 5)) +
  scale_y_continuous("wait time in minutes", limits = c(0, NA)) +
  theme_pearl_earring(axis.ticks.x = element_blank(),
                      legend.position = "none") +
  facet_wrap(~ cafe, ncol = 1)
```
### 14.1.3 The varying slopes model.
$$
\begin{align}
\text{wait}_i & \sim \text{Normal}(\mu_i, \sigma) \\
\mu_i & = \alpha_{\text{cafe[i]}} + \beta_{\text{cafe[i]}}\text{afternoon}_i \\
\begin{bmatrix}
\alpha_{\text{cafe}} \\
\beta_{\text{cafe}}
\end{bmatrix} & \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
\alpha \\
\beta
\end{bmatrix},
\Sigma
\end{pmatrix} \\

\Sigma & = 
\begin{bmatrix}
\sigma_\alpha & 0 \\
0 & \sigma_\beta
\end{bmatrix}
\text{R}
\begin{bmatrix}
\sigma_\alpha & 0 \\
0 & \sigma_\beta
\end{bmatrix} \\

\alpha & \sim \text{Normal}(5, 2) \\
\beta & \sim \text{Normal}(-1, 0.5) \\
\sigma & \sim \text{Exponential}(1) \\
\sigma_\alpha & \sim \text{Exponential}(1) \\
\sigma_\beta & \sim \text{Exponential}(1) \\
\text{R} & \sim \text{LKJcorr}(2),
\end{align}
$$

where
$$
\text{R} = 
\begin{bmatrix}
1 & \rho \\
\rho & 1
\end{bmatrix}.
$$

```{r}
library(rethinking)

n_sim <- 1e4

set.seed(14)
r_1 <- 
  rlkjcorr(n_sim, K = 2, eta = 1) %>% 
  data.frame()

set.seed(14)
r_2 <-
  rlkjcorr(n_sim, K = 2, eta = 2) %>% 
  data.frame()

set.seed(14)
r_4 <-
  rlkjcorr(n_sim, K = 2, eta = 4) %>% 
  data.frame()
```

```{r}
# for annotation
text <-
  tibble(x = c(.83, .625, .45),
         y = c(.56, .75, 1.07),
         label = c("eta = 1", "eta = 2", "eta = 4"))

# plot
r_1 %>% 
  ggplot(aes(x = X2)) +
  geom_density(color = "transparent", fill = "#394165",
               alpha = 2/3, adjust = 1/2) +
  geom_density(data = r_2,
               color = "transparent", fill = "#DCA258", alpha = 2/3, adjust = 1/2) +
  geom_density(data = r_4,
               color = "transparent", fill = "#FCF9F0", alpha = 2/3, adjust = 1/2) +
  geom_text(data = text,
            aes(x = x, y = y, label = label),
            color = "#A65141", family = "Courier") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(LKJcorr(eta)),
       x = "correlation")
```

```{r}
library(tidybayes)

crossing(k = 2:5,
         eta = 1:4) %>% 
  mutate(prior = str_c("lkjcorr_marginal(", k, ", ", eta, ")"),
         strip = str_c("K==", k)) %>% 
  parse_dist(prior) %>% 
  
  ggplot(aes(y = eta, dist = .dist, args = .args)) +
  stat_dist_halfeye(.width = c(.5, .95),
                    color = "#fcf9f0", fill = "#a65141") +
  scale_x_continuous(expression(rho), limits = c(-1, 1),
                     breaks = c(-1, -.5, 0, .5, 1),
                     labels = c("-1", "-.5", "0", ".5", "1")) +
  scale_y_continuous(expression(eta), breaks = 1:4) +
  ggtitle(expression("Marginal correlation for the LKJ prior relative to K and "*eta)) +
  facet_wrap(~ strip, labeller = label_parsed, ncol = 4)
```

```{r}
detach(package:rethinking, unload = T)
library(brms)
```

```{r}
b14.1 <-
  brm(data = d,
      family = gaussian,
      wait ~ 1 + afternoon + (1 + afternoon | cafe),
      prior = c(prior(normal(5, 2), class = Intercept),
                prior(normal(-1, 0.5), class = b),
                prior(exponential(1), class = sd),
                prior(exponential(1), class = sigma),
                prior(lkj(2), class = cor)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 867530,
      file = "fits/b14.01")
```

```{r}
post <- as_draws_df(b14.1)

post %>% 
  ggplot() +
  geom_density(data = r_2, aes(x = X2),
               color = "transparent", fill = "#eeda9d",
               alpha = 3/4) +
  geom_density(aes(x = cor_cafe__Intercept__afternoon),
               color = "transparent", fill = "#a65141",
               alpha = 9/10) +
  annotate(geom = "text",
           x = c(-0.15, 0), y = c(2.21, 0.85),
           label = c("posterior", "prior"),
           color = c("#a65141", "#eeda9d"), family = "Courier") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "Correlation between intercepts\nand slopes, prior and posterior",
  x = "correlation")
```

```{r}
coef(b14.1)
```

Extract element from the coef() list
```{r}
partially_pooled_params <-
  # with this line we select each of the 20 cafe's posterior 
  # mean (i.e., Estimate) for both `Intercept` and `afternoon`
  coef(b14.1)$cafe[, 1, 1:2] %>% 
  data.frame() %>% # convert the two vectors to a data frame
  rename(Slope = afternoon) %>% 
  mutate(cafe = 1:nrow(.)) %>% # add the `cafe` index
  select(cafe, everything()) # simply moving `cafe` to the leftmost position
```

```{r}
# compute unpooled estimates directly from data
un_pooled_params <-
  d %>% 
  # with these two lines, we compute the mean value for each
  # cafe's wait time in the morning and then the afternoon
  group_by(afternoon, cafe) %>% 
  summarise(mean = mean(wait)) %>% 
  ungroup() %>% # ungrouping allows us to alter afternoon, one of the grouping variables
  mutate(afternoon = ifelse(afternoon == 0, "Intercept", "Slope")) %>% 
  spread(key = afternoon, value = mean) %>% # use `spread()` just as in the previous block
  mutate(Slope = Slope - Intercept) # finally, here's our slope!

# here we combine the partially-pooled and unpooled means into # a single data object, which will make plotting easier
params <-
  # `bind_rows()` will stack the second tibble below the first
  bind_rows(partially_pooled_params, un_pooled_params) %>% 
  # index whether the estimates are pooled
  mutate(pooled = rep(c("partially", "not"), each = nrow(.)/2))

# here's a glimpse at what we've been working for
params %>% 
  slice(c(1:5, 36:40))
```

```{r}
p1 <-
  ggplot(data = params, aes(x = Intercept, y = Slope)) +
  stat_ellipse(geom = "polygon", type = "norm", level = 1/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 2/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 3/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 4/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 5/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 6/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 7/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 8/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = 9/10, linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  stat_ellipse(geom = "polygon", type = "norm", level = .99,  linewidth = 0, alpha = 1/20, fill = "#E7CDC2") +
  geom_point(aes(group = cafe, color = pooled)) +
  geom_line(aes(group = cafe), linewidth = 1/4) +
  scale_color_manual("Pooled?", values = c("#80A0C7", "#A65141")) +
  coord_cartesian(xlim = range(params$Intercept),
                  ylim = range(params$Slope))
p1
```

Prep for figure 14.5.b
```{r}
# retrieve the partially-pooled estimates with coef()
partially_pooled_estimates <-
  coef(b14.1)$cafe[, 1, 1:2] %>% 
  # convert the two vectors to a data frame
  data.frame() %>% 
  # the Intercept is the wait time for morning (i.e., afternoon == 0)
  rename(morning = Intercept) %>% 
  # afternoon wait time is the morning wait time plus the afternoon slope
  mutate(afternoon = morning + afternoon,
         cafe = 1:n()) %>% # add the cafe index
  select(cafe, everything())

# compute unpooled estimates directly from data
un_pooled_estimates <-
  d %>% 
  # as above, with these two lines, we compute each cafe's mean wait value by time of day
  group_by(afternoon, cafe) %>% 
  summarise(mean = mean(wait)) %>% 
  # ungrouping allows us to alter the grouping variable, afternoon
  ungroup() %>% 
  mutate(afternoon = ifelse(afternoon == 0, "morning", "afternoon")) %>% 
  # this separates out the values into morning and afternoon columns
  spread(key = afternoon, value = mean)

estimates <-
  bind_rows(partially_pooled_estimates, un_pooled_estimates) %>% 
  mutate(pooled = rep(c("partially", "not"), each = n() / 2))
```

```{r}
p2 <-
  ggplot(data = estimates, aes(x = morning, y = afternoon)) +
  # nesting stat_ellipse() within mapply() is a less redundant
  # way to produce ten-layered semitransparent ellipses we dit
  # with ten lines of stat_ellipse() functions in the previous
  # plot
  mapply(function(level){
    stat_ellipse(geom = "polygon", type = "norm",
                 linewidth = 0, alpha = 1/20, fill = "#e7cdc2",
                 level = level)
  },
  level = c(1:9 / 10, .99)) +
  geom_point(aes(group = cafe, color = pooled)) +
  geom_line(aes(group = cafe), linewidth = 1/4) +
  scale_color_manual("Pooled?", values = c("#80A0C7", "#A65141")) +
  labs(x = "morning wait (mins)",
       y = "afternoon wait (mins)") +
  coord_cartesian(xlim = range(estimates$morning),
                  ylim = range(estimates$afternoon))
```

```{r}
library(patchwork)

(p1 + theme(legend.position = "none")) + 
  p2 + 
  plot_annotation(title = "Shrinkage in two dimensions")
```

## 14.2. Advanced varying slopes
```{r}
data(chimpanzees, package = "rethinking")
d <- chimpanzees
rm(chimpanzees)

# wrangle
d <-
  d %>% 
  mutate(actor = factor(actor),
         block = factor(block),
         treatment = factor(1 + prosoc_left + 2 * condition),
         # this will come in handy, later
         labels = factor(treatment,
                         levels = 1:4, 
                         labels = c("r/n", "l/n", "r/p", "l/p")))

glimpse(d)
```
non-centered statistical model
$$
\begin{align}
\text{left_pull}_i & \sim \text{Binomial}(n_i = 1, p_i) \\
\text{logit}(p_i) & = \gamma_{\text{treatment}[i]} + \alpha_{\text{actor[i], treatment[i]}} + \beta_{\text{block[i], treatment[i]}} \\
\gamma_j & \sim \text{Normal}(0, 1), \hspace{2em} \text{for} \ j = 1, ..., 4 \\
\begin{bmatrix}
\alpha_{j, 1} \\ \alpha_{j, 2} \\ \alpha_{j, 3} \\\alpha_{j, 4}
\end{bmatrix} & \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0 \\ 0 \\ 0 \\ 0
\end{bmatrix},
\Sigma_{\text{actor}}
\end{pmatrix} \\

\begin{bmatrix}
\beta_{j, 1} \\ \beta_{j, 2} \\ \beta_{j, 3} \\\beta_{j, 4}
\end{bmatrix} & \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0 \\ 0 \\ 0 \\ 0
\end{bmatrix},
\Sigma_{\text{block}}
\end{pmatrix} \\
\Sigma_{\text{actor}} & = \text{S}_\alpha \text{R}_\alpha \text{S}_\alpha \\
\Sigma_{\text{block}} & = \text{S}_\beta \text{R}_\beta \text{S}_\beta \\
\sigma_{\alpha,[1]},...,\sigma_{\alpha,[4]} & \sim \text{Exponential}(1) \\
\sigma_{\beta,[1]},...,\sigma_{\beta,[4]} & \sim \text{Exponential}(1) \\
\text{R}_\alpha & \sim \text{LKJ}(2) \\
\text{R}_\beta & \sim \text{LKJ}(2) \\
\end{align}
$$
$\text{S}_\alpha$ is a 4 x 4 matrix,
$$
\text{S}_\alpha = 
\begin{bmatrix}
\sigma_{\alpha,[1]} & 0 & 0 & 0 \\
0 & \sigma_{\alpha,[2]} & 0 & 0 \\
0 & 0 & \sigma_{\alpha,[3]} & 0 \\
0 & 0 & 0 & \sigma_{\alpha,[4]}
\end{bmatrix}
$$

In a similar way, $\text{R}_\alpha$ is a 4 x 4 matrix,
$$
\text{R}_\alpha = 
\begin{bmatrix}
1 & \rho_{\alpha,[1,2]} & \rho_{\alpha,[1,3]} & \rho_{\alpha,[1,4]} \\
\rho_{\alpha,[2,1]} & 1 & \rho_{\alpha,[2,3]} & \rho_{\alpha,[2,4]} \\
\rho_{\alpha,[3,1]} & \rho_{\alpha,[3,2]} & 1 & \rho_{\alpha,[3,4]} \\
\rho_{\alpha,[4,1]} & \rho_{\alpha,[4,2]} & \rho_{\alpha,[4,3]} & 1 \\
\end{bmatrix}
$$

```{r}
b14.3 <-
  brm(data = d,
      family = binomial,
      pulled_left | trials(1) ~ 0 + treatment + 
        (0 + treatment | actor) + (0 + treatment | block),
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd, group = actor),
                prior(exponential(1), class = sd, group = block),
                prior(lkj(2), class = cor, group = actor),
                prior(lkj(2), class = cor, group = block)),
      iter = 4000, warmup = 1000, chains = 4, cores = 4,
      seed = 4387510,
      file = "fits/b14.03"
      )
```

```{r}
library(bayesplot)

# give the parameters fancy names
names <-
  c(str_c("treatment[", 1:4, "]"),
    str_c("sigma['actor[", 1:4, "]']"),
    str_c("sigma['block[", 1:4, "]']"),
    str_c("rho['actor:treatment[", c(1, 1:2, 1:3), ",",
          rep(2:4, times = 1:3), "]']"),
    str_c("rho['block:treatment[", c(1, 1:2, 1:3), ",",
          rep(2:4, times = 1:3), "]']"),
    "Chain")

# wrangle
as_draws_df(b14.3) %>% 
  select(b_treatment1:`cor_block__treatment3__treatment4`, .chain) %>% 
  set_names(names) %>% 
  
  # plot
  mcmc_rank_overlay() +
  scale_color_manual(values = c("#80a0c7", "#b1934a",
                                "#a65141", "#eeda9d")) +
  scale_x_continuous(NULL, breaks = 0:4 * 1e3,
                     labels = c(0, str_c(1:4, "K"))) +
  coord_cartesian(ylim = c(30, NA)) +
  ggtitle("McElreath would love this trank plot.") +
  theme(legend.position = "bottom") +
  facet_wrap(~ parameter, labeller = label_parsed, ncol = 4)
```

```{r}
library(posterior)

as_draws_df(b14.3) %>% 
  summarise_draws() %>% 
  pivot_longer(starts_with("ess")) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(binwidth = 250, fill = "#eeda9d", color = "#dca258") +
  xlim(0, NA) +
  facet_wrap(~ name)
```

```{r}
print(b14.3)
```

```{r}
b14.3 <- add_criterion(b14.3, "waic")

waic(b14.3)
```

```{r}
# for annotation
text <-
  distinct(d, labels) %>% 
  mutate(actor = 1,
         prop = c(.07, .8, .08, .795))

nd <-
  d %>% 
  distinct(actor, condition, labels, prosoc_left, treatment) %>% 
  mutate(block = 5)

# compute and wrangle the posterior predictions
nd %>% 
  add_epred_draws(b14.3) %>% 
  mean_hdci(.epred) %>% 
  left_join(
    d %>% 
      group_by(actor, treatment) %>% 
      mutate(proportion = mean(pulled_left)) %>%
      distinct(actor, treatment, proportion),
    by = c("actor", "treatment")
  ) %>% 
  mutate(condition = factor(condition)) %>% 

# plot!
ggplot(aes(x = labels)) +
  geom_hline(yintercept = 0.5, color = "#e8dccf", 
             alpha = 1/2, linetype = 2) +
  
  # empirical proportions
  geom_line(aes(y = proportion, group = prosoc_left),
            linewidth = 1/4, color = "#394165") +
  geom_point(aes(y = proportion, shape = condition), 
             color = "#394165", fill = "#100f14",
             size = 2.5, show.legend = F) +
  
  # posterior predictions
  geom_line(aes(y = .epred, group = prosoc_left),
            linewidth = 3/4, color = "#80a0c7") +
  geom_pointrange(aes(y = .epred, ymin = .lower, 
                      ymax = .upper, shape = condition),
                  color = "#80a0c7", fill = "#100f14",
                  fatten = 8, linewidth = 1/3,
                  show.legend = F) +
  
  # anotation for the conditions
  geom_text(data = text,
            aes(y = prop, label = labels),
            color = "#dca258", family = "Courier", size = 3) +
  scale_shape_manual(values = c(21, 19)) +
  scale_x_discrete(NULL, breaks = NULL) +
  scale_y_continuous("proportion left lever", breaks = 0:2 / 2,
                     labels = c("0", ".5", "1")) +
  labs(subtitle = "Posterior predictions, in light blue, against the raw data, in dark blue, for\nmodel b14.3, the cross-classified varying effects model.") +
  facet_wrap(~ actor, nrow = 1, labeller = label_both)
```

## 14.3 Instruments and causal designs
### 14.3.1 Instrumental variables.
```{r}
library(ggdag)

dag_coords <-
  tibble(name = c("E", "U", "W"),
         x = c(1, 2, 3),
         y = c(1, 2, 1))
```

```{r}
theme_pearl_dag <- function(...) {
  
  theme_pearl_earring() +
    theme_dag() + 
    theme(panel.background = element_rect(fill = "#100F14"),
          ...)
  
}

dagify(E ~ U,
       W ~ E + U,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 shape = 21, stroke = 2, fill = "#FCF9F0", size = 6, show.legend = F) +
  geom_dag_text(color = "#100F14", family = "Courier") +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c("#EEDA9D", "#A65141")) +
  theme_pearl_dag()
```

Instrumental variables will solve some of the difficulties we have in not being able to condition on $U$. Here we’ll call our instrumental variable $Q$. In the terms of the present example, the instrumental variable has the qualities that
- $Q$ is independent of $U$,
- $Q$ is not independent of $E$, and
- $Q$ can only influence $W$ through $E$ (i.e., the effect of $Q$ on $W$ is fully mediated by $E$).
There is what this looks like in a DAG.
```{r}
dag_coords <-
  tibble(name = c("Q", "E", "U", "W"),
         x    = c(0, 1, 2, 3),
         y    = c(2, 1, 2, 1))


dagify(E ~ Q + U,
       W ~ E + U,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 shape = 21, stroke = 2, fill = "#FCF9F0", size = 6, show.legend = F) +
  geom_dag_text(color = "#100F14", family = "Courier") +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c("#EEDA9D", "#A65141")) +
  theme_pearl_dag()
```

Sadly, our condition that $Q$ can only influence $W$ through $E$-often called the exclusion restriction-generally cannot be tested. Given $U$ is unmeasured, by definition, we also cannot test that $Q$ is independent of $U$. These are model assumptions.

Let's simulate data to get a sense of how this works.
```{r}
# make a standardizing function
standardize <- function(x){
  (x - mean(x)) / sd(x)
}

# simulate
set.seed(73)

n <- 500

dat_sim <-
  tibble(u_sim = rnorm(n, mean = 0, sd = 1),
         q_sim = sample(1:4, size = n, replace = T)) %>% 
  mutate(e_sim = rnorm(n, mean = u_sim + q_sim, sd = 1)) %>% 
  mutate(w_sim = rnorm(n, mean = u_sim + 0 * e_sim, sd = 1)) %>% 
  mutate(w = standardize(w_sim),
         e = standardize(e_sim),
         q = standardize(q_sim))

dat_sim
```

First we fit the univariable model corresponding to $E \rightarrow W$.
```{r}
b14.4 <-
  brm(data = dat_sim,
      family = gaussian,
      w ~ 1 + e,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.04")
```

```{r}
print(b14.4)
```

Because we have not conditioned on $U$, then model suggests a moderately large spurious causal relation for $E \rightarrow W$. Now see what happens when we also condition directly on $Q$, as in $Q \rightarrow W \leftarrow E$.
```{r}
b14.5 <-
  brm(data = dat_sim,
      family = gaussian,
      w ~ 1 + e + q,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.05")
```

```{r}
print(b14.5)
```

That's a mess because this model suggests both $E$ and $Q$ have moderate to strong causal effects on $W$, even though we know neither do based on the true data-generating model.

The statistical solution to this mess is to express the data-generating DAG as a multivariate statistical model following the form
$$
\begin{align}
\begin{bmatrix}
W_i \\
E_i
\end{bmatrix} & \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
\mu_{W,i} \\
\mu_{E,i}
\end{bmatrix}, \Sigma
\end{pmatrix} \\
\mu_{W, i} & = \alpha_W + \beta_{EW}E_i \\
\mu_{E, i} & = \alpha_E + \beta_{QE}Q_i \\
\Sigma & = 
\begin{bmatrix}
\sigma_W & 0 \\
0 & \sigma_E
\end{bmatrix} \text{R}
\begin{bmatrix}
\sigma_W & 0 \\
0 & \sigma_E
\end{bmatrix} \\
\text{R} & = \begin{bmatrix}
1 & \rho \\
\rho & 1
\end{bmatrix} \\
\alpha_W \ \text{and} \ \alpha_E & \sim \text{Normal}(0, 0.2) \\
\beta_{EW} \ \text{and} \ \beta_{QE} & \sim \text{Normal}(0, 0.5) \\
\sigma_W \ \text{and} \ \sigma_E & \sim \text{Exponential}(1) \\
\rho & \sim \text{LKJ}(2).
\end{align}
$$

```{r}
e_model <- bf(e ~ 1 + q)
w_model <- bf(w ~ 1 + e)

b14.6 <-
  brm(data = dat_sim,
      family = gaussian,
      e_model + w_model + set_rescor(TRUE),
      prior = c(# E model
                prior(normal(0, .2), class = Intercept, resp = e),
                prior(normal(0, .5), class = b, resp = e),
                prior(exponential(1), class = sigma, resp = e),
                
                # W model
                prior(normal(0, .2), class = Intercept, resp = w),
                prior(normal(0, .5), class = b, resp = w),
                prior(exponential(1), class = sigma, resp = w),
                
                # rho
                prior(lkj(2), class = rescor)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.06")
```

```{r}
print(b14.6)
```

Adjust the simulation and try other scenarios
```{r}
set.seed(73)

n <- 500

dat_sim <-
  tibble(u_sim = rnorm(n, mean = 0, sd = 1),
         q_sim = sample(1:4, size = n, replace = T)) %>% 
  mutate(e_sim = rnorm(n, mean = u_sim + q_sim, sd = 1)) %>% 
  mutate(w_sim = rnorm(n, mean = -u_sim + 0.2 * e_sim, sd = 1)) %>% 
  mutate(w = standardize(w_sim),
         e = standardize(e_sim),
         q = standardize(q_sim))

dat_sim
```

```{r}
b14.4x <-
  update(b14.4,
         newdata = dat_sim,
         iter = 2000, warmup = 1000, chains = 4, cores = 4,
         seed = 14,
         file = "fits/b14.04x")

b14.6x <-
  update(b14.6,
         newdata = dat_sim,
         iter = 2000, warmup = 1000, chains = 4, cores = 4,  
         seed = 14, 
         file = "fits/b14.06x")
```

```{r}
text <-
  tibble(Estimate = c(fixef(b14.4x)[2, 3], fixef(b14.6x)[4, 4]),
         y = c(4.35, 3.65),
         hjust = c(0, 1),
         fit = c("b14.4x", "b14.6x"))

bind_rows(
  # b_14.4x
  posterior_summary(b14.4x)[1:3, ] %>% 
    data.frame() %>% 
    mutate(param = c("alpha[W]", "beta[EW]", "sigma[W]"),
           fit = "b14.4x"),
  #b_14.6x
  posterior_summary(b14.6x)[c(1:6,9), ] %>%  
    data.frame() %>% 
    mutate(param = c("alpha[E]", "alpha[W]", "beta[QE]", "beta[EW]", "sigma[E]", "sigma[W]", "rho"),
           fit = "b14.6x")) %>% 
  mutate(param = factor(
    param,
    levels = c("rho", "sigma[W]", "sigma[E]", "beta[EW]",
               "beta[QE]", "alpha[W]", "alpha[E]"))) %>% 
  
  ggplot(aes(x = param, y = Estimate, color = fit)) +
  geom_hline(yintercept = 0, color = "#e8dccf", alpha = 1/4) +
  geom_pointrange(aes(ymin = Q2.5, ymax = Q97.5), fatten = 2,
                  position = position_dodge(width = 0.5)) +
  geom_text(data = text,
            aes(x = y, label = fit, hjust = hjust)) +
  scale_color_manual(NULL, values = c("#E7CDC2", "#A65141")) +
  scale_x_discrete(NULL, labels = ggplot2:::parse_safe) +
  ylab("marginal posterior") +
  coord_flip() +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

With the help from b14.6x, we found that E and W have a negative correlation in their residual variance, because the confound positively influences one and negatively influences the other.

One can use dagitty() and instrumentalVariables() functions from the dagitty package to first define a DAG and then query whether there are instrumental variables for a given exposure and outcome.
```{r}
library(dagitty)

dagIV <- dagitty("dag{Q -> E <- U -> W <- E}")

instrumentalVariables(dagIV, exposure = "E", outcome = "W")
```

### 14.3.2 Other designs.
```{r}
dag_coords <-
  tibble(name = c("X", "Z", "U", "Y"),
         x    = c(1, 2, 2, 3),
         y    = c(1, 1, 2, 1))

dagify(X ~ U,
       Z ~ X,
       Y ~ U + Z,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 shape = 21, stroke = 2, fill = "#FCF9F0", size = 6, show.legend = F) +
  geom_dag_text(color = "#100F14", family = "Courier") +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c("#EEDA9D", "#A65141")) +
  theme_pearl_dag()
```

## 14.4 Social relations as correlated varying effects
```{r}
library(rethinking)
data(KosterLeckie)
```

```{r}
kl_dyads %>% glimpse()
```

The variables hidA and hidB tell us the household IDs in each dyad, and did is a unique dyad ID number. To get a sense of the interrlation among those three ID variables, we'll make a tile plot
```{r}
kl_dyads %>% 
  ggplot(aes(x = hidA, y = hidB, label = did)) +
  geom_tile(aes(fill = did),
            show.legend = F) +
  geom_text(size = 2.25, family = "Courier") +
  geom_vline(xintercept = 0:24 + 0.5, color = "#394165",
             linewidth = 1/5) +
  geom_vline(xintercept = 1:25 + 0.5, color = "#394165",
             linewidth = 1/5) +
  scale_fill_gradient(low = "#dca258", high = "#eeda9d",
                      limits = c(1, NA)) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_continuous(breaks = 2:25) +
  theme(axis.text = element_text(size = 9),
        axis.ticks = element_blank())
```

Here is our version of Figure 14.8, the bivariate distribution of dyadic gifts, collapsing across dyads.
```{r}
kl_dyads %>% 
  ggplot(aes(x = giftsAB, y = giftsBA)) +
  geom_hex(bins = 70) +
  geom_abline(color = "#dca258", linetype = 3) +
  scale_fill_gradient(low = "#e7cdc2", high = "#a65141",
                      limits = c(1, NA)) +
  scale_x_continuous("gifts household A to household B", 
                     limits = c(0, 113)) +
  scale_y_continuous("gifts from B to A", limits = c(0, 113)) +
  ggtitle("Distribution of dyadic gifts") +
  coord_equal()
```

Here's the overall Pearson's correlation coefficient, collapsing across grouping levels
```{r}
kl_dyads %>% 
  summarise(rho = cor(giftsAB, giftsBA))
```

However, it would be a mistake to take this correlation seriously. It is a disentangled mixture of various kinds of associations, none of which are guaranteed to be even close to r = .24. Remember this as we move along with the analyses and let the consequences burn a methodological mark into your soul.
```{r}
kl_data <-
  list(
    N = nrow(kl_dyads),
    N_households = max(kl_dyads$hidB),
    did = kl_dyads$did,
    hidA = kl_dyads$hidA,
    hidB = kl_dyads$hidB,
    giftsAB = kl_dyads$giftsAB,
    giftsBA = kl_dyads$giftsBA
  )

m14.7 <-
  ulam(
    alist(
      giftsAB ~ poisson(lambdaAB),
      giftsBA ~ poisson(lambdaBA),
      log(lambdaAB) <- a + gr[hidA, 1] + gr[hidB, 2] + d[did, 1],
      log(lambdaBA) <- a + gr[hidB, 1] + gr[hidA, 2] + d[did, 2],
      a ~ normal(0, 1),
      
      ## gr matrix of varying effects
      vector[2]:gr[N_households] ~ multi_normal(0, Rho_gr, sigma_gr),
      Rho_gr ~ lkj_corr(4),
      sigma_gr ~ exponential(1),
      
      ## dyad effects
      transpars> matrix[N, 2]:d <-
        compose_noncentered(rep_vector(sigma_d, 2), L_Rho_d, z),
      matrix[2, N]:z ~ normal(0, 1),
      cholesky_factor_corr[2]:L_Rho_d ~ lkj_corr_cholesky(8),
      sigma_d ~ exponential(1),
      
      ## compute correlation matrix for dyads
      gq> matrix[2, 2]:Rho_d <<- Chol_to_Corr(L_Rho_d)
    ),
    data = kl_data,
    chains = 4, cores = 4,iter = 2000
  )
```

```{r}
precis(m14.7)
```

```{r}
precis(m14.7, depth = 3, pars = c("Rho_gr", "sigma_gr"))
```

Let's view $\sigma_g$, $\sigma_r$, and their correlation in a plot.
```{r}
post <- extract.samples(m14.7)

tibble("sigma[italic(g)]" = post$sigma_gr[, 1],
       "sigma[italic(r)]" = post$sigma_gr[, 2],
       "rho[italic(g)][italic(r)]" = post$Rho_gr[, 2, 1]) %>% 
  pivot_longer(everything()) %>% 
  
  ggplot(aes(x = value, y = name, fill = name)) +
  geom_vline(xintercept = 0, color = "#fcf9f0", alpha = 1/3) +
  stat_halfeye(.width = .89, color = "#fcf9f0", height = 1.5) +
  scale_fill_manual(values = c("#80A0C7", "#EEDA9D", "#DCA258")) +
  scale_y_discrete(NULL, labels = ggplot2:::parse_safe) +
  xlab("marginal posterior") +
  coord_cartesian(ylim = c(1.5, 3.9)) +
  theme(legend.position = "none")
```

“This implies that individuals who give more across all dyads tend to receive less[, and ] clear evidence that rates of giving are more variable than rates of receiving” (p. 465). McElreath suggested we “try plot(exp(g[,1]),exp(r[,1])) for example to show the posterior distribution of giving/receiving for household number 1” (p. 465). Here’s a tidyverse version of that plot.

```{r}
g <- sapply(1:25, function(i) post$a + post$gr[, i, 1])
r <- sapply(1:25, function(i) post$a + post$gr[, i, 2])

tibble(g = exp(g[, 1]),
       r = exp(r[, 1])) %>% 
  ggplot(aes(x = g, y = r)) +
  geom_abline(color = "#fcf9f0", linetype = 2, alpha = 1/3) +
  geom_point(color = "#b1934a", alpha = 1/3, size = 1/4) +
  stat_ellipse(type = "norm", level = .5, linewidth = 1/2,
               color = "#80a0c7") +
  stat_ellipse(type = "norm", level = .9, linewidth = 1/2,
               color = "#80a0c7") +
  labs(x = expression(giving[italic(i) == 1]),
       y = expression(receiving[italic(i) == 1])) +
  coord_equal(xlim = c(0, 5),
              ylim = c(0, 5))
```
```{r}
rbind(exp(g), exp(r)) %>% 
  data.frame() %>% 
  set_names(1:25) %>% 
  mutate(parameter = rep(c("g", "r"), each = n() / 2),
         iter = rep(1:4000, times = 2)) %>% 
  pivot_longer(-c(parameter, iter), names_to = "household") %>% 
  pivot_wider(names_from = parameter, values_from = value) %>%
  group_by(household) %>% 
  mutate(mu_g = mean(g),
         mu_r = mean(r)) %>% 
  nest(data = c("g", "r", "iter")) %>% 
  
  ggplot(aes(group = household)) +
  geom_abline(color = "#fcf9f0", linetype = 2, alpha = 1/3) +
  stat_ellipse(data = . %>% unnest(data),
               aes(x = g, y = r), type = "norm", level = .5,
               linewidth = 1/2, alpha = 1/2, 
               color = "#80a0c7") +
  geom_point(aes(x = mu_g, y = mu_r), color = "#dca258") +
  labs(x = "generalized giving", y = "generalized receiving") +
  coord_equal(xlim = c(0, 8.5), ylim = c(0, 8.5))
```

Here is a look at the covariance matrix for the dyadic effects, the posterior summaries for the part of the model McElreath defined in the middle of the page 462 as,
$$
\begin{bmatrix}
d_{ij} \\ d_{ji}
\end{bmatrix} \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0 \\0
\end{bmatrix},
\begin{bmatrix}
\sigma_d^2 & \sigma_d^2 \rho_d \\
\sigma_d^2 \rho_d & \sigma_d^2
\end{bmatrix}
\end{pmatrix},
$$

```{r}
precis(m14.7, depth = 3, pars = c("Rho_d", "sigma_d"))
```

Here they are in a plot
```{r}
tibble("sigma[italic(d)]" = post$sigma_d,
       "rho[italic(d)]" = post$Rho_d[, 2, 1]) %>% 
  pivot_longer(everything()) %>% 
  
  ggplot(aes(x = value, y = name, fill = name)) +
  geom_vline(xintercept = 0, color = "#fcf9f0", alpha = 1/3) +
  stat_halfeye(.width = .89, color = "#fcf9f0") +
  scale_fill_manual(values = c("#a65141", "#b1934a")) +
  scale_y_discrete(NULL, labels = ggplot2:::parse_safe) +
  xlab("marginal posterior") +
  coord_cartesian(ylim = c(1.5, 2)) +
  theme(legend.position = "none")
```

“The correlation here is positive and strong. And there is more variation among dyads than there is among household in giving rates” (p. 467). Now make the right hand panel of Figure 14.9.
```{r}
tibble(dy1 = apply(post$d[, , 1], 2, mean),
       dy2 = apply(post$d[, , 2], 2, mean)) %>% 
  
  ggplot(aes(x = dy1, y = dy2)) +
  geom_abline(color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_vline(xintercept = 0, color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_hline(yintercept = 0, color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_point(color = "#8B9DAF", alpha = 1/2, size = 1/2) +
  geom_text(x = mean(post$d[, 1, 1]),
            y = mean(post$d[, 1, 2]),
            label = "1",
            color = "#EEDA9D", family = "Courier") +
  labs(x = "household A in dyad",
       y = "household B in dyad") +
  coord_equal(xlim = c(-2, 3.5),
              ylim = c(-2, 3.5))
```

As McElreath pointed out, each dot is the posterior mean for one of the 300 levels of did. Do you see the yellow #1 toward the middle? It’s marking off the posterior mean for did == 1. To give a better sense of the uncertainty in each of the levels of did, here is the full bivariate distribution for did == 1.
```{r}
tibble(dy1 = post$d[, 1, 1],
       dy2 = post$d[, 1, 2]) %>% 
  
  ggplot(aes(x = dy1, y = dy2)) +
  geom_abline(color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_vline(xintercept = 0, color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_hline(yintercept = 0, color = "#FCF9F0", linetype = 2, alpha = 1/3) +
  geom_point(color = "#8B9DAF", alpha = 1/3, size = 1/4) +
  stat_ellipse(type = "norm", level = .5, linewidth = 1/2, color = "#EEDA9D") +
  stat_ellipse(type = "norm", level = .9, linewidth = 1/2, color = "#EEDA9D") +
  labs(x = expression("household A in dyad"[italic(i)==1]),
       y = expression("household B in dyad"[italic(i)==1])) +
  coord_equal(xlim = c(-2, 3.5),
              ylim = c(-2, 3.5))
```

## 14.5 Continuous categories and the Gaussian process
### 14.5.1 Example: Spatial autocorrelation in Oceanic tools.
```{r}
# load the distance matrix
library(rethinking)
data("islandsDistMatrix")

# display (measured in thousands of km)
d_mat <- islandsDistMatrix
colnames(d_mat) <- c("Ml", "Ti", "SC", "Ya", "Fi", "Tr", "Ch", "Mn", "To", "Ha")

round(d_mat, 1)
```

```{r}
d_mat %>% 
  data.frame() %>% 
  rownames_to_column("row") %>% 
  gather(column, distance, -row) %>% 
  mutate(column = factor(column, levels = colnames(d_mat)),
         row = factor(row, levels = rownames(d_mat)) %>% fct_rev(),
         label = formatC(distance, format = 'f', digits = 2)) %>% 
  
  ggplot(aes(x = column, y = row)) +
  geom_raster(aes(fill = distance)) +
  geom_text(aes(label = label),
            size = 3, family = "Courier", color = "#100F14") +
  scale_fill_gradient(low = "#FCF9F0", high = "#A65141") +
  scale_x_discrete(NULL, position = "top", expand = c(0, 0)) +
  scale_y_discrete(NULL, expand = c(0, 0)) +
  theme_pearl_earring(axis.text.y = element_text(hjust = 0)) +
  theme(axis.ticks = element_blank())
```

```{r}
tibble(x = seq(from = 0, to = 4, by = .01),
       linear = exp(-1 * x),
       squared = exp(-1 * x^2)) %>% 
  
  ggplot(aes(x = x)) +
  geom_line(aes(y = linear),
            color = "#b1934a", linetype = "dashed") +
  geom_line(aes(y = squared), color = "#dca258") +
  scale_x_continuous("distance", expand = c(0, 0)) +
  scale_y_continuous("correlation", breaks = c(0, .5, 1),
                     labels = c(0, ".5", 1))
```

```{r}
data(Kline2)

d <-
  Kline2 %>% 
  mutate(society = 1:10)

rm(Kline2)

d %>% glimpse()
```

Convert lat and lon2 from decimal degrees into thousands of km
```{r}
d <-
  d %>% 
  mutate(lat_adj = lat * 0.11132,
         lon2_adj = lon2 * 0.11132)

d %>% 
  select(culture, lat, lon2, lat_adj:lon2_adj)
```

```{r}
b14.8 <-
  brm(data = d,
      family = poisson(link = "identity"),
      bf(total_tools ~ exp(a) * population^b / g,
         a ~ 1 + gp(lat_adj, lon2_adj, scale = FALSE),
         b + g ~ 1,
         nl = TRUE),
      prior = c(prior(normal(0, 1), nlpar = a),
                prior(exponential(1), nlpar = b, lb = 0),
                prior(exponential(1), nlpar = g, lb = 0),
                prior(inv_gamma(2.874624, 2.941204),
                      class = lscale, 
                      coef = gplat_adjlon2_adj, nlpar = a),
                prior(exponential(1), class = sdgp, 
                      coef = gplat_adjlon2_adj, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      sample_prior = T,
      file = "fits/b14.08")
```

```{r}
print(b14.8)
```

```{r}
posterior_summary(b14.8)[1:15, ] %>% round(digits = 2)
```
Let’s focus on our three non-linear parameters, first. Happily, both our b_b_Intercept and b_g_Intercept summaries look a lot like those for McElreath’s b and g, respectively. Our b_a_Intercept might look distressingly small, but that’s just because of how we parameterized our model. It’s actually very close to McElreath’s a after you exponentiate.
```{r}
fixef(b14.8, probs = c(.055, .945))["a_Intercept", c(1, 3:4)] %>% 
  exp() %>% round(digits = 2)
```

Our Gaussian process parameters are different from McElreath’s. From the gp section of the brms reference manual (Bürkner, 2022i), we learn the brms parameterization follows the form
$$
k(x_i, x_j) = sdgp^2 \text{exp}(-||x_i - -x_j||^2/(2lscale^2)),
$$

where $k(x_i, x_j)$ is the same as McElreath's $\text{K}_{ij}$ and $||x_i - x_j||^2$ is the Euclidean distance, the same as McElreath's $D_{ij}^2$. Thus we could also express the brms parametrization as
$$
\text{K}_{ij} = sdgp^2 \text{exp}(-D_{ij}^2/(2lscale^2)), 
$$
which is much close to McElreath's

$$
\text{K}_{ij} = \eta^2 \text{exp}(-\rho^2D_{ij}^2)+\delta_{ij}\sigma^2
$$

On page 470, McElreath explained that the final $\delta_{ij}\sigma^2$ term is mute with the Oceanic societies data. Thus we won't consider it further. This reduces McElreath's equation to
$$
\text{K}_{ij} = \eta^2 \text{exp}(-\rho^2D_{ij}^2)
$$

Importantly, what McElreath called $\eta$, Burkner called $sdgp$. Whkle McElreath estimated $\eta^2$, brms simply estimated $sdgp$. So we'll have to square our ```sdgp(a_gplat_adjlon2_adj)``` before it's on the same scale as ```etasq``` in the text. Here it is.
```{r}
post <-
  as_draws_df(b14.8) %>% 
  mutate(etasq = sdgp_a_gplat_adjlon2_adj^2)

post %>% 
  mean_hdci(etasq, .width = .89) %>% 
  mutate_if(is.double, round, digits = 3)
```
Though our posterior is a little bit larger than McElreath’s, we’re in the ballpark. You may have noticed that in our model brm() code, above, we just went with the flow and kept the exponential(1) prior on sdgp. The brms default would have been student_t(3, 0, 15.6).

Now look at the denominator of the inner part of Bürkner’s equation, $2lscale^2$. This appears to be the brms equivalent to McElreath’s $\rho^2$. Or at least it’s what we’ve got. Anyway, also note that McElreath estimated $\rho^2$ directly as ```rhosq```. If I’m doing the algebra correctly, we might expect
$$
\begin{align}
\rho^2 & = 1/(2 \cdot lscale^2) \hspace{2em} \text{and} \ \text{thus} \\
lscale & = \sqrt{1 / (2 \cdot\rho^2)}.
\end{align}
$$

To get a sense of this relation, it might be helpful to plot
```{r}
p1 <-
  tibble("rho^2" = seq(from = 0, to = 11, by = .01)) %>% 
  mutate(lscale = sqrt(1 / (2 * `rho^2`))) %>% 
  
  ggplot(aes(x = `rho^2`, y = lscale)) +
  geom_hline(yintercept = 0, color = "#FCF9F0", linewidth = 1/4, linetype = 2) +
  geom_vline(xintercept = 0, color = "#FCF9F0", linewidth = 1/4, linetype = 2) +
  geom_line(color = "#A65141") +
  xlab(expression(rho^2)) +
  coord_cartesian(xlim = c(0, 10),
                  ylim = c(0, 10))

p2 <-
  tibble(lscale = seq(from = 0, to = 11, by = .01)) %>% 
  mutate(`rho^2` = 1 / (2 * lscale^2)) %>% 
  
  ggplot(aes(x = lscale, y = `rho^2`)) +
  geom_hline(yintercept = 0, color = "#FCF9F0", linewidth = 1/4, linetype = 2) +
  geom_vline(xintercept = 0, color = "#FCF9F0", linewidth = 1/4, linetype = 2) +
  geom_line(color = "#80A0C7") +
  ylab(expression(rho^2)) +
  coord_cartesian(xlim = c(0, 10),
                  ylim = c(0, 10))

p1 + p2
```

The two aren’t quite inverses of one another, but the overall pattern is when one is large, the other is small. Now we have a sense of how they compare and how to covert one to the other, let’s see how our posterior for $lscale$ looks when we convert it to the scale of McElreath’s $\rho^2$.

```{r}
post <-
  post %>% 
  mutate(rhosq = 1 / (2 * lscale_a_gplat_adjlon2_adj^2))

post %>% 
  mean_hdci(rhosq, .width = .89) %>% 
  mutate_if(is.double, round, digits = 3)
```
This is about a third of the size of the McElreath’s $\rho^2 = 1.31, 89\% \ \text{HDI}\ [0.08, 4.41]$. The plot deepens. If you look back, you’ll see we used a very different prior for $lscale$. Here it is: ```inv_gamma(2.874624, 2.941204)```. Use ```get_prior()``` to discover where that came from.
```{r}
get_prior(data = d,
          family = poisson(link = "identity"),
          bf(total_tools ~ exp(a) * population^b / g,
             a ~ 1 + gp(lat_adj, lon2_adj, scale = FALSE),
             b + g ~ 1,
             nl = TRUE))
```

That is, we used the brms default prior for $lscale$. In a GitHub exchange, Bürkner pointed out that brms uses special priors for $lscale$ parameters based on Michael Betancourt’s (2017) vignette, Robust Gaussian processes in Stan. We can use the dinvgamma() function from the well-named invgamma package (Kahle & Stamey, 2017) to get a sense of what that prior looks like.
```{r}
tibble(lscale = seq(from = .01, to = 9, by = .01)) %>% 
  mutate(density = invgamma::dinvgamma(lscale, shape = 2.874624, rate = 2.941204)) %>% 
  
  ggplot(aes(x = lscale, y = density)) +
  geom_area(fill = "#80A0C7") +
  annotate(geom = "text", 
           x = 4.75, y = 0.75,
           label = "inverse gamma(2.874624, 2.941204)",
           color = "#8B9DAF", family = "Courier") +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 8))
```

Anyways, let’s make the subplots for our version of Figure 14.11 to get a sense of what this all means. Start with the left panel, the prior predictive distribution for the covariance.
```{r}
# for slice_sample()
set.seed(14)

# wrangle
p1 <-
  prior_draws(b14.8) %>% 
  transmute(draw = 1:n(),
            etasq = sdgp_a^2,
            rhosq = 1 / (2 * lscale_a__1_gplat_adjlon2_adj^2)) %>% 
  slice_sample(n = 100) %>% 
  expand_grid(x = seq(from = 0, to = 10, by = .05)) %>% 
  mutate(covariance = etasq * exp(-rhosq * x^2)) %>% 
  
  # plot
  ggplot(aes(x = x, y = covariance)) +
  geom_line(aes(group = draw),
            linewidth = 1/4, alpha = 1/4, color = "#eeda9d") +
  scale_x_continuous("distance (thousand km)", expand = c(0, 0),
                     breaks = 0:5 * 2) +
  coord_cartesian(xlim = c(0, 10),
                  ylim = c(0, 2)) +
  labs(subtitle = "Gaussian process prior")
```

```{r}
# for `slice_sample()`
set.seed(14)

# wrangle
p2 <-
  post %>% 
  mutate(etasq = sdgp_a_gplat_adjlon2_adj^2,
         rhosq = 1 / (2 * lscale_a_gplat_adjlon2_adj^2)) %>% 
  slice_sample(n = 50) %>% 
  expand_grid(x = seq(from = 0, to = 10, by = .05)) %>% 
  mutate(covariance = etasq * exp(-rhosq * x^2)) %>% 
  
  # plot
  ggplot(aes(x = x, y = covariance)) +
  geom_line(aes(group = .draw),
            linewidth = 1/4, alpha = 1/4, color = "#EEDA9D") +
  stat_function(fun = function(x) mean(post$sdgp_a_gplat_adjlon2_adj)^2 *
                  exp(-(1 / (2 * mean(post$lscale_a_gplat_adjlon2_adj)^2)) * x^2),
                color = "#DCA258", linewidth = 1) +
  scale_x_continuous("distance (thousand km)", expand = c(0, 0),
                     breaks = 0:5 * 2) +
  coord_cartesian(xlim = c(0, 10),
                  ylim = c(0, 2)) +
  labs(subtitle = "Gaussian process posterior")
```

```{r}
p1 + p2
```

Though the Gaussian process parameters from our brms parameterization loked different from McElreath's they resulted in a simimlar decline in spatial covariance. 

Let's finish ths up and push the parameters back through the function for K, the covariance matrix
```{r}
# compute posterior median covariance among societies
k <- matrix(0, nrow = 10, ncol = 10)
for (i in 1:10)
  for(j in 1:10)
    k[i, j] <- median(post$etasq) * exp(-median(post$rhosq) * islandsDistMatrix[i, j]^2)

diag(k) <- median(post$etasq) + 0.01

k %>% round(2)
```

```{r}
# convert to correlation matrix
rho <- round(cov2cor(k), 2)

# add row/col names for convenience
colnames(rho) <- c("Ml", "Ti", "SC", "Ya", "Fi", "Tr", "Ch", "Mn", "To", "Ha")
rownames(rho) <- colnames(rho)

rho %>% round(2)
```

```{r}
rho %>%
  data.frame() %>% 
  mutate(row = d$culture) %>% 
  pivot_longer(-row, values_to = "distance") %>% 
  mutate(column = factor(name, levels = colnames(d_mat)),
         row    = factor(row, levels = rownames(d_mat)) %>% fct_rev(),
         label  = formatC(distance, format = 'f', digits = 2) %>% str_replace(., "0.", ".")) %>%
  # omit this line to keep the diagonal of 1's
  filter(distance != 1) %>% 
  
  ggplot(aes(x = column, y = row)) + 
  geom_raster(aes(fill = distance)) + 
  geom_text(aes(label = label),
            size = 2.75, family = "Courier", color = "#100F14") +
  scale_fill_gradient(expression(rho), low = "#FCF9F0", high = "#A65141", limits = c(0, 1)) +
  scale_x_discrete(NULL, position = "top", expand = c(0, 0)) +
  scale_y_discrete(NULL, expand = c(0, 0)) +
  theme_pearl_earring(axis.text.y = element_text(hjust = 0)) +
  theme(axis.ticks = element_blank())
```
```{r}
d %>% 
  transmute(psize = logpop / max(logpop)) %>% 
  transmute(psize = exp(psize * 1.5) - 2)
```

```{r}
tidy_rho <-
  rho %>% 
  data.frame() %>% 
  rownames_to_column() %>% 
  bind_cols(d %>% select(culture, logpop, total_tools, lon2, lat)) %>% 
  pivot_longer(Ml:Ha,
               names_to = "colname",
               values_to = "correlation") %>% 
  mutate(group = str_c(pmin(rowname, colname), pmax(rowname, colname))) %>% 
  select(rowname, colname, group, culture, everything())

head(tidy_rho)
```

```{r}
library(ggrepel)

p1 <-
  tidy_rho %>% 
  ggplot(aes(x = lon2, y = lat)) +
  geom_point(data = d,
             aes(size = logpop), color = "#dca258") +
  geom_line(aes(group = group, alpha = correlation^2),
            color = "#eeda9d") +
  geom_text_repel(data = d, aes(label = culture),
                  seed = 14, point.padding = .2, size = 2.75,
                  color = "#fcf9f0", family = "Courier") +
  scale_alpha_continuous(range = c(0, 1)) +
  labs(subtitle = "Among societies in geographic space\n",
       x = "longitude", y = "latitude") +
  coord_cartesian(xlim = range(d$lon2),
                  ylim = range(d$lat)) +
  theme(legend.position = "none")
```

```{r}
# compute the average posterior predictive relationship
# between log population and total tools, summarized by the
# median and 80% interval
f <-
  post %>% 
  expand_grid(logpop = seq(from = 6, to = 14, 
                           length.out = 30)) %>% 
  mutate(population = exp(logpop)) %>% 
  mutate(lambda = exp(b_a_Intercept) * 
           population^b_b_Intercept / b_g_Intercept) %>% 
  group_by(logpop) %>% 
  median_qi(lambda, .width = .8)

# plot
p2 <-
  tidy_rho %>% 
  ggplot(aes(x = logpop)) +
  geom_smooth(data = f,
              aes(y = lambda, ymin = .lower, ymax = .upper),
              stat = "identity", fill = "#394165",
              color = "#100f14", alpha = .5, linewidth = 1.1) +
  geom_point(data = d, aes(y = total_tools, size = logpop),
             color = "#dca258") +
  geom_line(aes(y = total_tools, group = group, 
                alpha = correlation^2), color = "#eeda9d") +
  geom_text_repel(data = d, aes(y = total_tools, label = culture), seed = 14, point.padding = .2, size = 2.75,
                  color = "#fcf9f0", family = "Courier") +
  scale_alpha_continuous(range = c(0, 1)) +
  labs(subtitle = "Shown against the relation between\ntotal tools and log pop",
       x = "log population", y = "total tools") +
  coord_cartesian(xlim = range(d$logpop), 
                  ylim = range(d$total_tools)) +
  theme(legend.position = "none")
```

```{r}
p1 + p2 +
  plot_annotation(title = "Posterior median correlations")
```

##### 14.5.1.0.1 Rethinking: Dispersion by other names.
```{r}
b14.8_nb <-
  brm(data = d,
      family = negbinomial(link = "identity"),
      bf(total_tools ~ exp(a) * population^b / g,
         a ~ 1 + gp(lat_adj, lon2_adj, scale = FALSE),
         b + g ~ 1,
         nl = TRUE),
      prior = c(prior(normal(0, 1), nlpar = a),
                prior(exponential(1), nlpar = b, lb = 0),
                prior(exponential(1), nlpar = g, lb = 0),
                prior(inv_gamma(2.874624, 2.941204), 
                      class = lscale, 
                      coef = gplat_adjlon2_adj, nlpar = a),
                prior(exponential(1), class = sdgp,
                      coef = gplat_adjlon2_adj, nlpar = a),
                # default prior
                prior(gamma(0.01, 0.01), class = shape)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      control = list(adapt_delta = .85),
      file = "fits/b14.08_nb")
```

```{r}
print(b14.8_nb)
```

This resulted in a slightly less pronounced correlation matrix
```{r}
post <- as_draws_df(b14.8_nb)

# compute posterior median covariance among societies
k <- matrix(0, nrow = 10, ncol = 10)
for (i in 1:10)
    for (j in 1:10)
        k[i, j] <- median(post$sdgp_a_gplat_adjlon2_adj^2) * 
  exp(-islandsDistMatrix[i, j]^2 / (2 * median(post$lscale_a_gplat_adjlon2_adj)^2))

diag(k) <- median(post$sdgp_a_gplat_adjlon2_adj^2) + 0.01

# convert to correlation matrix
rho <- round(cov2cor(k), 2)

# add row/col names for convenience
colnames(rho) <- c("Ml", "Ti", "SC", "Ya", "Fi", "Tr", "Ch", "Mn", "To", "Ha")
rownames(rho) <- colnames(rho)

rho %>% round(2)
```

```{r}
rho %>%
  data.frame() %>% 
  mutate(row = d$culture) %>% 
  pivot_longer(-row, values_to = "distance") %>% 
  mutate(column = factor(name, levels = colnames(d_mat)),
         row    = factor(row, levels = rownames(d_mat)) %>% fct_rev(),
         label  = formatC(distance, format = 'f', digits = 2) %>% str_replace(., "0.", ".")) %>%
  # omit this line to keep the diagonal of 1's
  filter(distance != 1) %>% 
  
  ggplot(aes(x = column, y = row)) + 
  geom_raster(aes(fill = distance)) + 
  geom_text(aes(label = label),
            size = 2.75, family = "Courier", color = "#100F14") +
  scale_fill_gradient(expression(rho), low = "#FCF9F0", high = "#A65141", limits = c(0, 1)) +
  scale_x_discrete(NULL, position = "top", expand = c(0, 0)) +
  scale_y_discrete(NULL, expand = c(0, 0)) +
  theme_pearl_earring(axis.text.y = element_text(hjust = 0)) +
  theme(axis.ticks = element_blank())
```

```{r}
# tidy up rho
tidy_rho <-
  rho %>%
  data.frame() %>% 
  rownames_to_column() %>% 
  bind_cols(d %>% select(culture, logpop, total_tools, lon2, lat)) %>% 
  pivot_longer(Ml:Ha,
               names_to = "colname", 
               values_to = "correlation") %>%
  mutate(group = str_c(pmin(rowname, colname), pmax(rowname, colname))) 

# left panel
p1 <-
  tidy_rho %>%       
  ggplot(aes(x = lon2, y = lat)) +
  geom_point(data = d, 
             aes(size = logpop), color = "#DCA258") +
  geom_line(aes(group = group, alpha = correlation^2),
            color = "#EEDA9D") +
  geom_text_repel(data = d, aes(label = culture), 
                  seed = 14, point.padding = .2, size = 2.75, color = "#FCF9F0", family = "Courier") +
  scale_alpha_continuous(range = c(0, 1)) +
  labs(subtitle = "Among societies in geographic space\n",
       x = "longitude",
       y = "latitude") +
  coord_cartesian(xlim = range(d$lon2),
                  ylim = range(d$lat)) +
  theme(legend.position = "none")

# compute the average posterior predictive relationship between 
# log population and total tools, summarized by the median and 80% interval
f <-
  post %>% 
  expand_grid(logpop = seq(from = 6, to = 14, length.out = 30)) %>%
  mutate(population = exp(logpop)) %>% 
  mutate(lambda = exp(b_a_Intercept) * population^b_b_Intercept / b_g_Intercept) %>%
  group_by(logpop) %>% 
  median_qi(lambda, .width = .8)

# right panel
p2 <-
  tidy_rho %>% 
  ggplot(aes(x = logpop)) +
  geom_smooth(data = f,
              aes(y = lambda, ymin = .lower, ymax = .upper),
              stat = "identity",
              fill = "#394165", color = "#100F14", alpha = .5, linewidth = 1.1) +
  geom_point(data = d, 
             aes(y = total_tools, size = logpop), 
             color = "#DCA258") +
  geom_line(aes(y = total_tools, group = group, alpha = correlation^2),
            color = "#EEDA9D") +
  geom_text_repel(data = d, 
                  aes(y = total_tools, label = culture), 
                  seed = 14, point.padding = .2, size = 2.75, color = "#FCF9F0", family = "Courier") +
  scale_alpha_continuous(range = c(0, 1)) +
  labs(subtitle = "Shown against the relation between\ntotal tools and log pop",
       x = "log population",
       y = "total tools") +
  coord_cartesian(xlim = range(d$logpop),
                  ylim = range(d$total_tools)) +
  theme(legend.position = "none")

# combine
p1 + p2 + 
  plot_annotation(title = "Posterior median correlations based on the gamma-Poisson model") 
```

### 14.5.2 Example: Phylogenetic distance.
```{r}
dag_coords <-
  tibble(name = c("G1", "B1", "U1", "G2", "B2", "U2"),
         x    = rep(1:2, each = 3),
         y    = rep(3:1, times = 2))

dagify(G2 ~ G1 + U1,
       B2 ~ G1 + B1 + U1,
       U2 ~ U1,
       coords = dag_coords) %>%
  tidy_dagitty() %>% 
  mutate(color = ifelse(name %in% c("G1", "B1"), "a",
                        ifelse(name %in% c("G2", "B2"), "b", "c"))) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = color),
                 shape = 21, stroke = 2, fill = "#FCF9F0", size = 7, show.legend = F) +
  geom_dag_text(color = "#100F14", family = "Courier", parse = T,
                label = c(expression(B[1]), expression(G[1]), expression(U[1]), 
                          expression(B[2]), expression(G[2]), expression(U[2]))) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c("#80A0C7", "#EEDA9D", "#A65141")) +
  theme_pearl_dag()
```

```{r}
dag_coords <-
  tibble(name = c("G", "U", "M", "P", "B"),
         x    = c(0, 1, 1, 2, 2),
         y    = c(3, 1, 2, 1, 3))

dagify(G ~ U + M,
       U ~ P,
       M ~ U,
       B ~ G + M + U,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "U"),
                 shape = 21, stroke = 2, fill = "#FCF9F0", size = 6, show.legend = F) +
  geom_dag_text(color = "#100F14", family = "Courier", parse = T) +
  geom_dag_edges(edge_colour = "#FCF9F0") +
  scale_color_manual(values = c("#EEDA9D", "#A65141")) +
  theme_pearl_dag()
```

```{r}
data(Primates301, package = "rethinking")
data(Primates301_nex)
```

```{r}
#devtools::install_github("GuangchuangYu/ggtree")
library(ggtree)

Primates301_nex %>% 
  ggtree(layout = "circular", color = "#394165", size = 1/4) +
  geom_tiplab(size = 5/3, color = "#100f14")
```

```{r}
d <-
  Primates301 %>% 
  mutate(name = as.character(name)) %>% 
  drop_na(group_size, body, brain) %>% 
  mutate(m = log(body) %>% standardize(),
         b = log(brain) %>% standardize(),
         g = log(group_size) %>% standardize())

glimpse(d)
```

```{r}
b14.9 <-
  brm(data = d,
      family = gaussian,
      b ~ 1 + m + g,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.09")
```

```{r}
print(b14.9)
```

```{r}
as_draws_df(b14.9) %>% 
  mutate(sigma_sq = sigma^2) %>% 
  mean_qi(sigma_sq) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
library(ape)

spp_obs <- d$name

tree_trimmed <- keep.tip(Primates301_nex, spp_obs)
Rbm <- corBrownian(phy = tree_trimmed)

V = vcv(Rbm)
Dmat <- cophenetic(tree_trimmed)
```

```{r}
full_join(
  Dmat %>% 
    as_tibble(rownames = "row") %>% 
    pivot_longer(-row,
                 names_to = "col",
                 values_to = "distance"),
  V %>% 
    as_tibble(rownames = "row") %>% 
    pivot_longer(-row,
                 names_to = "col",
                 values_to = "covariance"),
  by = c("row", "col")
) %>% 
  
  ggplot(aes(x = distance, y = covariance)) +
  geom_point(color = "#80A0C7", alpha = 1/10) +
  labs(subtitle = "These variables are the\ninverse of one another.",
       x = "phylogenetic distance", 
       y = "covariance")
```

```{r}
# headmap of Dmat
p1 <-
  Dmat %>% 
  as_tibble(rownames = "row") %>% 
  pivot_longer(-row,
               names_to = "col",
               values_to = "distance") %>%
  
  ggplot(aes(x = col, y = row, fill = distance)) +
  geom_tile() +
  scale_fill_gradient(low = "#100F14", high = "#EEDA9D") +
  scale_x_discrete(NULL, breaks = NULL) +
  scale_y_discrete(NULL, breaks = NULL) +
  theme(legend.position = "top")

# headmap of V
p2 <-
  V %>% 
  as_tibble(rownames = "row") %>% 
  pivot_longer(-row,
               names_to = "col",
               values_to = "covariance") %>% 
  
  ggplot(aes(x = col, y = row, fill = covariance)) +
  geom_tile() +
  scale_fill_gradient(low = "#100F14", high = "#EEDA9D") +
  scale_x_discrete(NULL, breaks = NULL) +
  scale_y_discrete(NULL, breaks = NULL) +
  theme(legend.position = "top")

# combine
(p1 | p2) + plot_annotation(subtitle = "Again, distance is the inverse of covariance.")
```

```{r}
R <- V[spp_obs, spp_obs] / max(V)

b14.10 <-
  brm(data = d,
      data2 = list(R = R),
      family = gaussian,
      b ~ 1 + m + g + fcor(R),
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.10")
```

```{r}
print(b14.10)
```

```{r}
as_draws_df(b14.10) %>% 
  mutate(sigma_sq = sigma^2) %>% 
  mean_hdi(sigma_sq, .width = .89) %>% 
  mutate_if(is.double, round, 2)
```

```{r}
dat_list <-
  list(
    N_spp = nrow(d),
    M     = standardize(log(d$body)),
    B     = standardize(log(d$brain)),
    G     = standardize(log(d$group_size)), Imat = diag(nrow(d)),
    V     = V[spp_obs, spp_obs],
    R     = V[spp_obs, spp_obs] / max(V[spp_obs, spp_obs]),
    Dmat  = Dmat[spp_obs, spp_obs] / max(Dmat))
```

```{r}
m14.11 <-
  ulam(
    alist(
      B ~ multi_normal(mu, SIGMA),
      mu <- a + bM * M + bG * G,
      matrix[N_spp, N_spp]: SIGMA <- cov_GPL1(Dmat, etasq, rhosq, 0.01),
      a ~ normal(0, 1),
      c(bM, bG) ~ normal(0, 0.5),
      etasq ~ half_normal(1, 0.25),
      rhosq ~ half_normal(3, 0.25)
    ),
    data = dat_list,
    chains = 4, cores = 4)
```

```{r}
precis(m14.11)
```

```{r}
post <- extract.samples(m14.11)

set.seed(14)

left_join(
  # posterior
  post %>% 
    data.frame() %>% 
    slice_sample(n = 30) %>% 
    mutate(draw = 1:n()) %>% 
    select(draw, etasq, rhosq) %>% 
    expand_grid(d_seq = seq(from = 0, to = 1, length.out = 50)),
  
  # prior
  tibble(eta = abs(rnorm(1e5, mean = 1, sd = 0.25)),
         rho = abs(rnorm(1e5, mean = 3, sd = 0.25))) %>% 
    expand_grid(d_seq = seq(from = 0, to = 1, length.out = 50)) %>% 
    mutate(k = eta * exp(-rho * d_seq)) %>% 
    group_by(d_seq) %>% 
    mean_hdi(k, .width = .89),
  # join them
  by = "d_seq") %>% 
  
  # plot!
  ggplot(aes(x = d_seq)) +
  geom_line(aes(y = etasq * exp(-rhosq * d_seq), group = draw),
            color = "#80A0C7", alpha = 1/2) +
  geom_lineribbon(data = . %>% filter(draw == 1),
            aes(y = k, ymin = .lower, ymax = .upper),
            color = "#A65141", fill = "#E7CDC2") +
  annotate(geom = "text",
           x = c(0.2, 0.5), y = c(0.1, 0.5),
           label = c("posterior", "prior"),
           color = c("#80A0C7", "#E7CDC2"),
           family = "Courier") +
  labs(x = "phylogenetic distance", 
       y = "covariance") +
  ylim(0, 1.5)
```

## 14.6 Bonus: Multilevel growth models and the MELSM
### 14.6.1 Borrow some data.
```{r}
# load the data from GitHub
load(url("https://github.com/ASKurz/Statistical_Rethinking_with_brms_ggplot2_and_the_tidyverse_2_ed/raw/master/data/m_melsm_dat.rda"))

# add a scaled time variable
dat <-
  dat %>% 
  mutate(day01 = (day - 2) / max((day - 2)))

# take a look
glimpse(dat)
```

```{r}
distinct(dat, record_id) %>% count()
```

```{r}
dat %>% 
  count(record_id) %>% 
  summarise(median = median(n),
            min = min(n),
            max = max(n))
```

```{r}
dat %>% 
  count(record_id) %>% 
  
  ggplot(aes(x = n)) +
  geom_bar(fill = "#b1934a") +
  scale_x_continuous("number of days", limits = c(0, NA)) +
  theme_pearl_earring()
```

```{r}
set.seed(14)

dat %>% 
  nest(data = c(P_A.std, day, P_A.lag, N_A.lag, steps.pm,
                steps.pmd, N_A.std, day01)) %>% 
  slice_sample(n = 16) %>% 
  unnest(data) %>% 
  
  ggplot(aes(x = day, y = N_A.lag)) +
  geom_line(color = "#80a0c7") +
  geom_point(color = "#fcf9f0", size = 1/2) +
  ylab("negative affect (standardized)") +
  facet_wrap(~ record_id)
```

### 14.6.2 Conventional multilevel growth model
In the social sciences, a typical way to analyze data like these is with a multilevel growth model in which participants vary in their intercepts (starting point) and time slopes (change over time). In the sample of the data, above, it looks like most participants have fairly constant levels of NA over time (i.e., near-zero slopes) but some (e.g., #128 and 147) show some evidence of systemic decreases in NA (i.e., negative slopes). There is also some variation in starting points, though most of the participants in this subset of the data seemed to have endorsed relatively low levels of NA both at baseline and throughout the study.

We want a model that can capture all these kinds of variation. Eventually, we will fit a model that accounts for both PA and NA. But to keep things simple while we’re warming up, we will restrict our focus to $\text{NA}$. If we let $\text{NA}_{ij}$ be the standardized NA score for the i-th participant on the  j-th day, our first Bayesian multilevel growth model will follow the form

$$
\begin{align}
\text{NA}_{ij} & \sim \text{Normal}(u_{ij}, \sigma) \\
\mu_{ij} & = \beta_0 + \beta_1\text{time}_{ij} + \mu_{0i} + \mu_{1i}\text{time}_{ij} \\
\sigma & = \sigma_\epsilon \\
\begin{bmatrix}
\mu_{0i} \\
\mu_{1i}
\end{bmatrix} & \sim
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0 \\ 0
\end{bmatrix},
\text{SRS}
\end{pmatrix} \\
\text{S} & = \begin{bmatrix}
\sigma_0 & 0 \\
0 & \sigma_1
\end{bmatrix} \\
\text{R} & = \begin{bmatrix}
1 & \rho_{12} \\
\rho_{21} & 1
\end{bmatrix} \\
\beta_0 & \sim \text{Normal}(0, 0.2) \\
\beta_1 & \sim \text{Normal}(0, 1) \\
\sigma_0\ \text{and}\ \sigma_1 & \sim \text{Exponential}(1) \\
\sigma_\epsilon & \sim \text{Exponential}(1) \\
\text{R} & \sim \text{LKJ}(2),
\end{align}
$$

```{r}
b14.12 <-
  brm(data = dat,
      family = gaussian,
      N_A.std ~ 1 + day01 + (1 + day01 | record_id),
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd),
                prior(exponential(1), class = sigma),
                prior(lkj(2), class = cor)),
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      seed = 14,
      file = "fits/b14.12")
```

```{r}
print(b14.12)
```

Hopefully it makes sense that the population-level intercept $(\beta_0)$ is near zero. It would be odd if it wasn’t given these are standardized data. The coefficient for day01 $(\beta_1)$ is mildly negative, suggesting an overall trend for participants to endorse lower NA scores over time.

The two group-level $\sigma$ parameters are fairly large given the scale of the data. They suggest participants varied quite a bit in terms of both intercepts and slopes. They also have a moderate negative correlation, suggesting that participants with higher intercepts tended to have more negative slopes.

To get a sense of the model, we’ll plot the posterior means for each participants’ fitted trajectory across time (thin lines), along with the population-average trajectory (thick line).

```{r}
nd <- dat %>% 
  distinct(record_id, day01)

fitted(b14.12,
       newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  ggplot(aes(x = day01, y = Estimate, group = record_id)) +
  geom_line(alpha = 1/3, linewidth = 1/3, color = "#8b9daf") +
  geom_segment(x = 0, xend = 1, y = fixef(b14.12)[1, 1],
               yend = fixef(b14.12)[1, 1] + fixef(b14.12)[2, 1], linewidth = 3, color = "#80a0c7") +
  scale_x_continuous(breaks = c(0, .5, 1)) +
  ylab("negative affect (standardized)")
```

If you look back up to the model summary from before the plot, the one parameter we didn’t focus on was the lone sigma parameter at the bottom. That’s our $\sigma_\epsilon$, which captures the variation in NA not accounted for by the intercepts, slopes, and their correlation. An important characteristic of the conventional multilevel growth model is that $\sigma_\epsilon$ does not vary across persons, occasions, or other variables. To give a sense of why this might not be the best assumption, let’s take a focused look at the model implied trajectories for two participants. Here we will take cues from some Figure 4.8 from way back in Section 4.4.3.5. We will plot the original data atop both the fitted lines and their 95% intervals, which expresses the mean structure, along with the 95% posterior predictive interval, which expresses the uncertainty of the $\sigma_\epsilon$ parameter.
```{r}
nd <-
  dat %>% 
  filter(record_id %in% c(30, 115)) %>% 
  select(record_id, N_A.std, day01)

bind_cols(
  # fitted
  fitted(b14.12, newdata = nd) %>% 
    data.frame(),
  # predict
  predict(b14.12, newdata = nd) %>% 
    data.frame() %>% 
    select(Q2.5:Q97.5) %>% 
    set_names("p_lower", "p_upper")
) %>% 
  bind_cols(nd) %>% 
  
  ggplot(aes(x = day01)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity", fill = "#8b9daf",
              color = "#8b9daf", alpha = 1/2, 
              linewidth = 1/2) +
  geom_ribbon(aes(ymin = p_lower, ymax = p_upper),
              fill = "#8b9daf", alpha = 1/2) +
  geom_point(aes(y = N_A.std), color = "#8b9daf") +
  scale_x_continuous(breaks = c(0, .5, 1)) +
  ylab("negative affect (standardized)") +
  facet_wrap(~ record_id)
```

Because of our fixed $\sigma_\epsilon$ parameter, the 95% posterior predictive interval is the same width for both participants. Yet look how closely participant the data points for participant 30 cluster not only within the center region of the posterior prediction intervals, but also almost completely within the 95% interval for the fitted line. In contrast, notice how much more spread out the data points for participant 115 are, and how many of them extend well beyond the posterior predictive interval. This difference in variability is ignored by conventional growth models. However, there’s no reason we can’t adjust our model to capture this kind of person-level variability, too. Enter the MELSM.

### 14.6.3 Learn more about your data with the MELSM
Mixed-effects location scale models (MELSMs) have their origins in the work of Donald Hedeker and colleagues (Hedeker et al., 2008, 2012). Rast et al. (2012) showcased an early application of the framework to the BUGS/JAGS software. More recently Philippe Rast and colleagues (particularly Donald Williams) have adapted this approach for use within the Stan/brms ecosystem (Donald R. Williams et al., 2019; Donald R. Williams et al., 2020, 2022; Donald R. Williams, Mulder, et al., 2021).

Within the brms framework, MELSMs apply a distributional modeling approach (see Bürkner, 2022a) to the multilevel growth model. Not only are parameters from the mean structure allowed to vary across groups, but parameters applied to $\sigma$ are allowed to vary across groups, too. Do you remember the little practice model b10.1 from Section 10.2.2? We simulated Gaussian data for two groups with the same mean parameter but different parameters for $\sigma$. If you check back in that section, you’ll see that the brms default was to model $\text{log}(\sigma)$ in that case. This is smart because when you define a model for $\sigma$, you want to use a link function that ensures the predictions will be stay at zero and above. The MELSM approach of Hedeker, Rast, Williams and friends applies this logic to $\sigma_\epsilon$ in multilevel growth models. However, not only can $\sigma_\epsilon$ vary across groups in a fixed-effects sort of way, we can use multilevel partial pooling, too.

To get a sense of what this looks like, we’ll augment our previous model, but this time allowing $\sigma_\epsilon$ to vary across participants. You might express the updated statistical model as
$$
\begin{align*}
NA_{ij} &\sim \text{Normal}(\mu_{ij}, \sigma_i) \\
\mu_{ij} &= \beta_0 + \beta_1 \text{time}_{ij} + u_{0i} + u_{1i} \text{time}_{ij} \\
\log(\sigma_i) &= \eta_0 + u_{2i} \\
\begin{bmatrix}
u_{0i} \\
u_{1i} \\
u_{2i}
\end{bmatrix} &\sim \text{MVNormal}\left(
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix},
\text{SRS}
\right) \\
\text{S} &= \begin{bmatrix}
\sigma_0 & 0 & 0 \\
0 & \sigma_1 & 0 \\
0 & 0 & \sigma_2
\end{bmatrix} \\
\text{R} &= \begin{bmatrix}
1 & \rho_{12} & \rho_{13} \\
\rho_{21} & 1 & \rho_{23} \\
\rho_{31} & \rho_{32} & 1
\end{bmatrix} \\
\beta_0 &\sim \text{Normal}(0, 0.2) \\
\beta_1 &\sim \text{Normal}(0, 1) \\
\eta_0 &\sim \text{Normal}(0, 1) \\
\sigma_0, \ldots, \sigma_2 &\sim \text{Exponential}(1) \\
R &\sim \text{LKJ}(2)
\end{align*}
$$

In the opening likelihood statement from the prior model, we simply set $NA_{ij} \sim \text{Normal}(\mu_{ij}, \sigma)$. For our first MELSM, we now refer to $\sigma_i$, meaning the levels of variation not accounted for by the mean structure can vary across participants (hence the $i$ subscript). Two lines down, we see the formula for $\log(\sigma_i)$ contains population-level intercept, $\eta_0$, and participant-specific deviations around that parameter, $u_{2i}$.

In the next three lines, the plot deepens. We see that all three participant-level deviations, $u_{0i}, \ldots, u_{2i}$, are multivariate normal with means set to zero and variation expressed in the parameters $\sigma_0, \ldots, \sigma_2$ of the $S$ matrix. In the $R$ matrix, we now have three correlation parameters, with $\rho_{31}$ and $\rho_{32}$ allowing us to assess the correlations among individual differences in variability and individual differences in starting points and change over time, respectively. Let’s fit the model.

```{r}
b14.13 <-
  brm(data = dat,
      family = gaussian,
      bf(N_A.std ~ 1 + day01 + (1 + day01 |i| record_id),
         sigma ~ 1 + (1 |i| record_id)),
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd),
                
                prior(normal(0, 1), class = Intercept, dpar = sigma),
                prior(exponential(1), class = sd, dpar = sigma),
                
                prior(lkj(2), class = cor)),
      iter = 3000, warmup = 1000, chains = 4, cores = 8,
      seed = 14,
      file = "fits/b14.13")
```
We should note a few things about the `brm()` syntax. First, because we modeled both $\mu_{ij}$ and $\sigma_i$, we nested both model formulas within the `bf()` function. Second, because the `brms` default is to use the log link when modeling $\sigma_i$, there was no need to explicitly set it that way in the `family` line. However, we could have if we wanted to. Third, notice our use of the `|i|` syntax within the parentheses in the formula lines. If we had used the conventional `|` syntax, that would have not allowed our $u_{2i}$ parameters to correlate with $u_{0i}$ and $u_{1i}$ from the mean structure. It would have effectively set $\rho_{31} = \rho_{32} = 0$. Finally, notice how within the `prior()` functions, we explicitly referred to those for the new $\sigma$ structure with the `dpar = sigma` operator.

Okay, time to check the model summary.

```{r}
print(b14.13)
```

The 'sigma_Intercept' lines in the 'Population-Level Effects' section is the summary for our $\eta_0$ parameter. To get a sense of what this means out of the log space, just exponentiate.
```{r}
fixef(b14.13)["sigma_Intercept", c(1, 3:4)] %>% exp()
```

To get a sense of the variation in that parameter across participants [i.e., $\text{exp}(\eta_0 + \mu_{2i})]$, it's best to plot.
```{r}
coef(b14.13)$record_id[, , "sigma_Intercept"] %>% 
  exp() %>% 
  data.frame() %>% 
  arrange(Estimate) %>% 
  mutate(rank = 1:n()) %>% 
  
  ggplot(aes(x = rank, y = Estimate, ymin = Q2.5, 
             ymax = Q97.5)) +
  geom_pointrange(linewidth = 0.4, fatten = 0.3,
                  color = "#eeda9d") +
  scale_x_continuous("participants ranked by posterior mean",
                     breaks = NULL) +
  ylab(expression(exp(eta[0] + italic(u)[2][italic(i)])))
```

Looks like there's a lot of variation in a parameter that was formerly fixed across participants as a single value $\sigma_\epsilon$. Here's what this looks like in terms of our posterior predictive distributions for participants 30 and 115, from before.
```{r}
nd <-
  dat %>% 
  filter(record_id %in% c(30, 115)) %>% 
  select(record_id, N_A.std, day01)

bind_cols(
  fitted(b14.13, newdata = nd) %>% 
    data.frame(),
  predict(b14.13, newdata = nd) %>% 
    data.frame() %>% 
    select(Q2.5:Q97.5) %>% 
    set_names("p_lower", "p_upper")
) %>% 
  bind_cols(nd) %>% 
  
  ggplot(aes(x = day01)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity", fill = "#8b9daf",
              color = "#8b9daf", alpha = 1/2,
              linewidth = 1/2) +
  geom_ribbon(aes(ymin = p_lower, ymax = p_upper),
              fill = "#8b9daf", alpha = 1/2) +
  geom_point(aes(y = N_A.std), color = "#8b9daf") +
  scale_x_continuous(breaks = c(0, .5, 1)) +
  ylab("negative affect (standardized)") +
  facet_wrap(~ record_id)
```

That's a big improvement. Let's expand our skill set. Within the MELSM paradigm, one can use multiple variables to model participant-specific variation. Here we'll add in our time variable.

$$
\begin{align*}
\text{NA}_{ij} & \sim \operatorname{Normal}(\mu_{ij}, \color{#A65141}{\sigma_{ij}}) \\
\mu_{ij} & = \beta_0 + \beta_1 \text{time}_{ij} + u_{0i} + u_{1i} \text{time}_{ij} \\
\log(\sigma_{i\color{#A65141}j}) & = \eta_0 + \color{#A65141}{\eta_1 \text{time}_{ij}} + u_{2i} + \color{#A65141}{u_{3i} \text{time}_{ij}} \\
\begin{bmatrix} u_{0i} \\ u_{1i} \\ u_{2i} \\ \color{#A65141}{u_{3i}} \end{bmatrix} & \sim \operatorname{MVNormal}\begin{pmatrix} \begin{bmatrix} 0 \\ 0 \\ 0 \\ 0 \end{bmatrix}, \mathbf S \mathbf R \mathbf S \end{pmatrix} \\
\mathbf S & = \begin{bmatrix} \sigma_0 & 0 & 0 & 0 \\ 0 & \sigma_1 & 0 & 0 \\ 0 & 0 & \sigma_2 & 0 \\ 0 & 0 & 0 & \sigma_3 \end{bmatrix} \\
\mathbf R & = \begin{bmatrix} 1 & \rho_{12} & \rho_{13} & \rho_{14} \\ \rho_{21} & 1 & \rho_{23} & \rho_{24} \\ \rho_{31} & \rho_{32} & 1 & \rho_{34} \\ \rho_{41} & \rho_{42} & \rho_{43} & 1 \end{bmatrix} \\
\beta_0   & \sim \operatorname{Normal}(0, 0.2) \\
\beta_1, \eta_0, \text{and } \eta_1 & \sim \operatorname{Normal}(0, 1) \\
\sigma_0,\dots, \sigma_3 & \sim \operatorname{Exponential}(1) \\
\mathbf R & \sim \operatorname{LKJ}(2).
\end{align*}
$$

Note how in the very first line we are not speaking in terms of $\sigma_{ij}$. Variation in the criterion $\text{NA}_{ij}$ not accounted for by the mean structrue can now vary across participants, $i$, and time, $j$. This results in four $\mu_{xi}$ terms, a $4 \times 4\ \text{S}$ matrix and $4 \times 4\ \text{R}$ matrix. Here's how you might fit the model with `brms::brm()`. Warning: it'll probably take a couple hours.

```{r}
b14.14 <-
  brm(data = dat,
      family = gaussian,
      bf(N_A.std ~ 1 + day01 + (1 + day01 |i| record_id),
         sigma ~ 1 + day01 + (1 + day01 |i| record_id)),
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(exponential(1), class = sd),
                
                prior(normal(0, 1), class = Intercept, dpar = sigma),
                prior(normal(0, 1), class = b, dpar = sigma),
                prior(exponential(1), class = sd, dpar = sigma),
                
                prior(lkj(2), class = cor)),
      iter = 3000, warmup = 1000, chains = 4, cores = 8,
      seed = 14,
      file = "fits/b14.14")
```

```{r}
print(b14.14)
```

Our new line for ‘sigma_day01’ suggests there is a general trend for less variation in negative affect ratings over time. However, the ‘sd(sigma_day01)’ line in the ‘Group-Level Effects’ section indicates even this varies a bit across participants. At this point, a lot of the action is now in the estimates for the R matrix. Here that is in a coefficient plot.
```{r}
posterior_summary(b14.14) %>% 
  data.frame() %>% 
  rownames_to_column("par") %>% 
  filter(str_detect(par, "cor_")) %>% 
  mutate(rho = str_c("(rho[", c(21, 31, 32, 41, 42, 43), "])")) %>% 
  mutate(par = str_c("'", par, "'~", rho)) %>% 
  
  ggplot(aes(x = Estimate, xmin = Q2.5, xmax = Q97.5, y = par)) +
  geom_vline(xintercept = c(-.5, 0, .5), linetype = c(2, 1, 2), linewidth = c(1/4, 1/2, 1/4), color = "#FCF9F0", alpha = 1/4) +
  geom_pointrange(color = "#B1934A") +
  scale_y_discrete(labels = ggplot2:::parse_safe) +
  xlim(-1, 1) +
  labs(x = "marginal posterior",
       y = NULL) +
  theme(axis.text.y = element_text(hjust = 0, size = 9),
        axis.ticks.y = element_blank())
```

```{r}
levels <- c("beta[0]", "beta[1]", "eta[0]", "eta[1]")

rho <-
  posterior_summary(b14.14) %>% 
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "cor_")) %>% 
  mutate(param = str_remove(param, "cor_record_id__")) %>% 
  separate(param, into = c("left", "right"), sep = "__") %>% 
  mutate(
    left = case_when(
      left == "Intercept" ~ "beta[0]",
      left == "day01" ~ "beta[1]",
      left == "sigma_Intercept" ~ "eta[0]"),
    right = case_when(
      right == "day01" ~ "beta[1]",
      right == "sigma_Intercept" ~ "eta[0]",
      right == "sigma_day01" ~ "eta[1]",
    )
  ) %>% 
  mutate(
    label = formatC(Estimate, digits = 2, format = "f") %>%
      str_replace(., "0.", ".")) %>% 
  mutate(left = factor(left, levels = levels),
         right = factor(right, levels = levels)) %>% 
  mutate(right = fct_rev(right))

rho
```

```{r}
rho %>% 
  ggplot(aes(x = left, y = right)) +
  geom_tile(aes(fill = Estimate)) +
  geom_text(aes(label = label),
            family = "Courier", size = 3) +
  scale_fill_gradient2(expression(rho), low = "#59708b",
                       mid = "#fcf9f0", high = "#a65141",
                       midpoint = 0, 
                       labels = c(-1, "", 0, "", 1),
                       limits = c(-1, 1)) +
  scale_x_discrete(NULL, drop = F, 
                   labels = ggplot2:::parse_safe,
                   position = "top") +
  scale_y_discrete(NULL, drop = F,
                   labels = ggplot2:::parse_safe) +
  ggtitle(expression("The lower triangle for "*bold(R)),
          subtitle = "Note, each cell is summarized by\nits posterior mean.") +
  theme(axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.text = element_text(hjust = 1))
```

```{r}
nd <-
  dat %>% 
  filter(record_id %in% c(30, 115)) %>% 
  select(record_id, N_A.std, day01)

bind_cols(
  fitted(b14.14, newdata = nd) %>% 
    data.frame(),
  predict(b14.14, newdata = nd) %>% 
    data.frame() %>% 
    select(Q2.5:Q97.5) %>% 
    set_names("p_lower", "p_upper")
) %>% 
  bind_cols(nd) %>% 
  
  ggplot(aes(x = day01)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "#8B9DAF", color = "#8B9DAF", alpha = 1/2, linewidth = 1/2) +
  geom_ribbon(aes(ymin = p_lower, ymax = p_upper),
              fill = "#8B9DAF", alpha = 1/2) +
  geom_point(aes(y = N_A.std),
             color = "#8B9DAF") +
  scale_x_continuous(breaks = c(0, .5, 1)) +
  ylab("negative affect (standardized)") +
  facet_wrap(~ record_id)
```

### 14.6.4 Time to go multivariate
For our final stage in this progression, we will fit what Donald R. Williams et al. (2020) called a M-MELSM, a multivariate mixed-effects location scale model. Recall these data have measures of both negative and positive affect. The standardized values for PA are waiting for us in the P_A.std column. Within the brms framework, this is a combination of sensibilities from Bürkner’s vignettes on distributional models (Bürkner, 2022a) and multivariate models (Bürkner, 2022d). We might express the statistical model as

$$
\begin{align*}
\color{#A65141}{\begin{bmatrix} \text{NA}_{ij} \\ \text{PA}_{ij} \end{bmatrix}} & \color{#A65141}\sim \color{#A65141}{\operatorname{MVNormal}\begin{pmatrix} \begin{bmatrix}
\mu_{ij}^\text{NA} \\ \mu_{ij}^\text{PA} \end{bmatrix}, \mathbf \Sigma \end{pmatrix}} \\

\mu_{ij}^{\color{#A65141}{\text{NA}}} & = \beta_0^\text{NA} + \beta_1^\text{NA} \text{time}_{ij} + u_{0i}^\text{NA} + u_{1i}^\text{NA} \\
\mu_{ij}^{\color{#A65141}{\text{PA}}} & = \beta_0^\text{PA} + \beta_1^\text{PA} \text{time}_{ij} + u_{0i}^\text{PA} + u_{1i}^\text{PA} \\
\log \left ( \sigma_{ij}^{\color{#A65141}{\text{NA}}} \right ) & = \eta_0^\text{NA} + \eta_1^\text{NA} \text{time}_{ij} + u_{2i}^\text{NA} + u_{3i}^\text{NA} \\
\log \left ( \sigma_{ij}^{\color{#A65141}{\text{PA}}} \right ) & = \eta_0^\text{PA} + \eta_1^\text{PA} \text{time}_{ij} + u_{2i}^\text{PA} + u_{3i}^\text{PA}  \\

\begin{bmatrix} u_{0i}^\text{NA}, u_{1i}^\text{NA}, u_{2i}^\text{NA}, u_{3i}^\text{NA}, u_{0i}^\text{PA}, u_{1i}^\text{PA}, u_{2i}^\text{PA}, u_{3i}^\text{PA} \end{bmatrix}' & \sim \operatorname{MVNormal}(\mathbf 0, \mathbf S \mathbf R \mathbf S) \\

\mathbf S & = \begin{bmatrix} \sigma_0^\text{NA} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ 0 & \sigma_1^\text{NA} & 0 & 0 & 0 & 0 & 0 & 0 \\ 0 & 0 & \sigma_2^\text{NA} & 0 & 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & \sigma_3^\text{NA} & 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 & \sigma_0^\text{PA} & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 & 0 & \sigma_1^\text{PA} & 0 & 0 \\ 0 & 0 & 0 & 0 & 0 & 0 & \sigma_2^\text{PA} & 0 \\ 0 & 0 & 0 & 0 & 0 & 0 & 0 & \sigma_3^\text{PA}  \end{bmatrix} \\

\mathbf R & = \begin{bmatrix}
1 & \rho_{12} & \rho_{13} & \rho_{14} & \rho_{15} & \rho_{16} & \rho_{17} & \rho_{18} \\
\rho_{21} & 1 & \rho_{23} & \rho_{24} & \rho_{25} & \rho_{26} & \rho_{27} & \rho_{28} \\
\rho_{31} & \rho_{32} & 1 & \rho_{34} & \rho_{35} & \rho_{36} & \rho_{37} & \rho_{38} \\
\rho_{41} & \rho_{42} & \rho_{43} & 1 & \rho_{45} & \rho_{46} & \rho_{47} & \rho_{48} \\
\rho_{51} & \rho_{52} & \rho_{53} & \rho_{54} & 1 & \rho_{56} & \rho_{57} & \rho_{58} \\
\rho_{61} & \rho_{62} & \rho_{63} & \rho_{64} & \rho_{65} & 1 & \rho_{67} & \rho_{68} \\
\rho_{71} & \rho_{72} & \rho_{73} & \rho_{74} & \rho_{75} & \rho_{76} & 1 & \rho_{78} \\
\rho_{81} & \rho_{82} & \rho_{83} & \rho_{84} & \rho_{85} & \rho_{86} & \rho_{87} & 1
\end{bmatrix} \\

\beta_0^\text{NA} \text{ and } \beta_0^\text{PA} & \sim \operatorname{Normal}(0, 0.2) \\
\beta_1^\text{NA} \text{ and } \beta_1^\text{PA} & \sim \operatorname{Normal}(0, 1) \\
\eta_0^\text{NA},\dots, \eta_1^\text{PA} & \sim \operatorname{Normal}(0, 1) \\
\sigma_0^\text{NA},\dots, \sigma_1^\text{PA} & \sim \operatorname{Exponential}(1) \\
\mathbf R & \sim \operatorname{LKJ}(2),
\end{align*}
$$

where the $\text{NA}$ and $\text{PA}$ superscripts indicate which variable is connected with which parameter. This is a straight multivariate generalization from the previous model, b14.14. Now we have eight parameters varying across participants, resulting in an $8 \times 8\ \text{S}$ matrix and an $8 \times 8\ \text{R}$ matrix. Here’s the `brms::brm()` code.

```{r}
b14.15 <-
  brm(data = dat,
      family = gaussian,
      bf(mvbind(N_A.std, P_A.std) ~ 1 + day01 + (1 + day01 |i| record_id),
         sigma ~ 1 + day01 + (1 + day01 |i| record_id)) + set_rescor(rescor = FALSE),
      prior = c(prior(normal(0, 0.2), class = Intercept, resp = NAstd),
                prior(normal(0, 1), class = b, resp = NAstd),
                prior(exponential(1), class = sd, resp = NAstd),
                
                prior(normal(0, 1), class = Intercept, dpar = sigma, resp = NAstd),
                prior(normal(0, 1), class = b, dpar = sigma, resp = NAstd),
                prior(exponential(1), class = sd, dpar = sigma, resp = NAstd),
                
                prior(normal(0, 0.2), class = Intercept, resp = PAstd),
                prior(normal(0, 1), class = b, resp = PAstd),
                prior(exponential(1), class = sd, resp = PAstd),
                
                prior(normal(0, 1), class = Intercept, dpar = sigma, resp = PAstd),
                prior(normal(0, 1), class = b, dpar = sigma, resp = PAstd),
                prior(exponential(1), class = sd, dpar = sigma, resp = PAstd),

                prior(lkj(2), class = cor)),
      iter = 3000, warmup = 1000, chains = 4, cores = 8,
      seed = 14,
      file = "fits/b14.15")
```

```{r}
print(b14.15)
```

```{r}
levels <- c("beta[0]^'NA'", "beta[1]^'NA'", "eta[0]^'NA'", "eta[1]^'NA'",
            "beta[0]^'PA'", "beta[1]^'PA'", "eta[0]^'PA'", "eta[1]^'PA'")

# two different options for ordering the parameters
# levels <- c("beta[0]^'NA'", "beta[1]^'NA'", "beta[0]^'PA'", "beta[1]^'PA'", "eta[0]^'NA'", "eta[1]^'NA'", "eta[0]^'PA'", "eta[1]^'PA'")
# levels <- c("beta[0]^'NA'", "beta[0]^'PA'", "beta[1]^'NA'", "beta[1]^'PA'","eta[0]^'NA'", "eta[0]^'PA'", "eta[1]^'NA'", "eta[1]^'PA'")

rho <-
  posterior_summary(b14.15) %>% 
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "cor_")) %>% 
  mutate(param = str_remove(param, "cor_record_id__")) %>% 
  separate(param, into = c("left", "right"), sep = "__") %>% 
  mutate(
    left = case_when(
      left == "NAstd_Intercept"       ~ "beta[0]^'NA'",
      left == "NAstd_day01"           ~ "beta[1]^'NA'",
      left == "sigma_NAstd_Intercept" ~ "eta[0]^'NA'",
      left == "sigma_NAstd_day01"     ~ "eta[1]^'NA'",
      left == "PAstd_Intercept"       ~ "beta[0]^'PA'",
      left == "PAstd_day01"           ~ "beta[1]^'PA'",
      left == "sigma_PAstd_Intercept" ~ "eta[0]^'PA'",
      left == "sigma_PAstd_day01"     ~ "eta[1]^'PA'"
      ),
    right = case_when(
      right == "NAstd_Intercept"       ~ "beta[0]^'NA'",
      right == "NAstd_day01"           ~ "beta[1]^'NA'",
      right == "sigma_NAstd_Intercept" ~ "eta[0]^'NA'",
      right == "sigma_NAstd_day01"     ~ "eta[1]^'NA'",
      right == "PAstd_Intercept"       ~ "beta[0]^'PA'",
      right == "PAstd_day01"           ~ "beta[1]^'PA'",
      right == "sigma_PAstd_Intercept" ~ "eta[0]^'PA'",
      right == "sigma_PAstd_day01"     ~ "eta[1]^'PA'"
    )
  ) %>% 
  mutate(label = formatC(Estimate, digits = 2, format = "f") %>% str_replace(., "0.", ".")) %>% 
  mutate(left  = factor(left, levels = levels),
         right = factor(right, levels = levels)) %>% 
  mutate(right = fct_rev(right))

rho %>% head()
```

```{r}
rho %>% 
  full_join(rename(rho, right = left, left = right),
            by = c("left", "right", "Estimate", "Est.Error", "Q2.5", "Q97.5", "label")) %>%
  
  ggplot(aes(x = left, y = right)) +
  geom_tile(aes(fill = Estimate)) +
  geom_hline(yintercept = 4.5, color = "#100F14") +
  geom_vline(xintercept = 4.5, color = "#100F14") +
  geom_text(aes(label = label),
            family = "Courier", size = 3) +
  scale_fill_gradient2(expression(rho),
                       low = "#59708b", mid = "#FCF9F0", high = "#A65141", midpoint = 0,
                       labels = c(-1, "", 0, "", 1), limits = c(-1, 1)) +
  scale_x_discrete(NULL, expand = c(0, 0), labels = ggplot2:::parse_safe, position = "top") +
  scale_y_discrete(NULL, expand = c(0, 0), labels = ggplot2:::parse_safe) +
  theme(axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.text = element_text(hjust = 1))
```

#### 14.6.4.1 Plot with uncertainty
```{r}
levels <- c("beta[0]^'NA'", "beta[1]^'NA'", "eta[0]^'NA'", "eta[1]^'NA'",
            "beta[0]^'PA'", "beta[1]^'PA'", "eta[0]^'PA'", "eta[1]^'PA'")

rho <-
  as_draws_df(b14.15) %>% 
  select(starts_with("cor_")) %>% 
  slice_sample(n = 50 * 50) %>% 
  bind_cols(crossing(x = 1:50, y = 1:50)) %>% 
  pivot_longer(cols = -c(x:y)) %>% 
  mutate(name = str_remove(name, "cor_record_id__")) %>% 
  separate(name, into = c("col", "row"), sep = "__") %>% 
  mutate(
    col = case_when(
      col == "NAstd_Intercept"       ~ "beta[0]^'NA'",
      col == "NAstd_day01"           ~ "beta[1]^'NA'",
      col == "sigma_NAstd_Intercept" ~ "eta[0]^'NA'",
      col == "sigma_NAstd_day01"     ~ "eta[1]^'NA'",
      col == "PAstd_Intercept"       ~ "beta[0]^'PA'",
      col == "PAstd_day01"           ~ "beta[1]^'PA'",
      col == "sigma_PAstd_Intercept" ~ "eta[0]^'PA'",
      col == "sigma_PAstd_day01"     ~ "eta[1]^'PA'"
    ),
    row = case_when(
      row == "NAstd_Intercept"       ~ "beta[0]^'NA'",
      row == "NAstd_day01"           ~ "beta[1]^'NA'",
      row == "sigma_NAstd_Intercept" ~ "eta[0]^'NA'",
      row == "sigma_NAstd_day01"     ~ "eta[1]^'NA'",
      row == "PAstd_Intercept"       ~ "beta[0]^'PA'",
      row == "PAstd_day01"           ~ "beta[1]^'PA'",
      row == "sigma_PAstd_Intercept" ~ "eta[0]^'PA'",
      row == "sigma_PAstd_day01"     ~ "eta[1]^'PA'"
    )
  ) %>% 
  mutate(col = factor(col, levels = levels),
         row = factor(row, levels = levels))

rho %>% 
  full_join(rename(rho, col = row, row = col),
            by = c("x", "y", "col", "row", "value")) %>%
  
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradient2(expression(rho),
                       low = "#59708b", mid = "#FCF9F0", high = "#A65141", midpoint = 0,
                       labels = c(-1, "", 0, "", 1), limits = c(-1, 1)) +
  scale_x_continuous(NULL, breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(0, 0)) +
  theme(strip.text = element_text(size = 12)) +
  facet_grid(row ~ col, labeller = label_parsed, switch = "y")
```

