---
title: "6 The Haunted dag & The Causal Terror"
format: html
---
```{r}
rm(list = ls())
```


##### 6.0.0.1 Overthinking: Simulated science distortion
```{r}
library(tidyverse)

set.seed(1914)
n <- 200
p <- 0.1

d <-
  # uncorrelated newsworthiness and trustworthiness
  tibble(newsworthiness = rnorm(n, mean = 0, sd = 1),
         trustworthiness = rnorm(n, mean = 0, sd = 1)) %>% 
  # total score
  mutate(total_score = newsworthiness + trustworthiness) %>% 
  # select top 10% of combined score
  mutate(selected = ifelse(total_score >= quantile(total_score, 1 - p), TRUE, FALSE))

head(d)
```
The correlation among those cases for which ```selected == TRUE```
```{r}
d %>% 
  filter(selected == TRUE) %>% 
  select(newsworthiness, trustworthiness) %>% 
  cor()
```

```{r}
theme_set(theme_minimal())

text <-
  tibble(newsworthiness = c(2, 1),
         trustworthiness = c(2.25, -2.5),
         selected = c(TRUE, FALSE),
         label = c("selected", "rejected"))

d %>% 
  ggplot(aes(x = newsworthiness, y = trustworthiness, color = selected)) +
  geom_point(aes(shape = selected), alpha = 3/4) +
  geom_text(data = text,
            aes(label = label)) +
  geom_smooth(data = . %>% filter(selected == TRUE),
              method = "lm", fullrange = T,
              color = "lightblue", se = F, linewidth = 1/2) +
  scale_color_manual(values = c("black", "lightblue")) +
  scale_shape_manual(values = c(1, 19)) +
  scale_x_continuous(limits = c(-3, 3.9), expand = c(0, 0)) +
  coord_cartesian(ylim = range(d$trustworthiness)) +
  theme(legend.position = "none")
```
Newsworthiness and trustworthiness are only negatively correlated if we stratify by selected paper

## 6.1 Multicollinearity
### 6.1.1 Multicollinear legs
```{r}
n <- 100
set.seed(909)

d <-
  tibble(height = rnorm(n, mean = 10, sd = 2),
         leg_prop = runif(n, min = 0.4, max = 0.5)) %>% 
  mutate(leg_left = leg_prop * height + rnorm(n, mean = 0, sd = .02),
         leg_right = leg_prop * height + rnorm(n, mean = 0, sd = .02))
```

Left leg and left right are very highly correlated (as expected)
```{r}
d %>% 
  summarize(r = cor(leg_left, leg_right) %>% round(digits = 4))
```

```{r}
d %>% 
  ggplot(aes(x = leg_left, y = leg_right)) +
  geom_point(alpha = 1/2, color = "forestgreen")
```

```{r}
library(brms)
```

```{r}
b6.1 <-
  brm(data = d,
      family = gaussian,
      height ~ 1 + leg_left + leg_right,
      prior = c(prior(normal(10, 100), class = Intercept),
                prior(normal(2, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.01")
```

```{r}
print(b6.1)
```
The est. error is very big and not looking too good

Coefficient plot
```{r}
library(bayesplot)

color_scheme_set("orange")

mcmc_plot(b6.1,
          type = "intervals",
          prob = .5,
          prob_outer = .95,
          point_est = "mean") +
  labs(title = "The coefficient plot for the two-leg model",
       subtitle = "Holy smokes; look at the widths of those betas!") +
  theme(axis.text.y = element_text(hjust = 0),
        panel.grid.minor = element_blank(),
        strip.text = element_text(hjust = 0))
```

```{r}
sim_and_fit <- function(seed, n = 100){
  # set up the parameters
  n <- n
  set.seed(seed)
  
  # simulate the new data
  d <- 
    tibble(height = rnorm(n, mean = 10, sd = 2),
           leg_prop = runif(n, min = 0.4, max = 0.5)) %>% 
    mutate(leg_left = leg_prop * height + rnorm(n, mean = 0, sd = .02),
           leg_right = leg_prop * height + rnorm(n, mean = 0, sd = .02))
  
  # update b6.1 to the new data
  fit <- update(b6.1, newdata = d, seed = 6)
}
```

```{r}
sim <-
  tibble(seed = 1:4) %>% 
  mutate(post = map(seed, ~ sim_and_fit(.) %>% 
                      as_draws_df()))
```

```{r}
head(sim)
```

```{r}
library(tidybayes)

sim %>% 
  unnest(post) %>% 
  pivot_longer(b_Intercept:sigma) %>% 
  mutate(seed = str_c("seed ", seed)) %>% 
  
  ggplot(aes(x = value, y = name)) +
  stat_pointinterval(.width = .95, color = "forestgreen") +
  labs(x = "posterior", y = NULL) +
  theme(axis.text.y = element_text(hjust = 0),
        panel.border = element_rect(color = "black", fill = "transparent"),
        panel.grid.minor = element_blank(),
        strip.text = element_text(hjust = 0)) +
  facet_wrap(~ seed, ncol = 1)
```

```{r}
pairs(b6.1, variable = variables(b6.1)[2:3])
```

nicer and more focused attempt at drawing pair
```{r}
post <- as_draws_df(b6.1)

post %>% 
  ggplot(aes(x = b_leg_left, y = b_leg_right)) +
  geom_point(color = "forestgreen", alpha = 1/10, size = 1/2)
```

Similar plot with the ```mcmc_scatter()```
```{r}
color_scheme_set("green")

post %>% 
  mcmc_scatter(pars = c("b_leg_left", "b_leg_right"),
               size = 1/2,
               alpha = 1/10)
```

```{r}
post %>% 
  summarise(rho = cor(b_leg_left, b_leg_right))
```
Model that adds two parameters
```{r}
post %>% 
  ggplot(aes(x = b_leg_left + b_leg_right, y = 0)) +
  stat_halfeye(point_interval = median_qi,
               fill = "steelblue", .width = .95) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Sum the multicollinear coefficients",
       subtitle = "Marked by the median and 95% PIs")
```

```{r}
b6.2 <-
  brm(data = d,
      family = gaussian,
      height ~ 1 + leg_left,
      prior = c(prior(normal(10, 100), class = Intercept),
                prior(normal(2, 10), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.02")
```

```{r}
print(b6.2)
```

Posterior SD for leg_left looks much better. Let's compare this density to the one in Figure 6.2b
```{r}
as_draws_df(b6.2) %>% 
  
  ggplot(aes(x = b_leg_left, y = 0)) +
  stat_halfeye(point_interval = median_qi,
               fill = "steelblue", .width = .95) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Just one coefficient needed",
       subtitle = "Marked by the median and 95% PIs",
       x = "only b_leg_left, this time")
```

### 6.1.2 Multicollinear milk
```{r}
data(milk, package = "rethinking")
d <- milk
rm(milk)
```

```{r}
d <-
  d %>% 
  mutate(k = rethinking::standardize(kcal.per.g),
         f = rethinking::standardize(perc.fat),
         l = rethinking::standardize(perc.lactose))
```

```{r}
# k regressed on f
b6.3 <-
  brm(data = d,
      family = gaussian,
      k ~ 1 + f,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.03")

# k regressed on l
b6.4 <-
  update(b6.3,
         newdata = d,
         formula = k ~ 1 + l,
         seed = 6,
         file = "fits/b06.04")
```

```{r}
posterior_summary(b6.3)[1:3, ] %>% round(digits = 2)
```

```{r}
posterior_summary(b6.4)[1:3, ] %>% round(digits = 2)
```
Using both predictors in the same regression model
```{r}
b6.5 <-
  update(b6.4,
         newdata = d,
         formula = k ~ 1 + f + l,
         seed = 6,
         file = "fits/b06.05")
```

```{r}
posterior_summary(b6.5)[1:3, ] %>% round(digits = 2)
```

```{r}
d %>% 
  select(kcal.per.g, perc.fat, perc.lactose) %>% 
  pairs(col = "forestgreen")
```

pairs using GGally version
```{r}
library(GGally)

# define a couple custom functions
my_diag <- function(data, mapping, ...){
  ggplot(data = data, mapping = mapping) +
    geom_density(fill = "steelblue", color = "black")
}

my_lower <- function(data, mapping, ...){
  ggplot(data = data, mapping = mapping) +
    geom_smooth(method = "lm", color = "orange", se = F) +
    geom_point(alpha = .8, size = 1/3, color = "blue")
}

# plug those custom functions into ggpairs()
ggpairs(data = d, columns = c(3:4, 6),
        upper = list(continuous = wrap("cor", family = "sans", color = "black")),
        diag = list(continuous = my_diag),
        lower = list(continuous = my_lower))
```

```{r}
library(ggdag)

dag_coords <-
  tibble(name = c("L", "D", "F", "K"),
         x = c(1, 2, 3, 2),
         y = c(2, 2, 2, 1))

dagify(L ~ D,
       F ~ D,
       K ~ L + F,
       coords = dag_coords) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "D"),
                 alpha = 1/2, size = 6.5, show.legend = F) +
  geom_point(x = 2, y = 2,
              size = 6.5, shape = 1, stroke = 1, color = "orange") +
  geom_dag_text(color = "black") +
  geom_dag_edges() +
  scale_color_manual(values = c("steelblue", "orange")) +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1))
```

#### 6.1.2.2 Overthinking: Simulating collinearity
```{r}
# define a custom function
sim_coll <- function(seed, rho){
  
  # simulate the data
  set.seed(seed)
  
  d <-
    d %>% 
    mutate(x = rnorm(n(),
                     mean = perc.fat * rho,
                     sd = sqrt((1 - rho^2) * var(perc.fat))))
  
  # fit an OLS model
  m <- lm(kcal.per.g ~ perc.fat + x, data = d)
  
  # extract the parameter SD
  sqrt(diag(vcov(m)))[2]
}

# how many simulations per 'rho'-value would you like?
n_seed <- 100

# how many 'rho'-values from 0, to .99 would you like to evaluate the process over?
n_rho <- 30

d <-
  crossing(seed = 1:n_seed,
           rho = seq(from = 0, to = .99, length.out = n_rho)) %>% 
  mutate(parameter_sd = purrr::map2_dbl(seed, rho, sim_coll)) %>% 
  group_by(rho) %>% 
  # we'll summarise() our output gby the mean and 95% intervals
  summarise(mean = mean(parameter_sd),
            ll = quantile(parameter_sd, prob = .025),
            ul = quantile(parameter_sd, prob = .975))
```

```{r}
d %>% 
  ggplot(aes(x = rho, y = mean)) +
  geom_smooth(aes(ymin = ll, ymax = ul),
              stat = "identity",
              fill = "orange", color = "orange", alpha = 1/3, linewidth = 2/3) +
  labs(x = expression(rho),
       y = "parameter SD") +
  coord_cartesian(ylim = c(0, .0072))
```

## 6.2 Post-treatment bias
```{r}
# how many plants would you like?
n <- 100

set.seed(71)
d <-
  tibble(h0 = rnorm(n, mean = 10, sd = 2),
         treatment = rep(0:1, each = n/2),
         fungus = rbinom(n, size = 1, prob = .5 - treatment * 0.4),
         h1 = h0 + rnorm(n, mean = 5 - 3 * fungus, sd = 1))
```

```{r}
d %>% head()
```

```{r}
d %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  mean_qi(.width = .89) %>% 
  mutate_if(is.double, round, digits = 2)
```

### 6.2.1 A prior is born
```{r}
set.seed(6)

# simulate
sim_p <-
  tibble(sim_p = rlnorm(1e4, meanlog = 0, sdlog = .25))

# wrangle
sim_p %>% 
  mutate("exp(sim_p)" = exp(sim_p)) %>% 
  gather() %>% 
  
  # plot
  ggplot(aes(x = value)) +
  geom_density(fill = "steelblue") +
  scale_x_continuous(breaks = c(0, .5, 1, 1.5, 2, 3, 5)) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 6)) +
  theme(panel.grid.minor.x = element_blank()) +
  facet_wrap(~ key, scale = "free_y", ncol = 1)
```

Summarize
```{r}
sim_p %>% 
  mutate("exp(sim_p)" = exp(sim_p)) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  mean_qi(.width = .89) %>% 
  mutate_if(is.double, round, digits = 2)
```
Create a model
```{r}
b6.6 <-
  brm(data = d,
      family = gaussian,
      h1 ~ 0 + h0,
      prior = c(prior(lognormal(0, .25), class = b, lb = 0),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.06")
```

```{r}
print(b6.6)
```

Add predictor variable treatment and fungus
```{r}
b6.7 <-
  brm(data = d,
      family = gaussian,
      bf(h1 ~ h0 * (a + t * treatment + f * fungus),
         a + t + f ~ 1,
         nl = TRUE),
      prior = c(prior(lognormal(0, .2), nlpar = a, lb = 0),
                prior(normal(0, .5), nlpar = t),
                prior(normal(0, .5), nlpar = f),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.07")
```

```{r}
print(b6.7)
```

### 6.2.2 Blocked by consequence
Previous model is wrong because we include fungus which is a post-treatment variable. To measure the treatment effect properly, we should omit fungus from model
```{r}
b6.8 <-
  brm(data = d,
      family = gaussian,
      bf(h1 ~ h0 * (a + t * treatment),
         a + t ~ 1,
         nl = TRUE),
      prior = c(prior(lognormal(0, .2), nlpar = a, lb = 0),
                prior(normal(0, .5), nlpar = t),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.08")
```

```{r}
print(b6.8)
```

### 6.2.3 Fungus and d-separation
```{r}
# define our coordinages
dag_coords <-
  tibble(name = c("H0", "T", "F", "H1"),
         x = c(1, 5, 4, 3),
         y = c(2, 2, 1.5, 1))

# save our DAG
dag <-
  dagify(F ~ T,
         H1 ~ H0 + F,
         coords = dag_coords)

# plot
dag %>% 
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "steelblue", alpha = 1/2, size = 6.5) +
  geom_dag_text(color = "black") +
  geom_dag_edges() +
  theme_dag()
```

```{r}
gg_simple_dag <- function(d){
  d %>% 
    ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_dag_point(color = "steelblue", alpha = 1/2, size = 6.5) +
    geom_dag_text(color = "black") +
    geom_dag_edges() +
    theme_dag()
}

# try it out
dag %>% 
  gg_simple_dag()
```

```{r}
library(dagitty)

dag %>% 
  dseparated("T", "H1")
```

```{r}
dag %>% 
  dseparated("T", "H1", "F")
```

```{r}
impliedConditionalIndependencies(dag)
```

Now consider a DAG of a different kind of causal structure
```{r}
# define our coordinates
dag_coords <-
  tibble(name = c("H0", "H1", "M", "F", "T"),
         x = c(1, 2, 2.5, 3, 4),
         y = c(2, 2, 1, 2, 2))

# save our DAG
dag <-
  dagify(F ~ M + T,
         H1 ~ H0 + M,
         coords = dag_coords)

# plot
dag %>% 
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = name == "M"),
                 alpha = 1/2, size = 6.5, show.legend = F) +
  geom_point(x = 2.5, y = 1,
             size = 6.5, shape = 1, stroke = 1, color = "orange") +
  geom_dag_text(color = "black") +
  geom_dag_edges() +
  scale_color_manual(values = c("steelblue", "orange")) +
  theme_dag()
```

```{r}
gg_fancy_dag <- function(d, x = 1, y = 1, circle = "U") {
  
  d %>% 
    ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_dag_point(aes(color = name == circle),
                   alpha = 1/2, size = 6.5, show.legend = F) +
    geom_point(x = x, y = y, 
               size = 6.5, shape = 1, stroke = 1, color = "orange") +
    geom_dag_text(color = "black") +
    geom_dag_edges() + 
    scale_color_manual(values = c("steelblue", "orange")) +
    theme_dag()
  
}

# check it out
dag %>% 
  gg_fancy_dag(x = 2.5, y = 1, circle = "M")
```

Simulate data on the new DAG
```{r}
set.seed(71)
n <- 1000

d2 <-
  tibble(h0 = rnorm(n, mean = 10, sd = 2),
         treatment = rep(0:1, each = n/2),
         m = rbinom(n, size = 1, prob = .5),
         fungus = rbinom(n, size = 1, prob = .5 - treatment * 0.4 + 0.4 * m),
         h1 = h0 + rnorm(n, mean = 5 + 3 * m, sd = 1))

head(d2)
```

Use update() to refit b6.7 and b6.8 to the new data
```{r}
b6.7b <-
  update(b6.7,
         newdata = d2,
         seed = 6,
         file = "fits/b06.07b")

b6.8b <-
  update(b6.8,
         newdata = d2,
         seed = 6,
         file = "fits/b06.08b")
```

```{r}
posterior_summary(b6.7b)[1:3, ] %>% round(digits = 2)
```

```{r}
posterior_summary(b6.8b)[1:3, ] %>% round(digits = 2)
```

# 6.3 Collider bias
```{r}
dag_coords <-
  tibble(name = c("T", "S", "N"),
         x = 1:3,
         y = 1)

dagify(S ~ T + N,
       coords = dag_coords) %>% 
  gg_simple_dag()
```

### 6.3.1 Collider of false sorrow
```{r}
dagify(M ~ H + A,
       coords = dag_coords %>% 
         mutate(name = c("H", "M", "A"))) %>% 
  gg_simple_dag()
```

```{r}
new_borns <- function(n = 20){
  tibble(a = 1,
         m = 0,
         h = seq(from = -2, to = 2, length.out = n))
}
```

```{r}
new_borns()
```

Make a function to update population
```{r}
update_population <- function(pop, n_births = 20, aom = 18, max_age = 65){
  pop %>% 
    mutate(a = a + 1, # everyone gets one year older
           # some people get marrid
           m = ifelse(m >= 1, 1, (a >= aom) * rbinom(n(), 1, rethinking::inv_logit(h - 4)))
           ) %>% 
    filter(a <= max_age) %>%  # old people die
    bind_rows(new_borns(n_births))
}
```

```{r}
new_borns() %>% 
  update_population()
```

Run the population simulation for 1000 years
```{r}
# this was McElreath's seed
set.seed(1977)

# year 1
d <- new_borns(n = 20)

# years 2 through 1000
for(i in 2:1000){
  d <- update_population(d, n_births = 20, aom = 18, max_age = 65)
}

# now rename()
d <-
  d %>% 
  rename(age = a, married = m, happiness = h)

# take a look
glimpse(d)
```

Summarize the variables
```{r}
d %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  mean_qi(value) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
d %>% 
  mutate(married = factor(married, labels = c("unmarried", "married"))) %>% 
  
  ggplot(aes(x = age, y = happiness, color = married)) +
  geom_point(size = 1.75) +
  scale_color_manual(NULL, values = c("grey85", "forestgreen")) +
  scale_x_continuous(expand = c(.015, .015)) +
  theme(panel.grid = element_blank())
```

Here's the likelihood for the simple Gaussian multivariable model predicting happiness
$$
\begin{align}
happiness_{i} &\sim Normal(\mu_{i}, \sigma) \\
\mu_{i} &= \alpha_{married[i]} + \beta_{1} age_{i},
\end{align}
$$

```{r}
d2 <-
  d %>% 
  filter(age > 17) %>% 
  mutate(a = (age - 18) / (65 - 18))

head(d2)
```

```{r}
d2 <-
  d2 %>% 
  mutate(mid = factor(married + 1, labels = c("single", "married")))

head(d2)
```

Fit the model
```{r}
b6.9 <-
  brm(data = d2,
      family = gaussian,
      happiness ~ 0 + mid + a,
      prior = c(prior(normal(0, 1), class = b, coef = midmarried),
                prior(normal(0, 1), class = b, coef = midsingle),
                prior(normal(0, 2), class = b, coef = a),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.09")
```

```{r}
print(b6.9)
```

Now drop marriage status
```{r}
b6.10 <-
  brm(data = d2,
      family = gaussian,
      happiness ~ 0 + Intercept + a,
      prior = c(prior(normal(0, 1), class = b, coef = Intercept),
                prior(normal(0, 2), class = b, coef = a),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.10")
```

```{r}
print(b6.10)
```

### 6.3.2 The haunted DAG
```{r}
dag_coords <-
  tibble(name = c("G", "P", "C"),
         x = c(1, 2, 2),
         y = c(2, 2, 1))

dagify(P ~ G,
       C ~ P + G,
       coords = dag_coords) %>% 
  gg_simple_dag()
```

Now we add the haunting variable, U
```{r}
dag_coords <-
  tibble(name = c("G", "P", "C", "U"),
         x = c(1, 2, 2, 2.5),
         y = c(2, 2, 1, 1.5))

dagify(P ~ G + U,
       C ~ P + G + U,
       coords = dag_coords) %>% 
  gg_fancy_dag(x = 2.5, y = 1.5, circle = "U")
```

Let's simulate some data
```{r}
# how many grandparent-parent-child triads would you like?
n <- 200

b_gp <- 1 # direct effect of G on P
b_gc <- 0 # direct effect of G on C
b_pc <- 1 # direct effect of P on C
b_u <- 2 # direct effect of U on P and C

# simulate triads
set.seed(1)
d <-
  tibble(u = 2 * rbinom(n, size = 1, prob = .5) - 1,
         g = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(p = rnorm(n, mean = b_gp * g + b_u * u, sd = 1)) %>% 
  mutate(c = rnorm(n, mean = b_pc * p + b_gc * g + b_u * u, sd = 1))

head(d)
```

Fit the model without u
```{r}
b6.11 <-
  brm(data = d,
      family = gaussian,
      c ~ 0 + Intercept + p + g,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.11")
```

```{r}
print(b6.11)
```

```{r}
d %>% 
  mutate(centile = ifelse(p >= quantile(p, prob = .45) & p <= quantile(p, prob = .60), "a", "b"),
         u = factor(u)) %>% 
  
  ggplot(aes(x = g, y = c)) +
  geom_point(aes(shape = centile, color = u),
             size = 2.5, stroke = 1/4) +
  stat_smooth(data = . %>% filter(centile == "a"),
              method = "lm", se = F, linewidth = 1/2, color = "black", fullrange = T) +
  scale_shape_manual(values = c(19, 1)) +
  scale_color_manual(values = c("black", "lightblue")) +
  theme(legend.position = "none")
```

Now fit the model including u
```{r}
b6.12 <-
  update(b6.11,
         newdata = d,
         formula = c ~ 0 + Intercept + p + g + u,
         seed = 6,
         file = "fits/b06.12")
```

```{r}
print(b6.12)
```

## 6.4 Confronting confounding
```{r}
dag_coords <-
  tibble(name = c("E", "U", "W"),
         x = c(1, 2, 3),
         y = c(1, 2, 1))

dagify(E ~ U,
       W ~ E + U,
       coords = dag_coords) %>% 
  gg_simple_dag()
```

```{r}
dagify( W ~ E + U,
        coords = dag_coords) %>% 
  gg_simple_dag()
```

### 6.4.1 Shutting the backdoor
Representations for four types of variable relations: the fork, pipe, collider, and descendant
```{r}
d1 <-
  dagify(X ~ Z,
         Y ~ Z,
         coords = tibble(name = c("X", "Y", "Z"),
                         x = c(1, 3, 2),
                         y = c(2, 2, 1)))

d2 <-
  dagify(Z ~ X,
         Y ~ Z,
         coords = tibble(name = c("X", "Y", "Z"),
                         x = c(1, 3, 2),
                         y = c(2, 1, 1.5)))

d3 <- 
  dagify(Z ~ X + Y,
         coords = tibble(name = c("X", "Y", "Z"),
                         x = c(1, 3, 2),
                         y = c(1, 1, 2)))

d4 <- 
  dagify(Z ~ X + Y,
         D ~ Z,
         coords = tibble(name = c("X", "Y", "Z", "D"),
                         x = c(1, 3, 2, 2),
                         y = c(1, 1, 2, 1.05)))

p1 <- gg_simple_dag(d1) + labs(subtitle = "The Fork")
p2 <- gg_simple_dag(d2) + labs(subtitle = "The Pipe")
p3 <- gg_simple_dag(d3) + labs(subtitle = "The Collider")
p4 <- gg_simple_dag(d4) + labs(subtitle = "The Descendant")
```

```{r}
library(patchwork)

(p1 | p2 | p3 | p4) &
  theme(plot.subtitle = element_text(hjust = 0.5)) &
  plot_annotation(title = "The four elemental confounds") 
```

### 6.4.2 Two roads
```{r}
dag_coords <-
  tibble(name = c("A", "B", "C", "U", "X", "Y"),
         x    = c(2, 2, 3, 1, 1, 3),
         y    = c(4, 2, 3, 3, 1, 1))

dagify(B ~ C + U,
       C ~ A,
       U ~ A,
       X ~ U,
       Y ~ C + X,
       coords = dag_coords) %>%
  gg_fancy_dag(x = 1, y = 3, circle = "U")
```

```{r}
dag_6.1 <- 
  dagitty(
    "dag {
    U [unobserved]
    X -> Y
    X <- U <- A -> C -> Y 
    U -> B <- C
    }"
  )

adjustmentSets(dag_6.1, exposure = "X", outcome = "Y")
```

```{r}
adjustmentSets(dag_6.1, exposure = "X", outcome = "Y", type = "all")
```

### 6.4.3 Backdoor waffles
```{r}
dag_coords <-
  tibble(name = c("A", "D", "M", "S", "W"),
         x = c(1, 3, 2, 1, 3),
         y = c(1, 1, 2, 3, 3))

dagify( A ~ S,
        D ~ A + M + W,
        M ~ A + S,
        W ~ S,
        coords = dag_coords) %>% 
  gg_simple_dag()
```

```{r}
dag_6.2 <-
  dagitty(
    "dag {
    A -> D
    A -> M -> D
    A <- S -> M
    S -> W -> D
    }"
  )

adjustmentSets(dag_6.2, exposure = "W", outcome = "D")
```

## 6.5 Summary [and a little more practice]
```{r}
data(WaffleDivorce, package = "rethinking")
d <- WaffleDivorce

# standardize the continuous focal variable
d <-
  d %>% 
  mutate(a = rethinking::standardize(MedianAgeMarriage),
         d = rethinking::standardize(Divorce),
         m = rethinking::standardize(Marriage),
         s = factor(South, levels = 0:1, labels = c("North", "South")),
         w = rethinking::standardize(WaffleHouses))

# tidy up
rm(WaffleDivorce)
```

```{r}
ggpairs(data = d, columns = c(14:16, 18, 17),
        upper = list(continuous = wrap("cor", family = "sans", color = "black", size = 3)),
        diag = list(continuous = my_diag),
        lower = list(continuous = my_lower),
        mapping = aes(color = s)) +
  scale_fill_manual(values = c("forestgreen", "lightblue"))
```

```{r}
b6.13 <-
  brm(data = d,
      family = gaussian,
      d ~ 1 + w,
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.13")

b6.14 <-
  brm(data = d,
      family = gaussian,
      d ~ 1 + w + s,
      prior = c(prior(normal(0, .2), class = Intercept),
                prior(normal(0, .5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.14")

b6.15 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + w + a + m,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.15")

b6.16 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + w + a,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.16")

b6.17 <- 
  brm(data = d, 
      family = gaussian,
      d ~ 1 + w + m,
      prior = c(prior(normal(0, 0.2), class = Intercept),
                prior(normal(0, 0.5), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 6,
      file = "fits/b06.17")
```

```{r}
formula <- c("d ~ 1 + w", 
             "d ~ 1 + w + s", 
             "d ~ 1 + w + a + m", 
             "d ~ 1 + w + a", 
             "d ~ 1 + w + m")

tibble(fit = str_c("b6.1", 3:7)) %>% 
  mutate(y    = str_c(fit, " (", formula, ")"),
         post = purrr::map(fit, ~ get(.) %>% 
                               as_draws_df() %>% 
                               select(b_w))) %>% 
  unnest(post) %>% 
  
  ggplot(aes(x = b_w, y = y, color = fit %in% c("b6.14", "b6.15"))) +
  stat_pointinterval(.width = .95) +
  scale_color_manual(values = c("grey50", "forestgreen")) +
  labs(x = expression(beta[w]),
       y = NULL) +
  coord_cartesian(xlim = c(-0.4, 0.6)) +
  theme(axis.text.y = element_text(hjust = 0),
        legend.position = "none")
```

