---
title: "10 Big Entropy and the Generalized Linear Model"
format: html
---

## 10.1 Maximum entropy
```{r}
library(tidyverse)

d <-
  tibble(a = c(0, 0, 10, 0, 0),
         b = c(0, 1, 8, 1, 0),
         c = c(0, 2, 6, 2, 0),
         d = c(1, 2, 4, 2, 1),
         e = 2)

# this is our analogue to McElreath's 'lapply()' code
d %>% 
  mutate_all(~ . / sum(.)) %>% 
  # the next few lines constitute our analogue to his sapply() code
  pivot_longer(everything(), names_to = "plot") %>% 
  group_by(plot) %>% 
  summarise(h = -sum(ifelse(value == 0, 0, value * log(value))))
```
```{r}
library(ghibli)

ghibli_palette("MarnieMedium1")

ghibli_palette("MarnieMedium1")[1:7]
```

```{r}
d %>% 
  mutate(bucket = 1:5) %>% 
  pivot_longer(-bucket,
               names_to = "letter",
               values_to = "pebbles") %>% 
  
  ggplot(aes(x = bucket, y = pebbles)) +
  geom_col(width = 1/5, fill = ghibli_palette("MarnieMedium1")[2]) +
  geom_text(aes(y = pebbles + 1, label = pebbles)) +
  geom_text(data = tibble(
    letter  = letters[1:5],
    bucket  = 5.5,
    pebbles = 10.5,
    label   = str_c(c(1, 90, 1260, 37800, 113400), 
                    rep(c(" way", " ways"), times = c(1, 4)))),
    aes(label = label), 
    hjust = 1) +
  scale_y_continuous(breaks = c(0, 5, 10), limits = c(0, 12)) +
  theme(panel.background = element_rect(fill = ghibli_palette("MarnieMedium1")[6]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("MarnieMedium1")[7])) +
  facet_wrap(~ letter, ncol = 2)
```

We might plot our version of the final panel like so
```{r}
d %>% 
  # the next four lines are the same from above
  mutate_all(~ . / sum(.)) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  summarise(h = -sum(ifelse(value == 0, 0, value * log(value)))) %>% 
  # here's the R code 9.4 stuff
  mutate(n_ways = c(1, 90, 1260, 37800, 113400)) %>% 
  group_by(name) %>% 
  mutate(log_ways = log(n_ways) / 10,
         text_y = ifelse(name < "c", h + .15, h - .15)) %>% 
  
  # plot
  ggplot(aes(x = log_ways, y = h)) +
  geom_abline(intercept = 0, slope = 1.37, color = "white") +
  geom_point(size = 2.5, color = ghibli_palette("MarnieMedium1")[7]) +
  geom_text(aes(y = text_y, label = name)) +
  labs(x = "log (ways) per pebble",
       y = "entropy") +
  theme(panel.background = element_rect(fill = ghibli_palette("MarnieMedium1")[6]),
        panel.grid = element_blank())
```

### 10.1.1 Gaussian.
```{r}
alpha_per_beta <- function(beta, variance = 1){
  sqrt((variance * gamma(1 / beta)) / gamma(3 / beta))
}

crossing(value = seq(from = -5, to = 5, by = .1),
         # I arrived at these values by trial and error
         beta = c(1, 1.5, 2, 4)) %>% 
  mutate(mu = 0,
         alpha = alpha_per_beta(beta)) %>% 
  # behold the formula for the generalized normal distribution in code!
  mutate(density = (beta / (2 * alpha * gamma(1 / beta))) *
           exp(1) ^ (-1 * (abs(value - mu) / alpha) ^ beta)) %>% 
  
  # plot
  ggplot(aes(x = value, y = density, group = beta)) +
  geom_line(aes(color = beta == 2, size = beta == 2)) +
  scale_color_manual(values = c(ghibli_palette("MarnieMedium2")[c(2, 4)])) +
  scale_size_manual(values = c(1/4, 1.25)) +
  labs(subtitle = "Guess which color denotes the Gaussian.") +
  coord_cartesian(xlim = c(-4, 4)) +
  theme(legend.position = "none",
        panel.background = element_rect(fill = ghibli_palette("MarnieMedium2")[7]),
        panel.grid = element_blank())
```

Once you have $\alpha$ and $\beta$, the entropy equation for the generalized normal distribution is
$$
\text{entropy} = \frac{1}{\beta} - \text{log}\left[\frac{\beta}{2\alpha\Gamma(1/\beta)}\right]
$$

Here's how we can use that equation to make our version of right panel
```{r}
tibble(beta = seq(from = 1, to = 4, length.out = 100)) %>% 
  mutate(alpha = alpha_per_beta(beta)) %>% 
  mutate(entropy = (1 / beta) - log((beta) / (2 * alpha * gamma(1 / beta)))) %>% 
  
  ggplot(aes(x = beta, y = entropy)) +
  geom_vline(xintercept = 2, color = "white") +
  geom_line(linewidth = 2, color = ghibli_palette("MarnieMedium2")[6]) +
  xlab(expression(beta~(i.e.*", "*shape))) +
  theme(panel.background = element_rect(fill = ghibli_palette("MarnieMedium2")[7]),
        panel.grid = element_blank())

```

### 10.1.2 Binomial
The binomial likelihood entails
$$
\text{Pr}(y|n, p) = \frac{n!}{y!(n - y)!}p^y(1 - p)^{n - y}
$$

```{r}
count_ways <- function(n, y){
  # n = the total number of trials (i.e., the number of rows in your vector)
  # y = the total number of 1s (i.e., successes) in your vector
  (factorial(n) / factorial(y) * factorial(n - y))
}
```
Now consider three sequences:
- 0, 0, 0, 0 (i.e., n = 4 and y = 0)
- 1, 0, 0, 0 (i.e., n = 4 and y = 1)
- 1, 1, 0, 0 (i.e., n = 4 and y = 2)

We can organize that information in a little tibble and then demo our count_ways() function
```{r}
tibble(sequence = 1:3,
       n = 4,
       y = c(0, 1, 2)) %>% 
  mutate(n_ways = count_ways(n = n, y = y))
```
Here's the pre-Figure 10.3 data McElreath presented on page 308.
```{r}
# data
d <-
  tibble(distribution = letters[1:4],
         ww = c(1/4, 2/6, 1/6, 1/8),
         bw = c(1/4, 1/6, 2/6, 4/8),
         wb = c(1/4, 1/6, 2/6, 2/8),
         bb = c(1/4, 2/6, 1/6, 1/8))

# table
d %>% 
  mutate_if(is.numeric, ~MASS::fractions(.) %>% as.character()) %>% 
  flextable::flextable()
```
Those data take just a tiny bit of wrangling before they're ready to plot in our version of Figure 10.3
```{r}
d <-
  d %>% 
  pivot_longer(-distribution,
               names_to = "sequence",
               values_to = "probability") %>% 
  mutate(sequence = factor(sequence, levels = c("ww", "bw", "wb", "bb")))

d %>% 
  ggplot(aes(x = sequence, y = probability, group = 1)) +
  geom_point(size = 2, color = ghibli_palette("PonyoMedium")[4]) +
  geom_line(color = ghibli_palette("PonyoMedium")[5]) +
  labs(x = NULL, y = NULL) +
  coord_cartesian(ylim = 0:1) +
  theme(axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = ghibli_palette("PonyoMedium")[2]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("PonyoMedium")[6])) +
  facet_wrap(~ distribution)
```

If we go step by step, we might count the expected value for each distribution like follows.
```{r}
d %>% 
  # str_count() will count the number of times "b" occurs within a given row of sequence
  mutate(n_b = str_count(sequence, "b")) %>% 
  mutate(product = probability * n_b) %>% 
  group_by(distribution) %>% 
  summarise(expected_value = sum(product))
```

We can use the same group_by strategy on the way to computing the entropy values
```{r}
d %>% 
  group_by(distribution) %>% 
  summarise(entropy = -sum(probability * log(probability)))
```

I’m going to alter McElreath’s simulation function from R code 10.9 to take a seed argument. In addition, I altered the names of the objects within the function and changed the output to a tibble that will also include the conditions “ww”, “bw”, “wb”, and “bb”.
```{r}
sim_p <- function(seed, g = 1.4){
  
  set.seed(seed)
  
  x_123 <- runif(3)
  x_4 <- ((g) * sum(x_123) - x_123[2] - x_123[3]) / (2 - g)
  z <- sum(c(x_123, x_4))
  p <- c(x_123, x_4) / z
  
  tibble(h = -sum(p * log(p)),
         p = p,
         key = factor(c("ww", "bw", "wb", "bb"), 
                      levels = c("ww", "bw", "wb", "bb")))
}
```

```{r}
sim_p(seed = 9.9, g = 1.4)
```

```{r}
# how many replications would you like?
n_rep <- 1e5

d <-
  tibble(seed = 1:n_rep) %>% 
  mutate(sim = map2(seed, 1.4, sim_p)) %>% 
  unnest(sim)
```

```{r}
head(d)
```

In order to intelligently choose which four replications we want to highlight in Figure 10.4, we’ll want to rank order them by entropy, h.
```{r}
ranked_d <-
  d %>% 
  group_by(seed) %>% 
  arrange(desc(h)) %>% 
  ungroup() %>%
  # here's the rank order step
  mutate(rank = rep(1:n_rep, each = 4))

head(ranked_d)
```

```{r}
subset_d <-
  ranked_d %>%
  # I arrived at these `rank` values by trial and error
  filter(rank %in% c(1, 87373, n_rep - 1500, n_rep - 10)) %>% 
  # I arrived at the `height` values by trial and error, too
  mutate(height       = rep(c(8, 2.25, .75, .5), each = 4),
         distribution = rep(letters[1:4], each = 4))
head(subset_d)
```

```{r}
p1 <-
  d %>% 
  ggplot(aes(x = h)) +
  geom_density(linewidth = 0, fill = ghibli_palette("LaputaMedium")[3],
               adjust = 1/4) +
  # note the data statements for the next two geoms
  geom_linerange(data = subset_d %>% group_by(seed) %>% slice(1),
                 aes(ymin = 0, ymax = height),
                 color = ghibli_palette("LaputaMedium")[5]) +
  geom_text(data = subset_d %>% group_by(seed) %>% slice(1),
            aes(y = height + .5, label = distribution)) +
  scale_x_continuous("Entropy", breaks = seq(from = .7, to = 1.2, by = .1)) +
  theme(panel.background = element_rect(fill = ghibli_palette("LaputaMedium")[7]),
        panel.grid = element_blank())
```

```{r}
p2 <-
  ranked_d %>%
  filter(rank %in% c(1, 87373, n_rep - 1500, n_rep - 10)) %>% 
  mutate(distribution = rep(letters[1:4], each = 4)) %>% 
  
  ggplot(aes(x = key, y = p, group = 1)) +
  geom_line(color = ghibli_palette("LaputaMedium")[5]) +
  geom_point(size = 2, color = ghibli_palette("LaputaMedium")[4]) +
  scale_y_continuous(NULL, breaks = NULL, limits = c(0, .75)) +
  xlab(NULL) +
  theme(axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = ghibli_palette("LaputaMedium")[7]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("LaputaMedium")[6])) +
  facet_wrap(~ distribution)

# combine and plot
library(patchwork)
p1 | p2
```

```{r}
ranked_d %>% 
  group_by(key) %>% 
  arrange(desc(h)) %>% 
  slice(1)
```

## 10.2 Generalized linear models
```{r}
tibble(x = seq(from = -1, to = 3, by = .01)) %>% 
  mutate(probability = .35 + x * .5) %>% 
  
  ggplot(aes(x = x, y = probability)) +
  geom_rect(xmin = -1, xmax = 3,
            ymin = 0, ymax = 1,
            fill = ghibli_palette("MononokeMedium")[5]) +
  geom_hline(yintercept = 0:1, linetype = 2,
             color = ghibli_palette("MononokeMedium")[7]) +
  geom_line(aes(linetype = probability > 1, 
                color = probability > 1),
            linewidth = 1) +
  geom_segment(x = 1.3, xend = 3,
               y = 1, yend = 1,
               linewidth = 2/3,
               color = ghibli_palette("MononokeMedium")[3]) +
  annotate(geom = "text",
           x = 1.28, y = 1.04, hjust = 1,
           label = "This is why we need link functions",
           color = ghibli_palette("MononokeMedium")[4], 
           size = 2.6) +
  scale_color_manual(values = c(ghibli_palette("MononokeMedium")[3:4])) +
  scale_y_continuous(breaks = c(0, .5, 1)) +
  coord_cartesian(xlim = c(0, 2),
                  ylim = c(0, 1.2)) +
  theme(legend.position = "none",
        panel.background = element_rect(fill = ghibli_palette("MononokeMedium")[1]),
        panel.grid = element_blank())
```

### 10.2.1 Meet the family
```{r}
length_out <- 100

tibble(x = seq(from = 0, to = 5, length.out = length_out)) %>% 
  mutate(Gamma = dgamma(x, 2, 2),
         Exponential = dexp(x)) %>% 
  pivot_longer(-x, values_to = "density") %>% 
  mutate(label = ifelse(name == "Gamma",
                        "y %~% Gamma(lambda, kappa)",
                        "y %~% Exponential(lambda)")) %>% 
  
  ggplot(aes(x = x, y = density)) +
  geom_area(fill = ghibli_palette("SpiritedMedium")[3]) +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 4)) +
  theme(panel.background = element_rect(fill = ghibli_palette("SpiritedMedium")[5]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("SpiritedMedium")[7])) +
  facet_wrap(~ label, scales = "free_y", labeller = label_parsed)
```
Gaussian
```{r}
tibble(x = seq(from = -5, to = 5, length.out = length_out)) %>%
  mutate(density = dnorm(x),
         strip = "y %~% Normal(mu, sigma)") %>% 
  
  ggplot(aes(x = x, y = density)) +
  geom_area(fill = ghibli_palette("SpiritedMedium")[3]) +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(-4, 4)) +
  theme(panel.background = element_rect(fill = ghibli_palette("SpiritedMedium")[5]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("SpiritedMedium")[7])) +
  facet_wrap(~ strip, labeller = label_parsed)
```

Poisson
```{r}
tibble(x = 0:20) %>% 
  mutate(density = dpois(x, lambda = 2.5),
         strip   = "y %~% Poisson(lambda)") %>% 
  
  ggplot(aes(x = x, y = density)) +
  geom_col(fill = ghibli_palette("SpiritedMedium")[2], width = 1/2) +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 10)) +
  theme(panel.background = element_rect(fill = ghibli_palette("SpiritedMedium")[5]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("SpiritedMedium")[7])) +
  facet_wrap(~ strip, labeller = label_parsed)
```

Binomial
```{r}
tibble(x = 0:10) %>% 
  mutate(density = dbinom(x, size = 10, prob = .85),
         strip   = "y %~% Binomial(n, p)") %>% 
  
  ggplot(aes(x = x, y = density)) +
  geom_col(fill = ghibli_palette("SpiritedMedium")[2], width = 1/2) +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 10)) +
  theme(panel.background = element_rect(fill = ghibli_palette("SpiritedMedium")[5]),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = ghibli_palette("SpiritedMedium")[7])) +
  facet_wrap(~ strip, labeller = label_parsed)
```

