---
title: "Homework - Week 1"
format: html
---

### Question 1
Suppose the globe tossing data had turned out to be 3 water and 11 land. Construct the posterior distribution

Gharda's code version
```{r}
library(ggplot2)

# create p_grid
p_grid <- seq(from = 0, to = 1, length.out = 100)

# create prior distributoin
prior <- rep(1, 20)

# compute likelihood at each value in the grid
# data for globe tossing with 3 water and 11 land
likelihood <- dbinom(3, size = 14, prob = p_grid)

# calculate posterior 
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)

ggplot(mapping = aes(x = p_grid, y = posterior)) +
  geom_line() +
  theme_minimal()
```
Homework solution's version 1
```{r}
library(ggplot2)

compute_posterior <- function(W, L, poss = c(0, 0.25, 0.5, 0.75, 1)){
  ways <- sapply(poss, function(q) q^W * (1-q)^L)
  post <- ways / sum(ways)
  data.frame(poss, ways, post = round(post, 3))
}

p_grid <- seq(from = 0, to = 1, len = 100)
post <- compute_posterior(3, 11, poss=p_grid)

ggplot(mapping = aes(x = p_grid, y = post$post)) +
  geom_line() +
  theme_minimal()
```
Homework solution's version 2
```{r}
curve(dbeta(x, 3+1, 11+1), from = 0, to = 1, xlab = "p")
```

### Question 2
Using the posterior distribution from 1, compute the posterior predictive distribution for the next 5 tosses of the same globe. I recommend you use the sampling method.

