---
title: "Homework Week 5"
format: html
---

```{r}
rm(list = ls())
library(tidyverse)
library(tidybayes)
library(brms)
library(dagitty)
library(ggdag)
library(ggrepel)
```

1. NWO Grants total causal effect
```{r}
data(NWOGrants, package = "rethinking")

nwo_dat <-
  NWOGrants %>% 
  mutate(gender = factor(gender, levels = c("m", "f")))

rm(NWOGrants)
```

```{r}
nwo_dag <- dagitty("dag { G -> D -> A <- G }")
coordinates(nwo_dag) <- list(
  x = c(G = 1, D = 2, A = 3),
  y = c(G = 1, D = 2, A = 1)
)
ggdag(nwo_dag)

adjustmentSets(nwo_dag, exposure = "G", outcome = "A")
```

```{r}
w5h1 <- 
  brm(data = nwo_dat,
      family = binomial,
      awards | trials(applications) ~ 0 + gender,
      prior = c(prior(normal(0, 1.5), class = b)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w5h1")
```

```{r}
summary(w5h1)
```

From the summary, it appears that females are less likely to be awarded grants, but we need to compute the contrast to know for certain. The plot below shows the contrast for males compared to females. On average, males are favored by about 3 percentage points. The contrast is faily reliably above 0, indicating some bias in favor of males.

```{r}
as_draws_df(w5h1) %>% 
  mutate(diff_male = inv_logit_scaled(b_genderm) - 
           inv_logit_scaled(b_genderf)) %>% 
  ggplot(aes(x = diff_male)) +
  stat_halfeye(.width = c(0.67, 0.89, 0.97),
               fill = "#009FB7") +
  labs(x = "beta_male - beta_female", y = "Density") +
  theme_minimal()
```
2. NWO Grants direct causal effect
```{r}
adjustmentSets(nwo_dag, exposure = "G", outcome = "A",
               effect = "direct")
```


```{r}
w5h2 <-
  brm(data = nwo_dat,
      family = binomial,
      bf(awards | trials(applications) ~ g + d + i,
         g ~ 0 + gender,
         d ~ 0 + discipline,
         i ~ 0 + gender:discipline,
         nl = TRUE),
      prior = c(prior(normal(0, 1.5), nlpar = g),
                prior(normal(0, 1.5), nlpar = d),
                prior(normal(0, 1.5), nlpar = i)),
      iter = 4000, warmup = 2000, chains = 4, cores = 4,
      seed = 1234,
      file = "fits/w5h2")
```

```{r}
summary(w5h2)
```

```{r}
apps_per_dept <-
  nwo_dat %>% 
  group_by(discipline) %>% 
  summarize(applications = sum(applications))

male_dat <-
  apps_per_dept %>% 
  mutate(gender = "m") %>% 
  uncount(applications) %>% 
  mutate(applications = 1L)

female_dat <-
  apps_per_dept %>% 
  mutate(gender = "f") %>% 
  uncount(applications) %>% 
  mutate(applications = 1L)
```

```{r}
marg_eff <-
  bind_rows(add_epred_draws(male_dat, w5h2),
            add_epred_draws(female_dat, w5h2)) %>% 
  pivot_wider(names_from = "gender", values_from = ".epred") %>% 
  mutate(diff = m - f)
```


```{r}
mean_qi(marg_eff$diff, .width = c(0.67, 0.89, 0.97))
```

```{r}
ggplot(marg_eff, aes(x = diff)) +
  stat_halfeye(.width = c(0.67, 0.89, 0.97), fill = "#009FB7") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Difference in Awards (Male - Female)",
       y = "Density") +
  theme_minimal()
```
3. Causes to the average difference of awards between women and men
```{r}
nwo_dat %>% 
  group_by(discipline) %>% 
  summarize(f = sum(applications[which(gender == "f")]),
            m = sum(applications[which(gender == "m")]),
            total_apps = sum(applications),
            total_awards = sum(awards)) %>% 
  mutate(female_pct = f / sum(f),
         male_pct = m / sum(m),
         award_pct = total_awards / total_apps) %>% 
  ggplot(aes(x = female_pct, y = male_pct)) +
  geom_point(aes(size = award_pct, color = abs(female_pct - male_pct) > 0.05)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_label_repel(data = ~filter(.x, abs(female_pct - male_pct) > 0.05),
                   aes(label = discipline),
                   max.overlaps = Inf, nudge_y = 0.03) +
  scale_size(breaks = seq(0.1, 0.3, by = 0.04)) +
  scale_color_manual(values = c("black", "#009FB7")) +
  guides(color = "none") +
  expand_limits(x = c(0, 0.35), y = c(0, 0.35)) +
  coord_equal() +
  labs(x = "Female Application Rate", y = "Male Application Rate", size = "Award Rate") +
  theme(legend.position = "right") +
  theme_minimal()
```

4. UFC lefties
```{r}
data(UFClefties, package = "rethinking")

ufc_dat <-
  UFClefties %>% 
  select(fighter1.win, fighter1.lefty, fighter2.lefty) 

rm(UFClefties)
```

```{r}
ufc_dag <- dagitty("dag { L -> Q <- A -> W <- L }")
coordinates(ufc_dag) <- list(
  x = c(Q = 1, L = 2, W = 3, A = 2),
  y = c(Q = 1, L = 1, W = 1, A = 2)
)
ggdag(ufc_dag) +
  theme_bw() +
  coord_cartesian(ylim = c(0.85, 2.15))

# adjustmentSets(ufc_dag, exposure = "L", outcome = "W")
```

Create simulated data
```{r}
set.seed(221)

n <- 5000
L <- as.integer(rbernoulli(n, 0.1))
A <- rnorm(n)

# qualify for UFC if high A, or lower A but left handed
Q <- ifelse(A > 2 | (A > 1.25 & L == 1L), 1L, 0L)

# filter to only qualified individuals
ufc_fighters <- tibble(l = L, a = A, q = Q) %>% 
  filter(q == 1L) %>% 
  select(-q) %>% 
  rowid_to_column(var = "fighter_id")

# create data
k <- 2.0    # importance of ability difference
b <- 0.5    # left-handedness advantage

ufc_sim <- 
  tibble(fighter_1 = ufc_fighters$fighter_id[ufc_fighters$fighter_id %% 2 == 1],
         fighter_2 = ufc_fighters$fighter_id[ufc_fighters$fighter_id %% 2 == 0]) %>% 
  left_join(ufc_fighters, by = c("fighter_1" = "fighter_id")) %>% 
  left_join(ufc_fighters, by = c("fighter_2" = "fighter_id")) %>% 
  rename(l1 = l.x, a1 = a.x, l2 = l.y, a2 = a.y) %>% 
  mutate(score1 = a1 + b * l1,
         score2 = a2 + b * l2,
         p = rethinking::inv_logit(k * (score1 - score2)),
         fighter1_win = as.integer(rbernoulli(1, p = p))) %>% 
  select(fighter_1, fighter_2, l1, l2, fighter1_win)
```

Fit the model and see the effect of left-handedness. Overall, we estimate a slight disadvantage of being left-handed, even though the data was simulated to have an advantage of 0.5. This is our collider bias in action.

```{r}
w5h4 <- 
  brm(
    data = ufc_sim, 
    family = bernoulli,
    bf(fighter1_win ~ 0 + b * (l1 - l2),
       b ~ 1,
       nl = TRUE),
    prior = c(prior(normal(0, 0.5), class = b, nlpar = b)),
    iter = 4000, warmup = 2000, chains = 4, cores = 4,
    seed = 1234,
    file = "fits/w5h4")
```

```{r}
summary(w5h4)
```


```{r}
as_draws_df(w5h4) %>% 
  ggplot(aes(x = b_b_Intercept)) +
  stat_halfeye(.width = 0.89, fill = "#009FB7") +
  labs(x = "beta_L", y = "Density") +
  theme_minimal()
```

