---
title: "12 Monsters and Mixtures"
format: html
---

## 12.1 Over-dispersed counts
### 12.1.1 Beta-binomial
```{r}
library(ggthemes)
```

```{r}
scales::show_col(canva_pal("Green fields")(4))
```

```{r}
canva_pal("Green fields")(4)
```

```{r}
canva_pal("Green fields")(4)[3]
```

```{r}
library(tidyverse)

theme_set(
  theme_hc() +
    theme(axis.ticks.y = element_blank(),
          plot.background = element_rect(fill = "grey92"))
)

pbar <- .5
theta <- 5

ggplot(data = tibble(x = seq(from = 0, to = 1, by = .01)),
       aes(x = x, y = rethinking::dbeta2(x, prob = pbar, theta = theta))) +
  geom_area(fill = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous("probability space", breaks = c(0, .5, 1)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle(expression(The~beta~distribution),
          subtitle = expression("Defined in terms of "*mu*" (i.e., pbar) and "*kappa*" (i.e., theta"))
```

```{r}
betaABfromMeanKappa <- function(mean, kappa) {
  if (mean <= 0 | mean >= 1) stop("must have 0 < mean < 1")
  if (kappa <= 0) stop("kappa must be > 0")
  a <- mean * kappa
  b <- (1.0 - mean) * kappa
  return(list(a = a, b = b))
}
```

```{r}
betaABfromMeanKappa(mean = pbar, kappa = theta)
```
Check that all of this works
```{r}
ggplot(data = tibble(x = seq(from = 0, to = 1, by = .01)),
       aes(x = x, y = dbeta(x, shape1 = 2.5, shape2 = 2.5))) +
  geom_area(fill = canva_pal("Green fields")(4)[4]) +
  scale_x_continuous("probability space", breaks = c(0, .5, 1)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle(expression(The~beta~distribution),
          subtitle = expression("This time defined in terms of "*alpha*" and "*beta))
```

```{r}
crossing(pbar = c(.25, .5, .75),
         theta = c(5, 15, 30)) %>% 
  expand_grid(x = seq(from = 0, to = 1, length.out = 100)) %>% 
  mutate(density = rethinking::dbeta2(x, prob = pbar, theta = theta),
         mu = str_c("mu == ", pbar %>% str_remove(., "0")),
         kappa = factor(str_c("kappa == ", theta),
                        levels = c("kappa == 30", "kappa == 15", "kappa == 5"))) %>% 
  
  ggplot(aes(x = x, y = density)) +
  geom_area(fill = canva_pal("Green fields")(4)[4]) +
  scale_x_continuous("probability space",
                     breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) +
  scale_y_continuous(NULL, labels = NULL) +
  theme(axis.ticks.y = element_blank()) +
  facet_grid(kappa ~ mu, labeller = label_parsed)
```

The statistical model we'll be fitting follows the form
$$
\begin{align}
\text{admit}_i & \sim \text{BetaBinomial}(n_i, \bar{p}_i, \phi) \\
\text{logit}_i & = \alpha_{gid[i]} \\
\alpha_j & \sim \text{Normal}(0, 1.5) \\
\phi & \sim \text{Exponential}(1).
\end{align}
$$

Load the data and brms
```{r}
data(UCBadmit, package = "rethinking")
d <-
  UCBadmit %>% 
  mutate(gid = ifelse(applicant.gender == "male", "1", "2"))

rm(UCBadmit)

library(brms)
```

1. Create custom_family() function to define the name and parameters of the beta-binomial distribution for use in brm()
2. Define some functions for Stan which are not defined in Stan itself. We'll save them as stan_funs.
3. We'll make a stanvar() statement which will allow us to pass our stan_funs to brm()
```{r}
beta_binomial2 <- custom_family(
  "beta_binomial2", dpars = c("mu", "phi"),
  links = c("logit", "log"),
  lb = c(0, 2), ub = c(1, NA),
  type = "int", vars = "vint1[n]"
)

stan_funs <- "
  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
  }
  int beta_binomial2_rng(real mu, real phi, int T) {
    return beta_binomial_rng(T, mu * phi, (1 - mu) * phi);
  }
"

stanvars <- stanvar(scode = stan_funs, block = "functions")
```

```{r}
b12.1 <-
  brm(data = d,
      family = beta_binomial2, # here's our custom likelihood
      admit | vint(applications) ~ 0 + gid,
      prior = c(prior(normal(0, 1.5), class = b),
                prior(exponential(1), class = phi)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      stanvars = stanvars, # note our stanvars
      seed = 12,
      file = "fits/b12.01"
      )
```

```{r}
print(b12.1)
```

```{r}
post <- as_draws_df(b12.1)
head(post)
```

```{r}
library(tidybayes)

post %>% 
  transmute(da = b_gid1 - b_gid2) %>% 
  mean_qi(.width = .89) %>% 
  mutate_if(is.double, round, digits = 3)
```

```{r}
set.seed(12)

lines <-
  post %>% 
  mutate(p_bar = inv_logit_scaled(b_gid2)) %>% 
  slice_sample(n = 100) %>% 
  select(.draw, p_bar, phi) %>% 
  expand_grid(x = seq(from = 0, to = 1, by = .005)) %>% 
  mutate(density = pmap_dbl(list(x, p_bar, phi), rethinking::dbeta2))

str(lines)
```

```{r}
lines %>% 
  ggplot(aes(x = x, y = density)) +
  stat_function(fun = rethinking::dbeta2,
                args = list(prob = mean(inv_logit_scaled(post %>% pull(b_gid2))),
                            theta = mean(post %>% pull(phi))),
                linewidth = 1.5,
                color = canva_pal("Green fields")(4)[4]) +
  geom_line(aes(group = .draw),
            alpha = .2, color = canva_pal("Green fields")(4)[4]) +
  scale_y_continuous(NULL, breaks = NULL, limits = c(0, 3)) +
  labs(subtitle = "distribution of female admission rates",
       x = "probability admit")
```

```{r}
expose_functions(b12.1, vectorize = TRUE)

# required to use predict()
log_lik_beta_binomial2 <- function(i, prep) {
  mu <- brms::get_dpar(prep, "mu", i = i)
  phi <- brms::get_dpar(prep, "phi", i = i)
  trials <- prep$data$vint1[i]
  y <- prep$da$Y[i]
  beta_binomial2_lpmf(y, mu, phi, trials)
}

posterior_predict_beta_binomial2 <- function(i, prep, ...) {
  mu <- brms::get_dpar(prep, "mu", i = i)
  phi <- brms::get_dpar(prep, "phi", i = i)
  trials <- prep$data$vint1[i]
  beta_binomial2_rng(mu, phi, trials)
}

# required to use fitted()
posterior_epred_beta_binomial2 <- function(prep) {
  mu <- brms::get_dpar(prep, "mu")
  trials <- prep$data$vint1
  trials <- matrix(trials, nrow = nrow(mu), ncol = ncol(mu), byrow = TRUE)
  mu * trials
}
```

```{r}
# the prediction intervals
predict(b12.1) %>% 
  data.frame() %>% 
  transmute(ll = Q2.5,
            ul = Q97.5) %>% 
  bind_cols(
    # the fitted intervals
    fitted(b12.1) %>% data.frame(),
    # the original data used to fit the model
    b12.1$data
  ) %>% 
  mutate(case = 1:12) %>% 
  
  # plot!
  ggplot(aes(x = case)) +
  geom_linerange(aes(ymin = ll / applications,
                     ymax = ul / applications),
                 color = canva_pal("Green fields")(4)[1],
                 linewidth = 2.5, alpha = 1/4) +
  geom_pointrange(aes(ymin = Q2.5 / applications,
                      ymax = Q97.5 / applications,
                      y = Estimate / applications),
                  color = canva_pal("Green fields")(4)[4],
                  linewidth = 1/2, shape = 1) +
  geom_point(aes(y = admit / applications),
             color = canva_pal("Green fields")(4)[2],
             size = 2) +
  scale_x_continuous(breaks = 1:12) +
  scale_y_continuous(breaks = 0:5 / 5, limits = c(0, 1)) +
  labs(subtitle = "Posterior validation check",
       caption = expression(italic(Note.)*" A = admittance probability"),
       y = "A") +
  theme(axis.ticks.x = element_blank(),
        legend.position = "none")
```

### 12.1.2 Negative-binomial or gamma-Poisson.
```{r}
data(Kline, package = "rethinking")
d <-
  Kline %>% 
  mutate(p = rethinking::standardize(log(population)),
         contact_id = ifelse(contact == "high", 2L, 1L),
         cid = contact)
rm(Kline)

print(d)
```

```{r}
get_prior(data = d,
          family = negbinomial,
          total_tools ~ 1)
```

```{r}
ggplot(data = tibble(x = seq(from = 0, to = 60, by = .1)),
       aes(x = x, y = dgamma(x, shape = .01, rate = .01))) +
  geom_area(color = "transparent",
            fill = canva_pal("Green fields")(4)[2]) +
  scale_x_continuous(NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 50)) +
  ggtitle(expression(brms~default~gamma(0.01*", "*0.01)~shape~prior))
```

Let's fit the model
```{r}
b12.2a <-
  brm(data = d,
      family = negbinomial,
      total_tools ~ 1,
      prior = c(prior(normal(3, 0.5), class = Intercept), # beta_0
                prior(gamma(.01, .01), class = shape)), # alpha
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.02a")
```

```{r}
print(b12.2a)
```

```{r}
p <-
  predict(b12.2a, summary = F)

p %>% str()
```

```{r}
p %>% 
  data.frame() %>% 
  set_names(d$culture) %>% 
  pivot_longer(everything(),
               names_to = "culture",
               values_to = "lambda") %>% 
  
  ggplot(aes(x = lambda)) +
  geom_density(color = "transparent", fill = canva_pal("Green fields")(4)[2]) +
  scale_x_continuous(expression(lambda["[culture]"]),
                     breaks = 0:2 * 100) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 210)) +
  facet_wrap(~ culture, nrow = 2)
```

```{r}
post <- as_draws_df(b12.2a)

post %>% 
  mutate(mu = exp(b_Intercept),
         alpha = shape) %>% 
  pivot_longer(mu:alpha, names_to = "parameter") %>% 
  
  ggplot(aes(x = value)) +
  geom_density(color = "transparent",
               fill = canva_pal("Green fields")(4)[2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Behold our gamma parameters!",
       x = "posterior") +
  facet_wrap(~ parameter, scales = "free", labeller = label_parsed)
```

```{r}
post %>% 
  mutate(mu = exp(b_Intercept),
         alpha = shape) %>% 
  mutate(theta = mu / alpha) %>% 
  
  ggplot(aes(x = theta)) +
  geom_density(color = "transparent",
               fill = canva_pal("Green fields")(4)[2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = expression(We~define~the~scale~as~theta == mu/alpha),
       x = "posterior") +
  coord_cartesian(xlim = c(0, 40))
```

```{r}
set.seed(12)

# wrangle to get 200 draws
post %>% 
  mutate(alpha = shape,
         theta = exp(b_Intercept) / shape) %>%
  slice_sample(n = 200) %>% 
  expand_grid(x = 0:250)%>% 
  mutate(density = dgamma(x, shape = alpha, scale = theta)) %>% 
  
  # plot
  ggplot(aes(x = x, y = density)) +
  geom_line(aes(group = .draw),
            alpha = .1, color = canva_pal("Green fields")(4)[4]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = expression("200 credible gamma densities for "*lambda),
       x = expression(lambda)) +
  coord_cartesian(xlim = c(0, 170),
                  ylim = c(0, 0.045))
```

```{r}
b12.2b <-
  brm(data = d, 
      family = negbinomial(link = "identity"),
      bf(total_tools ~ exp(b0) * population^b1 / g,
         b0 + b1 ~ 0 + cid,
         g ~ 1,
         nl = TRUE),
      prior = c(prior(normal(1, 1), nlpar = b0),
                prior(exponential(1), nlpar = b1, lb = 0),
                prior(exponential(1), nlpar = g, lb = 0),
                prior(exponential(1), class = shape)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.02b") 
```

```{r}
print(b12.2b)
```

```{r}
b12.2b <- add_criterion(b12.2b, "loo")
loo(b12.2b)
```

```{r}
d %>% 
  mutate(k = b12.2b$criteria$loo$diagnostics$pareto_k) %>% 
  dplyr::filter(k > .5) %>% 
  select(culture, k)
```

```{r}
b11.11 <- readRDS("fits/b11.11.rds")
```

```{r}
# the new data
nd <-
  distinct(d, cid) %>% 
  expand_grid(population = seq(from = 0, to = 300000, length.out = 100))

p1 <-
  # compute the expected trajectories
  fitted(b11.11,
         newdata = nd,
         probs = c(.055, .945)) %>%
  data.frame() %>%
  bind_cols(nd) %>%
  
  # plot
  ggplot(aes(x = population, group = cid, color = cid)) +
  geom_smooth(aes(y = Estimate, ymin = Q5.5, ymax = Q94.5, fill = cid),
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(data = bind_cols(d, b11.11$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k),
             alpha = 4/5) +
  labs(subtitle = "pure Poisson model",
       y = "total tools")
```

```{r}
# for the annotation
text <-
  distinct(d, cid) %>% 
  mutate(population  = c(150000, 110000),
         total_tools = c(57, 69),
         label       = str_c(cid, " contact"))

p2 <-
  fitted(b12.2b,
         newdata = nd,
         probs = c(.055, .945)) %>%
  data.frame() %>%
  bind_cols(nd) %>%
  
  ggplot(aes(x = population, group = cid, color = cid)) +
  geom_smooth(aes(y = Estimate, ymin = Q5.5, ymax = Q94.5, fill = cid),
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(data = bind_cols(d, b12.2b$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k),
             alpha = 4/5) +
  geom_text(data = text,
            aes(y = total_tools, label = label)) +
  scale_y_continuous(NULL, labels = NULL) +
  labs(subtitle = "gamma-Poisson model")
```

```{r}
library(patchwork)

(p1 | p2) &
  scale_fill_manual(values = canva_pal("Green fields")(4)[c(4, 1)]) &
  scale_color_manual(values = canva_pal("Green fields")(4)[c(4, 1)]) &
  scale_size(range = c(2, 5)) &
  scale_x_continuous("population", breaks = c(0, 50000, 150000, 250000)) &
  coord_cartesian(xlim = range(d$population),
                  ylim = range(d$total_tools)) &
  theme(axis.ticks = element_blank(),
        legend.position = "none")
```

```{r}
predict(b12.2b,
        summary = F) %>% 
  data.frame() %>% 
  set_names(d$culture) %>% 
  pivot_longer(everything(),
               names_to = "culture",
               values_to = "lambda") %>% 
  left_join(d, by = "culture") %>% 
  
  ggplot(aes(x = lambda, y = 0)) +
  stat_halfeye(point_interval = mean_qi, .width = .5,
               fill = canva_pal("Green fields")(4)[2],
               color = canva_pal("Green fields")(4)[1]) +
  geom_vline(aes(xintercept = total_tools),
             color = canva_pal("Green fields")(4)[3]) +
  scale_x_continuous(expression(lambda["[culture]"]), breaks = 0:2 * 100) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 210)) +
  facet_wrap(~ culture, nrow = 2)
```

### 12.1.3 Over-dispersion, entropy, and information criteria.

## 12.2 Zero-inflated outcomes
#### 12.2.0.1 Rethinking: Breaking the law.
### 12.2.1 Example: Zero-inflated Poisson.
```{r}
# define parameters
prob_drink <- 0.2 # 20% of days
rate_work <- 1 # average 1 manuscript per day

# sample one year of production
n <- 365

# simulate days monks drink
set.seed(365)
drink <- rbinom(n, size = 1, prob = prob_drink)

# simulate manuscripts completed
y <- (1 - drink) * rpois(n, lambda = rate_work)
```

```{r}
d <-
  tibble(drink = factor(drink, levels = 1:10),
         y = y)

d %>% 
  ggplot(aes(x = y)) +
  geom_histogram(aes(fill = drink), binwidth = 1,
                 linewidth = 1/10, color = "grey92") +
  scale_fill_manual(values = canva_pal("Green fields")(4)[1:2]) +
  xlab("Manuscripts completed") +
  theme(legend.position = "none")
```

With these data, the likelihood of observing zero on y, (i.e., the likelihood zero manuscripts were completed on a given occasion) is
$$
\begin{align}
\text{Pr}(0|p, \lambda) & = \text{Pr}(drink|p) + \text{Pr}(work|p) \times \text{Pr}(0|\lambda) \\
& = p + (1 - p) \text{exp}(-\lambda).
\end{align}
$$

Letting $p$ be the probability $y$ is zero and $\lambda$ be the shape of the distribution, the zero-inflated Poisson (ZIPoisson) regression model might take the basic form
$$
\begin{align}
y_i & \sim \text{ZIPoisson}(p_i, \lambda_i) \\
\text{logit}(p_i) & = \alpha_p + \beta_p x_i \\
\text{log}(\lambda_i) & = \alpha_\lambda + \beta_\lambda x_i,
\end{align}
$$

```{r}
b12.3 <-
  brm(data = d,
      family = zero_inflated_poisson,
      y ~ 1,
      prior = c(prior(normal(1, 0.5), class = Intercept),
                prior(beta(2, 6), class = zi)), # the brms default is beta(1, 1)
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.03")
```

```{r}
priors <-
  c(prior(beta(1, 1)),
    prior(beta(2, 6)))

priors %>% 
  parse_dist(prior)
```

```{r}
priors %>% 
  parse_dist(prior) %>% 
  ggplot(aes(y = prior, dist = .dist, args = .args, fill = prior)) +
  stat_dist_halfeye(,width = .95) +
  scale_fill_manual(values = canva_pal("Green fields")(4)[c(4, 1)]) +
  scale_x_continuous("zi", breaks = c(0, .5, 1)) +
  scale_y_discrete(expand = expansion(add = .1)) +
  ylab(NULL) +
  theme(legend.position = "none")
```

We might write our statistical model for b12.3 as
$$
\begin{align}
y_i & \sim \text{ZIPoisson}(p, \lambda) \\
p & = \alpha_p \\
\text{log} \lambda & = \alpha_\lambda \\
\alpha_p & \sim \text{Beta}(2, 6) \\
\alpha_\lambda & \sim \text{Normal}(1, 0.5).
\end{align}
$$

Here's that exponentiated $\alpha_\lambda$
```{r}
fixef(b12.3)
```


```{r}
fixef(b12.3)[1, ] %>% exp()
```

#### 12.2.1.1 Overthinking: Zero-inflated Poisson calculations in Stan.

```{r}
b12.3$model
```

## 12.3 Ordered categorical outcomes
### 12.3.1 Example: Moral intution.
```{r}
data(Trolley, package = "rethinking")
d <- Trolley
rm(Trolley)
```

```{r}
glimpse(d)
```

```{r}
d %>% 
  distinct(id) %>% 
  count()
```

### 12.3.2 Describing an ordered distribution with. intercepts.
```{r}
p1 <-
  d %>% 
  ggplot(aes(x = response, fill = after_stat(x))) +
  geom_histogram(binwidth = 1/4, linewidth = 0) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
p1
```

Our cumulative proportion plot, Figure 12.4.b, will require some pre-plot wrangling.
```{r}
p2 <-
  d %>%
  count(response) %>%
  mutate(pr_k     = n / nrow(d),
         cum_pr_k = cumsum(pr_k)) %>% 
  
  ggplot(aes(x = response, y = cum_pr_k, 
             fill = response)) +
  geom_line(color = canva_pal("Green fields")(4)[2]) +
  geom_point(shape = 21, color = "grey92", 
             size = 2.5, stroke = 1) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("cumulative proportion", 
                     breaks = c(0, .5, 1), limits = c(0, 1.001)) +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
p2
```

```{r}
logit <- function(x) log(x / (1 - x)) # convenience function

d %>% 
  count(response) %>% 
  mutate(pr_k = n / nrow(d),
         cum_pr_k = cumsum(n / nrow(d))) %>% 
  mutate(alpha = logit(cum_pr_k) %>% round(digits = 2))
```

```{r}
p3 <-
  d %>% 
  count(response) %>% 
  mutate(cum_pr_k = cumsum(n / nrow(d))) %>% 
  dplyr::filter(response < 7) %>% 
  
  # we can do the logit() conversion right in ggplot()
  ggplot(aes(x = response, y = logit(cum_pr_k), 
             fill = response)) +
  geom_line(color = canva_pal("Green fields")(4)[2]) +
  geom_point(shape = 21, colour = "grey92", size = 2.5,
             stroke = 1) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7, limits = c(1, 7)) +
  ylab("log-cumulative-odds") +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")

p3
```

```{r}
(p1 | p2 | p3) + 
  plot_annotation(title = "Re-describing a discrete distribution using log-cumulative-odds.")
```

```{r}
# primary data
d_plot <-
  d %>% 
  count(response) %>% 
  mutate(pr_k = n / nrow(d),
         cum_pr_k = cumsum(n / nrow(d))) %>% 
  mutate(discrete_probability = ifelse(response == 1, cum_pr_k, cum_pr_k - pr_k))

# annotation
text <-
  tibble(text = 1:7,
         response = seq(from = 1.25, to = 7.25, by = 1),
         cum_pr_k = d_plot$cum_pr_k - .065)

d_plot %>% 
  ggplot(aes(x = response, y = cum_pr_k,
             color = cum_pr_k, fill = cum_pr_k)) +
  geom_line(color = canva_pal("Green fields")(4)[1]) +
  geom_point(shape = 21, colour = "grey92", size = 2.5,
             stroke = 1) +
  geom_linerange(aes(ymin = 0, ymax = cum_pr_k),
                 alpha = 1/2, 
                 color = canva_pal("Green fields")(4)[1]) +
  geom_linerange(aes(x = response + .025,
                     ymin = ifelse(response == 1, 0,
                                   discrete_probability),
                     ymax = cum_pr_k),
                 color = "black") +
  # number annotation
  geom_text(data = text,
            aes(label = text),
            size = 4) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("cumulative proportion",
                     breaks = c(0, .5, 1),
                     limits = c(0, 1.001)) +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
```

A compact way to express the formula for this first type of statistical model is
$$
\begin{align}
\text{response}_i & \sim \text{Categorical}(p) \\
\text{logit}(p_k) & = \alpha_k - \phi \\
\phi & = 0 \\
\alpha_k & \sim \text{Normal}(0, 1.5),
\end{align}
$$

```{r}
# define the start values
inits <- list(`Intercept[1]` = -2,
              `Intercept[2]` = -1,
              `Intercept[3]` = 0,
              `Intercept[4]` = 1,
              `Intercept[5]` = 2,
              `Intercept[6]` = 2.5)

inits_list <- list(inits, inits, inits, inits)

b12.4 <-
  brm(data = d,
      family = cumulative,
      response ~ 1,
      prior(normal(0, 1.5), class = Intercept),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      init = inits_list, # here we add our start values
      file = "fits/b12.04")
```

```{r}
print(b12.4)
```

```{r}
b12.4 %>% 
  fixef() %>% 
  inv_logit_scaled() %>% 
  round(digits = 3)
```

```{r}
as_draws_df(b12.4) %>% 
  mutate_at(vars(starts_with("b_")), inv_logit_scaled) %>% 
  pivot_longer(starts_with("b_")) %>% 
  group_by(name) %>% 
  summarise(mean = mean(value),
            sd = sd(value),
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975))
```

```{r}
d_plot %>% 
  select(response, cum_pr_k)
```

```{r}
fixef(b12.4) %>% 
  data.frame() %>% 
  rownames_to_column("intercept") %>% 
  mutate(response = str_extract(intercept, "\\d") %>% as.double()) %>% 
  
  ggplot(aes(x = response, y = Estimate,
             ymin = Q2.5, ymax = Q97.5,
             fill = response)) +
  geom_line(color = canva_pal("Green fields")(4)[2]) +
  geom_point(shape = 21, color = "grey92",
             size = 1.5, stroke = 1) +
  geom_linerange(color = canva_pal("Green fields")(4)[2]) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7, limits = c(1, 7)) +
  ylab("log-cumulative-odds") +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
```

### 12.3.3 Adding predictor variables.
Now we define the generic linear model as $\phi_i = \beta x_i$. Accordingly, the formula for our cumulative logit model becomes
$$
\begin{align}
\text{log} \frac{\text{Pr}(y_i \le k)}{1 - \text{Pr}(y_i \le K)} & = \alpha_k - \phi_i \\
\phi_i & = \beta x_i.
\end{align}
$$

```{r}
logistic <- function(x) {
  p <- 1 / (1 + exp(-x))
  p <- ifelse(x == Inf, 1, p)
  p
}

# now we get down to it
dordlogit <- function(x, phi, a, log = FALSE) {
  a <- c(as.numeric(a), Inf)
  p <- logistic(a[x] - phi)
  na <- c(-Inf, a)
  np <- logistic(na[x] - phi)
  p <- p - np
  if(log == TRUE) p <- log(p)
  p
}
```

The dordlogit() function works like this:
```{r}
pk <- dordlogit(1:7, 0, fixef(b12.4)[, 1])

pk %>% 
  round(digits = 2)
```

```{r}
sum(pk * (1:7))
```

```{r}
(
  explicit_example <-
    tibble(probability_of_a_response = pk) %>% 
    mutate(the_response = 1:7) %>% 
    mutate(their_product = probability_of_a_response * the_response)
)
```

```{r}
explicit_example %>% 
  summarise(average_outcome_value = sum(their_product))
```

Now we'll try it by subtracting .5 from each.
```{r}
# the probabilities of a given response
pk <- dordlogit(1:7, phi = 0, a = fixef(b12.4)[, 1] - .5)

pk %>% 
  round(digits = 2)
```

```{r}
# the average rating
sum(pk * (1:7))
```

As to our upcoming model, we might express the statistical formula as
$$
\begin{align}
\text{response}_{i} & \sim \text{Categorical}(p) \\
\text{logit}(p_k) & = \alpha_k - \phi_i \\
\phi_i & = \beta_1 \text{action}_i + \beta_2 \text{contact}_i + (\beta_3 + \beta_4 \text{contact}_i)\text{intention}_i \\
\alpha_k & \sim \text{Normal}(0, 1.5) \\
\beta_1, ..., \beta_5 & \sim \text{Normal}(0, 0.5),
\end{align}
$$

The linear model for $\phi$ is
$$
\phi_i = \beta_1 \text{action}_i + \beta_2 \text{contact}_i + \beta_3 \text{intention}_i + \beta_4 (\text{action}_i \times \text{intention}_i) + \beta_5 (\text{contact}_i \times \text{intention}_i)
$$

```{r}
b12.5 <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + action + contact + intention + intention:action + intention:contact,
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(normal(0, 0.5), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.05")
```

```{r}
print(b12.5)
```

```{r}
labs <- str_c("beta[", 1:5, "]")

as_draws_df(b12.5) %>% 
  select(b_action: `b_contact:intention`) %>% 
  set_names(labs) %>% 
  pivot_longer(everything()) %>% 
  
  ggplot(aes(x = value, y = name)) +
  geom_vline(xintercept = 0, alpha = 1/5, linetype = 3) +
  stat_gradientinterval(.width = .5, size = 1, 
                        point_size = 3/2, shape = 21,
                        point_fill = canva_pal("Green fields")(4)[3],
                        fill = canva_pal("Green fields")(4)[1],
                        color = canva_pal("Green fields")(4)[2]) +
  scale_x_continuous("marginal posterior", breaks = -5:0 / 4) +
  scale_y_discrete(NULL, labels = parse(text = labs)) +
  coord_cartesian(xlim = c(-1.4, 0))
```

Our approach to making the op panels of Figure 12.6 will start with fitted()
```{r}
nd <-
  d %>% 
  distinct(action, contact, intention) %>% 
  mutate(combination = str_c(action, contact, intention, sep = "_"))

f <-
  fitted(b12.5,
         newdata = nd,
         summary = F)

# what have we done?
f %>% str()
```
```{r}
f <-
  rbind(f[, , 1],
        f[, , 2],
        f[, , 3],
        f[, , 4],
        f[, , 5],
        f[, , 6],
        f[, , 7]) %>% 
  data.frame() %>% 
  set_names(pull(nd, combination)) %>% 
  mutate(response = rep(1:7, each = n() / 7),
         draw = rep(1:4000, times = 7)) %>% 
  pivot_longer(-c(draw, response),
               names_to = c("action", "contact", "intention"),
               names_sep = "_",
               values_to = "pk") %>% 
  mutate(intention = intention %>% as.integer())

# how do the data look, now?
glimpse(f)
```

```{r}
# to order our factor levels for `facet`
levels <- c("action=0, contact=0", "action=1, contact=0", "action=0, contact=1")

p1 <-
  f %>% 
  # unnecessary for these plots
  dplyr::filter(response < 7) %>% 
  # this will help us define the three panels of the triptych
  mutate(facet = factor(str_c("action=", action, ", contact=", contact),
                        levels = levels)) %>% 
  # these next three lines allow us to compute the cumulative probabilities
  group_by(draw, facet, intention) %>% 
  arrange(draw, facet, intention, response) %>% 
  mutate(probability = cumsum(pk)) %>% 
  ungroup() %>% 
  # these next three lines are how we randomly selected 50 posterior draws
  nest(data = -draw) %>% 
  slice_sample(n = 50) %>%
  unnest(data) %>% 
  
  # plot!
  ggplot(aes(x = intention, y = probability)) +
  geom_line(aes(group = interaction(draw, response), color = probability),
            alpha = 1/10) +
  geom_point(data = d %>%  # wrangle the original data to make the dots
               group_by(intention, contact, action) %>% 
               count(response) %>% 
               mutate(probability = cumsum(n / sum(n)),
                      facet = factor(str_c("action=", action, ", contact=", contact),
                                     levels = levels)) %>% 
               dplyr::filter(response < 7),
             color = canva_pal("Green fields")(4)[2]) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[4],
                       high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous("intention", breaks = 0:1) +
  scale_y_continuous(breaks = c(0, .5, 1), limits = 0:1) +
  theme(legend.position = "none") +
  facet_wrap(~ facet)
```

```{r}
p <-
  predict(b12.5,
          newdata = nd,
          ndraws = 1000,
          scale = "response",
          summary = F)

p %>% str()
```


```{r}
p2 <-
  p %>% 
  data.frame() %>% 
  set_names(pull(nd, combination)) %>% 
  pivot_longer(everything(),
               names_to = c("action", "contact", "intention"),
               names_sep = "_",
               values_to = "response") %>% 
  mutate(facet = factor(str_c("action=", action, ", contact=", contact),
                        levels = levels)) %>% 
  
  ggplot(aes(x = response, fill = intention)) +
  geom_bar(width = 1/3, position = position_dodge(width = .4)) +
  scale_fill_manual(values = canva_pal("Green fields")(4)[2:1]) +
  scale_x_continuous("response", breaks = 1:7) +
  theme(legend.position = "none") +
  facet_wrap(~ facet)
```

```{r}
(p1 / p2) & theme(panel.background = element_rect(fill = "grey94"))
```

```{r}
p1 <-
  f %>% 
  mutate(facet = factor(str_c("action=", action, ", contact=", contact),
                        levels = levels)) %>% 
  group_by(draw, facet, intention) %>% 
  summarise(mean_response = sum(pk * response)) %>% 
  ungroup() %>% 
  nest(data = -draw) %>% 
  slice_sample(n = 50) %>%
  unnest(data) %>%

  ggplot(aes(x = intention, y = mean_response)) +
  geom_line(aes(group = draw, color = mean_response),
            alpha = 1/10) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[4],
                       high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous("intention", breaks = 0:1) +
  scale_y_continuous("response", breaks = 1:7, limits = c(1, 7)) +
  theme(legend.position = "none") +
  facet_wrap(~ facet)
```

```{r}
p2 <-
  f %>% 
  mutate(facet = factor(str_c("action=", action, ", contact=", contact),
                        levels = levels)) %>% 
  
  ggplot(aes(x = response, y = pk, fill = factor(intention))) +
  stat_ccdfinterval(.width = .95, justification = 1, size = 1/4, 
                    shape = 21, point_fill = canva_pal("Green fields")(4)[3], point_size = 1/3,
                    position = "dodge", width = .75) +
  scale_fill_manual(values = canva_pal("Green fields")(4)[2:1]) +
  scale_x_continuous("response", breaks = 1:7) +
  scale_y_continuous("count", breaks = 0:3 / 10, labels = 0:3 * 100, limits = c(0, NA)) +
  theme(legend.position = "none") +
  facet_wrap(~ facet)
```

```{r}
p1 / p2
```

#### 12.3.3.1 Rethinking: Staring into the abyss.

## 12.4 Ordered categorical predictors
```{r}
distinct(d, edu)
```

```{r}
d <-
  d %>% 
  mutate(edu_new = 
           recode(edu,
                  "Elementary School" = 1,
                  "Middle School" = 2,
                  "Some High School" = 3,
                  "High School Graduate" = 4,
                  "Some College" = 5,
                  "Bachelor's Degree" = 6,
                  "Master's Degree" = 7,
                  "Graduate Degree" = 8) %>% 
           as.integer())

# what did we do?
d %>% 
  distinct(edu, edu_new) %>% 
  arrange(edu_new)
```

```{r}
library(gtools)
set.seed(1805)
delta <- rdirichlet(10, alpha = rep(2, 7))

str(delta)
```

```{r}
delta %>% 
  data.frame() %>% 
  set_names(1:7) %>% 
  mutate(row = 1:n()) %>% 
  pivot_longer(-row, names_to = "index") %>% 
  
  ggplot(aes(x = index, y = value, group = row,
             alpha = row == 3, color = row == 3)) +
  geom_line() +
  geom_point() +
  scale_alpha_manual(values = c(1/3, 1)) +
  scale_color_manual(values = canva_pal("Green fields")(4)[1:2]) +
  ylab("probability") +
  theme(legend.position = "none")
```

The brms package has a rdirichlet() function, too.
```{r}
set.seed(12)

brms::rdirichlet(n = 1e4, alpha = rep(2, 7)) %>% 
  data.frame() %>% 
  set_names(1:7) %>% 
  pivot_longer(everything()) %>% 
  mutate(name = name %>% as.double(),
         alpha = str_c("alpha[", name, "]")) %>% 
  
  ggplot(aes(x = value, color = name, group = name, fill = name)) +
  geom_density(alpha = .8) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[2],
                      high = canva_pal("Green fields")(4)[3]) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[2],
                      high = canva_pal("Green fields")(4)[3]) +
  scale_x_continuous("probability", limits = c(0, 1),
                     breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1"), ) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = expression("Dirichlet"*(2*", "*2*", "*2*", "*2*", "*2*", "*2*", "*2))) +
  theme(legend.position = "none") +
  facet_wrap(~ alpha, labeller = label_parsed, nrow = 2)
```

```{r}
b12.6 <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + action + contact + intention + mo(edu_new), # note the mo() syntax
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(normal(0, 1), class = b),
                # note the new kinds of prior statements
                prior(normal(0, .143), class = b, coef = moedu_new),
                prior(dirichlet(2, 2, 2, 2, 2, 2, 2), class = simo, coef = moedu_new1)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.06")
```

```{r}
print(b12.6)
```

```{r}
as_draws_df(b12.6) %>% 
  transmute(bE = bsp_moedu_new * 7) %>% 
  median_qi(.width = .89) %>% 
  mutate_if(is.double, round, digits = 2)
```

Formal statistical notation of Categorical ordered variables
$$
\begin{align}
\text{response}_i & \sim \text{Categorical}(p) \\
\text{logit}(p_k) & = \alpha_k - \phi_i \\
\phi_i & = \beta_1 \text{action}_i + \beta_2 \text{contact}_i + \beta_3 \text{intention}_i + \beta_4\text{mo}(\text{edu_new}_i, \delta) \\
\alpha_k & \sim \text{Normal}(0, 1.5) \\
\beta_1, ..., \beta_3 & \sim \text{Normal}(0, 1) \\
\beta_4 & \sim \text{Normal}(0, 0.143) \\
\delta & \sim \text{Dirichlet}(2, 2, 2, 2, 2, 2, 2),
\end{align}
$$
where $mo(x, \delta)$ is an operator indicating some predictor $x$ has undergone the monotonic transform and $\delta$ is our vector of simplex parameters $\delta_1, ..., \delta_7$. That is, $\beta_4 \text{mo}(\text{edu_new}_i, \delta)$ is our alternative brms-oriented way of epxressing McElreath's $\beta_E \Sigma_{j=0}^{\text{E}_i-1} \delta_j$.

```{r}
my_lower <- function(data, mapping, ...) {
  # get the x and y data to use the other code
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  # compute the correlations
  corr <- cor(x, y, method = "p", use = "pairwise")
  abs_corr <- abs(corr)
  
  # plot the cor value
  ggally_text(
    label = formatC(corr, digits = 2, format = "f") %>% str_replace(., "0.", "."),
    mapping = aes(),
    size = 4,
    color = canva_pal("Green fields")(4)[2]) +
    scale_x_continuous(NULL, breaks = NULL) +
    scale_y_continuous(NULL, breaks = NULL)
}

my_diag <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_density(fill = canva_pal("Green fields")(4)[1],
                 linewidth = 0) +
    scale_x_continuous(NULL, breaks = NULL) +
    scale_y_continuous(NULL, breaks = NULL)
}

my_upper <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_hex(bins = 18) +
    scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                        high = canva_pal("Green fields")(4)[3]) +
    scale_x_continuous(NULL, breaks = NULL) +
    scale_y_continuous(NULL, breaks = NULL) +
    theme(panel.background = element_rect(fill = canva_pal("Green fields")(4)[2]))
}
```

```{r}
library(GGally)

delta_labels <- c("Elem", "MidSch", "SHS", "HSG", "SCol", "Bach", "Mast", "Grad")

as_draws_df(b12.6) %>% 
  select(contains("simo_moedu_new1")) %>% 
  set_names(str_c(delta_labels[2:8], "~(delta[", 1:7, "])")) %>% 
  ggpairs(upper = list(continuous = my_upper),
          diag = list(continuous = my_diag),
          lower = list(continuous = my_lower),
          labeller = label_parsed) +
  theme(strip.text = element_text(size = 8))
```

Here we add a normalized version of edu_new called edu_norm
```{r}
d <-
  d %>% 
  mutate(edu_norm = (edu_new - 1) / 7)

# what does this look like?
d %>% 
  distinct(edu, edu_new, edu_norm) %>% 
  arrange(edu_new)
```

Now fit the more conventional model in which we treat edu_norm as a simple continuous predictor
```{r}
b12.7 <-
  brm(data = d,
      family = cumulative,
      response ~ 1 + action + contact + intention + edu_norm,
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(normal(0, 1), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      file = "fits/b12.07")
```

```{r}
fixef(b12.7)[7:10, ]
```

```{r}
nd <-
  tibble(edu_new = 1:8,
         action = 0,
         contact = 0,
         intention = 0)

f <-
  fitted(b12.6,
         newdata = nd)

f %>% str()
```

```{r}
# b12.6
f12.6 <-
  rbind(f[, , 1],
        f[, , 2],
        f[, , 3],
        f[, , 4],
        f[, , 5],
        f[, , 6],
        f[, , 7]) %>% 
  data.frame() %>% 
  mutate(edu = factor(rep(1:8, times = 7)),
         response = rep(1:7, each = 8))

# b12.7
nd <-
  nd %>% 
  mutate(edu_norm = 1:8)

f <-
  fitted(b12.7, newdata = nd)

f12.7 <-
  rbind(f[, , 1],
        f[, , 2],
        f[, , 3],
        f[, , 4],
        f[, , 5],
        f[, , 6],
        f[, , 7]) %>% 
  data.frame() %>% 
  mutate(edu      = factor(rep(1:8, times = 7)),
         response = rep(1:7, each = 8))
```

```{r}
# this will help with scale_color_manual()
colors <-
  scales::seq_gradient_pal(canva_pal("Green fields")(4)[4],
                           canva_pal("Green fields")(4)[3])(seq(0, 1, length.out = 8))

bind_rows(f12.6, f12.7) %>% 
  mutate(fit = rep(c("b12.6 with `mo()` syntax",
                     "b12.7 with conventional syntax"),
                   each = n() / 2)) %>% 
  ggplot(aes(x = response, y = Estimate,
             ymin = Q2.5, ymax = Q97.5,
             color = edu, group = edu)) +
  geom_pointrange(fatten = 3/2, position = position_dodge(width = 3/4)) +
  scale_color_manual("education", values = colors, labels = delta_labels) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("probability", limits = c(0, .43)) +
  theme(legend.background = element_blank(),
        legend.position = "right") +
  facet_wrap(~ fit)
```

```{r}
b12.6 <- add_criterion(b12.6, "loo")
b12.7 <- add_criterion(b12.7, "loo")

loo_compare(b12.6, b12.7, criterion = "loo") %>% print(simplify = F)
```

```{r}
model_weights(b12.6, b12.7, weights = "loo") %>% round(digits = 2)
```

```{r}
as_draws_df(b12.6) %>% 
  select(contains("new1")) %>% 
  set_names(1:7) %>% 
  mutate(draw = 1:n(),
         `0` = 0) %>% 
  pivot_longer(-draw, names_to = "delta") %>% 
  arrange(delta) %>% 
  group_by(draw) %>% 
  mutate(cum_sum = cumsum(value)) %>% 
  
  ggplot(aes(x = delta, y = cum_sum)) +
  stat_pointinterval(.width = .95, size = 1,
                     color = canva_pal("Green fields")(4)[1]) +
  stat_pointinterval(.width = .5,
                     color = canva_pal("Green fields")(4)[4],
                     point_color = canva_pal("Green fields")(4)[2]) +
  scale_x_discrete(NULL, labels = parse(text = str_c("delta[", 0:7, "]"))) +
  ylab("cumulative sum")
```
This is another way to show the largest effects of education are when going from Elementary School to Middle School $(\delta_0 \rightarrow \delta_1)$ and when going from Some High School to High School Graduate $(\delta_2 \rightarrow \delta_3)$.
```{r}
d %>% 
  distinct(edu, edu_new) %>% 
  arrange(edu_new) %>% 
  mutate(`delta[j]` = edu_new - 1)
```

