---
title: "11 God Spiked the Integers"
format: html
---

## 11.1 Binomial regression
### 11.1.1 Logistic regression: Prosocial cimpanzees
```{r}
data(chimpanzees, package = "rethinking")
d <- chimpanzees
rm(chimpanz)
```

```{r}
library(tidyverse)
library(flextable)

d %>% 
  distinct(prosoc_left, condition) %>% 
  mutate(description = str_c("Two food items on ",
                             c("right and no partner",
                               "left and no partner",
                               "right and partner present",
                               "left and partner present"))) %>% 
  flextable() %>% 
  width(width = c(1, 1, 4))
```

```{r}
d <-
  d %>% 
  mutate(treatment = factor(1 + prosoc_left + 2 * condition)) %>% 
  # this will come in handy, later
  mutate(labels = factor(treatment,
                         levels = 1:4,
                         labels = c("r/n", "l/n", "r/p", "l/p")))
```

```{r}
d %>% 
  count(condition, treatment, prosoc_left)
```

```{r}
library(brms)
```

We start with the simple intercept-only logistic regression model, which follows the statistical formula

$$
\begin{align}
\text{pulled_left}_{i} & \sim \text{Binomial}(1, p_{i}) \\
\text{logit}(p_{i}) & = \alpha \\
\alpha & \sim \text{Normal}(0, w),
\end{align}
$$
where $w$ is the parameter for the $\sigma$ the value for which we have yet to choose. To start things off, we'll set $w$ = 10, fit a model with where we set sample_prior = TRUE, and get a sense of the prior on a plot.

```{r}
b11.1 <-
  brm(data = d,
      family = binomial,
      pulled_left | trials(1) ~ 1,
      prior(normal(0, 10), class = Intercept),
      seed = 11,
      sample_prior = T,
      file = "fits/b11.01")
```

```{r}
library(wesanderson)
wes_palette("Moonrise2")
```

```{r}
wes_palette("Moonrise2")[1:4]
```

```{r}
library(ggthemes)

theme_set(
  theme_default() +
    theme_tufte() +
    theme(plot.background = 
            element_rect(fill = wes_palette("Moonrise2")[3],
                         color = wes_palette("Moonrise2")[3]))
)
```

```{r}
prior_draws(b11.1) %>% 
  mutate(p = inv_logit_scaled(Intercept)) %>% 
  
  ggplot(aes(x = p)) +
  geom_density(fill = wes_palette("Moonrise2")[4],
               linewidth = 0, adjust = .1) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab("prior prob pull left")
```

```{r}
b11.1b <-
  brm(data = d,
      family = binomial,
      pulled_left | trials(1) ~ 1,
      prior(normal(0, 1.5), class = Intercept),
      seed = 11,
      sample_prior = T,
      file = "fits/b11.01b")
```

```{r}
# wrangle
bind_rows(prior_draws(b11.1),
          prior_draws(b11.1b)) %>% 
  mutate(p = inv_logit_scaled(Intercept),
         w = factor(rep(c(10, 1.5), each = n() / 2),
                    levels = c(10, 1.5))) %>% 
  
  # plot
  ggplot(aes(x = p, fill = w)) +
  geom_density(linewidth = 0, alpha = 3/4, adjust = .1) +
  scale_fill_manual(expression(italic(w)),
                    values = wes_palette("Moonrise2")[c(4, 1)]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(alpha%~%Normal(0*", "*italic(w))),
       x = "prior prob pull left")
```

```{r}
b11.2 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + b,
                              a ~ 1,
                              b ~ 0 + treatment,
                              nl = TRUE),
      prior = 
        c(prior(normal(0, 1.5), nlpar = a),
          prior(normal(0, 10), nlpar = b, coef = treatment1),
          prior(normal(0, 10), nlpar = b, coef = treatment2),
          prior(normal(0, 10), nlpar = b, coef = treatment3),
          prior(normal(0, 10), nlpar = b, coef = treatment4)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      sample_prior = T,
      file = "fits/b11.02")

# w = 0.5
b11.3 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + b,
                              a ~ 1,
                              b ~ 0 + treatment,
                              nl = TRUE),
      prior = 
        c(prior(normal(0, 1.5), nlpar = a),
          prior(normal(0, .5), nlpar = b, coef = treatment1),
          prior(normal(0, .5), nlpar = b, coef = treatment2),
          prior(normal(0, .5), nlpar = b, coef = treatment3),
          prior(normal(0, .5), nlpar = b, coef = treatment4)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      sample_prior = T,
      file = "fits/b11.03")
```

```{r}
# wrangle
prior <-
  bind_rows(prior_draws(b11.2),
            prior_draws(b11.3)) %>% 
  mutate(w = factor(rep(c(10, .5), each = n() / 2),
                    levels = c(10, .5)),
         p1 = inv_logit_scaled(b_a + b_b_treatment1),
         p2 = inv_logit_scaled(b_a + b_b_treatment2)) %>% 
  mutate(diff = abs(p1 - p2))

# plot
prior %>% 
  ggplot(aes(x = diff, fill = w)) +
  geom_density(linewidth = 0, alpha = 3/4, adjust = .1) +
  scale_fill_manual(expression(italic(w)), values = wes_palette("Moonrise2")[c(4, 2)]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(alpha%~%Normal(0*", "*italic(w))),
       x = "prior diff between treatments")
```

```{r}
prior %>% 
  group_by(w) %>% 
  summarise(mean = mean(diff))
```

```{r}
print(b11.3)
```

```{r}
d %>% 
  summarise(grand_mean = mean(pulled_left))
```

```{r}
library(tidybayes)

as_draws_df(b11.3) %>% 
  transmute(alpha = inv_logit_scaled(b_a_Intercept)) %>% 
  mean_qi()
```

```{r}
d %>% 
  group_by(treatment) %>% 
  summarise(mean = mean(pulled_left))
```
Here are the corresponding posteriors
```{r}
as_draws_df(b11.3) %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(treatment = str_remove(name, "b_b_treatment"),
         mean = inv_logit_scaled(b_a_Intercept + value)) %>% 
  group_by(treatment) %>% 
  mean_qi(mean)
```

Full model
$$
\begin{align}
\text{pulled_left}_{i} & \sim \text{Binomial}(1, p_{i}) \\
\text{logit}(p_{i}) & = \alpha_{\text{actor}[i]} + \beta_{\text{treatment}[i]} \\
\alpha_{j} & \sim \text{Normal}(0, 1.5) \\
\beta_{k} & \sim \text{Normal}(0, 0.5).
\end{align}
$$


```{r}
d <-
  d %>% 
  mutate(actor = factor(actor))
```

```{r}
b11.4 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + b,
         a ~ 0 + actor,
         b ~ 0 + treatment,
         nl = TRUE),
      prior = c(prior(normal(0, 1.5), nlpar = a),
                prior(normal(0, .5), nlpar = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.04")
```

```{r}
print(b11.4)
```

```{r}
library(tidybayes)

post <- as_draws_df(b11.4)

post %>% 
  pivot_longer(contains("actor")) %>% 
  mutate(probability = inv_logit_scaled(value),
         actor = factor(str_remove(name, "b_a_actor"),
                        levels = 7:1)) %>% 
  
  ggplot(aes(x = probability, y = actor)) +
  geom_vline(xintercept = .5, 
             color = wes_palette("Moonrise2")[1],
             linetype = 3) +
  stat_pointinterval(.width = .95, size = 1/2,
                     color = wes_palette("Moonrise2")[4]) +
  scale_x_continuous(expression(alpha[actor]), limits = 0:1) +
  ylab(NULL) +
  theme(axis.ticks.y = element_blank())
```

```{r}
tx <- c("R/N", "L/N", "R/P", "L/P")

post %>% 
  select(contains("treatment")) %>% 
  set_names("R/N", "L/N", "R/P", "L/P") %>% 
  pivot_longer(everything()) %>% 
  mutate(probability = inv_logit_scaled(value),
         treatment = factor(name, levels = tx)) %>% 
  mutate(treatment = fct_rev(treatment)) %>% 
  
  ggplot(aes(x = value, y = treatment)) +
  geom_vline(xintercept = 0,
             color = wes_palette("Moonrise2")[2],
             linetype = 3) +
  stat_pointinterval(.width = .95, size = 1/2,
                     color = wes_palette("Moonrise2")[4]) +
  labs(x = expression(beta[treatment]),
       y = NULL) +
  theme(axis.ticks.y = element_blank())
```

```{r}
post %>% 
  mutate(db13 = b_b_treatment1 - b_b_treatment3,
         db24 = b_b_treatment2 - b_b_treatment4) %>% 
  pivot_longer(db13:db24) %>% 
  mutate(diffs = factor(name, levels = c("db24", "db13"))) %>% 
  
  ggplot(aes(x = value, y = diffs)) +
  geom_vline(xintercept = 0, 
             color = wes_palette("Moonrise2")[2], linetype = 3) +
  stat_pointinterval(.width = .95, size = 1/2,
                     color = wes_palette("Moonrise2")[4]) +
  labs(x = "difference", y = NULL) +
  theme(axis.ticks.y = element_blank())
```

```{r}
d %>% 
  group_by(actor, treatment) %>% 
  summarise(proportion = mean(pulled_left)) %>% 
  dplyr::filter(actor == 1)
```

```{r}
p1 <-
  d %>% 
  group_by(actor, treatment) %>% 
  summarise(proportion = mean(pulled_left)) %>% 
  left_join(d %>% distinct(actor, treatment, labels,
                           condition, prosoc_left),
            by = c("actor", "treatment")) %>% 
  mutate(condition = factor(condition)) %>% 
  
  ggplot(aes(x = labels, y = proportion)) +
  geom_hline(yintercept = .5, 
             color = wes_palette("Moonrise2")[3]) +
  geom_line(aes(group = prosoc_left),
            linewidth = 1/4, 
            color = wes_palette("Moonrise2")[4]) +
  geom_point(aes(color = condition), size = 2.5,
             show.legend = F) +
  labs(subtitle = "observed proportions")
```

```{r}
nd <-
  d %>% 
  distinct(actor, treatment, labels, condition, prosoc_left)

p2 <-
  nd %>% 
  add_epred_draws(b11.4) %>% 
  mean_qi() %>% 
  mutate(condition = factor(condition)) %>% 
  
  ggplot(aes(x = labels, y = .epred, 
             ymin = .lower, ymax = .upper)) +
  geom_hline(yintercept = .5,
             color = wes_palette("Moonrise2")[3]) +
  geom_line(aes(group = prosoc_left), linewidth = 1/4,
            color = wes_palette("Moonrise2")[4]) +
  geom_pointrange(aes(color = condition),
                  fatten = 2.5, show.legend = F) +
  labs(subtitle = "posterior prediction")

# combine the two ggplots
library(patchwork)

(p1 / p2) &
  scale_color_manual(
    values = wes_palette("Moonrise2")[c(2:1)]) &
  scale_y_continuous("proportion left lever",
                     breaks = c(0, .5, 1),
                     limits = c(0, 1)) &
  xlab(NULL) &
  theme(axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = alpha("white", 1/10), linewidth = 0)) &
  facet_wrap(~ actor, nrow = 1, labeller = label_both)
```

```{r}
d <-
  d %>% 
  mutate(side = factor(prosoc_left + 1), # right 1, left 2
         cond = factor(condition + 1)) # no partner 1, partner 2
```

```{r}
b11.5 <-
  brm(data = d,
      family = binomial,
      bf(pulled_left | trials(1) ~ a + bs + bc,
         a ~ 0 + actor,
         bs ~ 0 + side,
         bc ~ 0 + cond,
         nl = TRUE),
      prior = c(prior(normal(0, 1.5), nlpar = a),
                prior(normal(0, .5), nlpar = bs),
                prior(normal(0, .5), nlpar = bc)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.05")
```

```{r}
b11.4 <- add_criterion(b11.4, c("loo", "waic"))
b11.5 <- add_criterion(b11.5, c("loo", "waic"))

loo_compare(b11.4, b11.5, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
loo_compare(b11.4, b11.5, criterion = "waic") %>% 
  print(simplify = F)
```

```{r}
model_weights(b11.4, b11.5, weights = "loo") %>% 
  round(digits = 2)
```

```{r}
model_weights(b11.4, b11.5, weights = "waic") %>% 
  round(digits = 2)
```

```{r}
print(b11.5)
```

```{r}
nd <-
  d %>% 
  distinct(actor, treatment, labels, cond, side)

nd %>% 
  add_epred_draws(b11.5) %>% 
  mean_qi() %>% 
  
  ggplot(aes(x = labels, y = .epred,
             ymin = .lower, ymax = .upper)) +
  geom_hline(yintercept = .5, 
             color = wes_palette("Moonrise2")[3]) +
  geom_line(aes(group = side), linewidth = 1/4,
            color = wes_palette("Moonrise2")[4]) +
  geom_pointrange(aes(color = cond), fatten = 2.5,
                  show.legend = F) +
  scale_color_manual(values = wes_palette("Moonrise2")[c(2:1)]) +
  scale_y_continuous("proportion left lever",
                     breaks = c(0, .5, 1),
                     limits = c(0, 1)) +
  labs(subtitle = "posterior prediction for b11.5",
       x = NULL) +
  theme(axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = alpha("white", 1/10), linewidth = 0)) +
  facet_wrap(~ actor, nrow = 1, labeller = label_both)
```
#### 11.1.1.1 Overthinking: Adding log-probability calculations to a Stan model.
```{r}
log_lik(b11.4) %>% str()
```

### 11.1.2 Relative shark and absolute deer.
```{r}
as_draws_df(b11.4) %>% 
  mutate(proportional_odds = exp(b_b_treatment4 - b_b_treatment2)) %>% 
  mean_qi(proportional_odds)
```

```{r}
tibble(disease_rate = 1/1e7,
       fold_increase = 5) %>% 
  mutate(new_disease_rate = disease_rate * fold_increase)
```

### 11.1.3 Aggregated binomial: Chimpanzees again, condensed.
```{r}
d_aggregated <-
  d %>% 
  group_by(treatment, actor, side, cond) %>% 
  summarise(left_pulls = sum(pulled_left)) %>% 
  ungroup()

d_aggregated %>% head(n = 8)
```

```{r}
b11.6 <-
  brm(data = d_aggregated,
      family = binomial,
      bf(left_pulls | trials(18) ~ a + b,
         a ~ 0 + actor,
         b ~ 0 + treatment,
         nl = TRUE),
      prior = c(prior(normal(0, 1.5), nlpar = a),
                prior(normal(0, .5), nlpar = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.06")
```

```{r}
print(b11.6)
```
```{r}
# this is just for fancy annotation
text <-
  tibble(value = c(1.4, 2.6),
         name = "b_a_actor7",
         fit = c("b11.6", "b11.4"))

# rope in the posterior draws and wrangle
bind_rows(as_draws_df(b11.4),
          as_draws_df(b11.6)) %>% 
  mutate(fit = rep(c("b11.4", "b11.6"), each = n() / 2)) %>% 
  pivot_longer(b_a_actor1:b_b_treatment4) %>% 
  
  # plot
  ggplot(aes(x = value, y = name, color = fit)) +
  stat_pointinterval(.width = .95, size = 2/3,
                     position = position_dodge(width = .5)) +
  scale_color_manual(values = wes_palette("Moonrise2")[2:1]) +
  geom_text(data = text, aes(label = fit), family = "Times",
            position = position_dodge(width = 2.25)) +
  labs(x = "posterior (log-odds scale)",
       y = NULL) +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
b11.4 <- add_criterion(b11.4, "loo")
b11.6 <- add_criterion(b11.6, "loo")
```

```{r}
# loo_compare(b11.4, b11.6, criterion = "loo") %>% print(simplify = F)
```

```{r}
loo(b11.4)
```

```{r}
loo(b11.6)
```

```{r}
# deviance of aggregated 6-in-9
-2 * dbinom(6, size = 9, prob = .2, log = TRUE)
```

```{r}
# deviance of dis-aggregated 
-2 * sum(dbinom(c(1, 1, 1, 1, 1, 1, 0, 0, 0), size = 1, prob = 0.2, log = TRUE))
```

```{r}
loo(b11.6) %>% 
  loo::pareto_k_table()
```

### 11.1.4 Aggregated binomial: Graduate school admissions.
```{r}
data(UCBadmit, package = "rethinking")
d <- UCBadmit
rm(UCBadmit)

d
```

```{r}
d <-
  d %>% 
  mutate(gid = factor(applicant.gender, levels = c("male", "female")),
         case = factor(1:n()))
```

The univariable logistic model with male as the sole predictor of admit follow the form
$$
\begin{align}
\text{admit}_{i} & \sim \text{Binomial}(n_{i}, p_{i}) \\
\text{logit}(p_{i}) & = \alpha_{\text{gid}[i]} \\
\alpha_{j} & \sim \text{Normal}(0, 1.5), \\
\end{align}
$$

```{r}
b11.7 <-
  brm(data = d,
      family = binomial,
      admit | trials(applications) ~ 0 + gid,
      prior(normal(0, 1.5), class = b),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.07")
```

```{r}
print(b11.7)
```

```{r}
as_draws_df(b11.7) %>% 
  mutate(diff_a = b_gidmale - b_gidfemale,
         diff_p = inv_logit_scaled(b_gidmale) - inv_logit_scaled(b_gidfemale)) %>% 
  pivot_longer(contains("diff")) %>% 
  group_by(name) %>% 
  mean_qi(value, .width = .89)
```

```{r}
p <-
  d %>% 
  add_predicted_draws(b11.7) %>% 
  mean_qi()

text <-
  d %>% 
  group_by(dept) %>% 
  summarise(case = mean(as.numeric(case)),
            admit = mean(admit / applications) + .05)

p %>% 
  ggplot(aes(x = case, y = admit / applications)) +
  geom_pointrange(aes(y = .prediction / applications,
                      ymin = .lower / applications,
                      ymax = .upper / applications),
                  color = wes_palette("Moonrise2")[1],
                  shape = 1, alpha = 1/3) +
  geom_point(color = wes_palette("Moonrise2")[2]) +
  geom_line(aes(group = dept),
            color = wes_palette("Moonrise2")[2]) +
  geom_text(data = text,
            aes(y = admit, label = dept),
            color = wes_palette("Moonrise2")[2],
            family = "serif") +
  scale_y_continuous("Proportion admitted", limits = 0:1) +
  ggtitle("Posterior validation check") +
  theme(axis.ticks.x = element_blank())
```

The model better suited to answer that question follows the form
$$
\begin{align}
\text{admit}_{i} & \sim \text{Binomial}(n_{i}, p_{i}) \\
\text{logit}(p_{i}) & = \alpha_{\text{gid}[i]} + \delta_{\text{dept}[i]} \\
\alpha_{j} & \sim \text{Normal}(0, 1.5) \\
\delta_{k} & \sim \text{Normal}(0, 1.5), \\
\end{align}
$$

```{r}
b11.8 <-
  brm(data = d,
      family = binomial,
      bf(admit | trials(applications) ~ a + d,
         a ~ 0 + gid,
         d ~ 0 + dept,
         nl = TRUE),
      prior = c(prior(normal(0, 1.5), nlpar = a),
                prior(normal(0, 1.5), nlpar = d)),
      iter = 4000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.08")
```

```{r}
print(b11.8)
```

```{r}
as_draws_df(b11.8) %>% 
  mutate(diff_a = b_a_gidmale - b_a_gidfemale,
         diff_p = inv_logit_scaled(b_a_gidmale) - inv_logit_scaled(b_a_gidfemale)) %>% 
  pivot_longer(contains("diff")) %>% 
  group_by(name) %>% 
  mean_qi(value, .width = .89)
```

```{r}
d %>% 
  group_by(dept) %>% 
  mutate(proportion = applications / sum(applications)) %>% 
  select(dept, gid, proportion) %>% 
  pivot_wider(names_from = dept,
              values_from = proportion) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
d %>% 
  group_by(dept) %>% 
  mutate(proportion = applications / sum(applications)) %>% 
  mutate(label = round(proportion, digits = 2),
         gid = fct_rev(gid)) %>% 
  
  ggplot(aes(x = dept, y = gid, fill = proportion,
             label = label)) +
  geom_tile() +
  geom_text(aes(color = proportion > .25),
            family = "serif") +
  scale_fill_gradient(low = wes_palette("Moonrise2")[4],
                      high = wes_palette("Moonrise2")[1],
                      limits = c(0, 1)) +
  scale_color_manual(values = wes_palette("Moonrise2")[c(1, 4)]) +
  scale_x_discrete(NULL, position = "top") +
  ylab(NULL) +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks = element_blank(),
        legend.position = "none")
```

```{r}
library(ggdag)

dag_coords <-
  tibble(name = c("G", "D", "A"),
         x = c(1, 2, 3),
         y = c(1, 2, 1))

dagify(D ~ G,
       A ~ D + G,
       coords = dag_coords) %>% 
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_text(color = wes_palette("Moonrise2")[4],
                family = "serif") +
  geom_dag_edges(edge_color = wes_palette("Moonrise2")[4]) +
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL)
```

```{r}
d %>% 
  add_predicted_draws(b11.8) %>% 
  mean_qi() %>% 
  
  ggplot(aes(x = case, y = admit / applications)) +
  geom_pointrange(aes(y = .prediction / applications,
                      ymin = .lower / applications,
                      ymax = .upper / applications),
                  color = wes_palette("Moonrise2")[1],
                  shape = 1, alpha = 1/3) +
  geom_point(color = wes_palette("Moonrise2")[2]) +
  geom_line(aes(group = dept),
            color = wes_palette("Moonrise2")[2]) +
  geom_text(data = text, aes(y = admit, label = dept),
            color = wes_palette("Moonrise2")[2],
            family = "serif") +
  scale_y_continuous("Proportion admitted", limits = 0:1) +
  labs(title = "Posterior validation check",
       subtitle = "Though imperfect, this model is a big improvement") +
  theme(axis.ticks.x = element_blank())
```

```{r}
dag_coords <-
  tibble(name = c("G", "D", "A", "U"),
         x    = c(1, 2, 3, 3),
         y    = c(1, 2, 1, 2))

dagify(D ~ G + U,
       A ~ D + G + U,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_point(x = 3, y = 2, 
             size = 5, color = wes_palette("Moonrise2")[2]) +
  geom_dag_text(color = wes_palette("Moonrise2")[4], family = "serif") +
  geom_dag_edges(edge_color = wes_palette("Moonrise2")[4]) + 
  scale_x_continuous(NULL, breaks = NULL) +
  scale_y_continuous(NULL, breaks = NULL)
```
```{r}
my_upper <- function(data, mapping, ...) {
  
  # get the x and y data to use the other code
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  r  <- unname(cor.test(x, y)$estimate)
  rt <- format(r, digits = 2)[1]
  tt <- as.character(rt)
  
  # plot the cor value
  ggally_text(
    label = tt, 
    mapping = aes(),
    size = 4,
    color = wes_palette("Moonrise2")[4], 
    alpha = 4/5,
    family = "Times") +
    theme_void()
}

my_diag <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) + 
    geom_density(fill = wes_palette("Moonrise2")[2], linewidth = 0) +
    theme_void()
}

my_lower <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) + 
    geom_point(color = wes_palette("Moonrise2")[1], 
               size = 1/10, alpha = 1/10) +
    theme_void()
}
```

```{r}
library(GGally)

as_draws_df(b11.8) %>% 
  select(starts_with("b_")) %>% 
  set_names(c("alpha[male]", "alpha[female]", 
              str_c("delta[", LETTERS[1:6], "]"))) %>% 
  ggpairs(upper = list(continuous = my_upper),
          diag = list(continuous = my_diag),
          lower = list(continuous = my_lower),
          labeller = "label_parsed") +
  labs(title = "Model: b11.8",
       subtitle = "The parameters are strongly correlated.") +
  theme(strip.text = element_text(size = 11))
```

## 11.2 Poisson regression
```{r}
set.seed(11)

tibble(y = rbinom(1e5, 1000, 1/1000)) %>% 
  summarise(y_mean = mean(y),
            y_variance = var(y))
```

Basic Poisson regression model
$$
\begin{align}
y_{i} & \sim \text{Poisson}(\lambda_{i}) \\
\text{log}(\lambda_{i}) & = \alpha + \beta(x_{i} - \bar{x}), 
\end{align}
$$

### 11.2.1 Example: Oceanic tool complexity.
```{r}
data(Kline, package = "rethinking")
d <- Kline
rm(Kline)

d
```

```{r}
d <-
  d %>% 
  mutate(log_pop_std = (log(population) - mean(log(population))) / sd(log(population)),
         cid = contact)
```

Statistical model
$$
\begin{align}
\text{total_tools} & \sim \text{Poisson}(\lambda_{i}) \\
\text{log}(\lambda_{i}) & = \alpha_{\text{cid}[i]} + \beta_{\text{cid}[i]} \text{log_pop_std}_{i} \\
\alpha_{j} & \sim \ ? \\
\beta_{j} & \sim \ ?
\end{align}
$$

```{r}
tibble(x = c(3, 22),
       y = c(.055, .04),
       meanlog = c(0, 3),
       sdlog = c(10, .5)) %>% 
  expand_grid(number = seq(from = 0, to = 100, length.out = 200)) %>% 
  mutate(density = dlnorm(number, meanlog, sdlog),
         group = str_c("alpha%~%Normal(", meanlog, ", ",
                       sdlog, ")")) %>% 
  
  ggplot(aes(fill = group, color = group)) +
  geom_area(aes(x = number, y = density),
            alpha = 3/4, linewidth = 0, position = "identity") +
  geom_text(data = . %>% group_by(group) %>% slice(1),
            aes(x = x, y = y, label = group),
            family = "Times", parse = T, hjust = 0) +
  scale_fill_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_color_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab("mean number of tools") +
  theme(legend.position = "none")
```
```{r}
exp(0 + 10^2 / 2)
```

```{r}
set.seed(11)

rnorm(1e4, 0, 10) %>% exp() %>% mean()
```

```{r}
exp(3 + 0.5^2 / 2)
```

```{r}
tibble(x = c(10, 4),
       y = c(.05, .5),
       mean = c(0, 3),
       sd = c(10, .5)) %>% 
  expand_grid(number = seq(from = -25, to = 25, length.out = 500)) %>% 
  mutate(density = dnorm(number, mean, sd),
         group = str_c("alpha%~%Normal(", mean, ", ", sd, ")")) %>% 
  
  ggplot(aes(fill = group, color = group)) +
  geom_area(aes(x = number, y = density),
            alpha = 3/4, linewidth = 0, position = "identity") +
  geom_text(data = . %>% group_by(group) %>% slice(1),
            aes(x = x, y = y, label = group),
            family = "Times", parse =T, hjust = 0) +
  scale_fill_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_color_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(lambda~scale)) +
  theme(legend.position = "none")
```

```{r}
set.seed(11)

# how many lines would you like?
n <- 100

# simulate and wrangle
tibble(i = 1:n,
       a = rnorm(n, mean = 3, sd = .5)) %>% 
  mutate(`beta%~%Normal(0*', '*10)` = rnorm(n, mean = 0, sd = 10),
         `beta%~%Normal(0*', '*0.2)` = rnorm(n, mean = 0, sd = .2)) %>% 
  pivot_longer(contains("beta"),
               values_to = "b",
               names_to = "prior") %>% 
  expand_grid(x = seq(from = -2, to = 2, length.out = 100)) %>% 
  
  # plot
  ggplot(aes(x = x, y = exp(a + b * x), group = i)) +
  geom_line(linewidth = 1/4, alpha = 2.3, 
            color = wes_palette("Moonrise2")[4]) +
  labs(x = "log population (std)",
       y = "total tools") +
  coord_cartesian(ylim = c(0, 100)) +
  facet_wrap(~ prior, labeller = label_parsed)
```

```{r}
set.seed(11)

prior <-
  tibble(i = 1:n,
         a = rnorm(n, mean = 3, sd = .5),
         b = rnorm(n, mean = 0, sd = .2)) %>% 
  expand_grid(x = seq(from = log(100), to = log(200000),
                      length.out = 100))

# left
p1 <-
  prior %>% 
  ggplot(aes(x = x, y = exp(a + b * x), group = i)) +
  geom_line(linewidth = 1/4, alpha = 2/3,
            color = wes_palette("Moonrise2")[4]) +
  labs(subtitle = expression(beta%~%Normal(0*', '*0.2)),
       x = "log population", y = "total tools") +
  coord_cartesian(xlim = c(log(100), log(200000)),
                  ylim = c(0, 500))

# right 
p2 <-
  prior %>% 
  ggplot(aes(x = exp(x), y = exp(a + b * x), group = i)) +
  geom_line(linewidth = 1/4, alpha = 2/3,
            color = wes_palette("Moonrise2")[4]) +
  labs(subtitle = expression(beta%~%Normal(0*', '*0.2)),
       x = "population",
       y = "total tools") +
  coord_cartesian(xlim = c(100, 200000),
                  ylim = c(0, 500))

# combine
p1 | p2
```

After settling on two priors, the updated model formula is
$$
\begin{align}
y_{i} & \sim \text{Poisson}(\lambda_{i}) \\
\text{log}(\lambda_{i}) & = \alpha + \beta(x_{i} - \bar{x}) \\
\alpha & \sim \text{Normal}(3, 0.5) \\
\beta & \sim \text{Normal}(0, 0.2).
\end{align}
$$

```{r}
# intercept only
b11.9 <-
  brm(data = d,
      family = poisson,
      total_tools ~ 1,
      prior(normal(3, .5), class = Intercept),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.09")

# interaction model
b11.10 <-
  brm(data = d,
      family = poisson,
      bf(total_tools ~ a + b * log_pop_std,
         a + b ~ 0 + cid,
         nl = TRUE),
      prior = c(prior(normal(3, .5), nlpar = a),
                prior(normal(0, .2), nlpar = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.10")
```

```{r}
print(b11.9)
```

```{r}
print(b11.10)
```

```{r}
b11.9 <- add_criterion(b11.9, "loo")
b11.10 <- add_criterion(b11.10, "loo")

loo_compare(b11.9, b11.10, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
model_weights(b11.9, b11.10, weights = "loo") %>% 
  round(digits = 2)
```

```{r}
loo(b11.10) %>% loo::pareto_k_table()
```

```{r}
tibble(culture = d$culture,
       k = b11.10$criteria$loo$diagnostics$pareto_k) %>% 
  arrange(desc(k)) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
cultures <- c("Hawaii", "Tonga", "Trobriand", "Yap")

library(ggrepel)

nd <-
  distinct(d, cid) %>% 
  expand_grid(log_pop_std = seq(from = -4.5, to = 2.5, length.out = 100))

f <- 
  nd %>% 
  add_epred_draws(b11.10) %>% 
  mean_qi()

p1 <-
  f %>%
  ggplot(aes(x = log_pop_std, group = cid, color = cid)) +
  geom_smooth(aes(y = .epred, ymin = .lower, 
                  ymax = .upper, fill = cid),
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(data = bind_cols(d, b11.10$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k),
             alpha = 4/5) +
  geom_text_repel(data = 
                    bind_cols(d, b11.10$criteria$loo$diagnostics) %>% 
                    filter(culture %in% cultures) %>% 
                    mutate(label = str_c(culture, " (", round(pareto_k, digits = 2), ")")),
                  aes(y = total_tools, label = label), 
                  size = 3, seed = 11, color = "black", family = "Times") +
  labs(x = "log population (std)",
       y = "total tools") +
  coord_cartesian(xlim = range(b11.10$data$log_pop_std),
                  ylim = c(0, 80))
```

```{r}
p2 <-
  f %>% 
  mutate(population = exp((log_pop_std * sd(log(d$population))) + mean(log(d$population)))) %>% 
  
  ggplot(aes(x = population, group = cid, color = cid)) +
  geom_smooth(aes(y = .epred, ymin = .lower, ymax = .upper,
                  fill = cid), stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(data = bind_cols(d, b11.10$criteria$loo$diagnostics), 
             aes(y = total_tools, size = pareto_k),
             alpha = 4/5) +
  scale_x_continuous("population",
                     breaks = c(0, 50000, 150000, 250000)) +
  ylab("total tools") +
  coord_cartesian(xlim = range(d$population),
                  ylim = c(0, 80))
```

```{r}
(p1 | p2) &
  scale_fill_manual(values = wes_palette("Moonrise2")[1:2]) &
  scale_color_manual(values = wes_palette("Moonrise2")[1:2]) &
  scale_size(range = c(2, 5)) &
  theme(legend.position = "none")
```

#### 11.2.1.1 Overthinking: Modeling tool innovation.
Scientific model for total_tools is
$$
\text{total_tools} = \frac{\alpha_{\text{cid}[i]}\text{population}^{\beta_{cid}[i]}}{\gamma}.
$$

Poisson likelihood to express the Bayesian model is
$$
\begin{align}
\text{total_tools} & \sim \text{Poisson}(\lambda_{i}) \\
\lambda_{i} & = [\text{exp}(\alpha_{cid[i]})\text{population}_{i}^{\beta_{cid[i]}} / \gamma \\
\alpha_{j} & \sim \text{Normal}(1, 1) \\
\beta_{j} & \sim \text{Exponential}(1) \\
\gamma & \sim \text{Exponential}(1),
\end{align}
$$

```{r}
b11.11 <-
  brm(data = d,
      family = poisson(link = "identity"),
      bf(total_tools ~ exp(a) * population ^ b / g,
         a + b ~ 0 + cid,
         g ~ 1,
         nl = TRUE),
      prior = c(prior(normal(1, 1), nlpar = a),
                prior(exponential(1), nlpar = b, lb = 0),
                prior(exponential(1), nlpar = g, lb = 0)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      control = list(adapt_delta = .85),
      file = "fits/b11.11")
```

```{r}
print(b11.11)
```

```{r}
b11.11 <- add_criterion(b11.11, criterion = "loo",
                        moment_match = T)

loo(b11.11)
```

```{r}
# for the annotation
text <-
  distinct(d, cid) %>% 
  mutate(population = c(210000, 72500),
         total_tools = c(59, 68),
         label = str_c(cid, " contact"))

# redefine the new data
nd <-
  distinct(d, cid) %>% 
  expand_grid(population = seq(from = 0, to = 300000, length.out = 100))

# compute the posterior prediction for lambda
nd %>% 
  add_epred_draws(b11.11) %>% 
  mean_qi(.width = .89) %>% 
  
  # plot!
  ggplot(aes(x = population, group = cid, color = cid)) +
  geom_smooth(aes(y = .epred, ymin = .lower, ymax = .upper,
                  fill = cid), 
              stat = "identity",
              alpha = 1/4, linewidth = 1/2) +
  geom_point(data = bind_cols(d, b11.11$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k),
             alpha = 4/5) +
  geom_text(data = text, aes(y = total_tools, label = label),
            family = "serif") +
  scale_fill_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_color_manual(values = wes_palette("Moonrise2")[1:2]) +
  scale_size(range = c(2 ,5)) +
  scale_x_continuous("population", breaks = c(0, 50000, 150000, 250000)) +
  ylab("total tools") +
  coord_cartesian(xlim = range(d$population),
                  ylim = range(d$total_tools)) +
  theme(legend.position = "none")
```

```{r}
loo_compare(b11.10, b11.11, criterion = "loo") %>% print(simplify = F)
```

```{r}
model_weights(b11.10, b11.11, weights = "loo") %>% round(digits = 3)
```
Comparison of the two models by the Pareto k values.
```{r}
tibble(b11.10 = b11.10$criteria$loo$diagnostics$pareto_k,
       b11.11 = b11.11$criteria$loo$diagnostics$pareto_k) %>% 
  pivot_longer(everything()) %>% 
  
  ggplot(aes(x = value, y = name)) +
  geom_vline(xintercept = c(.5, .7, 1), linetype = 3, color = wes_palette("Moonrise2")[2]) +
  stat_dots(slab_fill = wes_palette("Moonrise2")[1], 
            slab_color = wes_palette("Moonrise2")[1]) + 
  scale_x_continuous(expression(Pareto~italic(k)), breaks = c(.5, .7, 1)) +
  ylab(NULL) +
  coord_cartesian(ylim = c(1.5, 2.4))
```

### 11.2.3 Example: Exposure and the offset.
```{r}
set.seed(11)

num_days <- 30
y <- rpois(num_days, lambda = 1.5)

num_weeks <- 4
y_new <- rpois(num_weeks, lambda = 0.5 * 7)
```


```{r}
(
  d <-
    tibble(y = c(y, y_new),
           days = rep(c(1, 7), times = c(num_days, num_weeks)), # this is the exposure
           monastery = rep(0:1, times = c(num_days, num_weeks))) %>% 
    mutate(log_days = log(days))
)
```

Within the context of the Poisson likelihood, we can decompose $\lambda$ into two parts, $\mu$ (mean) and $\tau$ (exposure), like this:
$$
\begin{align}
y_{i} & \sim Poisson(\lambda_{i}) \\
\text{log}\lambda_{i} & = \text{log} \frac{\mu_{i}}{\tau_{i}} = \text{log} \ \mu_{i} - \text{log} \ \tau_{i}.
\end{align}
$$

You can rewrite the equation if the exposure varies in your data and you still want to model the mean. Here's what that might look like:

$$
\begin{align}
y_{i} & \sim \text{Poisson}(\mu_i) \\
\text{log}\mu_i & = \text{log} \ \tau_i + \alpha + \beta \text{monastery}_i \\
\alpha & \sim \text{Normal}(0, 1) \\
\beta & \sim \text{Normal}(0, 1),
\end{align}
$$

```{r}
b11.12 <-
  brm(data = d,
      family = poisson,
      y ~ 1 + offset(log_days) + monastery,
      prior = c(prior(normal(0, 1), class = Intercept),
                prior(normal(0, 1), class = b)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.12")
```

```{r}
print(b11.12)
```

To get the posterior distributions for average daily outpus for the old and new monasteries, respectively, we'll use the formulas
$$
\begin{align}
\lambda_{old} & = \text{exp}(\alpha) \ \text{and} \\
\lambda_{new} & = \text{exp}(\alpha + \beta_{\text{monestary}}).
\end{align}
$$

```{r}
as_draws_df(b11.12) %>% 
  mutate(lambda_old = exp(b_Intercept),
         lambda_new = exp(b_Intercept + b_monastery)) %>% 
  pivot_longer(contains("lambda")) %>% 
  mutate(name = factor(name, levels = c("lambda_old", "lambda_new"))) %>% 
  group_by(name) %>% 
  mean_hdi(value, .width = .89) %>% 
  mutate_if(is.double, round, digits = 2)
```

## 11.3 Multinomial and categorical models
### 11.3.1 Predictors matched to outcomes.
```{r}
library(rethinking)

# simulate career choices among 500 individuals
n      <- 500           # number of individuals
income <- c(1, 2, 5)    # expected income of each career
score  <- 0.5 * income  # scores for each career, based on income

# next line converts scores to probabilities
p <- softmax(score[1], score[2], score[3])

# now simulate choice
# outcome career holds event type values, not counts
career <- rep(NA, n)  # empty vector of choices for each individual

# sample chosen career for each individual
set.seed(34302)
# sample chosen career for each individual
for(i in 1:n) career[i] <- sample(1:3, size = 1, prob = p)
```

```{r}
print(score)
```

```{r}
# put them in a tibble
d <-
  tibble(career = career) %>% 
  mutate(career_income = ifelse(career == 3, 5, career))

# plot 
d %>%
  ggplot(aes(x = career)) +
  geom_bar(linewidth = 0, fill = wes_palette("Moonrise2")[2])
```

```{r}
d %>% 
  count(career) %>% 
  mutate(percent     = (100 * n / sum(n)),
         probability =        n / sum(n))
```

```{r}
tibble(income = c(1, 2, 5)) %>% 
  mutate(score = 0.5 * income) %>% 
  mutate(p = exp(score) / sum(exp(score)))
```

```{r}
tibble(income        = c(1, 2, 5), 
       some_constant = 11) %>% 
  mutate(score = (0.5 * income) + some_constant) %>% 
  mutate(p = exp(score) / sum(exp(score)))
```

```{r}
# define the model
code_m11.13 <- "
data{
  int N; // number of individuals
  int K; // number of possible careers 
  array[N] int career; // outcome
  vector[K] career_income;
}
parameters{
  vector[K - 1] a; // intercepts
  real<lower=0> b; // association of income with choice
}
model{
  vector[K] p;
  vector[K] s;
  a ~ normal(0, 1);
  b ~ normal(0, 0.5);
  s[1] = a[1] + b * career_income[1]; 
  s[2] = a[2] + b * career_income[2]; 
  s[3] = 0; // pivot
  p = softmax(s);
  career ~ categorical(p);
} 
"
# wrangle the data
dat_list <- 
  list(N = n, 
       K = 3, 
       career = career, 
       career_income = income)

# compile the model
code_m11.13_compiled <- stan_model(model_code = code_m11.13)

# fit the model
m11.13 <- sampling(
  code_m11.13_compiled,
  data = dat_list,
  chains = 4)
```

```{r}
precis(m11.13, depth = 2) %>% round(digits = 2)
```

```{r}
post <- extract.samples(m11.13)

# set up logit scores
s1 <- with(post, a[, 1] + b * income[1])
s2_orig <- with(post, a[, 2] + b * income[2])
s2_new <- with(post, a[, 2] + b * income[2] * 2)

# compute probabilities for original and counterfactual
p_orig <- sapply(1:length(post$b), function(i)
  softmax(c(s1[i], s2_orig[i], 0)))

p_new <- sapply(1:length(post$b), function(i)
  softmax(c(s1[i], s2_new[i], 0)))

# summarize
p_diff <- p_new[2, ] - p_orig[2, ]
precis(p_diff)
```

```{r}
data.frame(s1 = score[3] + s1,
           s2 = score[3] + s2_orig,
           s3 = score[3] + 0) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  mean_qi(value) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
get_prior(data = d,
          family = categorical(link = logit, refcat = 3),
          career ~ 1)
```

```{r}
b11.13io <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      career ~ 1,
      prior = c(prior(normal(0, 1), class = Intercept, dpar = mu1),
                prior(normal(0, 1), class = Intercept, dpar = mu2)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.13io")
```

```{r}
print(b11.13io)
```
```{r}
tibble(income = c(1, 2, 5)) %>% 
  mutate(score = 0.5 * income) %>% 
  mutate(rescaled_score = score - 2.5)
```

```{r}
fitted(b11.13io) %>% str()
```

```{r}
fitted(b11.13io)[1, , ] %>% round(digits = 2)
```

```{r}
fitted(b11.13io)[1, , ] %>% 
  round(digits = 2) %>% 
  t()
```

```{r}
tibble(income = c(1, 2, 5)) %>% 
  mutate(score = 0.5 * income) %>% 
  mutate(p = exp(score) / sum(exp(score)))
```

```{r}
as_draws_df(b11.13io) %>% 
  mutate(b_mu3_Intercept = 0) %>% 
  mutate(p1 = exp(b_mu1_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept)),
         p2 = exp(b_mu2_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept)),
         p3 = exp(b_mu3_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept))) %>% 
 pivot_longer(p1:p3) %>% 
  group_by(name) %>% 
  mean_qi(value) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
# verbose syntax
b11.13io_verbose <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         mu1 ~ 1,
         mu2 ~ 1),
      prior = c(prior(normal(0, 1), class = Intercept, dpar = mu1),
                prior(normal(0, 1), class = Intercept, dpar = mu2)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.13io_verbose")

# nonlinear syntax
b11.13io_nonlinear <-
  brm(data = d, 
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1),
         nlf(mu2 ~ a2),
         a1 + a2 ~ 1),
      prior = c(prior(normal(0, 1), class = b, nlpar = a1),
                prior(normal(0, 1), class = b, nlpar = a2)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.13io_nonlinear")
```

```{r}
crossing(b = factor(c("b1 & b2", "b"),
                    levels = c("b1 & b2", "b")),
         lb = factor(c("NA", 0), levels = c("NA", 0))) %>% 
  mutate(fit = str_c("b11.13", letters[1:n()])) %>% 
  select(fit, everything()) %>% 
  
  flextable() %>% 
  width(width = 1.25)
```

```{r}
b11.13a <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1 + b1 * 1),
         nlf(mu2 ~ a2 + b2 * 2),
         a1 + a2 + b1 + b2 ~ 1),
      prior = c(prior(normal(0, 1), class = b, nlpar = a1),
                prior(normal(0, 1), class = b, nlpar = a2),
                prior(normal(0, .5), class = b, nlpar = b1),
                prior(normal(0, .5), class = b, nlpar = b2)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.13a")

b11.13b <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1 + b1 * 1),
         nlf(mu2 ~ a2 + b2 * 2),
         a1 + a2 + b1 + b2 ~ 1),
      prior = c(prior(normal(0, 1), class = b, nlpar = a1),
                prior(normal(0, 1), class = b, nlpar = a2),
                prior(normal(0, .5), class = b, nlpar = b1, lb = 0),
                prior(normal(0, .5), class = b, nlpar = b2, lb = 0)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.13b")

b11.13c <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1 + b * 1),
         nlf(mu2 ~ a2 + b * 2),
         a1 + a2 + b ~ 1),
      prior = c(prior(normal(0, 1), class = b, nlpar = a1),
                prior(normal(0, 1), class = b, nlpar = a2),
                prior(normal(0, .5), class = b, nlpar = b)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      file = "fits/b11.13c")

b11.13d <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1 + b * 1),
         nlf(mu2 ~ a2 + b * 2),
         a1 + a2 + b ~ 1),
      prior = c(prior(normal(0, 1), class = b, nlpar = a1),
                prior(normal(0, 1), class = b, nlpar = a2),
                prior(normal(0, .5), class = b, nlpar = b, lb = 0)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 11,
      control = list(adapt_delta = .99),
      file = "fits/b11.13d")
```

```{r}
tibble(fit = str_c("b11.13", letters[1:4])) %>% 
  mutate(fixef = purrr::map(fit, ~ get(.) %>% 
                              fixef() %>% 
                              data.frame() %>% 
                              rownames_to_column("parameter"))) %>% 
  unnest(fixef) %>% 
  mutate(parameter = str_remove(parameter, "_Intercept"),
         fit = factor(fit, levels = str_c("b11.13", letters[4:1]))) %>% 
  
  ggplot(aes(x = Estimate, xmin = Q2.5, xmax = Q97.5, y = fit)) +
  geom_vline(xintercept = 0, color = wes_palette("Moonrise2")[3]) +
  geom_pointrange(fatten = 3/2, color = wes_palette("Moonrise2")[4]) +
  ylab(NULL) +
  theme(axis.ticks.y = element_blank(),
        panel.background = element_rect(fill = alpha("white", 1/8), linewidth = 0)) +
  facet_wrap(~ parameter, nrow = 1)
```

```{r}
precis(m11.13, depth = 2)
```

```{r}
fixef(b11.13d) %>% round(digits = 2)
```
```{r}
b11.13a <- add_criterion(b11.13a, "loo")
b11.13b <- add_criterion(b11.13b, "loo")
b11.13c <- add_criterion(b11.13c, "loo")
b11.13d <- add_criterion(b11.13d, "loo")

loo_compare(b11.13a, b11.13b, b11.13c, b11.13d, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
model_weights(b11.13a, b11.13b, b11.13c, b11.13d, weights = "loo") %>% 
  round(digits = 2)
```

```{r}
b11.13io <- add_criterion(b11.13io, "loo")

loo_compare(b11.13io, b11.13a, b11.13b, b11.13c, b11.13d, criterion = "loo") %>% 
  print(simplify = F)
```

```{r}
model_weights(b11.13io, b11.13a, b11.13b, b11.13c, b11.13d, weights = "loo") %>% 
  round(digits = 2)
```

```{r}
as_draws_df(b11.13d) %>% 
  transmute(s1      = b_a1_Intercept + b_b_Intercept * income[1],
            s2_orig = b_a2_Intercept + b_b_Intercept * income[2],
            s2_new  = b_a2_Intercept + b_b_Intercept * income[2] * 2) %>% 
  mutate(p_orig = purrr::map2_dbl(s1, s2_orig, ~softmax(.x, .y, 0)[2]),
         p_new  = purrr::map2_dbl(s1, s2_new, ~softmax(.x, .y, 0)[2])) %>% 
  mutate(p_diff = p_new - p_orig) %>% 
  mean_qi(p_diff) %>% 
  mutate_if(is.double, round, digits = 2)
```

### 11.3.2 Predictors matched to observations.
```{r}
n <- 500
set.seed(11)

# simulate family incomes for each individual
family_income <- runif(n)

# assign a unique coefficient for each type of event
b <- c(-2, 0, 2)
career <- rep(NA, n)
for (i in 1:n) {
  score <- 0.5 * (1:3) + b * family_income[i]
  p <- softmax(score[1], score[2], score[3])
  career[i] <- sample(1:3, size = 1, prob = p)
}
```

```{r}
# put the data in a tibble
d <-
  tibble(career = career) %>% 
  mutate(family_income = family_income)

p1 <-
  d %>% 
  mutate(career = as.factor(career)) %>% 
  
  ggplot(aes(x = family_income, fill = career)) +
  geom_density(linewidth = 0, alpha = 3/4) +
  scale_fill_manual(values = wes_palette("Moonrise2")[c(4, 2, 1)]) +
  theme(legend.position = "none")

p2 <-
  d %>% 
  mutate(career = as.factor(career)) %>% 
  
  mutate(fi = santoku::chop_width(family_income, width = .1, start = 0, labels = 1:10)) %>% 
  count(fi, career) %>% 
  group_by(fi) %>% 
  mutate(proportion = n / sum(n)) %>% 
  mutate(f = as.double(fi)) %>% 
  
  ggplot(aes(x = (f - 1) / 9, y = proportion, fill = career)) +
  geom_area() +
  scale_fill_manual(values = wes_palette("Moonrise2")[c(4, 2, 1)]) +
  xlab("family_income, descritized")

p1 + p2
```

```{r}
b11.14 <-
  brm(data = d,
      family = categorical(link = logit, refcat = 3),
      bf(career ~ 1,
         nlf(mu1 ~ a1 + b1 * family_income),
         nlf(mu2 ~ a2 + b2 * family_income),
         a1 + a2 + b1 + b2 ~ 1),
      prior = c(prior(normal(0, 1.5), class = b, nlpar = a1),
                prior(normal(0, 1.5), class = b, nlpar = a2),
                prior(normal(0, 1), class = b, nlpar = b1),
                prior(normal(0, 1), class = b, nlpar = b2)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 11,
      file = "fits/b11.14")
```

```{r}
print(b11.14)
```

```{r}
b11.14 <- add_criterion(b11.14, "loo")

loo(b11.14)
```

```{r}
nd <- tibble(family_income = seq(from = 0, to = 1, length.out = 60))

f <-
  fitted(b11.14, newdata = nd)
```

```{r}
# wrangle
rbind(f[, , 1],
      f[, , 2],
      f[, , 3]) %>% 
  data.frame() %>% 
  bind_cols(expand_grid(career = 1:3, nd)) %>%
  mutate(career = str_c("career: ", career)) %>% 
  
  # plot
  ggplot(aes(x = family_income, y = Estimate,
             ymin = Q2.5, ymax = Q97.5,
             fill = career, color = career)) +
  geom_ribbon(alpha = 2/3, linewidth = 0) +
  geom_line(linewidth = 3/4) +
  scale_fill_manual(values = wes_palette("Moonrise2")[c(4, 2, 1)]) +
  scale_color_manual(values = wes_palette("Moonrise2")[c(4, 2, 1)]) +
  scale_x_continuous(breaks = 0:2 / 2) +
  scale_y_continuous("probability", limits = c(0, 1),
                     breaks = 0:3 / 3,
                     labels = c("0", ".33", ".67", "1")) +
  theme(axis.text.y = element_text(hjust = 0),
        legend.position = "none") +
  facet_wrap(~ career)
```

