---
title: "Chapter 3 BRMS"
format: html
---

# 3 Sampling the Imaginary

If you would like to know the probability someone is a vampire given they test positive to the blood-based vampire test, you compute

$$
\text{Pr}(\text{vampire}|\text{positive}) = \frac{\text{Pr}(\text{positive}|\text{vampire})\text{Pr}(\text{vampire})}{\text{Pr}(\text{positive})}.
$$
We'll do so within a tibble.
```{r}
library(tidyverse)

tibble(pr_positive_vampire = .95,
       pr_positive_mortal = .01,
       pr_vampire = .001) %>% 
  mutate(pr_positive = pr_positive_vampire * pr_vampire + pr_positive_mortal * (1 - pr_vampire)) %>% 
  mutate(pr_vampire_positive = pr_positive_vampire * pr_vampire / pr_positive) %>% 
  glimpse()
```

Here's the other way of tackling the vampire problem, this time using the frequency format.
```{r}
tibble(pr_vampire = 10/10000,
       pr_positive_vampire = 95/100,
       pr_positive_mortal = 999/99900) %>% 
  mutate(pr_positive = 95 + 999) %>% 
  mutate(pr_vampire_positive = pr_positive_vampire * 100 / pr_positive) %>% 
  glimpse()
```

## 3.1 Sampling from a grid-approximate posterior
```{r}
# how many grid points would you like
n <- 1000
n_success <- 6
n_trials <- 9

(
  d <-
    tibble(p_grid = seq(from = 0, to = 1, length.out = n),
           # note we're still using a flat uniform prior
           prior = 1) %>% 
    mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
    mutate(posterior = (likelihood * prior) / sum(likelihood * prior))
)
```

```{r}
# how many samples would you like?
n_samples <- 1e4

# make it reproducible
set.seed(3)

samples <-
  d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)

glimpse(samples)
```
Now we can plot the left panel of Figure 3.1 with geom_point(). But before we do, we'll need to add a variable numbering the samples
```{r}
samples %>% 
  mutate(sample_number = 1:n()) %>% 
  
  ggplot(aes(x = sample_number, y = p_grid)) +
  geom_point(alpha = 1/10) +
  scale_y_continuous("proportion of water (p)", limits = c(0, 1)) +
  xlab("sample number")
```

We'll make the density in right panel with geom_density()
```{r}
samples %>% 
  ggplot(aes(x = p_grid)) +
  geom_density(fill = "black") +
  scale_x_continuous("proportion of water (p)", limits = c(0, 1))
```

Increase number of samples to 1e6
```{r}
set.seed(3)

d %>% 
  slice_sample(n = 1e6, weight_by = posterior, replace = T) %>% 
  ggplot(aes(x = p_grid)) +
  geom_density(fill = "black") +
  scale_x_continuous("proportion of water (p)", limits = c(0, 1))
```

## 3.2 Sampling to summarize
### 3.2.1 Intervals of defined boundaries

To get the proportion of water less than some value of p_grid within the tidyverse, you might first filter by that value and then take the sum() within summarise()
```{r}
d %>% 
  filter(p_grid < .5) %>% 
  summarise(sum = sum(posterior))
```

If what you want is a frequency based on filtering by samples, then you might use n() within summarise
```{r}
samples %>% 
  filter(p_grid < .5) %>% 
  summarise(sum = n() / n_samples)
```

A more explicit approach for the same computation is to follow up count() with mutate()
```{r}
samples %>% 
  count(p_grid < .5) %>% 
  mutate(probability = n / sum(n))
```

An even trickier approach for the same is to insert the logical statement p_grid < .5 within the mean function
```{r}
samples %>% 
  summarise(sum = mean(p_grid < .5))
```

To determine the posterior probability between 0.5 and 0.75, you can use & within filter()
```{r}
samples %>% 
  filter(p_grid > .5 & p_grid < .75) %>% 
  summarise(sum = n() / n_samples,
            percent = n() / n_samples * 100)
```

```{r}
# upper left panel
p1 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  geom_area(data = d %>% filter(p_grid < .5)) +
  labs(x = "proportion of water (p)",
       y = "density")

# upper right panel
p2 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  # note this next line is the only diffrence in code from the last plot
  geom_area(data = d %>% filter(p_grid < .75 & p_grid > .5)) +
  labs(x = "proportion of water (p)",
       y = "density")

# combine
library(patchwork)
p1 + p2
```

Since we saved our p_grid samples within the well-named samples tibble, we'll have to index with $ within quantile
```{r}
(q_80 <- quantile(samples$p_grid, prob = .8))
```

Alternative approach
```{r}
samples %>% 
  pull(p_grid) %>% 
  quantile(prob = .8)
```

We might also use quantile() within summarise()
```{r}
samples %>% 
  summarise('80th percentile' = quantile(p_grid, p = .8))
```

Here's the summarise() approach with two probabilities
```{r}
samples %>% 
  summarise('10th percentile' = quantile(p_grid, p = .1),
            '90th percentile' = quantile(p_grid, p = .9))
```
```{r}
(q_10_and_90 <- quantile(samples$p_grid, prob = c(.1, .9)))
```

```{r}
# lower left panel
p1 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  geom_area(data = d %>% filter(p_grid < q_80)) +
  annotate(geom = "text",
           x = .25, y = .0025,
           label = "lower 80%") +
  labs(x = "proportion of water (p)",
       y = "density")

# lower right panel
p2 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  geom_area(data = d %>% filter(p_grid > q_10_and_90[1] & p_grid < q_10_and_90[2])) +
  annotate(geom = "text",
           x = .25, y = .0025,
           label = "middle 80%") +
  labs(x = "proportion of water (p)",
       y = "density")

# combine 
p1 + p2
```

```{r}
# here we update the dbinom() parameters
n_success <- 3
n_trials <- 3

# update d
d <-
  d %>% 
  mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))

# make the next part reproducible
set.seed(3)

# here's our new samples tibble
(
  samples <-
    d %>% 
    slice_sample(n = n_samples, weight_by = posterior, replace = T)
)
```

The rethinking::PI() function works like a nice shorthand for quantile()
```{r}
quantile(samples$p_grid, prob = c(.25, .75))
```

```{r}
rethinking::PI(samples$p_grid, prob = .5)
```

```{r}
library(tidybayes)

median_qi(samples$p_grid, .width = .5)
```

```{r}
median_qi(samples$p_grid, .width = c(.5, .8, .99))
```

```{r}
rethinking::HPDI(samples$p_grid, prob = .5)
```

tidybayes also support HPDIs
```{r}
mode_hdi(samples$p_grid, .width = .5)
```

If you want are the intervals without the measure of central tendency or all that other techinical information, tidybayes also offers the handy qi() and hdi() functions
```{r}
qi(samples$p_grid, .width = .5)
```

```{r}
hdi(samples$p_grid, .width = .5)
```

Make Figure 3.3
```{r}
# lower left panel
p1 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  # check out our sweet `qi()` indexing
  geom_area(data = d %>% 
              filter(p_grid > qi(samples$p_grid, .width = .5)[1] & p_grid < qi(samples$p_grid, .width = .5)[2]),
            fill = "grey75") +
  geom_line() +
  labs(subtitle = "50% Percentile Interval",
       x = "proportion of water (p)",
       y = "density")

# lower right panel
p2 <-
  d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_area(data = d %>% 
              filter(p_grid > hdi(samples$p_grid, .width = .5)[1] & p_grid < hdi(samples$p_grid, .width = .5)[2]),
            fill = "grey75") +
  geom_line() + 
  labs(subtitle = "50% HPDI",
       x = "proportion of water (p)",
       y = "density")

# combine!
p1 | p2
```

Try using PI and HPDI in priobability masses that are symmetrical
```{r}
# "six waters in nine tosses"
n_success <- 6
n_trials <- 9

new_d <-
  d %>% 
  mutate(likelihood = dbinom(n_success, n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))
set.seed(3)

new_samples <-
  new_d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)
```

Here are the intervals by .width and type of .interval
```{r}
bind_rows(mean_hdi(new_samples$p_grid, .width = c(.8, .95)),
          mean_qi(new_samples$p_grid, .width = c(.8, .95))) %>% 
  select(.width, .interval, ymin:ymax) %>% 
  arrange(.width) %>% 
  mutate_if(is.double, round, digits = 2)
```
Take a look at the posterior
```{r}
new_d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  labs(subtitle = "Six waters in nine tosses made for a more symmetrical posterior",
       x = "proportion of water (p)",
       y = "density")
```

### 3.2.3 Point estimates.
```{r}
d %>% 
  arrange(desc(posterior)) %>% 
  round(digits = 5)
```

```{r}
d %>% 
  arrange(desc(posterior)) %>% 
  slice(1)
```

We can get the mode with mode_hdi() or mode_qi()
```{r}
samples %>% mode_hdi(p_grid)
```

```{r}
samples %>% mode_qi(p_grid)
```

If you want only the mode you can use tidybayes::Mode()
```{r}
Mode(samples$p_grid)
```

Median and means are typical measures of central tendency
```{r}
samples %>% 
  summarise(mean = mean(p_grid),
            median = median(p_grid)) %>% 
  round(digits = 3)
```

We can inspect the three types of point estimate in the left panel of Figure 3.4. First, we'll bundle the three point estimates together in a tibble.
```{r}
(
  point_estimates <-
    bind_rows(samples %>% mean_qi(p_grid),
              samples %>% median_qi(p_grid),
              samples %>% mode_qi(p_grid)) %>% 
    select(p_grid, .point) %>% 
    # these last two columns will help us annotate
    mutate(x = p_grid + c(-.03, .03, -.03),
           y = c(.0005, .0012, .002)) %>% 
    mutate_if(is.double, round, digits = 4)
)
```

Now plot
```{r}
d %>% 
  ggplot(aes(x = p_grid)) +
  geom_area(aes(y = posterior),
            fill = "grey75") +
  geom_vline(xintercept = point_estimates$p_grid) +
  geom_text(data = point_estimates,
            aes(x = x, y = y, label = .point),
            angle = 90) +
  labs(x = "proportion of water (p)",
       y = "density") +
  theme(panel.grid = element_blank())
```

```{r}
d %>% 
  summarise('expected loss' = sum(posterior * abs(0.5 - p_grid)))
```

```{r}
make_loss <- function(our_d) {
  d %>% 
    mutate(loss = posterior * abs(our_d - p_grid)) %>% 
    summarise(weighted_average_loss = sum(loss))
}

(
  l <-
    d %>% 
    select(p_grid) %>% 
    rename(decision = p_grid) %>% 
    mutate(weighted_average_loss = purrr::map(decision, make_loss)) %>% 
    unnest(weighted_average_loss)
)
```

```{r}
# this will help us find the x and y coordinates for the minimum value
min_loss <-
  l %>% 
  filter(weighted_average_loss == min(weighted_average_loss)) %>% 
  as.numeric()

# the plot
l %>% 
  ggplot(aes(x = decision, y = weighted_average_loss)) +
  geom_area(fill = "grey75") +
  geom_vline(xintercept = min_loss[1], color = "white", linetype = 3) +
  geom_hline(yintercept = min_loss[2], color = "white", linetype = 3) +
  ylab("expected proportional loss") +
  theme(panel.grid = element_blank())
```

```{r}
samples %>% 
  summarise(posterior_median = median(p_grid))
```

The quadratic loss (d - p)^2 suggests we should use the mean instead. Let's investigate
```{r}
# amend our loss function
make_loss <- function(our_d){
  d %>% 
    mutate(loss = posterior * (our_d - p_grid)^2) %>% 
    summarise(weighted_average_loss = sum(loss))
}

# remake our l data
l <-
  d %>% 
  select(p_grid) %>% 
  rename(decision = p_grid) %>% 
  mutate(weighted_average_loss = purrr::map(decision, make_loss)) %>% 
  unnest(weighted_average_loss)

# update to the new minimum loss coordinates
min_loss <-
  l %>% 
  filter(weighted_average_loss == min(weighted_average_loss)) %>% 
  as.numeric()

# update the plot
l %>% 
  ggplot(aes(x = decision, y = weighted_average_loss)) +
  geom_area(fill = "grey75") +
  geom_vline(xintercept = min_loss[1], color = "white", linetype = 3) +
  geom_hline(yintercept = min_loss[2], color = "white", linetype = 3) +
  ylab("expected proportional loss") +
  theme(panel.grid = element_blank())
```

Based on quadratic loss (d - p)^2, the exact minimum value is 0.8008008. Within sampling erorr, this is the posterior mean of our samples.
```{r}
samples %>% 
  summarise(posterior_mean = mean(p_grid))
```

## 3.3 Sampling to simulate prediction
McElerath's five good reasons for simulation were
1. Model design
2. Model checking
3. Software validation
4. Research design
5. Forecasting

### 3.3.1 Dummy data
Dummy data for the globe tossing model arise from the binomial likelihood. If you let w be a count of water and n be the number of tosses, the binomial likelihood is
$$
\text{Pr}(w|n, p) = \frac{n!}{w!(n - w)!}p^{w}(1 - p)^{n-w}.
$$
Letting n = 2, p(w) = .7, and $w_{observed) = 0 \text{through} 2$, the densities are:
```{r}
tibble(n = 2,
       `p(w)` = .7,
       w = 0:2) %>% 
  mutate(density = dbinom(w, size = n, prob = `p(w)`))
```

```{r}
set.seed(3)

rbinom(1, size = 2, prob = .7)
```

Here are ten reproducible draws.
```{r}
set.seed(3)

rbinom(10, size = 2, prob = .7)
```

Now generate 100,000 reproducible dummy observations
```{r}
# how many would you like?
n_draws <- 1e5

set.seed(3)

d <- tibble(draws = rbinom(n_draws, size = 2, prob = .7))

d %>% 
  count(draws) %>% 
  mutate(proportion = n / nrow(d))
```

Here's the simulation updated so n = 9, which we plot in our version of Figure 3.5
```{r}
set.seed(3)
d <- tibble(draws = rbinom(n_draws, size = 9, prob = .7))

# the histogram
d %>% 
  ggplot(aes(x = draws)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  scale_x_continuous("dummy water count", breaks = 0:4 * 2) +
  ylab("frequency") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(panel.grid = element_blank())
```

Let's play around with different values of size and prob
```{r}
n_draws <- 1e5

simulate_binom <- function(n, probability){
  set.seed(3)
  rbinom(n_draws, n, prob = probability)
}

d <-
  crossing(n = c(3, 6, 9),
           probability = c(.3, .6, .9)) %>% 
  mutate(draws = map2(n, probability, simulate_binom)) %>% 
  ungroup() %>% 
  mutate(n = str_c("n = ", n),
         probability = str_c("p = ", probability)) %>% 
  unnest(draws)

head(d)
```

Plot the simulation results
```{r}
d %>% 
  ggplot(aes(x = draws)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", linwidth = 1/10) +
  scale_x_continuous("dummy water count", breaks = 0:4 * 2) +
  ylab("frequency") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(panel.grid = element_blank()) +
  facet_grid(n ~ probability)
```

### 3.3.2 Model checking.
If you're new to applied statistics, you might be surprised how often mistakes arise.

#### 3.3.2.1 Did the software work?
#### 3.3.2.2 Is the model adequate?
```{r}
# how many grid points would you like?
n < -1001
n_success <- 6
n_trials <- 9

(
  d <-
    tibble(p_grid = seq(from = 0, to = 1, length.out = n),
           # note we're still using a flat uniform prior
           prior = 1) %>% 
    mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
    mutate(posterior = (likelihood * prior) / sum(likelihood * prior))
)
```

We can make our version of the top of Figure 3.6 with a little tricky filter in
```{r}
d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_area(color = "grey67", fill = "grey67") +
  geom_segment(data = . %>% 
                 filter(p_grid %in% c(seq(from = .1, to = .9, by = .1), 3 / 10)),
               aes(xend = p_grid, yend = 0, linewidth = posterior),
               color = "grey33", show.legend = F) +
  geom_point(data = . %>% 
               filter(p_grid %in% c(seq(from = .1, to = .9, by = .1), 3 / 10))) +
  annotate(geom = "text",
           x = .08, y = .0025,
           label = "Posterior probability") +
  scale_linewidth_continuous(range = c(0, 1)) +
  scale_x_continuous("probability of water", breaks = 0:10 / 10) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

Wranging before making middle panel plot of Figure 3.6
```{r}
n_draws <- 1e5

simulate_binom <- function(probability){
  set.seed(3)
  rbinom(n_draws, size = 9, prob = probability)
}

d_small <-
  tibble(probability = seq(from = .1, to = .9, by = .1)) %>% 
  mutate(draws = purrr::map(probability, simulate_binom)) %>% 
  unnest(draws) %>% 
  mutate(label = str_c("p = ", probability))

head(d_small)
```

We're ready to plot
```{r}
d_small %>% 
  ggplot(aes(x = draws)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  scale_x_continuous(NULL, breaks = 0:3 * 3) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "Sampling distributions") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ label, ncol = 9)
```

Make bottom Figure 3.6
```{r}
# how many samples would you like?
n_samples <- 1e4

# make it reproducible
set.seed(3)

samples <-
  d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T) %>% 
  mutate(w = purrr::map_dbl(p_grid, rbinom, n = 1, size = 9))

glimpse(samples)
```

Histogram
```{r}
samples %>% 
  ggplot(aes(x = w)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  ggtitle("Posterior predictive distribution") +
  scale_x_continuous("number of water samples", breaks = 0:3 * 3) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 9), ylim = c(0, 3000)) +
  theme(panel.grid = element_blank())
```

```{r}
rbinom(10, size = 9, prob = .6)
```

```{r}
rbinom(9, size = 1, prob = .6)
```

Simulate the disaggregated version of the numerals returned by rbinom()
```{r}
# make it reproducible
set.seed(3)

samples <-
  samples %>% 
  mutate(iter = 1:n(),
         draws = purrr::map(p_grid, rbinom, n = 9, size = 1)) %>% 
  unnest(draws)

glimpse(samples)
```
Now we have to count the longest sequences
```{r}
tosses <- c("w", "l", "w", "w", "w", "l", "w", "l", "w")
```

```{r}
rle(tosses)
```

For our purposes, we're interested in length. That tells us the length of each sequences of the same value. Use max function to get max length
```{r}
rle(tosses)$lengths %>% max()
```

Let's apply to the data and plot
```{r}
samples %>% 
  group_by(iter) %>% 
  summarise(longest_run_length = rle(draws)$lengths %>% max()) %>% 
  
  ggplot(aes(x = longest_run_length)) +
  geom_histogram(aes(fill = longest_run_length == 3),
                 binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  scale_fill_viridis_d(option = "D", end = .9) +
  scale_x_continuous("longest run length", breaks = 0:3 * 3) +
  ylab("frequency") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

```{r}
samples %>% 
  group_by(iter) %>% 
  summarise(longest_run_length = rle(draws)$lengths %>% length()) %>% 
  
  ggplot(aes(x = longest_run_length)) +
  geom_histogram(aes(fill = longest_run_length == 7),
                 binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  scale_fill_viridis_d(option = "D", end = .9) +
  scale_x_continuous("number of switches", breaks = 0:3 * 3) +
  ylab("frequency") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

## 3.4 Let's practice with brms
```{r}
library(brms)
```
```{r}
b3.1 <-
  brm(data = list(w = 6),
      family = binomial(link = "identity"),
      w | trials(9) ~ 0 + Intercept,
      # this is a flat prior
      prior(beta(1, 1), class = b, lb = 0, ub = 1),
      iter = 5000, warmup = 1000,
      seed = 3,
      file = "fits/b03.01")
```

```{r}
posterior_summary(b3.1)["b_Intercept",] %>% 
  round(digits = 2)
```

```{r}
f <-
  fitted(b3.1,
         summary = F,
         scale = "linear") %>% 
  data.frame() %>% 
  set_names("p")

glimpse(f)
```

By default, we have a generically named vector of 4,000 samples.
```{r}
f %>% 
  ggplot(aes(x = p)) +
  geom_density(fill = "grey50", color = "grey50") +
  annotate(geom = "text", x = .08, y = 2.5,
           label = "Posterior probability") +
  scale_x_continuous("probability of water", 
                     breaks = c(0, 0.5, 1),
                     limits = 0:1) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

```{r}
set.seed(3)

f <-
  f %>% 
  mutate(w = rbinom(n(), size = n_trials, prob = p))

f %>% 
  ggplot(aes(x = w)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", linewidth = 1/10) +
  scale_x_continuous("number of water samples", breaks = 0:3 * 3) +
  scale_y_continuous(NULL, breaks = NULL, limits = c(0, 5000)) +
  ggtitle("Posterior predictive distribution") +
  coord_cartesian(xlim = c(0, 9)) +
  theme(panel.grid = element_blank())
```

